### 2016-09-23 paper/ms.tex
latex stuff

\documentclass[manuscript, letterpaper]{aastex6}

% to-do list
% ----------

% style notes
% -----------
% - This file generates by Makefile; don't be typing ``pdflatex'' or some bullshit.
% - Line break between sentences to make the git diffs readable.
% - Use \, as a multiply operator.
% - Reserve () for function arguments; use [] or {} for outer shit.
% - Use \sectionname not Section, \figname not Figure, \documentname not Article or Paper or paper.

\include{gitstuff}
\include{aastexmods}

% packages
\definecolor{cbblue}{HTML}{3182bd}
\usepackage{microtype}  % ALWAYS!
\usepackage{amsmath}
\hypersetup{backref,breaklinks,colorlinks,urlcolor=cbblue,linkcolor=cbblue,citecolor=black}

% define macros for text
\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\acronym}[1]{{\small{#1}}}
\newcommand{\gaia}{\project{Gaia}}
\newcommand{\documentname}{\textsl{Article}}
\newcommand{\sectionname}{Section}
\newcommand{\figname}{Figure}

% define macros for math
\newcommand{\given}{\,|\,}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\transpose}[1]{{#1}^{\mathsf{T}}}
\newcommand{\inv}[1]{{#1}^{-1}}
\newcommand{\bs}[1]{\boldsymbol{#1}}

% TODO
\newcommand{\todo}[1]{{\color{red}TODO: #1}}

\begin{document}\sloppy\sloppypar\raggedbottom\frenchspacing % trust me

\title{}
\author{}

% Affiliations
% \newcommand{\pu}{1}
% \newcommand{\adrn}{2}
% \newcommand{\ccpp}{3}
% \newcommand{\mpia}{4}
% \newcommand{\uw}{5}
% \newcommand{\sagan}{6}

% \altaffiltext{\pu}{Department of Astrophysical Sciences,
%                    Princeton University, Princeton, NJ 08544, USA}
% \altaffiltext{\adrn}{To whom correspondence should be addressed:
%                      \texttt{adrn@princeton.edu}}
% \altaffiltext{\ccpp}{Center for Cosmology and Particle Physics,
%                      Department of Physics,
%                      New York University, 4 Washington Place,
%                      New York, NY 10003, USA}
% \altaffiltext{\mpia}{Max-Planck-Institut f\"ur Astronomie,
%                      K\"onigstuhl 17, D-69117 Heidelberg, Germany}
% \altaffiltext{\uw}{Astronomy Department, University of Washington,
%                    Seattle, WA 98195, USA}
% \altaffiltext{\sagan}{Sagan Fellow}

\begin{abstract}
% Context
% Aims
% Methods
% Results
\end{abstract}

\keywords{
  methods: data analysis
  ---
  methods: statistical
}

\section{Introduction} \label{sec:intro}

\acknowledgements

This research was partially supported by the \acronym{NSF} (grants
  \acronym{IIS-1124794}, \acronym{AST-1312863}, \acronym{AST-1517237}),
  \acronym{NASA} (grant \acronym{NNX12AI50G}),
  and the Moore-Sloan Data Science Environment at \acronym{NYU}. The data
analysis presented in this article was partially performed on computational
resources supported by the Princeton Institute for Computational Science and
Engineering (PICSciE) and the Office of Information Technology's High
Performance Computing Center and Visualization Laboratory at Princeton
University.

\software{The code used in this project is available from
\url{https://github.com/smoh/gaia-wide-binaries} under the MIT open-source
software license. This version was generated at git commit
\texttt{\githash\,(\gitdate)}.
This research additionally utilized:
    \texttt{Astropy} (\citealt{Astropy-Collaboration:2013}),
    %\texttt{emcee} (\citealt{Foreman-Mackey:2013}),
    \texttt{IPython} (\citealt{Perez:2007}),
    \texttt{matplotlib} (\citealt{Hunter:2007}),
    and \texttt{numpy} (\citealt{Van-der-Walt:2011}).}

% \facility{\sdssiii, \apogee}

\bibliographystyle{aasjournal}
\bibliography{refs}

\end{document}

### 2016-09-23 paper/ms.tex
add equations decribing the FML ratio

@@
 \newcommand{\dd}{\mathrm{d}}
-\newcommand{\transpose}[1]{{#1}^{\mathsf{T}}}
+\newcommand{\transp}[1]{{#1}^{\mathsf{T}}}
 \newcommand{\inv}[1]{{#1}^{-1}}
 \newcommand{\bs}[1]{\boldsymbol{#1}}
+\newcommand{\vperp}{\bs{v}^\perp}
+\newcommand{\propm}{\bs{\mu}}
+\newcommand{\matrx}[1]{\mathbf{#1}}
+\newcommand{\kms}{\rm km~s^{-1}}
 
@@
 
-\title{}
-\author{}
+\title{Wide binaries in Gaia DR1}
+\author{People}
 
@@
 \begin{abstract}
+Blerg.
 % Context
@@
 \section{Introduction} \label{sec:intro}
+
+\section{Methods} \label{sec:methods}
+
+We want to evaluate the likelihood ratio
+\begin{equation}
+  \frac{p(\propm_1, \propm_2)_{\vperp_1 = \vperp_2}}
+  {p(\propm_1, \propm_2)_{\vperp_1 \neq \vperp_2}}
+\end{equation}
+where
+\begin{multline}
+  p(\propm_1, \propm_2)_{\vperp_1 = \vperp_2} =
+    \int \, \dd d_1 \, \dd d_2 \, \dd \vperp \,
+    p(\propm_1 \given d_1, \vperp, \matrx{C}_{\mu,1}) \,
+    p(\propm_2 \given d_2, \vperp, \matrx{C}_{\mu,2}) \\
+    p(d_1 \given \varpi_1, \sigma_{\varpi_1}) \,
+    p(d_2 \given \varpi_2, \sigma_{\varpi_2}) \,
+    p(\vperp)
+\end{multline}
+and
+\begin{multline}
+  p(\propm_1, \propm_2)_{\vperp_1 \neq \vperp_2} =
+    \int \, \dd d_1 \, \dd d_2 \, \dd \vperp_1 \, \dd \vperp_2 \,
+    p(\propm_1 \given d_1, \vperp_1, \matrx{C}_{\mu,1}) \,
+    p(\propm_2 \given d_2, \vperp_2, \matrx{C}_{\mu,2}) \\
+    p(d_1 \given \varpi_1, \sigma_{\varpi_1}) \,
+    p(d_2 \given \varpi_2, \sigma_{\varpi_2}) \,
+    p(\vperp_1) \, p(\vperp_2)
+\end{multline}
+We'll assume the sky positions $(\alpha, \delta)$ are known perfectly, and that
+the uncertainties in proper motion components, $\propm=(\mu_\alpha,
+\mu_\delta)$, are Gaussian with known covariance matrix $\matrx{C}_\mu$ so that
+\begin{equation}
+  p(\propm \given d, \vperp, \matrx{C}) = \left[\det\left(\frac{\matrx{C}_\mu^{-1}}{2\pi}\right)\right]^{1/2} \, \exp \left[ -\frac{1}{2} \transp{\left(\propm - \frac{\vperp}{d}\right)} \, \matrx{C}_\mu^{-1} \, \left(\propm - \frac{\vperp}{d}\right) \right]
+\end{equation}
+
+Our distance posterior pdfs, $p(d \given \varpi, \sigma_\varpi)$, will come from
+Coryn Bailer-Jones.
+
+Our priors on tangential velocity components, $i$, will be something like:
+\begin{equation}
+  p(v^\perp_i) = \mathcal{N}(0, 30) \, \kms
+\end{equation}
 

### 2016-09-24 paper/ms.tex
infer distances too

--- 2016-09-23 paper/ms.tex
@@
 \section{Introduction} \label{sec:intro}
+
+\section{Methods} \label{sec:methods}
+
+For a pair of stars $(i,j)$, we want to ratio the fully marginalized
+likelihood (FML) of the stars having the same tangential velocity (numerator)
+compared to the FML of the stars having different tangential velocities:
+\begin{equation}
+  \frac{p(\propm_i, \propm_j)_{\vperp_i = \vperp_j}}
+  {p(\propm_i, \propm_j)_{\vperp_i \neq \vperp_j}}.
+\end{equation}
+When this ratio is large, the pair is a candidate binary star. These likelihood
+expressions are marginalized over true tangential velocity and distance for each
+star in the pair. For the numerator,
+\begin{equation}
+  p(\propm_i, \propm_j)_{\vperp_i = \vperp_j} =
+    \int \, \dd \vperp \, \dd d_i \, \dd d_j \,
+    p(\propm_i, \varpi_i, \vperp, d_i, \matrx{C}_i) \,
+    p(\propm_j, \varpi_j, \vperp, d_j, \matrx{C}_j)
+\end{equation}
+where
+\begin{align}
+  p(\propm, \varpi, \vperp, d, \matrx{C}) &=
+    p(\propm, \varpi \given \vperp, d, \matrx{C}) \, p(\vperp) \, p(d) \\
+  p(\propm, \varpi \given \vperp, d, \matrx{C}) &=
+    \left[\det\left(\frac{\matrx{C}^{-1}}{2\pi}\right)\right]^{1/2} \,
+    \exp \left[ -\frac{1}{2} \transp{\left(\bs{x} - \bs{x}_\theta \right)} \,
+    \matrx{C}^{-1} \,
+    \left(\bs{x} - \bs{x}_\theta \right) \right] \\
+  \bs{x} &= \transp{\left(\varpi, \mu_\alpha, \mu_\delta\right)} \\
+  \bs{x}_\theta &= \transp{\left(\frac{1}{d}, \frac{v^\perp_\alpha}{d}, \frac{v^\perp_\delta}{d}\right)}
+\end{align}
+and we assume Gaussian priors on the tangential velocity components, $\vperp=(v^\perp_\alpha,v^\perp_\delta)$, and take the ``Milky Way distance prior'' from Coryn Bailer-Jones
+\begin{align}
+  p(v^\perp_\alpha,v^\perp_\alpha) = \mathcal{N}(0,25~\kms) \, \mathcal{N}(0,25~\kms)
+\end{align}
+
+For the denominator of the likelihood ratio,
+\begin{equation}
+  p(\propm_i, \propm_j)_{\vperp_i \neq \vperp_j} =
+    \int \, \dd \vperp_i \, \dd \vperp_j \, \dd d_i \, \dd d_j \,
+    p(\propm_i, \varpi_i, \vperp_i, d_i, \matrx{C}_i) \,
+    p(\propm_j, \varpi_j, \vperp_j, d_j, \matrx{C}_j)
+\end{equation}
 

### 2016-09-25 paper/ms.tex
radial velocity

--- 2016-09-23 paper/ms.tex
@@
 \newcommand{\gaia}{\project{Gaia}}
+\newcommand{\rave}{\project{\acronym{RAVE}}}
+\newcommand{\apogee}{\project{\acronym{APOGEE}}}
 \newcommand{\documentname}{\textsl{Article}}
@@
 \section{Introduction} \label{sec:intro}
+
+\section{Methods} \label{sec:methods}
+
+For a pair of stars $(i,j)$, we want to ratio the fully marginalized likelihood
+(FML) of the stars having the same 3-space velocity, $\bs{v}=(v^\perp_\alpha,
+v^\perp_\delta, v_r)$, compared to the FML of the stars having different 3-space
+velocities:
+\begin{equation}
+  \frac{p(\propm_i, \varpi_i, \matrx{C}_i, \propm_j, \varpi_j, \matrx{C}_j)_{\bs{v}_i = \bs{v}_j}}
+  {p(\propm_i, \varpi_i, \matrx{C}_i, \propm_j, \varpi_j, \matrx{C}_j)_{\bs{v}_i \neq \bs{v}_j}}.
+\end{equation}
+When this ratio is large---that is, the likelihood that the stars have the same
+3-space velocity is large and the the likelihood that the stars have different
+3-space velicities is small---the pair is a candidate binary star. For \gaia\
+DR1, we only observe a parallax, $\varpi$, and two proper motion components,
+$\propm = (\mu_\alpha, \mu_\delta)$. Where there is overlap with \rave\ or
+\apogee, some number of observed radial velocities, $\tilde{v}_r$, with observed
+variances $\sigma^2_{v_r}$ (independent of the \gaia\ data) may also be
+available. The likelihood expressions above are marginalized over true 3-space
+velocity and distance for each star in the pair. For the numerator,
+\begin{equation}
+  p(\propm_i, \varpi_i, \matrx{C}_i, \propm_j, \varpi_j, \matrx{C}_j)_{\bs{v}_i = \bs{v}_j} =
+    \int \, \dd \bs{v} \, \dd d_i \, \dd d_j \,
+    p(\propm_i, \varpi_i, \bs{v}, d_i, \matrx{C}_i) \,
+    p(\propm_j, \varpi_j, \bs{v}, d_j, \matrx{C}_j)
+\end{equation}
+where
+\begin{align}
+  p(\propm, \varpi, \bs{v}, d, \matrx{C}) &=
+    p(\propm, \varpi \given \vperp, d, \matrx{C}) \, p(\vperp, v_r) \, p(\tilde{v}_r \given v_r, \sigma^2_{v_r}) \, p(d \given l,b) \\
+  p(\propm, \varpi \given \vperp, d, \matrx{C}) &=
+    \left[\det\left(\frac{\matrx{C}^{-1}}{2\pi}\right)\right]^{1/2} \,
+    \exp \left[ -\frac{1}{2} \transp{\left(\bs{x} - \bs{x}_\theta \right)} \,
+    \matrx{C}^{-1} \,
+    \left(\bs{x} - \bs{x}_\theta \right) \right] \\
+  \bs{x} &= \transp{\left(\varpi, \mu_\alpha, \mu_\delta\right)} \\
+  \bs{x}_\theta &= \transp{\left(\frac{1}{d}, \frac{v^\perp_\alpha}{d}, \frac{v^\perp_\delta}{d}\right)}
+\end{align}
+
+We assume Gaussian priors on the true velocity components
+\begin{align}
+  p(\bs{v}) = \mathcal{N}(0,25)~\kms
+\end{align}
+and take one of the distance priors from Coryn Bailer-Jones,
+
+For the denominator of the likelihood ratio,
+\begin{multline}
+  p(\propm_i, \varpi_i, \matrx{C}_i, \propm_j, \varpi_j, \matrx{C}_j)_{\bs{v}_i \neq \bs{v}_j} = \\
+    \int \, \dd \bs{v}_i \, \dd \bs{v}_j \, \dd d_i \, \dd d_j \,
+    p(\propm_i, \varpi_i, \bs{v}_i, d_i, \matrx{C}_i) \,
+    p(\propm_j, \varpi_j, \bs{v}_j, d_j, \matrx{C}_j)
+\end{multline}
 

### 2016-09-25 paper/ms.tex

--- 2016-09-24 paper/ms.tex
@@
 
-For a pair of stars $(i,j)$, we want to ratio the fully marginalized
-likelihood (FML) of the stars having the same tangential velocity (numerator)
-compared to the FML of the stars having different tangential velocities:
-\begin{equation}
-  \frac{p(\propm_i, \propm_j)_{\vperp_i = \vperp_j}}
-  {p(\propm_i, \propm_j)_{\vperp_i \neq \vperp_j}}.
-\end{equation}
-When this ratio is large, the pair is a candidate binary star. These likelihood
-expressions are marginalized over true tangential velocity and distance for each
-star in the pair. For the numerator,
-\begin{equation}
-  p(\propm_i, \propm_j)_{\vperp_i = \vperp_j} =
-    \int \, \dd \vperp \, \dd d_i \, \dd d_j \,
-    p(\propm_i, \varpi_i, \vperp, d_i, \matrx{C}_i) \,
-    p(\propm_j, \varpi_j, \vperp, d_j, \matrx{C}_j)
-\end{equation}
+For a pair of stars $(i,j)$, we want to ratio the fully marginalized likelihood
+(FML) of the stars having the same 3-space velocity (hypothesis 1,
+$\mathcal{L}_1$) over the FML of the stars having different 3-space velocities
+(hypothesis 2, $\mathcal{L}_2$). When this ratio is large, the pair is a
+candidate binary star. For \gaia\ DR1, we only observe parallax, $\varpi$, and
+two proper motion components, $\propm = (\mu_\alpha, \mu_\delta)$ for each star.
+Where there is overlap with \rave\ or \apogee, some number of observed radial
+velocities, $v_r$, with observed variances $\sigma^2_{v_r}$ (independent of the
+\gaia\ data) may also be available. We assume that the uncertainties in these
+observed quantites are known with covariance matrix $\matrx{C}$. The likelihoods
+$\mathcal{L}_1, \mathcal{L}_2$ are marginalized over true 3-space velocity and
+distance for each star in the pair.
+\begin{align}
+  \mathcal{L}_1 &=
+    \int \, \dd d_i \, \dd d_j \, \dd^3 \bs{v} \,
+    L_i(\bs{v}, d_i) \,
+    L_j(\bs{v}, d_j) \,
+    p(\bs{v}) \, p(d_i) \, p(d_j) \\
+  \mathcal{L}_2 &=
+    \int \, \dd d_i \, \dd d_j \, \dd^3 \bs{v}_i \, \dd^3 \bs{v}_j \,
+    L_i(\bs{v}_i, d_i) \,
+    L_j(\bs{v}_j, d_j) \,
+    p(\bs{v}_i) \, p(\bs{v}_j) \, p(d_i) \, p(d_j) \\
+\end{align}
 where
 \begin{align}
-  p(\propm, \varpi, \vperp, d, \matrx{C}) &=
-    p(\propm, \varpi \given \vperp, d, \matrx{C}) \, p(\vperp) \, p(d) \\
-  p(\propm, \varpi \given \vperp, d, \matrx{C}) &=
+  L(\bs{v}, d) &=
     \left[\det\left(\frac{\matrx{C}^{-1}}{2\pi}\right)\right]^{1/2} \,
@@
     \left(\bs{x} - \bs{x}_\theta \right) \right] \\
-  \bs{x} &= \transp{\left(\varpi, \mu_\alpha, \mu_\delta\right)} \\
-  \bs{x}_\theta &= \transp{\left(\frac{1}{d}, \frac{v^\perp_\alpha}{d}, \frac{v^\perp_\delta}{d}\right)}
+  \bs{x} &= \transp{\left(\varpi, \mu_\alpha, \mu_\delta, v_r \right)} \\
+  \bs{x}_\theta &= \transp{\left(\frac{1}{d}, \frac{v_\alpha}{d},
+    \frac{v_\delta}{d}, \tilde{v}_r \right)}
 \end{align}
-and we assume Gaussian priors on the tangential velocity components, $\vperp=(v^\perp_\alpha,v^\perp_\delta)$, and take the ``Milky Way distance prior'' from Coryn Bailer-Jones
+We assume Gaussian priors on the true velocity components and use a \todo{which
+one will we use?} distance prior \citep{Astraatmadja:2016} on heliocentric
+distance
 \begin{align}
-  p(v^\perp_\alpha,v^\perp_\alpha) = \mathcal{N}(0,25~\kms) \, \mathcal{N}(0,25~\kms)
+  p(\bs{v}) &= \mathcal{N}(0,25)~\kms \\
+  p(d) &= \frac{r^2}{2\,L^3} \, \exp\left(-r/L\right)
 \end{align}
 
-For the denominator of the likelihood ratio,
+For hypothesis 1, the integral over velocity is non-trivial because the same
+``true'' 3-space velocity vector projects into tangential velocity components
+differently depending on sky position when the sky separation is appreciable
+($v_\alpha = v_\alpha(\bs{v}, \alpha, \delta)$). For hypothesis 2, the integral
+can be split into two the product of two simpler integrals $\mathcal{L}_2 = Q_i
+\, Q_j$ where
 \begin{equation}
-  p(\propm_i, \propm_j)_{\vperp_i \neq \vperp_j} =
-    \int \, \dd \vperp_i \, \dd \vperp_j \, \dd d_i \, \dd d_j \,
-    p(\propm_i, \varpi_i, \vperp_i, d_i, \matrx{C}_i) \,
-    p(\propm_j, \varpi_j, \vperp_j, d_j, \matrx{C}_j)
+  Q = \int \, \dd d \, \dd^3 \bs{v} \, L(\bs{v}, d) \, p(\bs{v}) \, p(d)
 \end{equation}

### 2016-09-26 code/prob.py
minor bugs

@@ -1,5 +1,5 @@
 import numpy as np
-from numpy as sqrt, exp, pi, inf
+from numpy import sqrt, exp, pi, inf
 from scipy import integrate
 
 
@@ -90,11 +90,11 @@ def likelihood(d1, d2, v_ra, v_dec, p1, p2, sigp1, sigp2, mu1, mu2, C1, C2):
     # d1,d2 : scalars
     v = np.array([v_ra, v_dec])
     Pmu1 = likelihood_gaussian(v/d1/4.74, mu1, C1)
-    Pmu2 = likelihood_gaussian(v/d2/4.74, mu1, C2)
+    Pmu2 = likelihood_gaussian(v/d2/4.74, mu2, C2)
     Pd1 = post_distance(d1, p1, sigp1)
     Pd2 = post_distance(d2, p2, sigp2)
     Pv = prior_velocity_gaussian(v_ra, v_dec)
-    return Pmu1*Pmu2*Pv*Pd1*Pd2
+    return Pd1*Pd2*Pmu1*Pmu2*Pv
 
 def likelihood_null(d1, d2, v1_ra, v1_dec, v2_ra, v2_dec, p1, p2, sigp1, sigp2, mu1, mu2, C1, C2):
     # mu: [pmra, pmdec]
@@ -104,8 +104,8 @@ def likelihood_null(d1, d2, v1_ra, v1_dec, v2_ra, v2_dec, p1, p2, sigp1, sigp2, 
     Pmu1 = likelihood_gaussian(v1/d1/4.74, mu1, C1)
     v2 = np.array([v2_ra, v2_dec])
     Pmu2 = likelihood_gaussian(v2/d2/4.74, mu1, C2)
-    Pd1 = post_distance(d1, p1, sigp1, prior=prior_distance_expcutoff)
-    Pd2 = post_distance(d2, p2, sigp2, prior=prior_distance_expcutoff)
+    Pd1 = post_distance(d1, p1, sigp1, prior=prior_distance_uniform)
+    Pd2 = post_distance(d2, p2, sigp2, prior=prior_distance_uniform)
     Pv1 = prior_velocity_gaussian(v1_ra, v1_dec)
     Pv2 = prior_velocity_gaussian(v2_ra, v2_dec)
     return Pmu1*Pmu2*Pv1*Pv2*Pd1*Pd2

### 2016-09-28 gwb/likelihood.py
back to sph geom - faster with no units

@@ -1,7 +1,6 @@
 from __future__ import division, print_function
 
 # Third-party
-import astropy.coordinates as coord
 import astropy.units as u
 import numpy as np
 
@@ -27,8 +26,10 @@ def get_u_vec(lon, lat):
         Unit 3-vector.
 
     """
-    usph = coord.UnitSphericalRepresentation(lon=lon*u.rad, lat=lat*u.rad)
-    return usph.represent_as(coord.CartesianRepresentation).xyz.value.T
+    u_hat = np.array([np.cos(lon) * np.cos(lat),
+                      np.sin(lon) * np.cos(lat),
+                      np.sin(lat)])
+    return u_hat / np.sqrt(np.sum(u_hat**2))
 
 def get_tangent_basis(ra, dec, dra=0.5, ddec=0.5):
     """

### 2016-09-28 gwb/tests/test_likelihood.py
test many lon, lat

@@ -2,6 +2,7 @@ from __future__ import division, print_function
 
 # Third-party
 import astropy.units as u
+import astropy.coordinates as coord
 import numpy as np
 
 # Project
@@ -16,6 +17,12 @@ def test_u_vec():
 
     assert get_u_vec(ra[0], dec[0]).shape == (3,)
     assert get_u_vec(ra, dec).shape == (n_test,3)
+
+    for lon,lat in zip(ra,dec):
+        usph = coord.UnitSphericalRepresentation(lon=lon*u.rad, lat=lat*u.rad)
+        astropy_xyz = usph.represent_as(coord.CartesianRepresentation).xyz.value.T
+        assert np.allclose(astropy_xyz, get_u_vec(lon, lat))
+
 
 def test_tangent_basis():
     ra = np.random.uniform(0, 2*np.pi, size=n_test)

### 2016-09-28 paper/ms.tex
started writing some of hte marginalization down

@@
 
+\appendix
+
+Here we derive analytic expressions for the marginal likelihood integrals above.
+
+\todo{Hypothesis 1}
+
+The integrals for Hypothesis 2 are somewhat easier because the true velocity
+vector can be considered to be already in the orthogonal basis defined by the
+tangent plane at the coordinates of the source (we are assuming an isotropic
+velocity prior). We will analytically integrate over the velocity components
+$(v_\alpha,v_\delta,v_r)$ and leave the distance integral to be done
+numerically:
+\begin{equation}
+  \int \, \dd^3\bs{v} \, L(\bs{v}, d) \, p(\bs{v})
+\end{equation}
+The covariance matrix over parallax, proper motion components, and
+radial velocity (if they exist) is $\matrx{C}$ and the inverse variance matrix
+is $\matrx{V}$. Defining the constant
+\begin{equation}
+  \mathcal{Z} = \left[ \det \left(\frac{\matrx{V}}{2\pi} \right) \right]^{1/2}
+\end{equation}
+the integrand can be written
+\begin{equation}
+  L(\bs{v}, d) \, p(\bs{v}) = \mathcal{Z} \,
+    \exp \left[ -\frac{1}{2} \transp{\left(\bs{x} - \bs{x}_\theta \right)} \,
+    \matrx{V} \,
+    \left(\bs{x} - \bs{x}_\theta \right)
+    - \frac{|\bs{v}|^2}{2\,\sigma^2} \right] \\
+\end{equation}
+
 \acknowledgements

### 2016-09-28 paper/ms.tex
partitioned matrix

@@
 \newcommand{\figname}{Figure}
+\newcommand{\eqname}{Equation}
 
@@
     \left(\bs{x} - \bs{x}_\theta \right) \right] \\
-  \bs{x} &= \transp{\left(\varpi, \mu_\alpha, \mu_\delta, v_r \right)} \\
-  \bs{x}_\theta &= \transp{\left(\frac{1}{d}, \frac{v_\alpha}{d},
-    \frac{v_\delta}{d}, \tilde{v}_r \right)}
+  \bs{x} &= \transp{\left(\begin{array}{c c c c} \varpi & \mu_\alpha &
+    \mu_\delta & v_r \end{array}\right)} \\
+  \bs{x}_\theta &= \transp{\left(\begin{array}{c c c c} \frac{1}{d} & \frac{v_\alpha}{d} &
+    \frac{v_\delta}{d} & \tilde{v}_r \end{array}\right)}
 \end{align}
+\todo{Holy fuck that vector is UGLY}
 We assume Gaussian priors on the true velocity components and use a \todo{which
@@
 \end{equation}
+The integral over radial velocity is simple and can be performed without
+considering any other terms because uncertainties in these measurements will
+not, in general, be covariant with the Gaia data. The integrand over tangential
+velocities can then be written
+\begin{equation}
+  L(v_\alpha, v_\delta, d) \, p(v_\alpha, v_\delta) = \tilde{\mathcal{Z}} \,
+    \exp \left[ -\frac{1}{2}\,\transp{\left(\bs{x}' - \bs{x}'_\theta \right)} \,
+    \matrx{V}' \,
+    \left(\bs{x}' - \bs{x}'_\theta \right)
+    - \frac{(v_\alpha^2 + v_\delta^2)}{2\,\sigma^2} \right]
+    \label{eq:vtan-integrand}
+\end{equation}
+where the primed variables indicate that they are in the sub-space of just the
+astrometric quantities, the constant pre-factor now includes the integral over
+radial velocity
+\begin{align}
+  \tilde{\mathcal{Z}} = \mathcal{Z} \,
+    \exp\left[-\frac{1}{2}\frac{v_r^2}{\sigma_{v_r}^2 + \sigma^2}\right] \,
+    \left[ 2\pi \, (\sigma_{v_r}^2 + \sigma^2) \right]^{1/2}
+\end{align}
+and $\sigma_{v_r}$ is the observational uncertainty of the radial velocity
+measurement. To perform the integral over tangential velocities, we partition
+the inverse variance matrix and vectors so that
+\begin{align}
+  \bs{x}_A &= (\varpi)\\
+  \bs{x}_B &= (\mu_\alpha, \mu_\delta)\\
+  \bs{x}_{\theta, A} &= (d^{-1}) \\
+  \bs{x}_{\theta, B} &= (d^{-1}\,v_\alpha, d^{-1}\,v_\delta) \\
+  \matrx{V} &=
+    \left(\begin{array}{c| c c}
+      V_{11} & V_{12} & V_{13}\\\hline
+      V_{12} & V_{22} & V_{23}\\
+      V_{13} & V_{23} & V_{33}
+    \end{array}\right)
+    = \left(\begin{array}{c c}
+      \matrx{V}_{AA} & \matrx{V}_{AB}\\
+      \matrx{V}_{AB} & \matrx{V}_{BB}
+    \end{array}\right)
+\end{align}
+\todo{Holy fuck that matrix is UGLY - fix with new version of AASTex 6}
+and the argument of the Gaussian in \eqname~\ref{eq:vtan-integrand} can be
+written
+\begin{multline}
+  -\frac{1}{2}\,[
+  \transp{(\bs{x}_A - \bs{x}_{\theta, A})}\,
+    \matrx{V}_{AA}\,(\bs{x}_A - \bs{x}_{\theta, A})\\
+  + \transp{(\bs{x}_B - \bs{x}_{\theta, B})}\,
+    \matrx{V}_{BB}\,(\bs{x}_B - \bs{x}_{\theta, B})\\
+  + 2\,\transp{(\bs{x}_A - \bs{x}_{\theta, A})}\,
+    \matrx{V}_{AB}\,(\bs{x}_B - \bs{x}_{\theta, B})]
+\end{multline}
+\todo{prior in above}
 

### 2016-09-28 paper/ms.tex
start appendix

--- 2016-09-25 paper/ms.tex
@@
 
+\appendix
+\section{Some mathematics relevant to the above}
+
+In what follows, all vectors are column vectors.
+Properties of the squared exponential:
+\begin{eqnarray}
+  \ln\left[\int\exp(-\frac{1}{2}\,[x-\mu]\transp\,A\inv\,[x-\mu] - \Delta)\,\dd x\right]
+  &=& -\frac{d}{2}\,\ln 2\pi -\frac{1}{2}\ln ||A|| -\Delta
+  \quad ,
+\end{eqnarray}
+where $x$ and $\mu$ are $d$-dimensional vectors, $A$ is a symmetric
+matrix, $\Delta$ is a scalar, and the integral is implicitly over all
+of $d$-space.
+We will need to complete squares.
+If we equate
+\begin{eqnarray}
+  \frac{1}{2}\,[x-\mu]\transp\,A\inv\,[x-\mu] + \Delta
+  &=& \frac{1}{2}\,x\transp\,A\inv\,x + x\transp\,B\,b + C
+  \quad ,
+\end{eqnarray}
+where $B\,b$ is a $d$-vector, and $C$ is a scalar, then we find
+\begin{eqnarray}
+  \mu &=& A\,B\,b
+  \\
+  C & = & \Delta - \frac{1}{2}\,\mu\transp\,A\inv\,\mu
+  \quad .
+\end{eqnarray}
+
 \end{document}

### 2016-09-28 gwb/coords.py
what

@@ -29,23 +29,26 @@ def get_u_vec(lon, lat):
     u_hat = np.array([np.cos(lon) * np.cos(lat),
                       np.sin(lon) * np.cos(lat),
                       np.sin(lat)])
-    return u_hat / np.sqrt(np.sum(u_hat**2))
+    return u_hat / np.sqrt(np.sum(u_hat**2, axis=0))[None]
 
 def get_tangent_basis(ra, dec, dra=0.5, ddec=0.5):
     """
-    column vectors are the tangent-space basis at (alpha, delta, r)
+    row vectors are the tangent-space basis at (alpha, delta, r)
     """
+
+    ra = np.atleast_1d(ra)
+    dec = np.atleast_1d(dec)
 
     # unit vector pointing at the sky position of the target
     u_hat = get_u_vec(ra, dec)
 
     # unit vector offset in declination
-    if dec > np.pi/4:
-        v_hat = get_u_vec(ra, dec-ddec)
-        dec_hat_sign = -1.
-    else:
-        v_hat = get_u_vec(ra, dec+ddec)
-        dec_hat_sign = 1.
+    dec_hat_sign = np.ones_like(u_hat)
+    dec_hat_sign[:,dec > np.pi/4] = -1.
+
+    v_hat = np.zeros_like(u_hat)
+    v_hat[:,dec <= np.pi/4] = get_u_vec(ra[dec <= np.pi/4], dec[dec <= np.pi/4]+ddec)
+    v_hat[:,dec > np.pi/4] = get_u_vec(ra[dec > np.pi/4], dec[dec > np.pi/4]-ddec)
 
     dec_hat = dec_hat_sign * (v_hat - u_hat)
     ra_hat = get_u_vec(ra+dra, dec) - u_hat # always a positive offset in RA
@@ -54,13 +57,13 @@ def get_tangent_basis(ra, dec, dra=0.5, ddec=0.5):
     #  - u1 is the unit vector that points to (ra,dec)
     u1 = u_hat
 
-    u2 = dec_hat - dec_hat.dot(u1)*u1
-    u2 /= np.sqrt(np.sum(u2**2))
+    u2 = dec_hat - np.einsum('ij,ij->j', dec_hat, u1)*u1
+    u2 /= np.sqrt(np.sum(u2**2, axis=0))
 
[…]

### 2016-09-28 paper/ms.tex
tweak

--- 2016-09-28 paper/ms.tex
@@
 \newcommand{\dd}{\mathrm{d}}
-\newcommand{\transp}[1]{{#1}^{\mathsf{T}}}
+\newcommand{\transp}[1]{{#1}^{\!\mathsf{T}}}
 \newcommand{\inv}[1]{{#1}^{-1}}
@@
 
+\appendix
+\section{Some mathematics relevant to the above}
+
+In what follows, all vectors are column vectors.
+Properties of the squared exponential:
+\begin{eqnarray}
+  \ln\left[\int\exp(-\frac{1}{2}\,[x-\mu]\transp\,A\inv\,[x-\mu] - \Delta)\,\dd x\right]
+  &=& -\frac{d}{2}\,\ln 2\pi -\frac{1}{2}\ln ||A|| -\Delta
+  \quad ,
+\end{eqnarray}
+where $x$ and $\mu$ are $d$-dimensional vectors, $A$ is a symmetric
+matrix, $\Delta$ is a scalar, and the integral is implicitly over all
+of $d$-space.
+We will need to complete squares.
+If we equate
+\begin{eqnarray}
+  \frac{1}{2}\,[x-\mu]\transp\,A\inv\,[x-\mu] + \Delta
+  &=& \frac{1}{2}\,x\transp\,A\inv\,x + x\transp\,B\,b + C
+  \quad ,
+\end{eqnarray}
+where $B\,b$ is a $d$-vector, and $C$ is a scalar, then we find
+\begin{eqnarray}
+  \mu &=& A\,B\,b
+  \\
+  C & = & \Delta - \frac{1}{2}\,\mu\transp\,A\inv\,\mu
+  \quad .
+\end{eqnarray}
+
+Now in this case, we want to write down the probability of the
+
 \end{document}

### 2016-09-28 gwb/coords.py
support arrays

@@ -63,7 +63,7 @@ def get_tangent_basis(ra, dec, dra=0.5, ddec=0.5):
     u3 = ra_hat - np.einsum('ij,ij->j', ra_hat, u1)*u1 - np.einsum('ij,ij->j', ra_hat, u2)*u2
     u3 /= np.sqrt(np.sum(u3**2, axis=0))
 
-    return np.hstack((u3,u2,u1))
+    return np.squeeze(np.stack((u3,u2,u1), axis=1))
 
 def marg_likelihood_helper(v, d, data):
     """

### 2016-09-28 gwb/tests/test_coords.py

@@ -6,8 +6,7 @@ import astropy.coordinates as coord
 import numpy as np
 
 # Project
-from ..likelihood import (get_u_vec, get_tangent_basis,
-                          ln_H1_marg_likelihood, ln_H2_marg_likelihood)
+from ..coords import get_u_vec, get_tangent_basis
 
 n_test = 8
 
@@ -16,13 +15,12 @@ def test_u_vec():
     dec = np.pi/2. - np.arccos(2*np.random.uniform(size=n_test)-1.)
 
     assert get_u_vec(ra[0], dec[0]).shape == (3,)
-    assert get_u_vec(ra, dec).shape == (n_test,3)
+    assert get_u_vec(ra, dec).shape == (3,n_test)
 
     for lon,lat in zip(ra,dec):
         usph = coord.UnitSphericalRepresentation(lon=lon*u.rad, lat=lat*u.rad)
         astropy_xyz = usph.represent_as(coord.CartesianRepresentation).xyz.value.T
         assert np.allclose(astropy_xyz, get_u_vec(lon, lat))
-
 
 def test_tangent_basis():
     ra = np.random.uniform(0, 2*np.pi, size=n_test)
@@ -32,7 +30,7 @@ def test_tangent_basis():
         A = get_tangent_basis(a, d)
         assert A.shape == (3,3)
 
-        assert np.allclose(np.linalg.norm(A, axis=1), 1.)
+        assert np.allclose(np.linalg.norm(A, axis=0), 1.)
         assert np.allclose(np.dot(A[0], A[1]), 0.)
         assert np.allclose(np.dot(A[0], A[2]), 0.)
         assert np.allclose(np.dot(A[1], A[2]), 0.)
@@ -40,6 +38,18 @@ def test_tangent_basis():
         assert np.allclose(np.cross(A[0], A[1]), A[2])
         assert np.allclose(np.cross(A[0], A[2]), -A[1])
         assert np.allclose(np.cross(A[1], A[2]), A[0])
+
+    A = get_tangent_basis(ra[:4], dec[:4])
+    assert A.shape == (3,3,4)
+
+    assert np.allclose(np.linalg.norm(A, axis=0), 1.)
+    assert np.allclose(np.einsum('ij,ij->', A[0], A[1]), 0.)
+    assert np.allclose(np.einsum('ij,ij->', A[0], A[2]), 0.)
[…]

### 2016-09-28 gwb/tests/test_coords.py
remove old tests

@@ -50,127 +50,3 @@ def test_tangent_basis():
     assert np.allclose(np.cross(A[0], A[1], axisa=0, axisb=0, axisc=0), A[2])
     assert np.allclose(np.cross(A[0], A[2], axisa=0, axisb=0, axisc=0), -A[1])
     assert np.allclose(np.cross(A[1], A[2], axisa=0, axisb=0, axisc=0), A[0])
-
-def make_pair_same_v(d1=81.314, sep_pc=0.1):
-    # same velocity vector
-    true_v = np.random.normal(0, 30., size=3)
-
-    ra1 = np.random.uniform(0, 2*np.pi)
-    dec1 = np.pi/2. - np.arccos(2*np.random.uniform() - 1.)
-    v1 = get_tangent_basis(ra1, dec1).dot(true_v)
-
-    ra2 = ra1 + np.random.choice([1.,-1.])*sep_pc/np.sqrt(2.)/d1
-    dec_choice = np.random.choice([1.,-1.])
-    dec2 = dec1 + dec_choice*sep_pc/np.sqrt(2.)/d1
-    if dec2 > np.pi/2 or dec2 < -np.pi/2.:
-        dec2 += -2*dec_choice*sep_pc/np.sqrt(2.)/d1
-
-    d2 = d1 + np.random.choice([1.,-1.])*sep_pc/np.sqrt(2.)
-    v2 = get_tangent_basis(ra2, dec2).dot(true_v)
-
-    # picked at random from TGAS data
-    Cov1 = np.array([[0.066019, 0.0581179, -0.06287875],
-                     [0.0581179, 2.76016904, -1.31836247],
-                     [-0.06287875, -1.31836247, 0.83536227]])
-    Cov2 = np.array([[0.07143981, 0.31518988, -0.11283286],
-                     [0.31518988, 6.78049538, -2.21299533],
-                     [-0.11283286, -2.21299533, 0.9342514]])
-    # Cov1 = Cov2 = np.diag(np.zeros(3) + 1E-8)
-
-    plx1 = (d1*u.pc).to(u.mas,u.parallax()).value
-    plx2 = (d2*u.pc).to(u.mas,u.parallax()).value
-
-    pm1 = (v1[:2] / d1 * u.km/u.s/u.pc).to(u.mas/u.yr, u.dimensionless_angles()).value
-    pm2 = (v2[:2] / d2 * u.km/u.s/u.pc).to(u.mas/u.yr, u.dimensionless_angles()).value
-
[…]

### 2016-09-28 gwb/coords.py
line break

@@ -60,7 +60,8 @@ def get_tangent_basis(ra, dec, dra=0.5, ddec=0.5):
     u2 = dec_hat - np.einsum('ij,ij->j', dec_hat, u1)*u1
     u2 /= np.sqrt(np.sum(u2**2, axis=0))
 
-    u3 = ra_hat - np.einsum('ij,ij->j', ra_hat, u1)*u1 - np.einsum('ij,ij->j', ra_hat, u2)*u2
+    u3 = (ra_hat - np.einsum('ij,ij->j', ra_hat, u1)*u1 -
+          np.einsum('ij,ij->j', ra_hat, u2)*u2)
     u3 /= np.sqrt(np.sum(u3**2, axis=0))
 
     return np.squeeze(np.stack((u3,u2,u1), axis=1))

### 2016-09-28 paper/ms.tex
math

@@
 \newcommand{\given}{\,|\,}
+\newcommand{\normal}{N}
 \newcommand{\dd}{\mathrm{d}}
@@
 \newcommand{\kms}{\rm km~s^{-1}}
+\newcommand{\data}{\mathrm{data}}
 
@@
 In what follows, all vectors are column vectors.
+We indicate whether a symbol represents a scalar, vector, or tensor
+(matrix) by context, not typesetting.
 Properties of the squared exponential:
 \begin{eqnarray}
-  \ln\left[\int\exp(-\frac{1}{2}\,[x-\mu]\transp\,A\inv\,[x-\mu] - \Delta)\,\dd x\right]
+  \ln\left[\int\exp(-\frac{1}{2}\,\transp{[x-\mu]}\,\inv{A}\,[x-\mu] - \Delta)\,\dd x\right]
   &=& -\frac{d}{2}\,\ln 2\pi -\frac{1}{2}\ln ||A|| -\Delta
@@
 \begin{eqnarray}
-  \frac{1}{2}\,[x-\mu]\transp\,A\inv\,[x-\mu] + \Delta
-  &=& \frac{1}{2}\,x\transp\,A\inv\,x + x\transp\,B\,b + C
+  \frac{1}{2}\,\transp{[x-\mu]}\,\inv{A}\,[x-\mu] + \Delta
+  &=& \frac{1}{2}\,\transp{x}\,\inv{A}\,x + \transp{x}\,B\,b + C
   \quad ,
@@
   \\
-  C & = & \Delta - \frac{1}{2}\,\mu\transp\,A\inv\,\mu
+  C & = & \Delta - \frac{1}{2}\,\transp{\mu}\,\inv{A}\,\mu
   \quad .
@@
 
-Now in this case, we want to write down the probability of the
+Now in this case, we want to write down the probability of the data
+given the distance and velocity (the likelihood), and the prior pdf for
+the velocity, and their product.
+The prior pdf for the velocity is Gaussian:
+\begin{eqnarray}
+  p(v) &=& \normal(v\given 0,V)
+  \\
+  \ln p(v) &=& -\frac{3}{2}\,\ln\pi -\frac{1}{2}\,\ln||V|| -\frac{1}{2}\,\transp{v}\,\inv{V}\,v
+  \\
+  V &=& \sigma_v^2\,I
+  \quad ,
+\end{eqnarray}
+where $v$ is the three-vector velocity, and $V$ is an isotropic
+variance tensor.
+Similarly, the likelihood for the data is also a Gaussian.
+Before we write that down, we will make some definitions.
+We construct a vector $y$ as follows:
+\begin{eqnarray}
+  \transp{y} &\equiv& [d_1\,\varpi_1 - 1, d_1\,\mu_{\alpha,1}, d_1\,\mu_{\delta,1}
+                       d_1\,\varpi_2 - 1, d_2\,\mu_{\alpha,2}, d_1\,\mu_{\delta,2}]
+  \\
+  \transp{y} &\equiv& [d_1\,\varpi_1 - 1, d_1\,\mu_{\alpha,1}, d_1\,\mu_{\delta,1}, v_{r,1}
+                       d_1\,\varpi_2 - 1, d_2\,\mu_{\alpha,2}, d_1\,\mu_{\delta,2}, v_{r,2}
+  \quad ,
+\end{eqnarray}
+where we have given two versions, a 6-component and an 8-component
+version (depending on whether there are measured radial velocities),
+the subscripts 1, 2 refer to the first and second stars (arbitrarily
+ordered) of the pair, we have multiplied the strict observables by
+distances $d_1, d_2$, which we are permitted to do (because we are
+\emph{conditioning} on $d_1, d_2$). The vector $y$ is a column vector
+that we have transposed to make it easy to write down.  Fundamentally,
+our model is
+\begin{eqnarray}
+  y &=& M\,v + \mathrm{noise}
+  \\
+  \transp{M} &=& [0, \hat{\alpha_1}, \hat{\delta_1},
+                  0, \hat{\alpha_2}, \hat{\delta_2}]
+  \\
+  \transp{M} &=& [0, \hat{\alpha}_1, \hat{\delta}_2, \hat{r}_1
+                  0, \hat{\alpha}_2, \hat{\delta}_2, \hat{r}_2]
+  \quad,
+\end{eqnarray}
+where $M$ is a $6\times 3$ or $8\times 3$ design matrix, with zeros
+in two rows and unit vectors along the other rows.
+The noise is drawn from a Gaussian with ($6\times 6$ or $8\times 8$)
+covariance matrix $C$, constructed from the distances $d_1, d_2$ and
+the \gaia-provided individual-object covariance matrices.
+Given all these definitions, the likelihood function is
+\begin{eqnarray}
+  p(\data\given v,d_1,d_2) &=& d_1^3\,d_2^3\,\normal(y\given M\,v, C)
+  \\
+  \ln p(\data\given v,d_1,d_2) &=& 3\,\ln d_1 + 3\,\ln d_2 -\frac{6~\mathrm{or}~8}{2}\,\ln\pi
+  -\frac{1}{2}\,\ln||C|| -\frac{1}{2}\,\transp{[y-M\,v]}\,\inv{C}\,[y-M\,v]
+  \quad ,
+\end{eqnarray}
+where the factor of $d_1^2\,d_2^3$ converts to units of \gaia\ data
+from units of $y$.
 

### 2016-09-28 gwb/likelihood.py
new file for likelihood functions

@@ -1,0 +1,50 @@
+from __future__ import division, print_function
+
+# Third-party
+import astropy.units as u
+import numpy as np
+
+# Project
+from .coords import get_tangent_basis
+
+pc_mas_yr_to_km_s = (1*u.pc * u.mas/u.yr).to(u.km/u.s,u.dimensionless_angles()).value
+
+def get_y(d, data):
+    y = np.stack([d * data._parallax*1E-3 - 1.,
+                  d * data._pmra * pc_mas_yr_to_km_s,
+                  d * data._pmdec * pc_mas_yr_to_km_s,
+                  data._rv], axis=1)
+    return y
+
+def get_M(data):
+    M0 = np.zeros((3,) + data._ra.shape)
+    M_vel = get_tangent_basis(data._ra, data._dec)
+    return np.stack((M0[None], M_vel), axis=1)
+
+def get_Cinv(d, data):
+    Cinv = data.get_Cinv()
+    return Cinv[:3,:3] / d**2
+
+def get_A_mu_Delta(M, Cinv, y, Vinv):
+    # TODO: is A is actually the inverse of what you think it is?
+    Ainv = np.einsum('...ij,...jk,...ks->is', M.T, Cinv, M) + Vinv
+    A = np.linalg.inv(Ainv)
+
+    Bb = np.einsum('...ij,...jk,...k->i', M.T, Cinv, y)
+    mu = np.einsum('...ij,...j->i', A, Bb)
+
+    sgn,log_detCinv = np.slogdet(Cinv/(2*np.pi)) # TODO: do the right thing when Cinv[3,3] == 0
+    sgn,log_detVinv = np.slogdet(Vinv/(2*np.pi))
+    yT_Cinv_y = np.einsum('...i,...ji,...j->', y, Cinv, y)
+    muT_A_mu = np.einsum('...i,...ji,...j->', mu, A, mu)
+    Delta = 0.5*log_detCinv + 0.5*log_detVinv - 0.5*yT_Cinv_y - muT_A_mu
+
+    return A, mu, Delta
+
+def ln_marg_likelihood(d, data, Vinv):
+    y = get_y(d, data)
+    M = get_M(data)
+    Cinv = get_Cinv(d, data)
+    A,mu,Delta = get_A_mu_Delta(M, Cinv, y, Vinv)
+    _,log_Adet = np.linalg.slogdet(A/(2*np.pi))
+    return -Delta + 0.5*log_Adet # TODO: plus or minus? is A actually Ainv?

### 2016-09-29 gwb/tests/test_likelihood.py
tests for likelihood functions

@@ -1,0 +1,19 @@
+from __future__ import division, print_function
+
+# Third-party
+import astropy.units as u
+import astropy.coordinates as coord
+import numpy as np
+
+# Project
+from ..gaiatools import TGASData
+from ..likelihood import get_y, get_M, get_Cinv, get_A_mu_Delta
+
+# make fake data
+n_data = 128
+ra = np.random.uniform(0, 2*np.pi, size=n_data)
+dec = np.pi/2. - np.arccos(2*np.random.uniform(size=n_data)-1.)
+parallax = np.random.uniform(-0.1, 1000., size=n_data)
+pmra,pmdec = np.random.normal(-0, 100, size=(2,n_data))
+
+np.random.uniform([0.1,])

### 2016-09-29 paper/ms.tex

@@
 \begin{eqnarray}
-  \ln\left[\int\exp(-\frac{1}{2}\,\transp{[x-\mu]}\,\inv{A}\,[x-\mu] - \Delta)\,\dd x\right]
-  &=& -\frac{d}{2}\,\ln 2\pi -\frac{1}{2}\ln ||A|| -\Delta
+  \ln\left[\int\exp(-\frac{1}{2}\,\transp{[x-\nu]}\,\inv{A}\,[x-\nu] - \Delta)\,\dd x\right]
+  &=& \frac{D}{2}\,\ln 2\pi +\frac{1}{2}\ln ||A|| -\Delta
   \quad ,
 \end{eqnarray}
-where $x$ and $\mu$ are $d$-dimensional vectors, $A$ is a symmetric
+where $x$ and $\nu$ are $D$-dimensional vectors, $A$ is a symmetric
 matrix, $\Delta$ is a scalar, and the integral is implicitly over all
-of $d$-space.
+of $D$-dimensional $x$-space.
 We will need to complete squares.
@@
 \begin{eqnarray}
-  \frac{1}{2}\,\transp{[x-\mu]}\,\inv{A}\,[x-\mu] + \Delta
+  \frac{1}{2}\,\transp{[x-\nu]}\,\inv{A}\,[x-\nu] + \Delta
   &=& \frac{1}{2}\,\transp{x}\,\inv{A}\,x + \transp{x}\,B\,b + C
@@
 \begin{eqnarray}
-  \mu &=& A\,B\,b
+  \nu &=& A\,B\,b
   \\
-  C & = & \Delta - \frac{1}{2}\,\transp{\mu}\,\inv{A}\,\mu
+  C & = & \Delta - \frac{1}{2}\,\transp{\nu}\,\inv{A}\,\nu
   \quad .
@@
   \\
-  \ln p(v) &=& -\frac{3}{2}\,\ln\pi -\frac{1}{2}\,\ln||V|| -\frac{1}{2}\,\transp{v}\,\inv{V}\,v
+  \ln p(v) &=& -\frac{3}{2}\,\ln 2\pi -\frac{1}{2}\,\ln||V|| -\frac{1}{2}\,\transp{v}\,\inv{V}\,v
   \\
@@
 \begin{eqnarray}
-  \transp{y} &\equiv& [d_1\,\varpi_1 - 1, d_1\,\mu_{\alpha,1}, d_1\,\mu_{\delta,1}
+  \transp{y} &\equiv& [(d_1\,\varpi_1 - 1), d_1\,\mu_{\alpha,1}, d_1\,\mu_{\delta,1},
                        d_1\,\varpi_2 - 1, d_2\,\mu_{\alpha,2}, d_1\,\mu_{\delta,2}]
   \\
-  \transp{y} &\equiv& [d_1\,\varpi_1 - 1, d_1\,\mu_{\alpha,1}, d_1\,\mu_{\delta,1}, v_{r,1}
+  \transp{y} &\equiv& [(d_1\,\varpi_1 - 1), d_1\,\mu_{\alpha,1}, d_1\,\mu_{\delta,1}, v_{r,1}
                        d_1\,\varpi_2 - 1, d_2\,\mu_{\alpha,2}, d_1\,\mu_{\delta,2}, v_{r,2}
@@
 distances $d_1, d_2$, which we are permitted to do (because we are
-\emph{conditioning} on $d_1, d_2$). The vector $y$ is a column vector
-that we have transposed to make it easy to write down.  Fundamentally,
-our model is
+\emph{conditioning} on $d_1, d_2$), and note the wacky minus-ones.
+The vector $y$ is a column vector that we have transposed to make it
+easy to write down.
+Fundamentally, our model is
 \begin{eqnarray}
@@
   \\
-  \transp{M} &=& [0, \hat{\alpha}_1, \hat{\delta}_2, \hat{r}_1
+  \transp{M} &=& [0, \hat{\alpha}_1, \hat{\delta}_2, \hat{r}_1,
                   0, \hat{\alpha}_2, \hat{\delta}_2, \hat{r}_2]
@@
 where $M$ is a $6\times 3$ or $8\times 3$ design matrix, with zeros
-in two rows and unit vectors along the other rows.
-The noise is drawn from a Gaussian with ($6\times 6$ or $8\times 8$)
-covariance matrix $C$, constructed from the distances $d_1, d_2$ and
+in two rows, and unit vectors along the other rows.
+The noise (in the $y$ vector) is drawn from a Gaussian with ($6\times 6$ or $8\times 8$)
+covariance matrix $\Sigma$, constructed from the distances $d_1, d_2$ and
 the \gaia-provided individual-object covariance matrices.
@@
 \begin{eqnarray}
-  p(\data\given v,d_1,d_2) &=& d_1^3\,d_2^3\,\normal(y\given M\,v, C)
+  p(\data\given v,d_1,d_2) &=& d_1^3\,d_2^3\,\normal(y\given M\,v, \Sigma)
   \\
-  \ln p(\data\given v,d_1,d_2) &=& 3\,\ln d_1 + 3\,\ln d_2 -\frac{6~\mathrm{or}~8}{2}\,\ln\pi
-  -\frac{1}{2}\,\ln||C|| -\frac{1}{2}\,\transp{[y-M\,v]}\,\inv{C}\,[y-M\,v]
+  \ln p(\data\given v,d_1,d_2) &=& 3\,\ln d_1 + 3\,\ln d_2 -\frac{6~\mathrm{or}~8}{2}\,\ln 2\pi
+  -\frac{1}{2}\,\ln||\Sigma|| \\ && \quad -\frac{1}{2}\,\transp{[y-M\,v]}\,\inv{\Sigma}\,[y-M\,v]
   \quad ,
@@
 
+Now we multiply together (add in the log) and complete the square.
+We get, in the original parameterization:
+\begin{eqnarray}
+  \nu &=& \inv{[\transp{M}\,\inv{\Sigma}\,M+\inv{V}]}\,\transp{M}\,\inv{\Sigma}\,y
+  \\
+  \Delta &=& 3\,\ln d_1 + 3\,\ln d_2 -\frac{9~\mathrm{or}~11}{2}\,\ln 2\pi
+  -\frac{1}{2}\,\ln||\Sigma|| -\frac{1}{2}\,\ln||V|| \\
+  &&\quad -\frac{1}{2}\,\transp{y}\,\inv{\Sigma}\,y -\frac{1}{2}\,\transp{\nu}\,\inv{[\transp{M}\,\inv{\Sigma}\,M+\inv{V}]}\,\nu
+  \\
+  \ln p(\data\given d_1,d_2) &=& [HOGG]
+  \quad ,
+\end{eqnarray}
+where....HOGG
+
 \end{document}

### 2016-09-29 paper/ms.tex
fix signs, other bugs, tweaks

@@
 \newcommand{\given}{\,|\,}
-\newcommand{\normal}{N}
+\newcommand{\normal}{{\mathcal{N}}}
 \newcommand{\dd}{\mathrm{d}}
@@
 \begin{align}
-  p(\bs{v}) &= \mathcal{N}(0,25)~\kms \\
+  p(\bs{v}) &= \normal(0,25)~\kms \\
   p(d) &= \frac{r^2}{2\,L^3} \, \exp\left(-r/L\right)
@@
 \end{eqnarray}
-where $x$ and $\nu$ are $D$-dimensional vectors, $A$ is a symmetric
+where $x$ and $\nu$ are $D$-dimensional vectors, $A$ is a positive definite
 matrix, $\Delta$ is a scalar, and the integral is implicitly over all
@@
   \\
-  C & = & \Delta - \frac{1}{2}\,\transp{\nu}\,\inv{A}\,\nu
+  \Delta & = & C - \frac{1}{2}\,\transp{\nu}\,\inv{A}\,\nu
   \quad .
@@
 
-Now in this case, we want to write down the probability of the data
-given the distance and velocity (the likelihood), and the prior pdf for
+Now in this case, we want to write down the probability density for the data
+given the distances and velocity (the likelihood), and the prior pdf for
 the velocity, and their product.
@@
 \end{eqnarray}
-where the factor of $d_1^2\,d_2^3$ converts to units of \gaia\ data
-from units of $y$.
+where the factor of $d_1^3\,d_2^3$ converts to units of per \gaia\ data
+from units of per $y$.
 
@@
 \begin{eqnarray}
-  \nu &=& \inv{[\transp{M}\,\inv{\Sigma}\,M+\inv{V}]}\,\transp{M}\,\inv{\Sigma}\,y
+  \inv{A} &=& [\transp{M}\,\inv{\Sigma}\,M+\inv{V}]
   \\
-  \Delta &=& 3\,\ln d_1 + 3\,\ln d_2 -\frac{9~\mathrm{or}~11}{2}\,\ln 2\pi
-  -\frac{1}{2}\,\ln||\Sigma|| -\frac{1}{2}\,\ln||V|| \\
-  &&\quad -\frac{1}{2}\,\transp{y}\,\inv{\Sigma}\,y -\frac{1}{2}\,\transp{\nu}\,\inv{[\transp{M}\,\inv{\Sigma}\,M+\inv{V}]}\,\nu
+  \nu &=& A\,\transp{M}\,\inv{\Sigma}\,y
   \\
-  \ln p(\data\given d_1,d_2) &=& [HOGG]
+  \Delta &=& -3\,\ln d_1 -3\,\ln d_2 +\frac{9~\mathrm{or}~11}{2}\,\ln 2\pi
+  +\frac{1}{2}\,\ln||\Sigma|| +\frac{1}{2}\,\ln||V|| \\
+  &&\quad +\frac{1}{2}\,\transp{y}\,\inv{\Sigma}\,y -\frac{1}{2}\,\transp{\nu}\,\inv{A}\,\nu
   \quad ,
 \end{eqnarray}
-where....HOGG
+where there are sign issues that Hogg doesn't like to think about.
+Finally, we marginalize out the velocity to get the marginalized
+likelihood conditioned on the two distances $d_1, d_2$:
+\begin{eqnarray}
+  \ln p(\data\given d_1,d_2)
+  &=& \frac{3}{2}\,\ln 2\pi +\frac{1}{2}\ln ||A|| -\Delta
+  \quad .
+\end{eqnarray}
 

### 2016-09-29 gwb/data.py
delete gaiatools, make data for only one source

@@ -1,0 +1,138 @@
+from __future__ import division, print_function
+
+# Third-party
+import astropy.coordinates as coord
+import astropy.units as u
+import numpy as np
+
+class TGASStar(object):
+
+    def __init__(self, tgas_row, rv=None, rv_err=None, metadata=None):
+        self._row = tgas_row
+        self._cov = None # for caching
+        self._Cinv = None # for caching
+
+        # radial velocities
+        if rv is not None:
+            if rv_err is None:
+                raise ValueError("If radial velocity is provided, you must also "
+                                 "provide an error.")
+            self._rv = rv.to(u.km/u.s).value
+            self._rv_err = rv_err.to(u.km/u.s).value
+        else:
+            self._rv = 0.
+            self._rv_err = None
+
+    def __getitem__(self, slc):
+        return self._row[slc]
+
+    # Astrometric data as attributes:
+    @property
+    def _ra(self):
+        return np.radians(self['ra'])
+
+    @property
+    def ra(self):
+        return (self._ra*u.radian).to(u.degree)
+
+    @property
+    def _dec(self):
+        return np.radians(self['dec'])
+
+    @property
+    def dec(self):
+        return (self._dec*u.radian).to(u.degree)
+
+    @property
+    def _parallax(self):
+        return self['parallax'] # mas
+
+    @property
+    def parallax(self):
+        return self._parallax*u.mas
+
+    @property
+    def _pmra(self):
+        return self['pmra']
+
+    @property
+    def pmra(self):
+        return self._pmra*u.mas/u.yr
+
+    @property
+    def _pmdec(self):
+        return self['pmdec']
+
+    @property
+    def pmdec(self):
+        return self._pmdec*u.mas/u.yr
+
+    @property
+    def rv(self):
+        return self._rv*u.km/u.s
+
+    # Other useful things
+    def get_coord(self, with_parallax=False):
+        if with_parallax:
+            return coord.SkyCoord(ra=self['ra']*u.degree,
+                                  dec=self['dec']*u.degree,
+                                  distance=(1000./self['parallax'])*u.pc)
+
+        else:
+            return coord.SkyCoord(ra=self['ra']*u.degree,
+                                  dec=self['dec']*u.degree)
+
+    def get_cov(self):
+        """
+        The Gaia TGAS data table contains correlation coefficients and standard
[…]

### 2016-09-29 gwb/gaiatools.py

@@ -1,145 +1,0 @@
-from __future__ import division, print_function
-
-# Third-party
-import astropy.coordinates as coord
-import astropy.units as u
-import numpy as np
-
-class TGASData(object):
-
-    def __init__(self, tbl):
-        self.tbl = tbl
-        self._cov = None # for caching
-        self._Cinv = None # for caching
-
-    def __getitem__(self, slc):
-        return TGASData(self.tbl[slc])
-
-    # Astrometric data as attributes:
-    @property
-    def _ra(self):
-        return np.radians(self.tbl['ra'])
-
-    @property
-    def ra(self):
-        return (self._ra*u.radian).to(u.degree)
-
-    @property
-    def _dec(self):
-        return np.radians(self.tbl['dec'])
-
-    @property
-    def dec(self):
-        return (self._dec*u.radian).to(u.degree)
-
-    @property
-    def _parallax(self):
-        return self.tbl['parallax'] # mas
-
-    @property
-    def parallax(self):
-        return self._parallax*u.mas
-
-    @property
-    def _pmra(self):
-        return self.tbl['pmra']
-
-    @property
-    def pmra(self):
-        return self._pmra*u.mas/u.yr
-
-    @property
-    def _pmdec(self):
-        return self.tbl['pmdec']
-
-    @property
-    def pmdec(self):
-        return self._pmdec*u.mas/u.yr
-
-    @property
-    def _rv(self):
-        return 0. # TODO:
-
-    @property
-    def rv(self):
-        return self._rv*u.km/u.s
-
-    # Other useful things
-    def get_coord(self, with_parallax=False):
-        if with_parallax:
-            return coord.SkyCoord(ra=self.tbl['ra']*u.degree,
-                                  dec=self.tbl['dec']*u.degree,
-                                  distance=(1000./self.tbl['parallax'])*u.pc)
-
-        else:
-            return coord.SkyCoord(ra=self.tbl['ra']*u.degree,
-                                  dec=self.tbl['dec']*u.degree)
-
-    def get_cov(self, units=None):
-        """
-        The Gaia TGAS data table contains correlation coefficients and standard
-        deviations for (ra, dec, parallax, pm_ra, pm_dec), but for most analysis
-        we need covariance matrices. This converts the Gaia table into covariance
-        matrix.
-
-        Parameters
-        ----------
-        units : list (optional)
-            If None, returns the terms in the native Gaia release units
[…]

### 2016-09-29 gwb/likelihood.py

@@ -10,10 +10,11 @@ from .coords import get_tangent_basis
 pc_mas_yr_to_km_s = (1*u.pc * u.mas/u.yr).to(u.km/u.s,u.dimensionless_angles()).value
 
 def get_y(d, data):
+    d = np.atleast_1d(d)
     y = np.stack([d * data._parallax*1E-3 - 1.,
                   d * data._pmra * pc_mas_yr_to_km_s,
                   d * data._pmdec * pc_mas_yr_to_km_s,
-                  data._rv], axis=1)
+                  np.atleast_1d(data._rv)], axis=1)
     return y
 
 def get_M(data):

### 2016-09-29 gwb/tests/test_likelihood.py

@@ -6,14 +6,41 @@ import astropy.coordinates as coord
 import numpy as np
 
 # Project
-from ..gaiatools import TGASData
+from ..gaiatools import TGASStar
 from ..likelihood import get_y, get_M, get_Cinv, get_A_mu_Delta
 
-# make fake data
-n_data = 128
-ra = np.random.uniform(0, 2*np.pi, size=n_data)
-dec = np.pi/2. - np.arccos(2*np.random.uniform(size=n_data)-1.)
-parallax = np.random.uniform(-0.1, 1000., size=n_data)
-pmra,pmdec = np.random.normal(-0, 100, size=(2,n_data))
+def make_data():
+    # make fake data
+    n_data = 128
+    ra = np.random.uniform(0, 2*np.pi, size=n_data)
+    dec = np.pi/2. - np.arccos(2*np.random.uniform(size=n_data)-1.)
+    parallax = np.random.uniform(-0.1, 1000., size=n_data)
+    pmra,pmdec = np.random.normal(0, 100, size=(2,n_data))
 
-np.random.uniform([0.1,])
+    _M = np.random.uniform(0.1, 0.3, size=(n_data,5,5))
+    Cov = 0.5 * np.array([M.dot(M.T) for M in _M])
+    assert Cov.shape == (n_data,5,5)
+
+    all_data = []
+    for r,d,plx,pmr,pmd,C in zip(ra,dec,parallax,pmra,pmdec,Cov):
+        row = {
+            'ra': r, 'dec': d,
+            'parallax': plx, 'pmra': pmr, 'pmdec': pmd
+        }
+        data = TGASStar(row)
+        data._cov = C
+        all_data.append(data)
+
+    return all_data
+
+def test_y():
+    all_data = make_data()
+    for data in all_data:
+        d = 1000./data._parallax
+        y = get_y(d, data)
+        print(y.shape)
+        break
+
+# def test_M():
+#     data = make_data()
+#     M = get_M(data)

### 2016-09-29 gwb/likelihood.py
some dumb ass unit tests

@@ -9,35 +9,52 @@ from .coords import get_tangent_basis
 
 pc_mas_yr_to_km_s = (1*u.pc * u.mas/u.yr).to(u.km/u.s,u.dimensionless_angles()).value
 
-def get_y(d, data):
+def get_y(d, star):
     d = np.atleast_1d(d)
-    y = np.stack([d * data._parallax*1E-3 - 1.,
-                  d * data._pmra * pc_mas_yr_to_km_s,
-                  d * data._pmdec * pc_mas_yr_to_km_s,
-                  np.atleast_1d(data._rv)], axis=1)
+    y = np.stack([d * star._parallax*1E-3 - 1.,
+                  d * star._pmra * pc_mas_yr_to_km_s,
+                  d * star._pmdec * pc_mas_yr_to_km_s,
+                  np.atleast_1d(star._rv)], axis=1)
     return y
 
-def get_M(data):
-    M0 = np.zeros((3,) + data._ra.shape)
-    M_vel = get_tangent_basis(data._ra, data._dec)
-    return np.stack((M0[None], M_vel), axis=1)
+def get_M(star):
+    ra = np.atleast_1d(star._ra)
+    dec = np.atleast_1d(star._dec)
+    M0 = np.zeros((3,) + ra.shape)
+    M_vel = get_tangent_basis(ra, dec)
+    return np.hstack((M0, M_vel)).T[None]
 
-def get_Cinv(d, data):
-    Cinv = data.get_Cinv()
-    return Cinv[:3,:3] / d**2
+def get_Cinv(d, star):
+    Cinv = star.get_sub_cov_inv()
+    Cinv[:3,:3] /= d**2
+    return Cinv[None]
 
 def get_A_mu_Delta(M, Cinv, y, Vinv):
+    if Cinv.ndim > 2:
+        n = Cinv.shape[0]
+    else:
+        n = 1
+
+    # using ji vs. ij does the transpose of M
     # TODO: is A is actually the inverse of what you think it is?
-    Ainv = np.einsum('...ij,...jk,...ks->is', M.T, Cinv, M) + Vinv
+    Ainv = np.einsum('...ji,...jk,...ks->...is', M, Cinv, M) + Vinv
     A = np.linalg.inv(Ainv)
 
[…]

### 2016-09-29 gwb/tests/test_likelihood.py

@@ -6,8 +6,8 @@ import astropy.coordinates as coord
 import numpy as np
 
 # Project
-from ..gaiatools import TGASStar
-from ..likelihood import get_y, get_M, get_Cinv, get_A_mu_Delta
+from ..data import TGASStar
+from ..likelihood import get_y, get_M, get_Cinv, get_A_mu_Delta, ln_marg_likelihood
 
 def make_data():
     # make fake data
@@ -17,9 +17,11 @@ def make_data():
     parallax = np.random.uniform(-0.1, 1000., size=n_data)
     pmra,pmdec = np.random.normal(0, 100, size=(2,n_data))
 
-    _M = np.random.uniform(0.1, 0.3, size=(n_data,5,5))
+    _M = np.random.uniform(0.1, 0.3, size=(n_data,6,6))
     Cov = 0.5 * np.array([M.dot(M.T) for M in _M])
-    assert Cov.shape == (n_data,5,5)
+    _M[5] = 0.
+    _M[:,5] = 0.
+    assert Cov.shape == (n_data,6,6)
 
     all_data = []
     for r,d,plx,pmr,pmd,C in zip(ra,dec,parallax,pmra,pmdec,Cov):
@@ -38,9 +40,48 @@ def test_y():
     for data in all_data:
         d = 1000./data._parallax
         y = get_y(d, data)
-        print(y.shape)
-        break
+        assert y.shape == (1,4)
+        assert np.allclose(y[:,0], 0)
 
-# def test_M():
-#     data = make_data()
-#     M = get_M(data)
+def test_M():
+    all_data = make_data()
+    for data in all_data:
+        M = get_M(data)
+        assert M.shape == (1,4,3)
+        assert np.allclose(np.linalg.norm(M[0], axis=1), [0, 1, 1, 1])
+
+def test_Cinv():
+    all_data = make_data()
+    for data in all_data:
+        d = 1000./data._parallax
+        Cinv = get_Cinv(d, data)
+        assert Cinv.shape == (1,4,4)
+        assert np.allclose(Cinv[0], Cinv[0].T)
+
+def test_A_mu_Delta():
+    all_data = make_data()
+    for data in all_data:
+        d = 1000./data._parallax
+
+        M = get_M(data)
+        Cinv = get_Cinv(d, data)
+        y = get_y(d, data)
+        Vinv = np.diag([1/25.**2]*3)
+
+        A, mu, Delta = get_A_mu_Delta(M, Cinv, y, Vinv)
[…]

### 2016-09-29 paper/ms.tex
parent
comma
parens for clarity
Q case

@@
 
-In what follows, all vectors are column vectors.
+In what follows, all vectors are column vectors, unless we have transposed them.
 We indicate whether a symbol represents a scalar, vector, or tensor
@@
   \ln\left[\int\exp(-\frac{1}{2}\,\transp{[x-\nu]}\,\inv{A}\,[x-\nu] - \Delta)\,\dd x\right]
-  &=& \frac{D}{2}\,\ln 2\pi +\frac{1}{2}\ln ||A|| -\Delta
+  &=& +\frac{1}{2}\ln ||2\pi\,A|| -\Delta
   \quad ,
@@
   \\
-  \ln p(v) &=& -\frac{3}{2}\,\ln 2\pi -\frac{1}{2}\,\ln||V|| -\frac{1}{2}\,\transp{v}\,\inv{V}\,v
+  \ln p(v) &=& -\frac{1}{2}\,\ln||2\pi\,V|| -\frac{1}{2}\,\transp{v}\,\inv{V}\,v
   \\
@@
 \begin{eqnarray}
-  \transp{y} &\equiv& [(d_1\,\varpi_1 - 1), d_1\,\mu_{\alpha,1}, d_1\,\mu_{\delta,1},
-                       d_1\,\varpi_2 - 1, d_2\,\mu_{\alpha,2}, d_1\,\mu_{\delta,2}]
+  \transp{y} &\equiv& [(d_1\,\varpi_1 - 1), (d_1\,\mu_{\alpha1}), (d_1\,\mu_{\delta1}),
+                       (d_2\,\varpi_2 - 1), (d_2\,\mu_{\alpha2}), (d_2\,\mu_{\delta2})] \quad \mbox{or}
   \\
-  \transp{y} &\equiv& [(d_1\,\varpi_1 - 1), d_1\,\mu_{\alpha,1}, d_1\,\mu_{\delta,1}, v_{r,1}
-                       d_1\,\varpi_2 - 1, d_2\,\mu_{\alpha,2}, d_1\,\mu_{\delta,2}, v_{r,2}
+  \transp{y} &\equiv& [(d_1\,\varpi_1 - 1), (d_1\,\mu_{\alpha1}), (d_1\,\mu_{\delta1}), (v_{r1}),
+                       (d_2\,\varpi_2 - 1), (d_2\,\mu_{\alpha2}), (d_2\,\mu_{\delta2}), (v_{r2})]
   \quad ,
@@
 easy to write down.
-Fundamentally, our model is
+Fundamentally, our H1 model (same velocity) is
 \begin{eqnarray}
@@
   \transp{M} &=& [0, \hat{\alpha_1}, \hat{\delta_1},
-                  0, \hat{\alpha_2}, \hat{\delta_2}]
+                  0, \hat{\alpha_2}, \hat{\delta_2}] \quad \mbox{or}
   \\
@@
   \\
-  \ln p(\data\given v,d_1,d_2) &=& 3\,\ln d_1 + 3\,\ln d_2 -\frac{6~\mathrm{or}~8}{2}\,\ln 2\pi
-  -\frac{1}{2}\,\ln||\Sigma|| \\ && \quad -\frac{1}{2}\,\transp{[y-M\,v]}\,\inv{\Sigma}\,[y-M\,v]
+  \ln p(\data\given v,d_1,d_2) &=& 3\,\ln d_1 + 3\,\ln d_2
+  -\frac{1}{2}\,\ln||2\pi\,\Sigma|| \nonumber \\ && \quad
+  -\frac{1}{2}\,\transp{[y-M\,v]}\,\inv{\Sigma}\,[y-M\,v]
   \quad ,
@@
   \\
-  \Delta &=& -3\,\ln d_1 -3\,\ln d_2 +\frac{9~\mathrm{or}~11}{2}\,\ln 2\pi
-  +\frac{1}{2}\,\ln||\Sigma|| +\frac{1}{2}\,\ln||V|| \\
-  &&\quad +\frac{1}{2}\,\transp{y}\,\inv{\Sigma}\,y -\frac{1}{2}\,\transp{\nu}\,\inv{A}\,\nu
+  \Delta &=& -3\,\ln d_1 -3\,\ln d_2
+  +\frac{1}{2}\,\ln||2\pi\,\Sigma|| +\frac{1}{2}\,\ln||2\pi\,V|| \nonumber \\ && \quad
+  +\frac{1}{2}\,\transp{y}\,\inv{\Sigma}\,y -\frac{1}{2}\,\transp{\nu}\,\inv{A}\,\nu
   \quad ,
@@
   \ln p(\data\given d_1,d_2)
-  &=& \frac{3}{2}\,\ln 2\pi +\frac{1}{2}\ln ||A|| -\Delta
+  &=& \frac{1}{2}\ln ||2\pi\,A|| -\Delta
   \quad .
@@
 
+The story for the H2 model (independent velocities) is very
+similar. In this case, the marginalized likelihood is a product of
+two independent integrals $Q$ each constructed as follows:
+\begin{eqnarray}
+  \transp{y} &\equiv& [(d\,\varpi - 1), (d\,\mu_{\alpha}), (d\,\mu_{\delta})] \quad \mbox{or}
+  \\
+  \transp{y} &\equiv& [(d\,\varpi - 1), (d\,\mu_{\alpha}), (d\,\mu_{\delta}), (v_{r})]
+  \\
+  \transp{M} &=& [0, \hat{\alpha}, \hat{\delta}] \quad \mbox{or}
+  \\
+  \transp{M} &=& [0, \hat{\alpha}, \hat{\delta}_2, \hat{r}]
+  \\
+  \inv{A} &=& [\transp{M}\,\inv{\Sigma}\,M+\inv{V}]
+  \\
+  \nu &=& A\,\transp{M}\,\inv{\Sigma}\,y
+  \\
+  \Delta &=& -3\,\ln d
+  +\frac{1}{2}\,\ln||2\pi\,\Sigma|| +\frac{1}{2}\,\ln||2\pi\,V|| \nonumber \\ && \quad
+  +\frac{1}{2}\,\transp{y}\,\inv{\Sigma}\,y -\frac{1}{2}\,\transp{\nu}\,\inv{A}\,\nu
+  \\
+  Q &=& \frac{1}{2}\ln ||2\pi\,A|| -\Delta
+  \quad ,
+\end{eqnarray}
+where again there are two options for the dimensionality of $y$ and
+$M$, and once the vectors and $\Sigma$ are properly truncated to one
+star, everything else follows.
+
 \end{document}

### 2016-09-29 gwb/likelihood.py
change to nu and fix signs

@@ -29,7 +29,7 @@ def get_Cinv(d, star):
     Cinv[:3,:3] /= d**2
     return Cinv[None]
 
-def get_A_mu_Delta(M, Cinv, y, Vinv):
+def get_A_nu_Delta(d, M, Cinv, y, Vinv):
     if Cinv.ndim > 2:
         n = Cinv.shape[0]
     else:
@@ -42,7 +42,7 @@ def get_A_mu_Delta(M, Cinv, y, Vinv):
 
     # using ji vs. ij does the transpose
     Bb = np.einsum('...ji,...jk,...k->...i', M, Cinv, y)
-    mu = np.einsum('...ij,...j->...i', A, Bb)
+    nu = np.einsum('...ij,...j->...i', A, Bb)
 
     # do the right thing when Cinv[3,3] == 0
     idx = np.isclose(Cinv[...,3,3], 0)
@@ -54,15 +54,15 @@ def get_A_mu_Delta(M, Cinv, y, Vinv):
     sgn,log_detVinv = np.linalg.slogdet(Vinv/(2*np.pi))
 
     yT_Cinv_y = np.einsum('...i,...ji,...j->...', y, Cinv, y)
-    muT_A_mu = np.einsum('...i,...ji,...j->...', mu, A, mu)
-    Delta = 0.5*log_detCinv + 0.5*log_detVinv - 0.5*yT_Cinv_y - muT_A_mu
+    nuT_A_nu = np.einsum('...i,...ji,...j->...', nu, A, nu)
+    Delta = -3*np.log(d) - 0.5*log_detCinv - 0.5*log_detVinv + 0.5*yT_Cinv_y - nuT_A_nu
 
-    return A, mu, Delta
+    return A, nu, Delta
 
-def ln_marg_likelihood(d, data, Vinv):
+def ln_marg_likelihood_helper(d, data, Vinv):
     y = get_y(d, data)
     M = get_M(data)
     Cinv = get_Cinv(d, data)
-    A,mu,Delta = get_A_mu_Delta(M, Cinv, y, Vinv)
-    _,log_Adet = np.linalg.slogdet(A/(2*np.pi))
-    return -Delta + 0.5*log_Adet # TODO: plus or minus? is A actually Ainv?
+    A,nu,Delta = get_A_nu_Delta(d, M, Cinv, y, Vinv)
+    _,log_detA = np.linalg.slogdet(A/(2*np.pi))
+    return -Delta + 0.5*log_detA # TODO: plus or minus? is A actually Ainv?

### 2016-09-29 gwb/tests/test_likelihood.py

@@ -1,13 +1,11 @@
 from __future__ import division, print_function
 
 # Third-party
-import astropy.units as u
-import astropy.coordinates as coord
 import numpy as np
 
 # Project
 from ..data import TGASStar
-from ..likelihood import get_y, get_M, get_Cinv, get_A_mu_Delta, ln_marg_likelihood
+from ..likelihood import get_y, get_M, get_Cinv, get_A_nu_Delta, ln_marg_likelihood_helper
 
 def make_data():
     # make fake data
@@ -58,7 +56,7 @@ def test_Cinv():
         assert Cinv.shape == (1,4,4)
         assert np.allclose(Cinv[0], Cinv[0].T)
 
-def test_A_mu_Delta():
+def test_A_nu_Delta():
     all_data = make_data()
     for data in all_data:
         d = 1000./data._parallax
@@ -68,20 +66,20 @@ def test_A_mu_Delta():
         y = get_y(d, data)
         Vinv = np.diag([1/25.**2]*3)
 
-        A, mu, Delta = get_A_mu_Delta(M, Cinv, y, Vinv)
+        A, nu, Delta = get_A_nu_Delta(d, M, Cinv, y, Vinv)
         assert A.shape == (1,3,3)
         assert np.isfinite(A).all()
 
-        assert mu.shape == (1,3)
-        assert np.isfinite(mu).all()
+        assert nu.shape == (1,3)
+        assert np.isfinite(nu).all()
 
         assert Delta.shape == (1,)
         assert np.isfinite(Delta).all()
 
-def test_marg_likelihood():
+def test_ln_marg_likelihood_helper():
     all_data = make_data()
     for data in all_data:
         d = 1000./data._parallax
         Vinv = np.diag([1/25.**2]*3)
-        ll = ln_marg_likelihood(d, data, Vinv)
+        ll = ln_marg_likelihood_helper(d, data, Vinv)
         assert np.isfinite(ll).all()

### 2016-09-29 gwb/likelihood.py
fix 2pi

@@ -36,7 +36,6 @@ def get_A_nu_Delta(d, M, Cinv, y, Vinv):
         n = 1
 
     # using ji vs. ij does the transpose of M
-    # TODO: is A is actually the inverse of what you think it is?
     Ainv = np.einsum('...ji,...jk,...ks->...is', M, Cinv, M) + Vinv
     A = np.linalg.inv(Ainv)
 
@@ -64,5 +63,5 @@ def ln_marg_likelihood_helper(d, data, Vinv):
     M = get_M(data)
     Cinv = get_Cinv(d, data)
     A,nu,Delta = get_A_nu_Delta(d, M, Cinv, y, Vinv)
-    _,log_detA = np.linalg.slogdet(A/(2*np.pi))
-    return -Delta + 0.5*log_detA # TODO: plus or minus? is A actually Ainv?
+    _,log_detA = np.linalg.slogdet(2*np.pi*A)
+    return 0.5*log_detA - Delta

### 2016-09-29 paper/ms.tex
title, abstract, keywords

@@
 
-\title{Wide binaries in Gaia DR1}
-\author{People}
+\title{Widely separated, comoving stars in \textsl{Gaia DR1}}
+\author{APW, SO, DWH, DNS, DSTN, others}
 
@@
 \begin{abstract}
-Blerg.
 % Context
+The primary sample of the \gaia\ \textsl{First Data Release} is the brand-new
+\textsl{Tycho-Gaia Astrometric Solution}. The precision and size of this sample
+creates an opportunity to find new binary stars and moving groups, especially
+rare binaries or sparsely populated groups.
 % Aims
+Here we seize this opportunity.
 % Methods
+We use a justified marginalized likelihood ratio test to separate
+pairs of stars with surprisingly similar three-space velocities from
+those consistent with being drawn from the field population.  Although
+we perform some visualizations using a (bias-corrected) inverse
+parallax as a distance, the likelihood test works in the observable
+space and uses the \gaia\ noise model responsibly.
 % Results
+We find pairs of comoving stars in abundance out to very wide
+separations, including separations of a few parsec! There does not
+seem to be any tidal-disruption feature in the separation distribution
+at sub-parsec scales. Pairs at separations beyond a
+parsec---whether they are bound in a multiple system or drifting
+apart---must be very short-lived. The photometric properties of the
+members of these pairs are consistent with youth. The prospects for
+testing stellar models and chemical abundance measurements are
+superficially discussed.
 \end{abstract}
@@
 \keywords{
-  methods: data analysis
+  binaries: visual
   ---
   methods: statistical
+  ---
+  open clusters and associations: general
+  ---
+  parallaxes
+  ---
+  proper motions
+  ---
+  stars: formation
 }

### 2016-09-29 paper/ms.tex

@@
 
-\title{Widely separated, comoving stars in \textsl{Gaia DR1}}
+\title{Widely separated, comoving stars in the \textsl{Gaia DR1 TGAS} sample}
 \author{APW, SO, DWH, DNS, DSTN, others}
@@
 pairs of stars with surprisingly similar three-space velocities from
-those consistent with being drawn from the field population.  Although
-we perform some visualizations using a (bias-corrected) inverse
-parallax as a distance, the likelihood test works in the observable
-space and uses the \gaia\ noise model responsibly.
+those consistent with being drawn independently from the field
+population.  Although we perform some visualizations using a
+(bias-corrected) inverse parallax as a distance, the likelihood test
+works in the observable space and uses the \gaia\ noise model
+responsibly.
 % Results

### 2016-09-29 gwb/likelihood.py
still not confident about the signs...

@@ -40,7 +40,7 @@ def get_A_nu_Delta(d, M, Cinv, y, Vinv):
     A = np.linalg.inv(Ainv)
 
     # using ji vs. ij does the transpose
-    Bb = np.einsum('...ji,...jk,...k->...i', M, Cinv, y)
+    Bb = -np.einsum('...ji,...jk,...k->...i', M, Cinv, y)
     nu = np.einsum('...ij,...j->...i', A, Bb)
 
     # do the right thing when Cinv[3,3] == 0
@@ -53,8 +53,8 @@ def get_A_nu_Delta(d, M, Cinv, y, Vinv):
     sgn,log_detVinv = np.linalg.slogdet(Vinv/(2*np.pi))
 
     yT_Cinv_y = np.einsum('...i,...ji,...j->...', y, Cinv, y)
-    nuT_A_nu = np.einsum('...i,...ji,...j->...', nu, A, nu)
-    Delta = -3*np.log(d) - 0.5*log_detCinv - 0.5*log_detVinv + 0.5*yT_Cinv_y - nuT_A_nu
+    nuT_Ainv_nu = np.einsum('...i,...ji,...j->...', nu, Ainv, nu)
+    Delta = -3*np.log(d) - 0.5*log_detCinv - 0.5*log_detVinv + 0.5*yT_Cinv_y - 0.5*nuT_Ainv_nu
 
     return A, nu, Delta
 

### 2016-09-29 gwb/tests/test_coords.py

@@ -15,7 +15,7 @@ def test_u_vec():
     dec = np.pi/2. - np.arccos(2*np.random.uniform(size=n_test)-1.)
 
     assert get_u_vec(ra[0], dec[0]).shape == (3,)
-    assert get_u_vec(ra, dec).shape == (n_test,3)
+    assert get_u_vec(ra, dec).shape == (3,n_test)
 
     for lon,lat in zip(ra,dec):
         usph = coord.UnitSphericalRepresentation(lon=lon*u.rad, lat=lat*u.rad)

### 2016-09-29 gwb/tests/test_likelihood.py

@@ -7,7 +7,7 @@ import numpy as np
 from ..data import TGASStar
 from ..likelihood import get_y, get_M, get_Cinv, get_A_nu_Delta, ln_marg_likelihood_helper
 
-def make_data():
+def make_random_data():
     # make fake data
     n_data = 128
     ra = np.random.uniform(0, 2*np.pi, size=n_data)
@@ -17,8 +17,8 @@ def make_data():
 
     _M = np.random.uniform(0.1, 0.3, size=(n_data,6,6))
     Cov = 0.5 * np.array([M.dot(M.T) for M in _M])
-    _M[5] = 0.
     _M[:,5] = 0.
+    _M[...,5] = 0.
     assert Cov.shape == (n_data,6,6)
 
     all_data = []
@@ -33,8 +33,29 @@ def make_data():
 
     return all_data
 
+def make_data_increase_uncertainties():
+    n_data = 8
+    ra = np.full(n_data, np.random.uniform(0, 2*np.pi))
+    dec = np.full(n_data, np.pi/2. - np.arccos(2*np.random.uniform()-1.))
+    parallax = np.ones(n_data)*10. # 100 pc
+    pmra = 25/4.740470463496208 * parallax
+    pmdec = -25/4.740470463496208 * parallax
+
+    all_data = []
+    for i,a in enumerate(np.linspace(-8,-1,n_data)):
+        Cov = np.diag(np.full(6, 2**a))**2
+        row = {
+            'ra': ra[i], 'dec': dec[i],
+            'parallax': parallax[i], 'pmra': pmra[i], 'pmdec': pmdec[i]
+        }
+        data = TGASStar(row)
+        data._cov = Cov
+        all_data.append(data)
+
+    return all_data
+
 def test_y():
-    all_data = make_data()
+    all_data = make_random_data()
     for data in all_data:
         d = 1000./data._parallax
         y = get_y(d, data)
@@ -42,44 +63,93 @@ def test_y():
         assert np.allclose(y[:,0], 0)
 
 def test_M():
-    all_data = make_data()
+    all_data = make_random_data()
     for data in all_data:
         M = get_M(data)
         assert M.shape == (1,4,3)
         assert np.allclose(np.linalg.norm(M[0], axis=1), [0, 1, 1, 1])
 
+    # make sure the projection matrix M is the same for all of these data
+    all_data = make_data_increase_uncertainties()
+    comp_M = None
+    for data in all_data:
[…]

### 2016-09-29 gwb/tests/test_likelihood.py
remove comments

@@ -101,23 +101,23 @@ def test_Cinv():
 def test_A_nu_Delta():
     Vinv = np.diag([1/25.**2]*3)
 
-    # all_data = make_random_data()
-    # for data in all_data:
-    #     d = 1000./data._parallax
+    all_data = make_random_data()
+    for data in all_data:
+        d = 1000./data._parallax
 
-    #     M = get_M(data)
-    #     Cinv = get_Cinv(d, data)
-    #     y = get_y(d, data)
+        M = get_M(data)
+        Cinv = get_Cinv(d, data)
+        y = get_y(d, data)
 
-    #     A, nu, Delta = get_A_nu_Delta(d, M, Cinv, y, Vinv)
-    #     assert A.shape == (1,3,3)
-    #     assert np.isfinite(A).all()
+        A, nu, Delta = get_A_nu_Delta(d, M, Cinv, y, Vinv)
+        assert A.shape == (1,3,3)
+        assert np.isfinite(A).all()
 
-    #     assert nu.shape == (1,3)
-    #     assert np.isfinite(nu).all()
+        assert nu.shape == (1,3)
+        assert np.isfinite(nu).all()
 
-    #     assert Delta.shape == (1,)
-    #     assert np.isfinite(Delta).all()
+        assert Delta.shape == (1,)
+        assert np.isfinite(Delta).all()
 
     # as uncertainties increase, nu^T A^-1 nu should *increase*
     all_data = make_data_increase_uncertainties()

### 2016-09-29 gwb/likelihood.py
added H1 and H2 likelihood funcs

@@ -58,10 +58,33 @@ def get_A_nu_Delta(d, M, Cinv, y, Vinv):
 
     return A, nu, Delta
 
-def ln_marg_likelihood_helper(d, data, Vinv):
+def ln_H1_marg_likelihood(d1, d2, data1, data2, Vinv):
+    d = np.array([d1, d2])
+
+    y1 = get_y(d1, data1)
+    y2 = get_y(d2, data2)
+    y = np.vstack((y1, y2))
+
+    M1 = get_M(data1)
+    M2 = get_M(data2)
+    M = np.vstack((M1,M2))
+
+    Cinv1 = get_Cinv(d1, data1)
+    Cinv2 = get_Cinv(d2, data2)
+    Cinv = np.vstack((Cinv1,Cinv2))
+
+    A,nu,Delta = get_A_nu_Delta(d, M, Cinv, y, Vinv)
+    _,log_detA = np.linalg.slogdet(2*np.pi*A)
+    return (0.5*log_detA - Delta).sum()
+
+def ln_H2_marg_likelihood_helper(d, data, Vinv):
     y = get_y(d, data)
     M = get_M(data)
     Cinv = get_Cinv(d, data)
     A,nu,Delta = get_A_nu_Delta(d, M, Cinv, y, Vinv)
     _,log_detA = np.linalg.slogdet(2*np.pi*A)
     return 0.5*log_detA - Delta
+
+def ln_H2_marg_likelihood(d1, d2, data1, data2, Vinv):
+    return (ln_H2_marg_likelihood_helper(d1, data1, Vinv) +
+            ln_H2_marg_likelihood_helper(d2, data2, Vinv))[0]

### 2016-09-29 gwb/tests/test_likelihood.py

@@ -5,7 +5,10 @@ import numpy as np
 
 # Project
 from ..data import TGASStar
-from ..likelihood import get_y, get_M, get_Cinv, get_A_nu_Delta, ln_marg_likelihood_helper
+from ..coords import get_tangent_basis
+from ..likelihood import (get_y, get_M, get_Cinv, get_A_nu_Delta,
+                          ln_H1_marg_likelihood, ln_H2_marg_likelihood_helper,
+                          ln_H2_marg_likelihood)
 
 def make_random_data():
     # make fake data
@@ -53,6 +56,38 @@ def make_data_increase_uncertainties():
         all_data.append(data)
 
     return all_data
+
+def make_random_pairs():
+    n_data = 128
+
+    for i in range(n_data):
+        ra1,ra2 = np.random.uniform(0, 2*np.pi, size=2)
+        dec1,dec2 = np.pi/2. - np.arccos(2*np.random.uniform(size=2)-1.)
+        plx1,plx2 = np.random.uniform(1, 1000., size=2)
+
+        true_v = np.random.normal(0, 25, size=3)
+        T1 = get_tangent_basis(ra1, dec1)
+        v1 = T1.dot(true_v)
+        pmra1,pmdec1 = v1[:2] * plx1 / 4.740470463496208
+        # TODO: ignoring RV for now
+
+        T2 = get_tangent_basis(ra2, dec2)
+        v2 = T2.dot(true_v)
+        pmra2,pmdec2 = v2[:2] * plx2 / 4.740470463496208
+        # TODO: ignoring RV for now
+
+        Cov = np.random.uniform(0.1, 0.3, size=(6,6))
+        Cov[5] = 0. # TODO: ignoring RV for now
+        Cov[:,5] = 0. # TODO: ignoring RV for now
+
+        # HACK: they have the same covariance matrix
+        star1 = TGASStar({'ra': ra1, 'dec': dec1, 'parallax': plx1, 'pmra': pmra1, 'pmdec': pmdec1})
+        star1._cov = Cov
+
+        star2 = TGASStar({'ra': ra2, 'dec': dec2, 'parallax': plx2, 'pmra': pmra2, 'pmdec': pmdec2})
+        star2._cov = Cov
+
+        yield [star1, star2]
 
 def test_y():
     all_data = make_random_data()
@@ -142,14 +177,42 @@ def test_A_nu_Delta():
[…]

### 2016-09-29 gwb/tests/test_likelihood.py
before refactor

@@ -216,3 +216,5 @@ def test_ln_likelihood_ratio():
         H1 = ln_H1_marg_likelihood(d1, d2, pair[0], pair[1], Vinv)
         H2 = ln_H2_marg_likelihood(d1, d2, pair[0], pair[1], Vinv)
         print(H1, H2)
+
+        break

### 2016-09-29 gwb/likelihood.py
refactor to do what we wrote on the board

@@ -3,78 +3,86 @@ from __future__ import division, print_function
 # Third-party
 import astropy.units as u
 import numpy as np
+from scipy.linalg import block_diag
 
 # Project
 from .coords import get_tangent_basis
 
 pc_mas_yr_to_km_s = (1*u.pc * u.mas/u.yr).to(u.km/u.s,u.dimensionless_angles()).value
 
-def get_y(d, star):
-    d = np.atleast_1d(d)
-    y = np.stack([d * star._parallax*1E-3 - 1.,
-                  d * star._pmra * pc_mas_yr_to_km_s,
-                  d * star._pmdec * pc_mas_yr_to_km_s,
-                  np.atleast_1d(star._rv)], axis=1)
+def get_y(ds, stars):
+    ds = np.atleast_1d(ds)
+    stars = np.atleast_1d(stars)
+
+    y = np.hstack([[d * star._parallax*1E-3 - 1.,
+                    d * star._pmra * pc_mas_yr_to_km_s,
+                    d * star._pmdec * pc_mas_yr_to_km_s,
+                    star._rv] for d,star in zip(ds,stars)])
     return y
 
-def get_M(star):
-    ra = np.atleast_1d(star._ra)
-    dec = np.atleast_1d(star._dec)
-    M0 = np.zeros((3,) + ra.shape)
-    M_vel = get_tangent_basis(ra, dec)
-    return np.hstack((M0, M_vel)).T[None]
+def get_M(stars):
+    stars = np.atleast_1d(stars)
+    M0 = np.zeros(3)
 
-def get_Cinv(d, star):
-    Cinv = star.get_sub_cov_inv()
-    Cinv[:3,:3] /= d**2
-    return Cinv[None]
+    M = [np.vstack((M0,get_tangent_basis(star._ra, star._dec)))
+         for star in stars]
+
+    return np.vstack(M)
+
+def get_Cinv(ds, stars):
+    ds = np.atleast_1d(ds)
+    stars = np.atleast_1d(stars)
+
+    Cinvs = []
+    for d,star in zip(ds,stars):
+        Cinv = star.get_sub_cov_inv()
+        Cinv[:3,:3] /= d**2
[…]

### 2016-09-29 gwb/tests/test_likelihood.py

@@ -27,7 +27,7 @@ def make_random_data():
     all_data = []
     for r,d,plx,pmr,pmd,C in zip(ra,dec,parallax,pmra,pmdec,Cov):
         row = {
-            'ra': r, 'dec': d,
+            'ra': np.degrees(r), 'dec': np.degrees(d),
             'parallax': plx, 'pmra': pmr, 'pmdec': pmd
         }
         data = TGASStar(row)
@@ -48,7 +48,7 @@ def make_data_increase_uncertainties():
     for i,a in enumerate(np.linspace(-8,-1,n_data)):
         Cov = np.diag(np.full(6, 2**a))**2
         row = {
-            'ra': ra[i], 'dec': dec[i],
+            'ra': np.degrees(ra[i]), 'dec': np.degrees(dec[i]),
             'parallax': parallax[i], 'pmra': pmra[i], 'pmdec': pmdec[i]
         }
         data = TGASStar(row)
@@ -61,9 +61,17 @@ def make_random_pairs():
     n_data = 128
 
     for i in range(n_data):
-        ra1,ra2 = np.random.uniform(0, 2*np.pi, size=2)
-        dec1,dec2 = np.pi/2. - np.arccos(2*np.random.uniform(size=2)-1.)
-        plx1,plx2 = np.random.uniform(1, 1000., size=2)
+        # ra1,ra2 = np.random.uniform(0, 2*np.pi, size=2)
+        # dec1,dec2 = np.pi/2. - np.arccos(2*np.random.uniform(size=2)-1.)
+        # plx1,plx2 = np.random.uniform(1, 1000., size=2)
+        ra1 = 0.1
+        ra2 = 0.11
+
+        dec1 = -0.57
+        dec2 = -0.56
+
+        plx1 = 1000.
+        plx2 = 1000.
 
         true_v = np.random.normal(0, 25, size=3)
         T1 = get_tangent_basis(ra1, dec1)
@@ -76,7 +84,7 @@ def make_random_pairs():
         pmra2,pmdec2 = v2[:2] * plx2 / 4.740470463496208
         # TODO: ignoring RV for now
 
-        Cov = np.random.uniform(0.1, 0.3, size=(6,6))
+        Cov = np.diag(np.random.uniform(0.1, 0.3, size=6))**2
         Cov[5] = 0. # TODO: ignoring RV for now
         Cov[:,5] = 0. # TODO: ignoring RV for now
 
@@ -89,20 +97,29 @@ def make_random_pairs():
 
         yield [star1, star2]
 
[…]

### 2016-09-29 setup.py
Add setup script and travis config

@@ -1,0 +1,44 @@
+#! /usr/bin/env python
+
+# Standard library
+import sys
+
+try:
+    from setuptools import setup
+except ImportError:
+    from distutils.core import setup
+
+# Thanks DFM: Handle encoding
+major, minor1, minor2, release, serial = sys.version_info
+if major >= 3:
+    def rd(filename):
+        f = open(filename, encoding="utf-8")
+        r = f.read()
+        f.close()
+        return r
+else:
+    def rd(filename):
+        f = open(filename)
+        r = f.read()
+        f.close()
+        return r
+
+setup(
+    name="gwb",
+    version='v0.1',
+    author="Semyeong Oh, Adrian Price-Whelan, David W. Hogg",
+    author_email="adrn@princeton.edu",
+    packages=["gwb", "gwb.tests"],
+    url="https://github.com/adrn/schwimmbad",
+    license="MIT",
+    description="Gaia wide binaries.",
+    long_description=rd("README.md"),
+    package_data={"": ["LICENSE"]},
+    include_package_data=True,
+    classifiers=[
+        "Intended Audience :: Science/Research",
+        "License :: OSI Approved :: MIT License",
+        "Operating System :: OS Independent",
+        "Programming Language :: Python",
+    ],
+)

### 2016-09-29 gwb/likelihood.py
switch back to using einsum

@@ -47,15 +47,12 @@ def get_A_nu_Delta(d, M, Cinv, y, Vinv):
     d = np.atleast_1d(d)
 
     # using ji vs. ij does the transpose of M
-    # Ainv = np.einsum('ji,jk,ks->is', M, Cinv, M) + Vinv
-    Ainv = (M.T @ (Cinv @ M)) + Vinv
+    Ainv = np.einsum('ji,jk,ks->is', M, Cinv, M) + Vinv
     A = np.linalg.inv(Ainv)
 
     # using ji vs. ij does the transpose
-    # Bb = np.einsum('ji,jk,k->i', M, Cinv, y)
-    # nu = np.einsum('ij,j->i', A, Bb)
-    Bb = M.T @ (Cinv @ y)
-    nu = A @ Bb
+    Bb = np.einsum('ji,jk,k->i', M, Cinv, y)
+    nu = np.einsum('ij,j->i', A, Bb)
 
     # do the right thing when Cinv == 0 for RV's
     idx, = np.where(np.isclose(np.diag(Cinv), 0.))

### 2016-09-29 gwb/tests/test_likelihood.py
added shitty tests

@@ -57,23 +57,59 @@ def make_data_increase_uncertainties():
 
     return all_data
 
-def make_random_pairs():
+def make_bad_pairs():
     n_data = 128
 
     for i in range(n_data):
-        # ra1,ra2 = np.random.uniform(0, 2*np.pi, size=2)
-        # dec1,dec2 = np.pi/2. - np.arccos(2*np.random.uniform(size=2)-1.)
-        # plx1,plx2 = np.random.uniform(1, 1000., size=2)
-        ra1 = 0.1
-        ra2 = 0.11
-
-        dec1 = -0.57
-        dec2 = -0.56
-
-        plx1 = 1000.
-        plx2 = 1000.
+        ra1,ra2 = np.random.uniform(0, 2*np.pi, size=2)
+        dec1,dec2 = np.pi/2. - np.arccos(2*np.random.uniform(size=2)-1.)
+        plx1,plx2 = np.exp(np.random.uniform(0,2,size=2))
 
         true_v = np.random.normal(0, 25, size=3)
+        T1 = get_tangent_basis(ra1, dec1)
+        v1 = T1.T.dot(true_v)
+        pmra1,pmdec1 = v1[:2] * plx1 / 4.740470463496208
+        # TODO: ignoring RV for now
+
+        true_v = np.random.normal(0, 25, size=3)
+        T2 = get_tangent_basis(ra2, dec2)
+        v2 = T2.T.dot(true_v)
+        pmra2,pmdec2 = v2[:2] * plx2 / 4.740470463496208
+        # TODO: ignoring RV for now
+
+        Cov = np.diag(np.random.uniform(0.1, 0.3, size=6))**2
+        Cov[5] = 0. # TODO: ignoring RV for now
+        Cov[:,5] = 0. # TODO: ignoring RV for now
+
+        # HACK: they have the same covariance matrix
+        star1 = TGASStar({'ra': np.degrees(ra1), 'dec': np.degrees(dec1),
+                          'parallax': plx1, 'pmra': pmra1, 'pmdec': pmdec1})
+        star1._cov = Cov
+
+        star2 = TGASStar({'ra': np.degrees(ra2), 'dec': np.degrees(dec2),
+                          'parallax': plx2, 'pmra': pmra2, 'pmdec': pmdec2})
+        star2._cov = Cov
+
+        yield [star1, star2]
+
+def make_good_pairs():
+    n_data = 128
+
[…]

### 2016-09-29 gwb/tests/test_likelihood.py
print difference

@@ -120,8 +120,7 @@ def make_good_pairs():
         pmra2,pmdec2 = v2[:2] * plx2 / 4.740470463496208
         # TODO: ignoring RV for now
 
-        # Cov = np.diag(np.random.uniform(0.1, 0.3, size=6))**2
-        Cov = np.diag(np.full(6,1E-6))**2
+        Cov = np.diag(np.random.uniform(0.1, 0.3, size=6))**2
         Cov[5] = 0. # TODO: ignoring RV for now
         Cov[:,5] = 0. # TODO: ignoring RV for now
 
@@ -274,7 +273,7 @@ def test_ln_likelihood_ratio():
 
         H1 = ln_H1_marg_likelihood(d1, d2, pair[0], pair[1], Vinv)
         H2 = ln_H2_marg_likelihood(d1, d2, pair[0], pair[1], Vinv)
-        print(H1, H2)
+        print(H1 - H2)
 
         break
 
@@ -284,6 +283,6 @@ def test_ln_likelihood_ratio():
 
         H1 = ln_H1_marg_likelihood(d1, d2, pair[0], pair[1], Vinv)
         H2 = ln_H2_marg_likelihood(d1, d2, pair[0], pair[1], Vinv)
-        print(H1, H2)
+        print(H1 - H2)
 
         break

### 2016-09-30 gwb/data.py
fixing typos

@@ -102,7 +102,7 @@ class TGASStar(object):
         # pre-load the diagonal
         for i,name in enumerate(names):
             full_name = "{}_error".format(name)
-            C[:,i,i] = self.tbl[full_name]**2
+            C[i,i] = self._row[full_name]**2
 
         for i,name1 in enumerate(names):
             for j,name2 in enumerate(names):

### 2016-09-30 gwb/likelihood.py

@@ -14,7 +14,7 @@ def get_y(ds, stars):
     ds = np.atleast_1d(ds)
     stars = np.atleast_1d(stars)
 
-    y = np.hstack([[d * star._parallax*1E-3 - 1.,
+    y = np.hstack([[d * star._parallax - 1000., # units
                     d * star._pmra * pc_mas_yr_to_km_s,
                     d * star._pmdec * pc_mas_yr_to_km_s,
                     star._rv] for d,star in zip(ds,stars)])
@@ -41,33 +41,35 @@ def get_Cinv(ds, stars):
 
     return block_diag(*Cinvs)
 
-def get_A_nu_Delta(d, M, Cinv, y, Vinv):
-    assert Cinv.ndim == 2
+def get_Ainv_nu_Delta(d, M_dirty, Cinv_dirty, y_dirty, Vinv):
+    assert Cinv_dirty.ndim == 2
 
     d = np.atleast_1d(d)
 
+    # do the right thing when Cinv == 0 for RV's
+    idx, = np.where(np.isclose(np.diag(Cinv_dirty), 0.))
+    Cinv = np.delete(Cinv_dirty, idx, axis=0)
+    Cinv = np.delete(Cinv, idx, axis=1)
+    _,log_detCinv = np.linalg.slogdet(Cinv/(2*np.pi))
+
+    M = np.delete(M_dirty, idx, axis=0)
+    y = np.delete(y_dirty, idx, axis=0)
+
     # using ji vs. ij does the transpose of M
     Ainv = np.einsum('ji,jk,ks->is', M, Cinv, M) + Vinv
-    A = np.linalg.inv(Ainv)
 
     # using ji vs. ij does the transpose
     Bb = np.einsum('ji,jk,k->i', M, Cinv, y)
-    nu = np.einsum('ij,j->i', A, Bb)
-
-    # do the right thing when Cinv == 0 for RV's
-    idx, = np.where(np.isclose(np.diag(Cinv), 0.))
-    Cinv_clean = np.delete(Cinv, idx, axis=0)
-    Cinv_clean = np.delete(Cinv_clean, idx, axis=1)
-    _,log_detCinv = np.linalg.slogdet(Cinv_clean/(2*np.pi))
+    nu = -np.linalg.solve(Ainv, Bb)
 
[…]

### 2016-09-30 gwb/tests/test_likelihood.py

@@ -121,6 +121,7 @@ def make_good_pairs():
         # TODO: ignoring RV for now
 
         Cov = np.diag(np.random.uniform(0.1, 0.3, size=6))**2
+        # Cov = np.diag(np.full(6,1E-2))**2
         Cov[5] = 0. # TODO: ignoring RV for now
         Cov[:,5] = 0. # TODO: ignoring RV for now
 
@@ -254,6 +255,8 @@ def test_ln_H1_marg_likelihood():
         d2 = 1000/pair[1]._parallax
 
         ll = ln_H1_marg_likelihood(d1, d2, pair[0], pair[1], Vinv)
+        print(ll)
+        return
         assert np.isfinite(ll)
 
 def test_ln_H2_marg_likelihood():
@@ -264,6 +267,23 @@ def test_ln_H2_marg_likelihood():
 
         ll = ln_H2_marg_likelihood(d1, d2, pair[0], pair[1], Vinv)
         assert np.isfinite(ll)
+
+def test_plot_shape_of_Q():
+    Vinv = np.diag([1/25.**2]*3)
+    for pair in make_good_pairs():
+        d1 = 1000/pair[0]._parallax
+        print(d1, pair[0]._parallax)
+        d_grid = np.linspace(0.1, 8*d1, 256)
+
+        ll = np.zeros_like(d_grid)
+        for i,d in enumerate(d_grid):
+            ll[i] = ln_H2_marg_likelihood_helper(d, pair[0], Vinv)
+
+        import matplotlib.pyplot as plt
+        plt.plot(d_grid, np.exp(ll-ll.max()))
+        plt.show()
+
+        return
 
 def test_ln_likelihood_ratio():
     Vinv = np.diag([1/25.**2]*3)
@@ -284,5 +304,26 @@ def test_ln_likelihood_ratio():
         H1 = ln_H1_marg_likelihood(d1, d2, pair[0], pair[1], Vinv)
         H2 = ln_H2_marg_likelihood(d1, d2, pair[0], pair[1], Vinv)
         print(H1 - H2)
+
+        d1_grid,d2_grid = np.meshgrid(np.linspace(-3, np.log10(16*d1), 128),
+                                      np.logspace(-3, np.log10(16*d2), 128))
+        d1_grid = d1_grid.ravel()
+        d2_grid = d2_grid.ravel()
+
+        HH1 = np.zeros_like(d1_grid)
+        HH2 = np.zeros_like(d1_grid)
[…]

### 2016-09-30 gwb/likelihood.py
we've now removed distance/parallax from all vectors/matrices

@@ -14,17 +14,15 @@ def get_y(ds, stars):
     ds = np.atleast_1d(ds)
     stars = np.atleast_1d(stars)
 
-    y = np.hstack([[d * star._parallax - 1000., # units
-                    d * star._pmra * pc_mas_yr_to_km_s,
+    y = np.hstack([[d * star._pmra * pc_mas_yr_to_km_s,
                     d * star._pmdec * pc_mas_yr_to_km_s,
                     star._rv] for d,star in zip(ds,stars)])
     return y
 
 def get_M(stars):
     stars = np.atleast_1d(stars)
-    M0 = np.zeros(3)
 
-    M = [np.vstack((M0,get_tangent_basis(star._ra, star._dec)))
+    M = [get_tangent_basis(star._ra, star._dec)
          for star in stars]
 
     return np.vstack(M)
@@ -37,7 +35,7 @@ def get_Cinv(ds, stars):
     for d,star in zip(ds,stars):
         Cinv = star.get_sub_cov_inv()
         Cinv[:3,:3] /= d**2
-        Cinvs.append(Cinv)
+        Cinvs.append(Cinv[1:,1:])
 
     return block_diag(*Cinvs)
 
@@ -66,7 +64,7 @@ def get_Ainv_nu_Delta(d, M_dirty, Cinv_dirty, y_dirty, Vinv):
 
     yT_Cinv_y = np.einsum('i,ji,j->', y, Cinv, y)
     nuT_Ainv_nu = np.einsum('i,ji,j->', nu, Ainv, nu)
-    Delta = (-sum([3*np.log(dd) for dd in d]) - 0.5*log_detCinv - 0.5*log_detVinv
+    Delta = (-sum([2*np.log(dd) for dd in d]) - 0.5*log_detCinv - 0.5*log_detVinv
              + 0.5*yT_Cinv_y - 0.5*nuT_Ainv_nu)
 
     return Ainv, nu, Delta

### 2016-09-30 gwb/__init__.py
imports

@@ -1,0 +1,4 @@
+from .coords import *
+from .data import *
+from .fml import *
+from .likelihood import *

### 2016-09-30 gwb/data.py
modify to only return velocity submatrix

@@ -5,6 +5,8 @@ import astropy.coordinates as coord
 import astropy.units as u
 import numpy as np
 
+__all__ = ['TGASStar']
+
 class TGASStar(object):
 
     def __init__(self, tgas_row, rv=None, rv_err=None, metadata=None):
@@ -123,16 +125,16 @@ class TGASStar(object):
         Return the sub-matrix of the full Gaia covariance matrix that
         just contains parallax, proper motion, and radial velocity terms.
         """
-        return self.get_cov()[2:,2:]
+        return self.get_cov()[3:,3:]
 
     def get_sub_cov_inv(self):
         """
-        Return the inverse of the sub-matrix over parallax, proper motion,
-        and radial velocity terms.
+        Return the inverse of the sub-matrix over velocity terms
+        (proper motion and radial velocity).
         """
-        Cinv = np.zeros((4,4))
+        Cinv = np.zeros((3,3))
         if self._rv_err is None:
-            Cinv[:3,:3] = np.linalg.inv(self.get_sub_cov()[:-1,:-1])
+            Cinv[:2,:2] = np.linalg.inv(self.get_sub_cov()[:-1,:-1])
         else:
             Cinv = np.linalg.inv(self.get_sub_cov())
         return Cinv

### 2016-09-30 gwb/likelihood.py
rename some functions and split likelihood tests

@@ -8,6 +8,9 @@ from scipy.linalg import block_diag
 # Project
 from .coords import get_tangent_basis
 
+__all__ = ['get_y', 'get_M', 'get_Cinv', 'get_Ainv_nu_Delta',
+           'ln_H1_marg_v_likelihood', 'ln_Q', 'ln_H2_marg_v_likelihood']
+
 pc_mas_yr_to_km_s = (1*u.pc * u.mas/u.yr).to(u.km/u.s,u.dimensionless_angles()).value
 
 def get_y(ds, stars):
@@ -34,14 +37,12 @@ def get_Cinv(ds, stars):
     Cinvs = []
     for d,star in zip(ds,stars):
         Cinv = star.get_sub_cov_inv()
-        Cinv[:3,:3] /= d**2
-        Cinvs.append(Cinv[1:,1:])
+        Cinv[:2,:2] /= d**2
+        Cinvs.append(Cinv)
 
     return block_diag(*Cinvs)
 
 def get_Ainv_nu_Delta(d, M_dirty, Cinv_dirty, y_dirty, Vinv):
-    assert Cinv_dirty.ndim == 2
-
     d = np.atleast_1d(d)
 
     # do the right thing when Cinv == 0 for RV's
@@ -69,10 +70,7 @@ def get_Ainv_nu_Delta(d, M_dirty, Cinv_dirty, y_dirty, Vinv):
 
     return Ainv, nu, Delta
 
-def ln_H1_marg_likelihood(d1, d2, data1, data2, Vinv):
-    ds = np.array([d1, d2])
-    data = [data1, data2]
-
+def _marg_likelihood_helper(ds, data, Vinv):
     y = get_y(ds, data)
     M = get_M(data)
     Cinv = get_Cinv(ds, data)
@@ -83,15 +81,13 @@ def ln_H1_marg_likelihood(d1, d2, data1, data2, Vinv):
     assert sgn > 0
     return 0.5*log_detA - Delta
 
-def ln_H2_marg_likelihood_helper(d, data, Vinv):
-    y = get_y(d, data)
-    M = get_M(data)
-    Cinv = get_Cinv(d, data)
-    Ainv,nu,Delta = get_Ainv_nu_Delta(d, M, Cinv, y, Vinv)
-    sgn,log_detAinv = np.linalg.slogdet(Ainv/(2*np.pi))
-    log_detA = -log_detAinv
-    return 0.5*log_detA - Delta
[…]

### 2016-09-30 gwb/tests/test_likelihood_f.py

@@ -1,0 +1,334 @@
+"""
+
+    Functional tests for gwb/likelihood.py
+
+"""
+
+from __future__ import division, print_function
+
+# Third-party
+import numpy as np
+
+# Project
+from ..data import TGASStar
+from ..coords import get_tangent_basis
+from ..likelihood import (get_y, get_M, get_Cinv, get_Ainv_nu_Delta,
+                          ln_H1_marg_v_likelihood, ln_Q, ln_H2_marg_v_likelihood)
+
+def make_random_data():
+    # make fake data
+    n_data = 128
+    ra = np.random.uniform(0, 2*np.pi, size=n_data)
+    dec = np.pi/2. - np.arccos(2*np.random.uniform(size=n_data)-1.)
+    parallax = np.random.uniform(-0.1, 1000., size=n_data)
+    pmra,pmdec = np.random.normal(0, 100, size=(2,n_data))
+
+    _M = np.random.uniform(0.1, 0.3, size=(n_data,6,6))
+    Cov = 0.5 * np.array([M.dot(M.T) for M in _M])
+    _M[:,5] = 0.
+    _M[...,5] = 0.
+    assert Cov.shape == (n_data,6,6)
+
+    all_data = []
+    for r,d,plx,pmr,pmd,C in zip(ra,dec,parallax,pmra,pmdec,Cov):
+        row = {
+            'ra': np.degrees(r), 'dec': np.degrees(d),
+            'parallax': plx, 'pmra': pmr, 'pmdec': pmd
+        }
+        data = TGASStar(row)
+        data._cov = C
+        all_data.append(data)
+
+    return all_data
+
+def make_data_increase_uncertainties():
+    n_data = 8
+    ra = np.full(n_data, np.random.uniform(0, 2*np.pi))
+    dec = np.full(n_data, np.pi/2. - np.arccos(2*np.random.uniform()-1.))
+    parallax = np.ones(n_data)*10. # 100 pc
+    pmra = 25/4.740470463496208 * parallax
+    pmdec = -25/4.740470463496208 * parallax
+
+    all_data = []
+    for i,a in enumerate(np.linspace(-8,-1,n_data)):
+        Cov = np.diag(np.full(6, 2**a))**2
+        row = {
+            'ra': np.degrees(ra[i]), 'dec': np.degrees(dec[i]),
+            'parallax': parallax[i], 'pmra': pmra[i], 'pmdec': pmdec[i]
+        }
[…]

### 2016-09-30 gwb/tests/test_likelihood_u.py

@@ -1,3 +1,9 @@
+"""
+
+    Unit tests for gwb/likelihood.py
+
+"""
+
 from __future__ import division, print_function
 
 # Third-party
@@ -6,9 +12,8 @@ import numpy as np
 # Project
 from ..data import TGASStar
 from ..coords import get_tangent_basis
-from ..likelihood import (get_y, get_M, get_Cinv, get_A_nu_Delta,
-                          ln_H1_marg_likelihood, ln_H2_marg_likelihood_helper,
-                          ln_H2_marg_likelihood)
+from ..likelihood import (get_y, get_M, get_Cinv, get_Ainv_nu_Delta,
+                          ln_H1_marg_v_likelihood, ln_Q, ln_H2_marg_v_likelihood)
 
 def make_random_data():
     # make fake data

### 2016-09-30 gwb/tests/test_likelihood_u.py
fix tests to use new shapes

@@ -11,7 +11,6 @@ import numpy as np
 
 # Project
 from ..data import TGASStar
-from ..coords import get_tangent_basis
 from ..likelihood import (get_y, get_M, get_Cinv, get_Ainv_nu_Delta,
                           ln_H1_marg_v_likelihood, ln_Q, ln_H2_marg_v_likelihood)
 
@@ -41,106 +40,6 @@ def make_random_data():
 
     return all_data
 
-def make_data_increase_uncertainties():
-    n_data = 8
-    ra = np.full(n_data, np.random.uniform(0, 2*np.pi))
-    dec = np.full(n_data, np.pi/2. - np.arccos(2*np.random.uniform()-1.))
-    parallax = np.ones(n_data)*10. # 100 pc
-    pmra = 25/4.740470463496208 * parallax
-    pmdec = -25/4.740470463496208 * parallax
-
-    all_data = []
-    for i,a in enumerate(np.linspace(-8,-1,n_data)):
-        Cov = np.diag(np.full(6, 2**a))**2
-        row = {
-            'ra': np.degrees(ra[i]), 'dec': np.degrees(dec[i]),
-            'parallax': parallax[i], 'pmra': pmra[i], 'pmdec': pmdec[i]
-        }
-        data = TGASStar(row)
-        data._cov = Cov
-        all_data.append(data)
-
-    return all_data
-
-def make_bad_pairs():
-    n_data = 128
-
-    for i in range(n_data):
-        ra1,ra2 = np.random.uniform(0, 2*np.pi, size=2)
-        dec1,dec2 = np.pi/2. - np.arccos(2*np.random.uniform(size=2)-1.)
-        plx1,plx2 = np.exp(np.random.uniform(0,2,size=2))
-
-        true_v = np.random.normal(0, 25, size=3)
-        T1 = get_tangent_basis(ra1, dec1)
-        v1 = T1.T.dot(true_v)
-        pmra1,pmdec1 = v1[:2] * plx1 / 4.740470463496208
-        # TODO: ignoring RV for now
-
-        true_v = np.random.normal(0, 25, size=3)
-        T2 = get_tangent_basis(ra2, dec2)
-        v2 = T2.T.dot(true_v)
-        pmra2,pmdec2 = v2[:2] * plx2 / 4.740470463496208
-        # TODO: ignoring RV for now
-
-        Cov = np.diag(np.random.uniform(0.1, 0.3, size=6))**2
[…]

### 2016-09-30 gwb/tests/test_likelihood_u.py
add unit test for H1 H2

@@ -14,29 +14,38 @@ from ..data import TGASStar
 from ..likelihood import (get_y, get_M, get_Cinv, get_Ainv_nu_Delta,
                           ln_H1_marg_v_likelihood, ln_Q, ln_H2_marg_v_likelihood)
 
-def make_random_data():
-    # make fake data
-    n_data = 128
-    ra = np.random.uniform(0, 2*np.pi, size=n_data)
-    dec = np.pi/2. - np.arccos(2*np.random.uniform(size=n_data)-1.)
-    parallax = np.random.uniform(-0.1, 1000., size=n_data)
-    pmra,pmdec = np.random.normal(0, 100, size=(2,n_data))
+Vinv = np.diag([1/25.**2]*3)
 
-    _M = np.random.uniform(0.1, 0.3, size=(n_data,6,6))
-    Cov = 0.5 * np.array([M.dot(M.T) for M in _M])
-    _M[:,5] = 0.
-    _M[...,5] = 0.
-    assert Cov.shape == (n_data,6,6)
-
+def make_random_data(n_batch=1, size=128):
     all_data = []
-    for r,d,plx,pmr,pmd,C in zip(ra,dec,parallax,pmra,pmdec,Cov):
-        row = {
-            'ra': np.degrees(r), 'dec': np.degrees(d),
-            'parallax': plx, 'pmra': pmr, 'pmdec': pmd
-        }
-        data = TGASStar(row)
-        data._cov = C
-        all_data.append(data)
+    for i in range(size):
+        data = []
+        for j in range(n_batch):
+            # make fake data
+            ra = np.random.uniform(0, 2*np.pi)
+            dec = np.pi/2. - np.arccos(2*np.random.uniform()-1.)
+            parallax = np.random.uniform(-0.1, 1000.)
+            pmra,pmdec = np.random.normal(0, 100, size=2)
+
+            _C = np.random.uniform(0.1, 0.3, size=(6,6))
+            Cov = 0.5 * _C.dot(_C.T)
+            Cov[5] = 0. # TODO: ignoring RV
+            Cov[:,5] = 0. # TODO: ignoring RV
+            assert Cov.shape == (6,6)
+
+            row = {
+                'ra': np.degrees(ra), 'dec': np.degrees(dec),
+                'parallax': parallax, 'pmra': pmra, 'pmdec': pmdec
+            }
+            star = TGASStar(row)
+            star._cov = Cov
+
+            data.append(star)
+
[…]

### 2016-09-30 gwb/tests/test_likelihood_f.py
dumb ass functional tests

@@ -15,34 +15,9 @@ from ..coords import get_tangent_basis
 from ..likelihood import (get_y, get_M, get_Cinv, get_Ainv_nu_Delta,
                           ln_H1_marg_v_likelihood, ln_Q, ln_H2_marg_v_likelihood)
 
-def make_random_data():
-    # make fake data
-    n_data = 128
-    ra = np.random.uniform(0, 2*np.pi, size=n_data)
-    dec = np.pi/2. - np.arccos(2*np.random.uniform(size=n_data)-1.)
-    parallax = np.random.uniform(-0.1, 1000., size=n_data)
-    pmra,pmdec = np.random.normal(0, 100, size=(2,n_data))
+Vinv = np.diag([1/25.**2]*3)
 
-    _M = np.random.uniform(0.1, 0.3, size=(n_data,6,6))
-    Cov = 0.5 * np.array([M.dot(M.T) for M in _M])
-    _M[:,5] = 0.
-    _M[...,5] = 0.
-    assert Cov.shape == (n_data,6,6)
-
-    all_data = []
-    for r,d,plx,pmr,pmd,C in zip(ra,dec,parallax,pmra,pmdec,Cov):
-        row = {
-            'ra': np.degrees(r), 'dec': np.degrees(d),
-            'parallax': plx, 'pmra': pmr, 'pmdec': pmd
-        }
-        data = TGASStar(row)
-        data._cov = C
-        all_data.append(data)
-
-    return all_data
-
-def make_data_increase_uncertainties():
-    n_data = 8
+def make_data_increase_uncertainties(n_data=128):
     ra = np.full(n_data, np.random.uniform(0, 2*np.pi))
     dec = np.full(n_data, np.pi/2. - np.arccos(2*np.random.uniform()-1.))
     parallax = np.ones(n_data)*10. # 100 pc
@@ -143,28 +118,7 @@ def make_good_pairs():
 
 # ----------------------------------------------------------------------------
 
-def test_y():
-    all_data = make_random_data()
-    for data in all_data:
-        d = 1000./data._parallax
-        y = get_y(d, data)
-        assert y.shape == (4,)
-        assert np.allclose(y[0], 0)
-
-    ds = [1000./data._parallax for data in all_data[:2]]
-    y = get_y(ds, all_data[:2])
-    assert y.shape == (8,)
-
 def test_M():
[…]

### 2016-09-30 gwb/prob.py
whitespace - oops

@@ -2,8 +2,6 @@ import numpy as np
 from numpy import sqrt, exp, pi, inf
 from scipy import integrate
 
-
-
 def construct_cov_pm(gaia_data):
     """
     Construct covariance matrix from gaia table for proper motion
@@ -20,7 +18,7 @@ def construct_cov_pm(gaia_data):
 
     for i,name1 in enumerate(names):
         for j,name2 in enumerate(names):
-            if j <= i: continue    
+            if j <= i: continue
             full_name = "{}_{}_corr".format(name1, name2)
             C[...,i,j] = gaia_data[full_name]*np.sqrt(C[...,i,i]*C[...,j,j])
             C[...,j,i] = gaia_data[full_name]*np.sqrt(C[...,i,i]*C[...,j,j])

### 2016-09-30 gwb/likelihood.py
added dumb comments

@@ -14,6 +14,9 @@ __all__ = ['get_y', 'get_M', 'get_Cinv', 'get_Ainv_nu_Delta',
 pc_mas_yr_to_km_s = (1*u.pc * u.mas/u.yr).to(u.km/u.s,u.dimensionless_angles()).value
 
 def get_y(ds, stars):
+    """
+    Construct the vector y, which should have length 3*n_stars
+    """
     ds = np.atleast_1d(ds)
     stars = np.atleast_1d(stars)
 
@@ -23,6 +26,9 @@ def get_y(ds, stars):
     return y
 
 def get_M(stars):
+    """
+    Construct the matrix M, which should have shape (3*n_stars, 3*n_stars)
+    """
     stars = np.atleast_1d(stars)
 
     M = [get_tangent_basis(star._ra, star._dec)
@@ -31,6 +37,9 @@ def get_M(stars):
     return np.vstack(M)
 
 def get_Cinv(ds, stars):
+    """
+    Construct the matrix Cinv, which should have shape (3*n_stars, 3*n_stars)
+    """
     ds = np.atleast_1d(ds)
     stars = np.atleast_1d(stars)
 

### 2016-09-30 paper/ms.tex
tweak
fix nu minus signs
remove parallax from y

@@
 % Results
-We find pairs of comoving stars in abundance out to very wide
-separations, including separations of a few parsec! There does not
+We find pairs of comoving stars out to very wide
+separations, including separations of a few parsecs! There does not
 seem to be any tidal-disruption feature in the separation distribution
@@
 Before we write that down, we will make some definitions.
-We construct a vector $y$ as follows:
+We construct a velocity-space data vector $y$ as follows:
 \begin{eqnarray}
-  \transp{y} &\equiv& [(d_1\,\varpi_1 - 1), (d_1\,\mu_{\alpha1}), (d_1\,\mu_{\delta1}),
-                       (d_2\,\varpi_2 - 1), (d_2\,\mu_{\alpha2}), (d_2\,\mu_{\delta2})] \quad \mbox{or}
+  \transp{y} &\equiv& [(d_1\,\mu_{\alpha1}), (d_1\,\mu_{\delta1}),
+                       (d_2\,\mu_{\alpha2}), (d_2\,\mu_{\delta2})] \quad \mbox{or}
   \\
-  \transp{y} &\equiv& [(d_1\,\varpi_1 - 1), (d_1\,\mu_{\alpha1}), (d_1\,\mu_{\delta1}), (v_{r1}),
-                       (d_2\,\varpi_2 - 1), (d_2\,\mu_{\alpha2}), (d_2\,\mu_{\delta2}), (v_{r2})]
+  \transp{y} &\equiv& [(d_1\,\mu_{\alpha1}), (d_1\,\mu_{\delta1}), (v_{r1}),
+                       (d_2\,\mu_{\alpha2}), (d_2\,\mu_{\delta2}), (v_{r2})]
   \quad ,
 \end{eqnarray}
-where we have given two versions, a 6-component and an 8-component
+where we have given two versions, a 4-component and an 6-component
 version (depending on whether there are measured radial velocities),
@@
   \\
-  \transp{M} &=& [0, \hat{\alpha_1}, \hat{\delta_1},
-                  0, \hat{\alpha_2}, \hat{\delta_2}] \quad \mbox{or}
+  \transp{M} &=& [\hat{\alpha_1}, \hat{\delta_1},
+                  \hat{\alpha_2}, \hat{\delta_2}] \quad \mbox{or}
   \\
-  \transp{M} &=& [0, \hat{\alpha}_1, \hat{\delta}_2, \hat{r}_1,
-                  0, \hat{\alpha}_2, \hat{\delta}_2, \hat{r}_2]
+  \transp{M} &=& [\hat{\alpha}_1, \hat{\delta}_2, \hat{r}_1,
+                  \hat{\alpha}_2, \hat{\delta}_2, \hat{r}_2]
   \quad,
@@
 \begin{eqnarray}
-  \transp{y} &\equiv& [(d\,\varpi - 1), (d\,\mu_{\alpha}), (d\,\mu_{\delta})] \quad \mbox{or}
+  \transp{y} &\equiv& [(d\,\mu_{\alpha}), (d\,\mu_{\delta})] \quad \mbox{or}
   \\
-  \transp{y} &\equiv& [(d\,\varpi - 1), (d\,\mu_{\alpha}), (d\,\mu_{\delta}), (v_{r})]
+  \transp{y} &\equiv& [(d\,\mu_{\alpha}), (d\,\mu_{\delta}), (v_{r})]
   \\
-  \transp{M} &=& [0, \hat{\alpha}, \hat{\delta}] \quad \mbox{or}
+  \transp{M} &=& [\hat{\alpha}, \hat{\delta}] \quad \mbox{or}
   \\
-  \transp{M} &=& [0, \hat{\alpha}, \hat{\delta}_2, \hat{r}]
+  \transp{M} &=& [\hat{\alpha}, \hat{\delta}_2, \hat{r}]
   \\

### 2016-09-30 gwb/coords.py
u_hat is unit vector by construction

@@ -29,7 +29,7 @@ def get_u_vec(lon, lat):
     u_hat = np.array([np.cos(lon) * np.cos(lat),
                       np.sin(lon) * np.cos(lat),
                       np.sin(lat)])
-    return u_hat / np.sqrt(np.sum(u_hat**2, axis=0))[None]
+    return u_hat
 
 def get_tangent_basis(ra, dec, dra=0.5, ddec=0.5):
     """

### 2016-09-30 gwb/coords.py
vastly simplify get_tangent_basis

@@ -31,38 +31,12 @@ def get_u_vec(lon, lat):
                       np.sin(lat)])
     return u_hat
 
-def get_tangent_basis(ra, dec, dra=0.5, ddec=0.5):
+# TODO: stack properly for array ra, decs
+def get_tangent_basis(ra, dec):
     """
     row vectors are the tangent-space basis at (alpha, delta, r)
     """
-
-    ra = np.atleast_1d(ra)
-    dec = np.atleast_1d(dec)
-
-    # unit vector pointing at the sky position of the target
-    u_hat = get_u_vec(ra, dec)
-
-    # unit vector offset in declination
-    dec_hat_sign = np.ones_like(u_hat)
-    dec_hat_sign[:,dec > np.pi/4] = -1.
-
-    v_hat = np.zeros_like(u_hat)
-    v_hat[:,dec <= np.pi/4] = get_u_vec(ra[dec <= np.pi/4], dec[dec <= np.pi/4]+ddec)
-    v_hat[:,dec > np.pi/4] = get_u_vec(ra[dec > np.pi/4], dec[dec > np.pi/4]-ddec)
-
-    dec_hat = dec_hat_sign * (v_hat - u_hat)
-    ra_hat = get_u_vec(ra+dra, dec) - u_hat # always a positive offset in RA
-
-    # define the orthogonal basis using gram-schmidt orthonormalization
-    #  - u1 is the unit vector that points to (ra,dec)
-    u1 = u_hat
-
-    u2 = dec_hat - np.einsum('ij,ij->j', dec_hat, u1)*u1
-    u2 /= np.sqrt(np.sum(u2**2, axis=0))
-
-    u3 = (ra_hat - np.einsum('ij,ij->j', ra_hat, u1)*u1 -
-          np.einsum('ij,ij->j', ra_hat, u2)*u2)
-    u3 /= np.sqrt(np.sum(u3**2, axis=0))
-
-    T = np.stack((u3,u2,u1))
-    return np.squeeze(np.moveaxis(T,-1,0))
+    return np.array([
+            [-np.sin(ra),  np.cos(ra), 0.],
+            [-np.sin(dec)*np.cos(ra), -np.sin(dec)*np.sin(ra), np.cos(dec)],
+            [np.cos(dec)*np.cos(ra),  np.cos(dec)*np.sin(ra), np.sin(dec)]])

### 2016-10-01 gwb/coords.py
stack properly for arrays of ra, dec

@@ -31,12 +31,18 @@ def get_u_vec(lon, lat):
                       np.sin(lat)])
     return u_hat
 
-# TODO: stack properly for array ra, decs
 def get_tangent_basis(ra, dec):
     """
     row vectors are the tangent-space basis at (alpha, delta, r)
     """
-    return np.array([
-            [-np.sin(ra),  np.cos(ra), 0.],
-            [-np.sin(dec)*np.cos(ra), -np.sin(dec)*np.sin(ra), np.cos(dec)],
-            [np.cos(dec)*np.cos(ra),  np.cos(dec)*np.sin(ra), np.sin(dec)]])
+    ra = np.atleast_1d(ra)
+    dec = np.atleast_1d(dec)
+    s = []
+    for cra, cdec in zip(ra, dec):
+        s.append(
+            np.array([
+                [-np.sin(cra),  np.cos(cra), 0.],
+                [-np.sin(cdec)*np.cos(cra), -np.sin(cdec)*np.sin(cra), np.cos(cdec)],
+                [np.cos(cdec)*np.cos(cra),  np.cos(cdec)*np.sin(cra), np.sin(cdec)]
+                ]))
+    return np.array(s).squeeze()

### 2016-10-03 gwb/fml.py
fix a number of bugs with units and update tests

@@ -71,16 +71,16 @@ def get_posterior_distance_samples(star, size=1, min_parallax=0.1):
 
     return 1000/samples[:size]
 
-def ln_H1_FML(star1, star2, Vinv, n_dist_samples=128):
+def ln_H1_FML(star1, star2, Vinv, n_dist_samples=128, v_scatter=0.):
     dist1 = get_posterior_distance_samples(star1, size=n_dist_samples)
     dist2 = get_posterior_distance_samples(star2, size=n_dist_samples)
-    ll_H1_at_samples = np.array([ln_H1_marg_v_likelihood(d1, d2, star1, star2, Vinv)
+    ll_H1_at_samples = np.array([ln_H1_marg_v_likelihood(d1, d2, star1, star2, Vinv, v_scatter)
                                  for d1,d2 in zip(dist1, dist2)])
     return logsumexp(ll_H1_at_samples) - np.log(float(ll_H1_at_samples.size))
 
-def ln_H2_FML(star1, star2, Vinv, n_dist_samples=128):
+def ln_H2_FML(star1, star2, Vinv, n_dist_samples=128, v_scatter=0.):
     dist1 = get_posterior_distance_samples(star1, size=n_dist_samples)
     dist2 = get_posterior_distance_samples(star2, size=n_dist_samples)
-    ll_H2_at_samples = np.array([ln_H2_marg_v_likelihood(d1, d2, star1, star2, Vinv)
+    ll_H2_at_samples = np.array([ln_H2_marg_v_likelihood(d1, d2, star1, star2, Vinv, v_scatter)
                                  for d1,d2 in zip(dist1, dist2)])
     return logsumexp(ll_H2_at_samples) - np.log(float(ll_H2_at_samples.size))

### 2016-10-03 gwb/likelihood.py

@@ -8,22 +8,48 @@ from scipy.linalg import block_diag
 # Project
 from .coords import get_tangent_basis
 
-__all__ = ['get_y', 'get_M', 'get_Cinv', 'get_Ainv_nu_Delta',
+__all__ = ['get_y_Cinv', 'get_M', 'get_Ainv_nu_Delta',
            'ln_H1_marg_v_likelihood', 'ln_Q', 'ln_H2_marg_v_likelihood']
 
-pc_mas_yr_to_km_s = (1*u.pc * u.mas/u.yr).to(u.km/u.s,u.dimensionless_angles()).value
+pc_mas_yr_per_km_s = (1*u.km/u.s).to(u.pc*u.mas/u.yr, u.dimensionless_angles()).value
+km_s_per_pc_mas_yr = 1/pc_mas_yr_per_km_s
 
-def get_y(ds, stars):
+def get_y_Cinv(ds, stars, v_scatter=0.):
     """
     Construct the vector y, which should have length 3*n_stars
     """
     ds = np.atleast_1d(ds)
     stars = np.atleast_1d(stars)
 
-    y = np.hstack([[d * star._pmra * pc_mas_yr_to_km_s,
-                    d * star._pmdec * pc_mas_yr_to_km_s,
+    y = np.hstack([[d * star._pmra * km_s_per_pc_mas_yr,
+                    d * star._pmdec * km_s_per_pc_mas_yr,
                     star._rv] for d,star in zip(ds,stars)])
-    return y
+
+    # Construct the matrix Cinv, which should have shape (3*n_stars, 3*n_stars)
+    ds = np.atleast_1d(ds)
+    stars = np.atleast_1d(stars)
+    assert len(ds) == len(stars)
+
+    Cinvs = []
+    for d,star in zip(ds,stars):
+        Cov = star.get_cov().copy() # returns full 6x6 covariance matrix from Gaia data
+        Cov = Cov[3:,3:]
+
+        Cov[0,:] *= d * km_s_per_pc_mas_yr # pmra
+        Cov[1,:] *= d * km_s_per_pc_mas_yr # pmdec
+        Cov[:,0] *= d * km_s_per_pc_mas_yr # pmra
+        Cov[:,1] *= d * km_s_per_pc_mas_yr # pmdec
+        Cov += np.eye(Cov.shape[0]) * v_scatter**2
+
+        Cinv = np.zeros((3,3))
[…]

### 2016-10-03 gwb/tests/test_likelihood_f.py

@@ -37,85 +37,6 @@ def make_data_increase_uncertainties(n_data=128):
 
     return all_data
 
-def make_bad_pairs():
-    n_data = 128
-
-    for i in range(n_data):
-        ra1,ra2 = np.random.uniform(0, 2*np.pi, size=2)
-        dec1,dec2 = np.pi/2. - np.arccos(2*np.random.uniform(size=2)-1.)
-        plx1,plx2 = np.exp(np.random.uniform(0,2,size=2))
-
-        true_v = np.random.normal(0, 25, size=3)
-        T1 = get_tangent_basis(ra1, dec1)
-        v1 = T1.T.dot(true_v)
-        pmra1,pmdec1 = v1[:2] * plx1 / 4.740470463496208
-        # TODO: ignoring RV for now
-
-        true_v = np.random.normal(0, 25, size=3)
-        T2 = get_tangent_basis(ra2, dec2)
-        v2 = T2.T.dot(true_v)
-        pmra2,pmdec2 = v2[:2] * plx2 / 4.740470463496208
-        # TODO: ignoring RV for now
-
-        Cov = np.diag(np.random.uniform(0.1, 0.3, size=6))**2
-        Cov[5] = 0. # TODO: ignoring RV for now
-        Cov[:,5] = 0. # TODO: ignoring RV for now
-
-        # HACK: they have the same covariance matrix
-        star1 = TGASStar({'ra': np.degrees(ra1), 'dec': np.degrees(dec1),
-                          'parallax': plx1, 'pmra': pmra1, 'pmdec': pmdec1})
-        star1._cov = Cov
-
-        star2 = TGASStar({'ra': np.degrees(ra2), 'dec': np.degrees(dec2),
-                          'parallax': plx2, 'pmra': pmra2, 'pmdec': pmdec2})
-        star2._cov = Cov
-
-        yield [star1, star2]
-
-def make_good_pairs():
-    n_data = 128
-
-    for i in range(n_data):
-        ra1 = np.random.uniform(0, 2*np.pi)
-        dec1 = np.pi/2. - np.arccos(2*np.random.uniform()-1.)
-        plx1 = 10 ** np.random.uniform(0,2)
-
-        ra2 = ra1 + np.random.uniform(-0.02, 0.02)
-        if dec1 > 45:
-            dec2 = dec1 - np.random.uniform(0, 0.03)
-        else:
-            dec2 = dec1 + np.random.uniform(0, 0.03)
-
[…]

### 2016-10-03 scripts/run-likelihood-ratio.py
script to compute likelihood ratio for pairs

@@ -1,0 +1,259 @@
+# Standard library
+import os
+import sys
+
+# Third-party
+from astropy import log as logger
+import astropy.units as u
+import h5py
+import numpy as np
+import six
+from schwimmbad import choose_pool
+
+# Project
+
+tgas_stacked = ""
+pair_indices = ""
+
+def main(pool, seed=42, cache_filename=None, overwrite=False):
+    n_distance_samples = 128
+
+    full_path = os.path.abspath(data_file)
+    if cache_filename is None:
+        if hdf5_key is None:
+            basename = os.path.basename(full_path)
+            name = os.path.splitext(basename)[0]
+            output_filename = os.path.join(paths.root, "cache", "{}.h5".format(name))
+        else:
+            output_filename = os.path.join(paths.root, "cache", "{}.h5".format(hdf5_key))
+
+    else:
+        output_filename = "{}.h5".format(os.path.splitext(cache_filename)[0])
+        _path,_basename = os.path.split(output_filename)
+        if not _path:
+            output_filename = os.path.join(paths.root, "cache", output_filename)
+
+    # container for hyper-parameter values
+    hyperpars = dict(jitter=None, P_min=None, P_max=None)
+
+    # see if we already did this:
+    if os.path.exists(output_filename):
+        if not overwrite and not continue_sampling:
+            logger.info("Sampling already performed. Use --overwrite to redo or --continue "
+                        "to keep sampling.")
+            pool.close()
+            sys.exit(0)
+
+        elif continue_sampling: # we need to increment the random number seed appropriately
+            mode = 'a' # append to output file
+
+            with h5py.File(output_filename, 'r') as f:
+                if 'rerun' not in f.attrs:
+                    rerun = 0
+                else:
+                    rerun = f.attrs['rerun'] + 1
+
+                if rerun != 0:
+                    logger.debug("Reading hyperparameters from cache file.")
+                    for name in ['jitter_m/s', 'P_min_day', 'P_max_day']:
+                        # HACK
+                        pieces = name.split("_")
+                        unit = u.Unit(pieces[-1])
+                        short_name = "_".join(pieces[:-1])
+                        var = hyperpars_strs[short_name]
+
+                        if var is not None:
+                            logger.warning("'{}' specified at command line, but using value "
[…]

### 2016-10-03 gwb/data.py
improvements to data handling

@@ -1,16 +1,125 @@
 from __future__ import division, print_function
 
 # Third-party
+from astropy.io import fits
 import astropy.coordinates as coord
 import astropy.units as u
 import numpy as np
+import six
 
-__all__ = ['TGASStar']
+__all__ = ['TGASData', 'TGASStar']
 
-class TGASStar(object):
+class TGASData(object):
+    _unit_map = {
+        'ra': u.degree,
+        'dec': u.degree,
+        'parallax': u.milliarcsecond,
+        'pmra': u.milliarcsecond/u.year,
+        'pmdec': u.milliarcsecond/u.year,
+    }
+    for name,unit in _unit_map:
+        _unit_map['{}_error'.format(name)] = unit
 
-    def __init__(self, tgas_row, rv=None, rv_err=None, metadata=None):
-        self._row = tgas_row
+    def __init__(self, filename_or_data, rv=None, rv_err=None):
+
+        # radial velocities
+        if rv is not None:
+            if rv_err is None:
+                raise ValueError("If radial velocity is provided, you must also "
+                                 "provide an error.")
+
+            if not hasattr(rv, 'unit') or not hasattr(rv_err, 'unit'):
+                raise TypeError("Input radial velocity and error must be an Astropy "
+                                "Quantity object.")
+
+            elif not rv.unit.is_equivalent(u.km/u.s) or not rv_err.unit.is_equivalent(u.km/u.s):
+                raise u.UnitsError("Radial velocity unit is not convertible to km/s!")
+
+            self._rv = rv.to(u.km/u.s).value
+            self._rv_err = rv_err.to(u.km/u.s).value
+
+        else:
+            self._rv = None
+            self._rv_err = None
+
+        # TODO: maybe support memory-mapping here?
+        if isinstance(filename_or_data, six.string_types):
+            self._data = fits.getdata(filename_or_data, 1)
+
+        else:
+            self._data = filename_or_data
+
+    def __getattr__(self, name):
+        if name in self._unit_map:
+            return self._data[name] * self._unit_map[name]
+
+        elif name in self._data.dtype.names:
+            return self._data[name]
+
+        else:
+            raise AttributeError("Object {} has no attribute '{}' and source data "
+                                 "table has no column with that name.")
+
+    def __getitem__(self, slc):
+        sliced = self._data[slc]
+
+        if self._rv is not None:
[…]

### 2016-10-03 gwb/tests/test_coords.py
remove old transpose

@@ -61,5 +61,5 @@ def test_projections():
     ]
     for ra,dec,v in all_radial[-1:]:
         T = get_tangent_basis(np.radians(ra), np.radians(dec))
-        assert np.allclose(T.T.dot(v), [0,0,100])
+        assert np.allclose(T.dot(v), [0,0,100])
 

### 2016-10-03 gwb/data.py
parallax snr

@@ -115,6 +115,10 @@ class TGASData(object):
         return coord.SkyCoord(ra=self.ra, dec=self.dec,
                               distance=self.get_distance(lutz_kelker=lutz_kelker))
 
+    @property
+    def parallax_snr(self):
+        return self.parallax / self.parallax_error
+
 
 class TGASStar(TGASData):
 

### 2016-10-03 setup.py
add six

@@ -34,6 +34,7 @@ setup(
     description="Gaia wide binaries.",
     long_description=rd("README.md"),
     package_data={"": ["LICENSE"]},
+    install_requires=["six"],
     include_package_data=True,
     classifiers=[
         "Intended Audience :: Science/Research",

### 2016-10-03 gwb/data.py
add len and fix dict bug

@@ -16,9 +16,12 @@ class TGASData(object):
         'parallax': u.milliarcsecond,
         'pmra': u.milliarcsecond/u.year,
         'pmdec': u.milliarcsecond/u.year,
+        'ra_error': u.degree,
+        'dec_error': u.degree,
+        'parallax_error': u.milliarcsecond,
+        'pmra_error': u.milliarcsecond/u.year,
+        'pmdec_error': u.milliarcsecond/u.year,
     }
-    for name,unit in _unit_map:
-        _unit_map['{}_error'.format(name)] = unit
 
     def __init__(self, filename_or_data, rv=None, rv_err=None):
 
@@ -76,6 +79,9 @@ class TGASData(object):
         else: # many rows
             return TGASData(sliced, rv=rv, rv_err=rv_err)
 
+    def __len__(self):
+        return len(self._data)
+
     @property
     def rv(self):
         return self._rv*u.km/u.s
@@ -88,7 +94,7 @@ class TGASData(object):
 
         if lutz_kelker:
             snr = self._data['parallax'] / self._data['parallax_error']
-            if snr < 4:
+            if np.any(snr < 4):
                 raise ValueError("S/N is smaller than 4!")
             tmp = self._data['parallax'] * (0.5 + 0.5*np.sqrt(1 - 16/snr**2))
 
@@ -137,6 +143,9 @@ class TGASStar(TGASData):
         else:
             self._rv = 0.
             self._rv_err = None
+
+    def __len__(self):
+        return 1
 
     def get_cov(self):
         """

### 2016-10-03 setup.py
add some basic unit tests for new data class structure and add package data

@@ -33,7 +33,7 @@ setup(
     license="MIT",
     description="Gaia wide binaries.",
     long_description=rd("README.md"),
-    package_data={"": ["LICENSE"]},
+    package_data={"": ["LICENSE"], "gwb": "tests/data/tgas_subset.fits"},
     install_requires=["six"],
     include_package_data=True,
     classifiers=[

### 2016-10-03 gwb/data.py
making it explicit doesn't take much; removes RuntimeError

@@ -16,9 +16,12 @@ class TGASData(object):
         'parallax': u.milliarcsecond,
         'pmra': u.milliarcsecond/u.year,
         'pmdec': u.milliarcsecond/u.year,
+        'ra_error': u.degree,
+        'dec_error': u.degree,
+        'parallax_error': u.milliarcsecond,
+        'pmra_error': u.milliarcsecond/u.year,
+        'pmdec_error': u.milliarcsecond/u.year,
     }
-    for name,unit in _unit_map:
-        _unit_map['{}_error'.format(name)] = unit
 
     def __init__(self, filename_or_data, rv=None, rv_err=None):
 

### 2016-10-03 gwb/data.py
clean up inconsistencies

@@ -157,7 +157,7 @@ class TGASStar(TGASData):
         # pre-load the diagonal
         for i,name in enumerate(names):
             full_name = "{}_error".format(name)
-            C[i,i] = self._row[full_name]**2
+            C[i,i] = self._data[full_name]**2
 
         for i,name1 in enumerate(names):
             for j,name2 in enumerate(names):

### 2016-10-03 gwb/fml.py

@@ -57,7 +57,7 @@ def get_posterior_distance_samples(star, size=1, min_parallax=0.1):
     samples = np.array([])
     iter = 0
     while len(samples) < size or iter == maxiter:
-        batch = np.random.normal(star._parallax, np.sqrt(star.get_cov()[0,0]), size=batch_size)
+        batch = np.random.normal(star.parallax, np.sqrt(star.get_cov()[0,0]), size=batch_size)
         tmps = ln_parallax_prior(batch, min_parallax=min_parallax)
         tmps -= tmps.max()
 

### 2016-10-03 gwb/likelihood.py
convert to radians

@@ -57,7 +57,7 @@ def get_M(stars):
     """
     stars = np.atleast_1d(stars)
 
-    M = [get_tangent_basis(star._ra, star._dec)
+    M = [get_tangent_basis(np.radians(star._ra), np.radians(star._dec))
          for star in stars]
 
     return np.vstack(M)

### 2016-10-03 gwb/fml.py
fix _parallax references

@@ -57,7 +57,7 @@ def get_posterior_distance_samples(star, size=1, min_parallax=0.1):
     samples = np.array([])
     iter = 0
     while len(samples) < size or iter == maxiter:
-        batch = np.random.normal(star._parallax, np.sqrt(star.get_cov()[0,0]), size=batch_size)
+        batch = np.random.normal(star._data['parallax'], np.sqrt(star.get_cov()[0,0]), size=batch_size)
         tmps = ln_parallax_prior(batch, min_parallax=min_parallax)
         tmps -= tmps.max()
 

### 2016-10-03 gwb/likelihood.py

@@ -21,8 +21,8 @@ def get_y_Cinv(ds, stars, v_scatter=0.):
     ds = np.atleast_1d(ds)
     stars = np.atleast_1d(stars)
 
-    y = np.hstack([[d * star._pmra * km_s_per_pc_mas_yr,
-                    d * star._pmdec * km_s_per_pc_mas_yr,
+    y = np.hstack([[d * star._data['pmra'] * km_s_per_pc_mas_yr,
+                    d * star._data['pmdec'] * km_s_per_pc_mas_yr,
                     star._rv] for d,star in zip(ds,stars)])
 
     # Construct the matrix Cinv, which should have shape (3*n_stars, 3*n_stars)
@@ -57,7 +57,7 @@ def get_M(stars):
     """
     stars = np.atleast_1d(stars)
 
-    M = [get_tangent_basis(np.radians(star._ra), np.radians(star._dec))
+    M = [get_tangent_basis(np.radians(star._data['ra']), np.radians(star._data['dec']))
          for star in stars]
 
     return np.vstack(M)

### 2016-10-03 gwb/tests/test_likelihood_u.py

@@ -54,11 +54,11 @@ def make_random_data(n_batch=1, size=128):
 def test_y():
     all_data = make_random_data()
     for data in all_data:
-        d = 1000./data._parallax
+        d = 1000./data._data['parallax']
         y = get_y(d, data)
         assert y.shape == (3,)
 
-    ds = [1000./data._parallax for data in all_data[:2]]
+    ds = [1000./data._data['parallax'] for data in all_data[:2]]
     y = get_y(ds, all_data[:2])
     assert y.shape == (6,)
 
@@ -75,12 +75,12 @@ def test_M():
 def test_Cinv():
     all_data = make_random_data()
     for data in all_data:
-        d = 1000./data._parallax
+        d = 1000./data._data['parallax']
         Cinv = get_Cinv(d, data)
         assert Cinv.shape == (3,3)
         assert np.allclose(Cinv[0], Cinv[0].T)
 
-    ds = [1000./data._parallax for data in all_data[:2]]
+    ds = [1000./data._data['parallax'] for data in all_data[:2]]
     Cinv = get_Cinv(ds, all_data[:2])
     assert Cinv.shape == (6,6)
     assert np.allclose(Cinv[0], Cinv[0].T)
@@ -88,7 +88,7 @@ def test_Cinv():
 def test_Ainv_nu_Delta():
     all_data = make_random_data()
     for data in all_data:
-        d = 1000./data._parallax
+        d = 1000./data._data['parallax']
 
         M = get_M(data)
         Cinv = get_Cinv(d, data)
@@ -103,7 +103,7 @@ def test_Ainv_nu_Delta():
 
         assert np.isfinite(Delta)
 
-    ds = [1000./data._parallax for data in all_data[:2]]
+    ds = [1000./data._data['parallax'] for data in all_data[:2]]
     M = get_M(all_data[:2])
     Cinv = get_Cinv(ds, all_data[:2])
     y = get_y(ds, all_data[:2])
@@ -116,14 +116,14 @@ def test_Ainv_nu_Delta():
 
 def test_H1_marg():
     for star1,star2 in make_random_data(n_batch=2):
-        d1 = 1000/star1._parallax
-        d2 = 1000/star1._parallax
+        d1 = 1000/star1._data['parallax']
+        d2 = 1000/star1._data['parallax']
         ll = ln_H1_marg_v_likelihood(d1, d2, star1, star2, Vinv)
         assert np.isfinite(ll)
 
[…]

### 2016-10-03 gwb/data.py
fix _row in get_cov

@@ -47,10 +47,10 @@ class TGASData(object):
 
         # TODO: maybe support memory-mapping here?
         if isinstance(filename_or_data, six.string_types):
-            self._data = fits.getdata(filename_or_data, 1)
+            self._data = np.array(fits.getdata(filename_or_data, 1))
 
         else:
-            self._data = filename_or_data
+            self._data = np.array(filename_or_data)
 
     def __getattr__(self, name):
         if name in self._unit_map:
@@ -61,7 +61,7 @@ class TGASData(object):
 
         else:
             raise AttributeError("Object {} has no attribute '{}' and source data "
-                                 "table has no column with that name.")
+                                 "table has no column with that name.".format(self, name))
 
     def __getitem__(self, slc):
         sliced = self._data[slc]
@@ -73,7 +73,7 @@ class TGASData(object):
             rv = None
             rv_err = None
 
-        if hasattr(sliced, 'row'): # this is only one row
+        if sliced.ndim == 0: # this is only one row
             return TGASStar(row=sliced, rv=rv, rv_err=rv_err)
 
         else: # many rows
@@ -167,7 +167,7 @@ class TGASStar(TGASData):
         # pre-load the diagonal
         for i,name in enumerate(names):
             full_name = "{}_error".format(name)
-            C[i,i] = self._row[full_name]**2
+            C[i,i] = self._data[full_name]**2
 
         for i,name1 in enumerate(names):
             for j,name2 in enumerate(names):

### 2016-10-03 gwb/data.py
damn cov

@@ -174,8 +174,8 @@ class TGASStar(TGASData):
                 if j <= i:
                     continue
                 full_name = "{}_{}_corr".format(name1, name2)
-                C[i,j] = self[full_name] * np.sqrt(C[i,i]*C[j,j])
-                C[j,i] = self[full_name] * np.sqrt(C[i,i]*C[j,j])
+                C[i,j] = self._data[full_name] * np.sqrt(C[i,i]*C[j,j])
+                C[j,i] = self._data[full_name] * np.sqrt(C[i,i]*C[j,j])
 
         if self._rv_err is not None:
             C[5,5] = self._rv_err**2

### 2016-10-03 gwb/data.py
fix data tests

@@ -42,7 +42,7 @@ class TGASData(object):
             self._rv_err = rv_err.to(u.km/u.s).value
 
         else:
-            self._rv = None
+            self._rv = 0. # need to do this so y can have float type
             self._rv_err = None
 
         # TODO: maybe support memory-mapping here?
@@ -66,7 +66,7 @@ class TGASData(object):
     def __getitem__(self, slc):
         sliced = self._data[slc]
 
-        if self._rv is not None:
+        if self._rv_err is not None:
             rv = self._rv[slc]
             rv_err = self._rv_err[slc]
         else:
@@ -141,11 +141,14 @@ class TGASStar(TGASData):
             self._rv = rv.to(u.km/u.s).value
             self._rv_err = rv_err.to(u.km/u.s).value
         else:
-            self._rv = 0.
+            self._rv = 0. # need to do this so y can have float type
             self._rv_err = None
 
     def __len__(self):
         return 1
+
+    def __getitem__(self, slc):
+        object.__getitem__(self, slc)
 
     def get_cov(self):
         """

### 2016-10-03 gwb/tests/test_data.py

@@ -3,6 +3,7 @@ from __future__ import division, print_function
 # Third-party
 from astropy.io import fits
 from astropy.utils.data import get_pkg_data_filename
+import numpy as np
 
 # Project
 from ..data import TGASData, TGASStar

### 2016-10-03 gwb/tests/test_likelihood_u.py

@@ -11,7 +11,7 @@ import numpy as np
 
 # Project
 from ..data import TGASStar
-from ..likelihood import (get_y, get_M, get_Cinv, get_Ainv_nu_Delta,
+from ..likelihood import (get_y_Cinv, get_M, get_Ainv_nu_Delta,
                           ln_H1_marg_v_likelihood, ln_Q, ln_H2_marg_v_likelihood)
 
 Vinv = np.diag([1/25.**2]*3)
@@ -55,12 +55,17 @@ def test_y():
     all_data = make_random_data()
     for data in all_data:
         d = 1000./data._data['parallax']
-        y = get_y(d, data)
+        y,Cinv = get_y_Cinv(d, data)
         assert y.shape == (3,)
+        assert y.dtype == np.float64
+        assert Cinv.shape == (3,3)
+        assert np.allclose(Cinv[0], Cinv[0].T)
 
     ds = [1000./data._data['parallax'] for data in all_data[:2]]
-    y = get_y(ds, all_data[:2])
+    y,Cinv = get_y_Cinv(ds, all_data[:2])
     assert y.shape == (6,)
+    assert Cinv.shape == (6,6)
+    assert np.allclose(Cinv[0], Cinv[0].T)
 
 def test_M():
     all_data = make_random_data()
@@ -72,27 +77,13 @@ def test_M():
     M = get_M(all_data[:2])
     assert M.shape == (6,3)
 
-def test_Cinv():
-    all_data = make_random_data()
-    for data in all_data:
-        d = 1000./data._data['parallax']
-        Cinv = get_Cinv(d, data)
-        assert Cinv.shape == (3,3)
-        assert np.allclose(Cinv[0], Cinv[0].T)
-
-    ds = [1000./data._data['parallax'] for data in all_data[:2]]
-    Cinv = get_Cinv(ds, all_data[:2])
-    assert Cinv.shape == (6,6)
-    assert np.allclose(Cinv[0], Cinv[0].T)
-
 def test_Ainv_nu_Delta():
     all_data = make_random_data()
     for data in all_data:
         d = 1000./data._data['parallax']
 
         M = get_M(data)
-        Cinv = get_Cinv(d, data)
-        y = get_y(d, data)
+        y,Cinv = get_y_Cinv(d, data)
 
[…]

### 2016-10-03 gwb/likelihood.py
don't use object array

@@ -7,6 +7,7 @@ from scipy.linalg import block_diag
 
 # Project
 from .coords import get_tangent_basis
+from .data import TGASStar
 
 __all__ = ['get_y_Cinv', 'get_M', 'get_Ainv_nu_Delta',
            'ln_H1_marg_v_likelihood', 'ln_Q', 'ln_H2_marg_v_likelihood']
@@ -19,15 +20,14 @@ def get_y_Cinv(ds, stars, v_scatter=0.):
     Construct the vector y, which should have length 3*n_stars
     """
     ds = np.atleast_1d(ds)
-    stars = np.atleast_1d(stars)
+    if isinstance(stars, TGASStar):
+        stars = [stars]
 
     y = np.hstack([[d * star._data['pmra'] * km_s_per_pc_mas_yr,
                     d * star._data['pmdec'] * km_s_per_pc_mas_yr,
                     star._rv] for d,star in zip(ds,stars)])
 
     # Construct the matrix Cinv, which should have shape (3*n_stars, 3*n_stars)
-    ds = np.atleast_1d(ds)
-    stars = np.atleast_1d(stars)
     assert len(ds) == len(stars)
 
     Cinvs = []
@@ -55,7 +55,8 @@ def get_M(stars):
     """
     Construct the matrix M, which should have shape (3*n_stars, 3*n_stars)
     """
-    stars = np.atleast_1d(stars)
+    if isinstance(stars, TGASStar):
+        stars = [stars]
 
     M = [get_tangent_basis(np.radians(star._data['ra']), np.radians(star._data['dec']))
          for star in stars]

### 2016-10-03 gwb/data.py
whoa, avoid recursion shite

@@ -53,8 +53,13 @@ class TGASData(object):
             self._data = np.array(filename_or_data)
 
     def __getattr__(self, name):
-        if name in self._unit_map:
-            return self._data[name] * self._unit_map[name]
+        # to prevent recursion errors:
+        #   http://nedbatchelder.com/blog/201010/surprising_getattr_recursion.html
+        if name == '_data':
+            raise AttributeError()
+
+        if name in TGASData._unit_map:
+            return self._data[name] * TGASData._unit_map[name]
 
         elif name in self._data.dtype.names:
             return self._data[name]

### 2016-10-03 scripts/compute-likelihood-ratio.py
script to compute likelihood ratio and cache to a file

@@ -3,6 +3,7 @@ import os
 import sys
 
 # Third-party
+from astropy.io import fits
 from astropy import log as logger
 import astropy.units as u
 import h5py
@@ -11,173 +12,86 @@ import six
 from schwimmbad import choose_pool
 
 # Project
+from gwb.data import TGASData, TGASStar
+from gwb.fml import ln_H1_FML, ln_H2_FML
 
-tgas_stacked = ""
-pair_indices = ""
+class Worker(object):
 
-def main(pool, seed=42, cache_filename=None, overwrite=False):
+    def __init__(self, Vinv, v_scatter, n_distance_samples, output_filename):
+        self.Vinv = np.array(Vinv)
+        self.v_scatter = v_scatter
+        self.n_distance_samples = n_distance_samples
+        self.output_filename = output_filename
+
+    def work(self, i, star1, star2):
+        h1 = ln_H1_FML(star1, star2, Vinv=self.Vinv, v_scatter=self.v_scatter,
+                       n_dist_samples=self.n_distance_samples)
+        h2 = ln_H2_FML(star1, star2, Vinv=self.Vinv, v_scatter=self.v_scatter,
+                       n_dist_samples=self.n_distance_samples)
+        return i, h1, h2
+
+    def __call__(self, task):
+        i, star1, star2 = task
+        return self.work(i, star1, star2)
+
+    def callback(self, result):
+        if result is None:
+            pass
+
+        else:
+            i,h1,h2 = result
+            with h5py.File(self.output_filename, 'a') as f:
+                f['lnH1'][i] = h1
+                f['lnH2'][i] = h2
+
+def main(pool, stacked_tgas_path, pair_indices_path, signal_to_noise_cut,
+         output_path="../data", seed=42, overwrite=False):
+
+    # MAGIC NUMBERs
     n_distance_samples = 128
+    v_scatter = 1. # km/s
+    Vinv = np.diag(np.full(3, 1/25)**2) # 3x3 inverse variance matrix for disk stars
 
-    full_path = os.path.abspath(data_file)
-    if cache_filename is None:
-        if hdf5_key is None:
-            basename = os.path.basename(full_path)
-            name = os.path.splitext(basename)[0]
-            output_filename = os.path.join(paths.root, "cache", "{}.h5".format(name))
-        else:
-            output_filename = os.path.join(paths.root, "cache", "{}.h5".format(hdf5_key))
+    if not os.path.exists(pair_indices_path):
[…]

### 2016-10-03 gwb/tests/test_likelihood_f.py
update test to use y_Cinv

@@ -12,7 +12,7 @@ import numpy as np
 # Project
 from ..data import TGASStar
 from ..coords import get_tangent_basis
-from ..likelihood import (get_y, get_M, get_Cinv, get_Ainv_nu_Delta,
+from ..likelihood import (get_y_Cinv, get_M, get_Ainv_nu_Delta,
                           ln_H1_marg_v_likelihood, ln_Q, ln_H2_marg_v_likelihood)
 
 Vinv = np.diag([1/25.**2]*3)
@@ -55,7 +55,7 @@ def test_Cinv():
     all_det_Cinv = []
     for data in all_data:
         d = 1000./data._data['parallax']
-        Cinv = get_Cinv(d, data)
+        y,Cinv = get_y_Cinv(d, data)
         sgn,det = np.linalg.slogdet(Cinv[:2,:2])
         all_det_Cinv.append(det)
 
@@ -70,8 +70,7 @@ def test_Ainv_nu_Delta():
     for data in all_data:
         d = 1000./data._data['parallax']
         M = get_M(data)
-        Cinv = get_Cinv(d, data)
-        y = get_y(d, data)
+        y,Cinv = get_y_Cinv(d, data)
 
         Ainv, nu, Delta = get_Ainv_nu_Delta(d, M, Cinv, y, Vinv)
         all_log_detA.append(-np.linalg.slogdet(Ainv)[1])

### 2016-10-03 gwb/tests/test_fml.py
more test tweaks

@@ -15,7 +15,7 @@ from ..fml import ln_H1_FML, ln_H2_FML
 Vinv = np.diag([1/25.**2]*3)
 
 def make_bad_pairs():
-    n_data = 128
+    n_data = 16
 
     for i in range(n_data):
         ra1,ra2 = np.random.uniform(0, 2*np.pi, size=2)
@@ -56,7 +56,7 @@ def make_bad_pairs():
         yield [star1, star2]
 
 def make_good_pairs():
-    n_data = 128
+    n_data = 16
 
     for i in range(n_data):
         ra1 = np.random.uniform(0, 2*np.pi)
@@ -107,7 +107,6 @@ def _compute_ratio(pair):
 def test_ln_likelihood_ratio():
     np.random.seed(12345)
     for pair in make_bad_pairs():
-        assert _compute_ratio(pair) < 0
         print(_compute_ratio(pair))
 
     # for pair in make_good_pairs():

### 2016-10-03 scripts/compute-likelihood-ratio.py
remove test and add option for vscatter

@@ -9,7 +9,7 @@ import numpy as np
 from schwimmbad import choose_pool
 
 # Project
-from gwb.data import TGASData, TGASStar
+from gwb.data import TGASData
 from gwb.fml import ln_H1_FML, ln_H2_FML
 
 class Worker(object):
@@ -42,17 +42,18 @@ class Worker(object):
                 f['lnH2'][i] = h2
 
 def main(pool, stacked_tgas_path, pair_indices_path, signal_to_noise_cut,
-         output_path="../data", seed=42, overwrite=False):
+         v_scatter, output_path="../data", seed=42, overwrite=False):
 
     # MAGIC NUMBERs
     n_distance_samples = 128
-    v_scatter = 1. # km/s
+    # v_scatter = 1. # km/s
     Vinv = np.diag(np.full(3, 1/25)**2) # 3x3 inverse variance matrix for disk stars
 
     if not os.path.exists(pair_indices_path):
         raise IOError("Path to pair indices file '{}' does not exist!".format(pair_indices_path))
 
-    output_file = '{}-likelihood-ratio.h5'.format(os.path.splitext(pair_indices_path)[0])
+    output_file = '{}_vscatter{:.0f}-lratio.h5'.format(os.path.splitext(pair_indices_path,
+                                                                        v_scatter)[0])
 
     # load the pair indices
     pair_idx = fits.getdata(pair_indices_path, 0)
@@ -76,7 +77,6 @@ def main(pool, stacked_tgas_path, pair_indices_path, signal_to_noise_cut,
                          "are not consistent!")
 
     all_pairs = [[k,tgas[i],tgas[j]] for k,(i,j) in enumerate(pair_idx)]
-    all_pairs = all_pairs[:8] # HACK: testing
 
     worker = Worker(Vinv=Vinv, v_scatter=v_scatter, n_distance_samples=n_distance_samples,
                     output_filename=output_file)
@@ -115,6 +115,8 @@ if __name__ == "__main__":
                         type=str, help="Path to pair indices file.")
     parser.add_argument("--snr-cut", dest="signal_to_noise_cut", default=8,
                         type=float, help="Minimum signal-to-noise ratio in parallax.")
+    parser.add_argument("--vscatter", dest="v_scatter", default=1,
+                        type=float, help="TODO")
     parser.add_argument("--output-path", dest="output_path", default="../data/",
                         type=str, help="Path to write output.")
 
@@ -142,4 +144,5 @@ if __name__ == "__main__":
 
[…]

### 2016-10-03 scripts/compute-likelihood-ratio.py
typo [ci skip]

@@ -52,8 +52,8 @@ def main(pool, stacked_tgas_path, pair_indices_path, signal_to_noise_cut,
     if not os.path.exists(pair_indices_path):
         raise IOError("Path to pair indices file '{}' does not exist!".format(pair_indices_path))
 
-    output_file = '{}_vscatter{:.0f}-lratio.h5'.format(os.path.splitext(pair_indices_path,
-                                                                        v_scatter)[0])
+    output_file = '{}_vscatter{:.0f}-lratio.h5'.format(os.path.splitext(pair_indices_path)[0],
+                                                       v_scatter)
 
     # load the pair indices
     pair_idx = fits.getdata(pair_indices_path, 0)

### 2016-10-03 paper/ms.tex
no parallax
change title and add authors

@@
 
-\title{Widely separated, comoving stars in the \textsl{Gaia DR1 TGAS} sample}
-\author{APW, SO, DWH, DNS, DSTN, others}
+\title{Co-moving stars in \textsl{Gaia DR1}}
+\author{Adrian M. Price-Whelan\altaffilmark{\pu,\adrn},
+        Semyeong Oh\altaffilmark{\pu},
+        Timothy D. Morton\altaffilmark{\pu},
+        David W. Hogg\altaffilmark{\ccpp,\mpia},
+        David N. Spergel\altaffilmark{\pu,\cca}
+}
 
 % Affiliations
-% \newcommand{\pu}{1}
-% \newcommand{\adrn}{2}
-% \newcommand{\ccpp}{3}
-% \newcommand{\mpia}{4}
-% \newcommand{\uw}{5}
-% \newcommand{\sagan}{6}
+\newcommand{\pu}{1}
+\newcommand{\adrn}{2}
+\newcommand{\ccpp}{3}
+\newcommand{\mpia}{4}
+\newcommand{\cca}{5}
 
-% \altaffiltext{\pu}{Department of Astrophysical Sciences,
-%                    Princeton University, Princeton, NJ 08544, USA}
-% \altaffiltext{\adrn}{To whom correspondence should be addressed:
-%                      \texttt{adrn@princeton.edu}}
-% \altaffiltext{\ccpp}{Center for Cosmology and Particle Physics,
-%                      Department of Physics,
-%                      New York University, 4 Washington Place,
-%                      New York, NY 10003, USA}
-% \altaffiltext{\mpia}{Max-Planck-Institut f\"ur Astronomie,
-%                      K\"onigstuhl 17, D-69117 Heidelberg, Germany}
-% \altaffiltext{\uw}{Astronomy Department, University of Washington,
-%                    Seattle, WA 98195, USA}
-% \altaffiltext{\sagan}{Sagan Fellow}
+\altaffiltext{\pu}{Department of Astrophysical Sciences,
+                   Princeton University, Princeton, NJ 08544, USA}
+\altaffiltext{\adrn}{To whom correspondence should be addressed:
+                     \texttt{adrn@princeton.edu}}
+\altaffiltext{\ccpp}{Center for Cosmology and Particle Physics,
+                     Department of Physics,
+                     New York University, 4 Washington Place,
+                     New York, NY 10003, USA}
+\altaffiltext{\mpia}{Max-Planck-Institut f\"ur Astronomie,
+                     K\"onigstuhl 17, D-69117 Heidelberg, Germany}
+\altaffiltext{\cca}{Simons Center for Computational Astrophysics, ...,
+                    New York, NY XXXXX, USA}
 
@@
 \section{Introduction} \label{sec:intro}
+
+\section{Data} \label{sec:data}
 
@@
 \begin{eqnarray}
-  \nu &=& A\,B\,b
+  \nu &=& -A\,B\,b
   \\
@@
 \begin{eqnarray}
-  p(\data\given v,d_1,d_2) &=& d_1^3\,d_2^3\,\normal(y\given M\,v, \Sigma)
+  p(\data\given v,d_1,d_2) &=& d_1^2\,d_2^2\,\normal(y\given M\,v, \Sigma)
   \\
@@
 \end{eqnarray}
-where the factor of $d_1^3\,d_2^3$ converts to units of per \gaia\ data
+where the factor of $d_1^2\,d_2^2$ converts to units of per \gaia\ data
 from units of per $y$.
@@
   \\
-  \nu &=& A\,\transp{M}\,\inv{\Sigma}\,y
+  \nu &=& -A\,\transp{M}\,\inv{\Sigma}\,y
   \\
-  \Delta &=& -3\,\ln d_1 -3\,\ln d_2
+  \Delta &=& -2\,\ln d_1 -2\,\ln d_2
   +\frac{1}{2}\,\ln||2\pi\,\Sigma|| +\frac{1}{2}\,\ln||2\pi\,V|| \nonumber \\ && \quad
@@
   \\
-  \Delta &=& -3\,\ln d
+  \Delta &=& -2\,\ln d
   +\frac{1}{2}\,\ln||2\pi\,\Sigma|| +\frac{1}{2}\,\ln||2\pi\,V|| \nonumber \\ && \quad

### 2016-10-03 scripts/generate-pair-sample.py
print how many have more neighbors

@@ -58,6 +58,10 @@ def main(stacked_tgas_path, signal_to_noise_cut, n_neighbors, delta_v_cut,
     all_pair_idx = all_pair_idx[cut]
     logger.info("{} pairs before trimming duplicates".format(len(all_pair_idx)))
 
+    hitting_edge = np.bincount(all_pair_idx[:,0]) == n_neighbors
+    logger.info("{} stars likely have more than {} neighbors".format(hitting_edge.sum(),
+                                                                     n_neighbors))
+
     all_pair_idx = np.sort(all_pair_idx, axis=1)
     str_pairs = np.array(["{}{}".format(i,j) for i,j in all_pair_idx])
     _, unq_idx = np.unique(str_pairs, return_index=True)

### 2016-10-03 scripts/compute-likelihood-ratio.py
remove old test statement and reformat

@@ -78,11 +78,9 @@ def main(pool, stacked_tgas_path, pair_indices_path, signal_to_noise_cut,
 
     all_pairs = [[k,tgas[i],tgas[j]] for k,(i,j) in enumerate(pair_idx)]
 
-    worker = Worker(Vinv=Vinv, v_scatter=v_scatter, n_distance_samples=n_distance_samples,
+    worker = Worker(Vinv=Vinv, v_scatter=v_scatter,
+                    n_distance_samples=n_distance_samples,
                     output_filename=output_file)
-
-    worker.callback(worker(all_pairs[0]))
-
     pool.map(worker, all_pairs, callback=worker.callback)
     pool.close()
 

### 2016-10-03 scripts/compute-likelihood-ratio.py
save dv and sep for all pairs, compute v_scatter with orbital velocity

@@ -2,6 +2,8 @@
 import os
 
 # Third-party
+import astropy.units as u
+from astropy.constants import G, M_sun
 from astropy.io import fits
 from astropy import log as logger
 import h5py
@@ -14,22 +16,21 @@ from gwb.fml import ln_H1_FML, ln_H2_FML
 
 class Worker(object):
 
-    def __init__(self, Vinv, v_scatter, n_distance_samples, output_filename):
+    def __init__(self, Vinv, n_distance_samples, output_filename):
         self.Vinv = np.array(Vinv)
-        self.v_scatter = v_scatter
         self.n_distance_samples = n_distance_samples
         self.output_filename = output_filename
 
-    def work(self, i, star1, star2):
-        h1 = ln_H1_FML(star1, star2, Vinv=self.Vinv, v_scatter=self.v_scatter,
+    def work(self, i, star1, star2, v_scatter):
+        h1 = ln_H1_FML(star1, star2, Vinv=self.Vinv, v_scatter=v_scatter,
                        n_dist_samples=self.n_distance_samples)
-        h2 = ln_H2_FML(star1, star2, Vinv=self.Vinv, v_scatter=self.v_scatter,
+        h2 = ln_H2_FML(star1, star2, Vinv=self.Vinv, v_scatter=v_scatter,
                        n_dist_samples=self.n_distance_samples)
         return i, h1, h2
 
     def __call__(self, task):
-        i, star1, star2 = task
-        return self.work(i, star1, star2)
+        i, star1, star2, v_scatter = task
+        return self.work(i, star1, star2, v_scatter)
 
     def callback(self, result):
         if result is None:
@@ -46,7 +47,6 @@ def main(pool, stacked_tgas_path, pair_indices_path, signal_to_noise_cut,
 
     # MAGIC NUMBERs
     n_distance_samples = 128
-    # v_scatter = 1. # km/s
     Vinv = np.diag(np.full(3, 1/25)**2) # 3x3 inverse variance matrix for disk stars
 
     if not os.path.exists(pair_indices_path):
@@ -56,7 +56,10 @@ def main(pool, stacked_tgas_path, pair_indices_path, signal_to_noise_cut,
                                                        v_scatter)
 
     # load the pair indices
-    pair_idx = fits.getdata(pair_indices_path, 0)
+    tbl = fits.getdata(pair_indices_path, 1)
[…]

### 2016-10-03 scripts/generate-pair-sample.py

@@ -23,9 +23,9 @@ def main(stacked_tgas_path, signal_to_noise_cut, n_neighbors, delta_v_cut,
     if not os.path.exists(os.path.abspath(output_path)):
         raise IOError("Specified output path '{}' does not exist.".format(output_path))
 
-    output_file = os.path.join(output_path, "snr{}_n{}_dv{:.0f}.fits".format(signal_to_noise_cut,
-                                                                             n_neighbors,
-                                                                             delta_v_cut))
+    output_file = os.path.join(output_path, "snr{:.0f}_n{}_dv{:.0f}.fits".format(signal_to_noise_cut,
+                                                                                 n_neighbors,
+                                                                                 delta_v_cut))
     if os.path.exists(output_file) and not overwrite:
         raise IOError("Output file '{}' already exists. Use --overwrite to overwrite.")
 
@@ -56,6 +56,8 @@ def main(stacked_tgas_path, signal_to_noise_cut, n_neighbors, delta_v_cut,
     cut = dv < delta_v_cut
 
     all_pair_idx = all_pair_idx[cut]
+    dv = dv[cut]
+    sep = tree_d.ravel()[cut]
     logger.info("{} pairs before trimming duplicates".format(len(all_pair_idx)))
 
     hitting_edge = np.bincount(all_pair_idx[:,0]) == n_neighbors
@@ -66,9 +68,15 @@ def main(stacked_tgas_path, signal_to_noise_cut, n_neighbors, delta_v_cut,
     str_pairs = np.array(["{}{}".format(i,j) for i,j in all_pair_idx])
     _, unq_idx = np.unique(str_pairs, return_index=True)
     all_pair_idx = all_pair_idx[unq_idx]
+    dv = dv[unq_idx]
+    sep = sep[unq_idx]
     logger.info("{} pairs after trimming duplicates".format(len(all_pair_idx)))
 
-    hdu = fits.PrimaryHDU(all_pair_idx)
+    rows = [(i1,i2,x,y) for i1,i2,x,y in zip(all_pair_idx[:,0], all_pair_idx[:,1], dv, sep)]
+    tbl = np.array(rows, dtype=[('star1', 'i8'), ('star2', 'i8'),
+                                ('delta_v', 'f8'), ('sep', 'f8')])
+
+    hdu = fits.BinTableHDU(tbl)
     hdu.writeto(output_file, clobber=True)
 
 if __name__ == "__main__":

### 2016-10-03 scripts/compute-likelihood-ratio.py
indicies are all in the entire stacked tgas

@@ -42,7 +42,7 @@ class Worker(object):
                 f['lnH1'][i] = h1
                 f['lnH2'][i] = h2
 
-def main(pool, stacked_tgas_path, pair_indices_path, signal_to_noise_cut,
+def main(pool, stacked_tgas_path, pair_indices_path,
          v_scatter, output_path="../data", seed=42, overwrite=False):
 
     # MAGIC NUMBERs
@@ -73,11 +73,6 @@ def main(pool, stacked_tgas_path, pair_indices_path, signal_to_noise_cut,
 
     # read in TGAS data
     tgas = TGASData(os.path.abspath(stacked_tgas_path))
-    tgas = tgas[tgas.parallax_snr > signal_to_noise_cut]
-
-    if len(tgas) < pair_idx.max():
-        raise ValueError("Number of surviving stars after S/N cut and pair indices "
-                         "are not consistent!")
 
     assumed_mass = 2*M_sun # HACK, MAGIC NUMBER
     orb_v = np.sqrt(G*assumed_mass / sep).to(u.km/u.s).value
@@ -116,8 +111,6 @@ if __name__ == "__main__":
                         type=str, help="Path to stacked TGAS data file.")
     parser.add_argument("--pairs-file", dest="pair_indices_path", required=True,
                         type=str, help="Path to pair indices file.")
-    parser.add_argument("--snr-cut", dest="signal_to_noise_cut", default=8,
-                        type=float, help="Minimum signal-to-noise ratio in parallax.")
     parser.add_argument("--vscatter", dest="v_scatter", default=1,
                         type=float, help="TODO")
     parser.add_argument("--output-path", dest="output_path", default="../data/",
@@ -146,6 +139,6 @@ if __name__ == "__main__":
     logger.debug("Using pool: {}".format(pool.__class__))
 
     # use a context manager so the prior samples file always gets deleted
-    main(pool, args.stacked_tgas_path, args.pair_indices_path, args.signal_to_noise_cut,
+    main(pool, args.stacked_tgas_path, args.pair_indices_path,
          v_scatter=args.v_scatter, output_path=args.output_path,
          seed=args.seed, overwrite=args.overwrite)

### 2016-10-03 scripts/generate-pair-sample.py

@@ -32,14 +32,15 @@ def main(stacked_tgas_path, signal_to_noise_cut, n_neighbors, delta_v_cut,
     if os.path.exists(output_file) and not overwrite:
         raise IOError("Output file '{}' already exists. Use --overwrite to overwrite.")
 
-    tgas = TGASData(os.path.abspath(stacked_tgas_path))
-    n_full_tgas = len(tgas)
+    tgas0 = TGASData(os.path.abspath(stacked_tgas_path))
+    n_full_tgas = len(tgas0)
+    index0 = np.arange(n_full_tgas)
 
     # first, do a signal-to-noise cut
-    tgas = tgas[tgas.parallax_snr > signal_to_noise_cut]
+    tgas = tgas0[tgas0.parallax_snr > signal_to_noise_cut]  # this should return a copy of tgas0
+    index0_snr = index0[tgas0.parallax_snr > signal_to_noise_cut]
     logger.info("{}/{} targets left after S/N cut".format(len(tgas), n_full_tgas))
 
-    # next, built the KD Tree using the XYZ positions
     c = tgas.get_coord()
     X = c.cartesian.xyz.T
 
@@ -48,7 +49,9 @@ def main(stacked_tgas_path, signal_to_noise_cut, n_neighbors, delta_v_cut,
         sep = np.linalg.norm(X[all_pair_idx[:,0]]-X[all_pair_idx[:,1]], axis=1)
         vtan = tgas.get_vtan().value
         dv = np.sqrt(np.sum((vtan[all_pair_idx[:,0]] - vtan[all_pair_idx[:,1]])**2, axis=1))
+        index0_out = np.vstack([index0_snr[all_pair_idx[:,0]], index0_snr[all_pair_idx[:,1]]]).T
     else:
+        # next, built the KD Tree using the XYZ positions
         # separation in position
         tree = KDTree(X)
         tree_d,tree_i = tree.query(X, k=n_neighbors+1) # 0th match is always self
@@ -77,11 +80,12 @@ def main(stacked_tgas_path, signal_to_noise_cut, n_neighbors, delta_v_cut,
         str_pairs = np.array(["{}{}".format(i,j) for i,j in all_pair_idx])
         _, unq_idx = np.unique(str_pairs, return_index=True)
         all_pair_idx = all_pair_idx[unq_idx]
[…]

### 2016-10-03 paper/ms.tex
words in data section

@@
 \newcommand{\eqname}{Equation}
+\newcommand{\dr}{\acronym{DR1}}
 
@@
 \newcommand{\data}{\mathrm{data}}
+\newcommand{\snr}{[S/N]_\varpi}
 
@@
         Semyeong Oh\altaffilmark{\pu},
+        David W. Hogg\altaffilmark{\ccpp,\mpia},
         Timothy D. Morton\altaffilmark{\pu},
-        David W. Hogg\altaffilmark{\ccpp,\mpia},
         David N. Spergel\altaffilmark{\pu,\cca}
@@
 
+\todo{Some bullshit about binary stars...}
+
 \section{Data} \label{sec:data}
+
+The primary data set used in this \documentname\ is the Tycho-Gaia Astrometric
+Solution (TGAS), released as a part of Data Release 1 of the Gaia mission
+(\citealt{many}). TGAS contains astrometric measurements (sky position,
+parallax, and proper motions) and associated covariance matrices for a large
+fraction of the \project{Tycho-2} catalog (\citealt{tycho2}) with median
+astrometric precision equivalent to the \project{Hipparcos} catalog ($\approx
+0.3~{\rm mas}$; \citealt{hipparcos}). In terms of parallax signal-to-noise
+($\snr = \varpi/\sigma_\varpi$), the TGAS catalog contains 1240114 stars
+with $\snr > 4$ and 619618 stars with $\snr > 8$. The increase in precision over
+Tycho-2 and increase in volume over Hipparcos makes the TGAS catalog a powerful
+data set to search for widely-separated, co-moving stars.
+
+In the sections that follow, we construct a statistical model that use the TGAS
+data and associated covariances to handle the non-trivial uncertainties [...].
+Here, we instead visualize simple quantities constructed directly from the
+astrometric measurements to emphasize the quality of the data. For each star
+with $\snr > 8$, we use the reported parallax and parallax error to compute a
+distance corrected for the Lutz-Kelker bias (\citealt{lutzkelker}):
+\begin{equation}
+  d = 1000 \, \left[\frac{\varpi}{2} \, \left(1 + \sqrt{1 - \frac{16}{\snr^2}} \right) \right]^{-1} \, {\rm pc}
+\end{equation}
 
@@
 \end{equation}
-
-\appendix
-
-Here we derive analytic expressions for the marginal likelihood integrals above.
-
-\todo{Hypothesis 1}
-
-The integrals for Hypothesis 2 are somewhat easier because the true velocity
-vector can be considered to be already in the orthogonal basis defined by the
-tangent plane at the coordinates of the source (we are assuming an isotropic
-velocity prior). We will analytically integrate over the velocity components
-$(v_\alpha,v_\delta,v_r)$ and leave the distance integral to be done
-numerically:
-\begin{equation}
-  \int \, \dd^3\bs{v} \, L(\bs{v}, d) \, p(\bs{v})
-\end{equation}
-The covariance matrix over parallax, proper motion components, and
-radial velocity (if they exist) is $\matrx{C}$ and the inverse variance matrix
-is $\matrx{V}$. Defining the constant
-\begin{equation}
-  \mathcal{Z} = \left[ \det \left(\frac{\matrx{V}}{2\pi} \right) \right]^{1/2}
-\end{equation}
-the integrand can be written
-\begin{equation}
-  L(\bs{v}, d) \, p(\bs{v}) = \mathcal{Z} \,
-    \exp \left[ -\frac{1}{2} \transp{\left(\bs{x} - \bs{x}_\theta \right)} \,
-    \matrx{V} \,
-    \left(\bs{x} - \bs{x}_\theta \right)
-    - \frac{|\bs{v}|^2}{2\,\sigma^2} \right] \\
-\end{equation}
-The integral over radial velocity is simple and can be performed without
-considering any other terms because uncertainties in these measurements will
-not, in general, be covariant with the Gaia data. The integrand over tangential
-velocities can then be written
-\begin{equation}
-  L(v_\alpha, v_\delta, d) \, p(v_\alpha, v_\delta) = \tilde{\mathcal{Z}} \,
-    \exp \left[ -\frac{1}{2}\,\transp{\left(\bs{x}' - \bs{x}'_\theta \right)} \,
-    \matrx{V}' \,
-    \left(\bs{x}' - \bs{x}'_\theta \right)
-    - \frac{(v_\alpha^2 + v_\delta^2)}{2\,\sigma^2} \right]
-    \label{eq:vtan-integrand}
-\end{equation}
-where the primed variables indicate that they are in the sub-space of just the
-astrometric quantities, the constant pre-factor now includes the integral over
-radial velocity
-\begin{align}
-  \tilde{\mathcal{Z}} = \mathcal{Z} \,
-    \exp\left[-\frac{1}{2}\frac{v_r^2}{\sigma_{v_r}^2 + \sigma^2}\right] \,
-    \left[ 2\pi \, (\sigma_{v_r}^2 + \sigma^2) \right]^{1/2}
-\end{align}
-and $\sigma_{v_r}$ is the observational uncertainty of the radial velocity
-measurement. To perform the integral over tangential velocities, we partition
-the inverse variance matrix and vectors so that
-\begin{align}
-  \bs{x}_A &= (\varpi)\\
-  \bs{x}_B &= (\mu_\alpha, \mu_\delta)\\
-  \bs{x}_{\theta, A} &= (d^{-1}) \\
-  \bs{x}_{\theta, B} &= (d^{-1}\,v_\alpha, d^{-1}\,v_\delta) \\
-  \matrx{V} &=
-    \left(\begin{array}{c| c c}
-      V_{11} & V_{12} & V_{13}\\\hline
-      V_{12} & V_{22} & V_{23}\\
-      V_{13} & V_{23} & V_{33}
-    \end{array}\right)
-    = \left(\begin{array}{c c}
-      \matrx{V}_{AA} & \matrx{V}_{AB}\\
-      \matrx{V}_{AB} & \matrx{V}_{BB}
-    \end{array}\right)
-\end{align}
-\todo{Holy fuck that matrix is UGLY - fix with new version of AASTex 6}
-and the argument of the Gaussian in \eqname~\ref{eq:vtan-integrand} can be
-written
-\begin{multline}
-  -\frac{1}{2}\,[
-  \transp{(\bs{x}_A - \bs{x}_{\theta, A})}\,
-    \matrx{V}_{AA}\,(\bs{x}_A - \bs{x}_{\theta, A})\\
-  + \transp{(\bs{x}_B - \bs{x}_{\theta, B})}\,
-    \matrx{V}_{BB}\,(\bs{x}_B - \bs{x}_{\theta, B})\\
-  + 2\,\transp{(\bs{x}_A - \bs{x}_{\theta, A})}\,
-    \matrx{V}_{AB}\,(\bs{x}_B - \bs{x}_{\theta, B})]
-\end{multline}
-\todo{prior in above}
 

### 2016-10-03 scripts/compute-likelihood-ratio.py
float division

@@ -47,7 +47,7 @@ def main(pool, stacked_tgas_path, pair_indices_path,
 
     # MAGIC NUMBERs
     n_distance_samples = 128
-    Vinv = np.diag(np.full(3, 1/25)**2) # 3x3 inverse variance matrix for disk stars
+    Vinv = np.diag(np.full(3, 1./25.)**2) # 3x3 inverse variance matrix for disk stars
 
     if not os.path.exists(pair_indices_path):
         raise IOError("Path to pair indices file '{}' does not exist!".format(pair_indices_path))

### 2016-10-04 scripts/compute-likelihood-ratio.py
fix pool.map for serialpool

@@ -47,6 +47,7 @@ def main(pool, stacked_tgas_path, pair_indices_path,
 
     # MAGIC NUMBERs
     n_distance_samples = 128
+    assumed_mass = 2*M_sun
     Vinv = np.diag(np.full(3, 1./25.)**2) # 3x3 inverse variance matrix for disk stars
 
     if not os.path.exists(pair_indices_path):
@@ -74,14 +75,18 @@ def main(pool, stacked_tgas_path, pair_indices_path,
     # read in TGAS data
     tgas = TGASData(os.path.abspath(stacked_tgas_path))
 
-    assumed_mass = 2*M_sun # HACK, MAGIC NUMBER
     orb_v = np.sqrt(G*assumed_mass / sep).to(u.km/u.s).value
     v_scatter = np.sqrt(v_scatter**2 + orb_v**2)
     all_pairs = [[k,tgas[i],tgas[j],v_scatter[k]] for k,(i,j) in enumerate(pair_idx)]
 
     worker = Worker(Vinv=Vinv, n_distance_samples=n_distance_samples,
                     output_filename=output_file)
-    pool.map(worker, all_pairs, callback=worker.callback)
+
+    for result in pool.map(worker, all_pairs, callback=worker.callback):
+        # returns a generator, so need to explicitly loop to do the processing, but
+        #   we ignore the results because the callback function caches them.
+        pass
+
     pool.close()
 
 if __name__ == "__main__":

### 2016-10-04 scripts/generate-pair-sample.py
comments

@@ -32,15 +32,18 @@ def main(stacked_tgas_path, signal_to_noise_cut, n_neighbors, delta_v_cut,
     if os.path.exists(output_file) and not overwrite:
         raise IOError("Output file '{}' already exists. Use --overwrite to overwrite.")
 
+    # the full TGAS data table
     tgas0 = TGASData(os.path.abspath(stacked_tgas_path))
     n_full_tgas = len(tgas0)
     index0 = np.arange(n_full_tgas)
 
-    # first, do a signal-to-noise cut
+    # do a signal-to-noise cut and preserve the indices of the surviving targets
+    #   from the original stacked_tgas file
     tgas = tgas0[tgas0.parallax_snr > signal_to_noise_cut]  # this should return a copy of tgas0
     index0_snr = index0[tgas0.parallax_snr > signal_to_noise_cut]
     logger.info("{}/{} targets left after S/N cut".format(len(tgas), n_full_tgas))
 
+    # convert the sky position and Lutz-Kelker corrected distance into a 3D cartesian position
     c = tgas.get_coord()
     X = c.cartesian.xyz.T
 

### 2016-10-05 paper/ms.tex
add a bunch of words [ci skip]

@@
 
-\todo{Some bullshit about binary stars...}
+Two stars moving with similar full-space velocity vectors (``co-moving stars'')
+are either remnants of a dispersing stellar system or are a widely-separated
+binary star system. In either case, co-moving stars are of great interest for
+studying chemical uniformity and star formation processes, the destruction of
+star clusters and dwarf galaxies, and the tidal field and small-scale properties
+of the mass distribution in the Milky Way. To date, thousands of candidate
+co-moving star pairs have been identified by searching for stars with common
+proper motions (\citealt{Luyten:1979,Poveda:1994,Allen:2000,Gould:2003,
+Chaname:2004,Lepine:2007,Alonso-Floriano:2015}).
+% Cite:
+% https://ui.adsabs.harvard.edu/#abs/2012ApJ...757..170A/abstract
+% https://ui.adsabs.harvard.edu/#abs/2012AJ....144..102T/abstract
+% https://ui.adsabs.harvard.edu/#abs/2014ApJ...790..158A/abstract
+
+% Two stars moving with full-space velocity vector differences $\lesssim 1~\kms$
+% with orbital semi-major axes $a \gtrsim 10^-3~{\rm pc}$.
+
+Useful for:
+- star cluster disruption and dispersal
+- chemical abundance comparison (no interaction)
+- dynamics
 
@@
 0.3~{\rm mas}$; \citealt{hipparcos}). In terms of parallax signal-to-noise
-($\snr = \varpi/\sigma_\varpi$), the TGAS catalog contains 1240114 stars
-with $\snr > 4$ and 619618 stars with $\snr > 8$. The increase in precision over
-Tycho-2 and increase in volume over Hipparcos makes the TGAS catalog a powerful
-data set to search for widely-separated, co-moving stars.
+($\snr = \varpi/\sigma_\varpi$), the TGAS catalog contains 1240114 stars with
+$\snr > 4$, 619618 stars with $\snr > 8$, and 197105 stars with $\snr > 16$. The
+increase in precision over Tycho-2 and increase in volume over Hipparcos makes
+the TGAS catalog a powerful data set to search for widely-separated, co-moving
+stars.
 
-In the sections that follow, we construct a statistical model that use the TGAS
-data and associated covariances to handle the non-trivial uncertainties [...].
-Here, we instead visualize simple quantities constructed directly from the
-astrometric measurements to emphasize the quality of the data. For each star
-with $\snr > 8$, we use the reported parallax and parallax error to compute a
-distance corrected for the Lutz-Kelker bias (\citealt{lutzkelker}):
+In the sections that follow, we construct a statistical model that propagates
+the non-trivial uncertainties in the data to our beliefs about the likelihood
+that a given pair are co-moving. Here, we instead compute quantities directly
+from the astrometric measurements to visualize and emphasize the quality of the
+data. For each star with $\snr > 16$, we use the reported parallax and parallax
+error to compute a distance corrected for the Lutz-Kelker bias
+(\citealt{lutzkelker}):
 \begin{equation}
-  d = 1000 \, \left[\frac{\varpi}{2} \, \left(1 + \sqrt{1 - \frac{16}{\snr^2}} \right) \right]^{-1} \, {\rm pc}
+  d_{\rm LK} = 1000 \, \left[\frac{\varpi}{2} \,
+    \left(1 + \sqrt{1 - \frac{16}{\snr^2}} \right) \right]^{-1} \, {\rm pc}
 \end{equation}
+We use this distance and the sky positions of each source to compute 3D
+Cartesian positions, $\boldsymbol{x}$, for all stars and find the nearest
+neighbor to each source in 3-space using the KDTree implementation in
+\project{scikit-learn} (\citealt{scikit-learn}). We remove duplicates from
+this sample and are left with 138962 pairs of stars. For all pairs, we then
+compute the tangential velocity difference
+\begin{equation}
+  |\Delta \bs{v}_{\rm t}| = d_{\rm LK} \,
+                            \sqrt{(\mu_{\alpha,1}-\mu_{\alpha,2})^2 +
+                                  (\mu_{\delta,1}-\mu_{\delta,2})^2}
+\end{equation}
+\figname~\ref{fig:dv-sep} shows these quantities plotted for all pairs that pass
+the above signal-to-noise cut. Pairs with small physical separation also tend to
+have small velocity differences. Intriguingly, there appear to be an abundance
+of pairs with small velocity separation at separations $|\delta\bs{x}| \gtrsim
+1~{\rm pc}$.
+
+\begin{figure*}[p]
+\begin{center}
+\includegraphics[width=\textwidth]{figures/dv-sep.pdf}
+\end{center}
+\caption{%
+Difference in tangential velocity vs. physical separation for pairs of stars
+selected to be nearest neighbors in separation. Dashed line shows zero velocity
+difference. Note the points with small velocity difference: these are likely
+widely-separated binary stars.
+\label{fig:dv-sep}}
+\end{figure*}
+
+\todo{More naive visualizations, or save for the full sample?}
 
@@
 
-For a pair of stars $(i,j)$, we want to ratio the fully marginalized likelihood
-(FML) of the stars having the same 3-space velocity (hypothesis 1,
-$\mathcal{L}_1$) over the FML of the stars having different 3-space velocities
-(hypothesis 2, $\mathcal{L}_2$). When this ratio is large, the pair is a
-candidate binary star. For \gaia\ DR1, we only observe parallax, $\varpi$, and
-two proper motion components, $\propm = (\mu_\alpha, \mu_\delta)$ for each star.
-Where there is overlap with \rave\ or \apogee, some number of observed radial
-velocities, $v_r$, with observed variances $\sigma^2_{v_r}$ (independent of the
-\gaia\ data) may also be available. We assume that the uncertainties in these
-observed quantites are known with covariance matrix $\matrx{C}$. The likelihoods
-$\mathcal{L}_1, \mathcal{L}_2$ are marginalized over true 3-space velocity and
-distance for each star in the pair.
+For a given pair of stars $(i,j)$, we compute the fully marginalized likelihood
+(FML) that the pair has the same 3-space velocity (hypothesis 1,
+$\mathcal{L}_1$) and the FML of the stars having different 3-space velocities
+(hypothesis 2, $\mathcal{L}_2$). We use the ratio of these FMLs as a scalar for
+selecting candidate widely-separated binary stars. \gaia\ \dr\ only contains
+parallax, $\varpi$, and two proper motion components, $\propm = (\mu_\alpha,
+\mu_\delta)$ for each star, so for the majority of sources we only have
+information about the tangential velocities of the stars. Where there is overlap
+with \rave\ or \apogee, some number of observed radial velocities, $v_r$, with
+observed variances $\sigma^2_{v_r}$ (independent of the \gaia\ data) may also be
+available. We assume that the uncertainties in these observed quantites are
+known with covariance matrix $\matrx{C}$. The likelihood functions $L_1, L_2$
+are marginalized over true 3-space velocity and distance for each star in the
+pair.
 \begin{align}
@@
     L_j(\bs{v}, d_j) \,
-    p(\bs{v}) \, p(d_i) \, p(d_j) \\
+    p(\bs{v}) \, p(d_i \given \varpi_i) \, p(d_j \given \varpi_j) \\
   \mathcal{L}_2 &=
@@
     L_j(\bs{v}_j, d_j) \,
-    p(\bs{v}_i) \, p(\bs{v}_j) \, p(d_i) \, p(d_j) \\
+    p(\bs{v}_i) \, p(\bs{v}_j) \, p(d_i \given \varpi_i) \, p(d_j \given \varpi_j) \label{eq:hyp2}
 \end{align}
@@
     \left(\bs{x} - \bs{x}_\theta \right) \right] \\
-  \bs{x} &= \transp{\left(\begin{array}{c c c c} \varpi & \mu_\alpha &
-    \mu_\delta & v_r \end{array}\right)} \\
-  \bs{x}_\theta &= \transp{\left(\begin{array}{c c c c} \frac{1}{d} & \frac{v_\alpha}{d} &
-    \frac{v_\delta}{d} & \tilde{v}_r \end{array}\right)}
+  \bs{x} &= \transp{(\begin{array}{c c c} \mu_\alpha &
+    \mu_\delta & v_r \end{array})} \\
+  \bs{x}_\theta &= \transp{\left(\begin{array}{c c c} d^{-1}\,v_\alpha &
+    d^{-1}\,v_\delta & \tilde{v}_r \end{array}\right)}
 \end{align}
-\todo{Holy fuck that vector is UGLY}
-We assume Gaussian priors on the true velocity components and use a \todo{which
-one will we use?} distance prior \citep{Astraatmadja:2016} on heliocentric
+and we compute the posterior pdfs for the true distances, $p(d\given\varpi)$,
+using a uniform space density prior \citep{Astraatmadja:2016} on heliocentric
 distance
-\begin{align}
-  p(\bs{v}) &= \normal(0,25)~\kms \\
-  p(d) &= \frac{r^2}{2\,L^3} \, \exp\left(-r/L\right)
-\end{align}
+\begin{equation}
+  p(d) \propto d^2
+\end{equation}
+We assume Gaussian priors on the true velocity components
+\begin{equation}
+  p(\bs{v}) = \normal(0,25)~\kms
+\end{equation}
 
@@
 ``true'' 3-space velocity vector projects into tangential velocity components
-differently depending on sky position when the sky separation is appreciable
-($v_\alpha = v_\alpha(\bs{v}, \alpha, \delta)$). For hypothesis 2, the integral
-can be split into two the product of two simpler integrals $\mathcal{L}_2 = Q_i
-\, Q_j$ where
+differently depending on sky position ($v_\alpha = v_\alpha(\bs{v}, \alpha,
+\delta)$). The marginalization integral for hypothesis 2 is simpler because the
+integral in \eqname~\ref{eq:hyp2} can be split into the product of two simpler
+integrals $\mathcal{L}_2 = Q_i \, Q_j$ where
 \begin{equation}
-  Q = \int \, \dd d \, \dd^3 \bs{v} \, L(\bs{v}, d) \, p(\bs{v}) \, p(d)
+  Q = \int \, \dd d \, \dd^3 \bs{v} \, L(\bs{v}, d) \, p(\bs{v}) \, p(d\given\varpi)
 \end{equation}
+In either case, the marginalization over true 3-space velocity can be performed
+analytically; we derive the relevant expressions in
+\sectionname~\ref{sec:appendix}. After marginalizing over velocity, the
+likelihood integrands only depend on distance; we numerically compute the
+integrals over the true distances to each star in a pair, $d_1,d_2$, using Monte
+Carlo integration with $K$ samples from the distance posterior pdfs:
+\begin{equation}
+  \int \, \dd d \, \tilde{L}(d) \, p(d\given\varpi) \approx
+    \frac{1}{K} \, \sum_k^K \, \tilde{L}(d_k)
+\end{equation}
+Through experimentation, we have found that $K=128$ samples are sufficient for
+estimating the above integrals for stars with a wide range in parallax
+signal-to-noise.
 
@@
 \appendix
-\section{Some mathematics relevant to the above}
+
+\section{Some mathematics relevant to the above} \label{sec:appendix}
+
+\todo{clean this up}
 

### 2016-10-05 paper/ms.tex
more words in the intro

@@
 Two stars moving with similar full-space velocity vectors (``co-moving stars'')
-are either remnants of a dispersing stellar system or are a widely-separated
-binary star system. In either case, co-moving stars are of great interest for
-studying chemical uniformity and star formation processes, the destruction of
-star clusters and dwarf galaxies, and the tidal field and small-scale properties
-of the mass distribution in the Milky Way. To date, thousands of candidate
-co-moving star pairs have been identified by searching for stars with common
-proper motions (\citealt{Luyten:1979,Poveda:1994,Allen:2000,Gould:2003,
-Chaname:2004,Lepine:2007,Alonso-Floriano:2015}).
+are either members of a widely-separated stellar multiplet, or remnants of a
+dispersing stellar system.
+
+Widely-separated binary stars are weakly bound binary star systems with
+semi-major axes larger than $a \gtrsim 10^{-3}~{\rm pc}$. It is generally
+assumed that the stars in a wide binary form from gas with the same chemical
+composition and are separated enough that there is no interaction or minimal
+cross-pollution as the more massive star evolves (\citealt{todo}). Because the
+binding energy of these systems is so low, if wide binary stars survive for many
+Gigayears their existence alone places limits on the tidal field of the Galaxy
+(\citealt{todo}) and the graininess of the gravitational field (\citealt{todo}).
+In this picture, a remaining puzzle is how these systems survive the dense
+stellar environments of the birth clusters in which they form. Wide binaries are
+therefore important tools for studying small-scale star formation and the
+large-scale Galactic mass distribution. These stars will appear to be co-moving
+with velocity differences comparable to the binary orbital velocity at the
+given separation.
+
+At the largest separations, $a \gtrsim 1~{\rm pc}$, co-moving stars are likely
+either remnants of disrupting stellar systems or young wide binary stars: at
+these extreme separations, the Galactic tidal field would disrupt binary stars
+within $t_{\rm disrupt} \lesssim XX~{\rm Gyr}$. Unbound open clusters with
+velocity dispersions $\sigma_v \lesssim 1~\kms$ in the Galactic disk disrupt
+quickly leaving small streams of stars with the same chemical abundances
+(\citealt{chemicaltaggingstuff}). Disrupting or disrupted globular clusters and
+dwarf galaxies also produce streams of stars with similar chemical properties
+but with larger velocity differences (\citealt{streampeeps}). Stars from such
+systems at similar orbital phases will appear to be co-moving with velocity
+differences comparable to the internal velocity dispersions, $\sigma_v \approx
+1$--$10~\kms$.
+
+To date, thousands of candidate co-moving star pairs have been identified by
+searching for stars with common proper motions (\citealt{Luyten:1979,
+Poveda:1994, Allen:2000, Gould:2003, Chaname:2004, Lepine:2007,
+Alonso-Floriano:2015}).
+\todo{how is our search different?
+(1) we have distance, (2) bigger volume than Hipparcos, (3) we use the data properly, (4) we don't cut on distance or on magnitude of proper motion -- we cut on parallax S/N}
 % Cite:
+% Tokovinin & Lepine 2012
 % https://ui.adsabs.harvard.edu/#abs/2012ApJ...757..170A/abstract
@@
 % https://ui.adsabs.harvard.edu/#abs/2014ApJ...790..158A/abstract
+
+% Many searches find things close in angular separation, but need pms / distance
 
@@
 - dynamics
+
+In either case, co-moving stars are of great interest
+for studying the destruction of star clusters and dwarf galaxies, the tidal
+field and small-scale properties of the mass distribution in the Milky Way, and
+chemical uniformity and star formation processes.
 
@@
 signal-to-noise.
+
+\section{Results 1: A catalog of candidate wide binaries}
+
+\todo{Galactic context / orbits}
+
+\section{Results 2: Star clusters and cluster remnants}
+
+\section{Conclusions}
 

### 2016-10-05 gwb/likelihood.py
add support for mixture of gaussians velocity prior

@@ -4,6 +4,7 @@ from __future__ import division, print_function
 import astropy.units as u
 import numpy as np
 from scipy.linalg import block_diag
+from scipy.misc import logsumexp
 
 # Project
 from .coords import get_tangent_basis
@@ -130,13 +131,20 @@ def _marg_likelihood_helper(ds, data, Vinv, v_scatter):
     assert sgn > 0
     return 0.5*log_detA - Delta
 
-def ln_H1_marg_v_likelihood(d1, d2, data1, data2, Vinv, v_scatter=0.):
+def ln_H1_marg_v_likelihood(d1, d2, data1, data2, Vinvs, v_scatter=0.):
+    Vinvs = np.array(Vinvs)
+    if Vinvs.ndim < 3:
+        Vinvs = Vinvs[None]
+
     ds = np.array([d1, d2])
     data = [data1, data2]
-    return _marg_likelihood_helper(ds, data, Vinv, v_scatter)
+    return logsumexp([_marg_likelihood_helper(ds, data, Vinv, v_scatter) for Vinv in Vinvs])
 
-def ln_Q(d, data, Vinv, v_scatter=0.):
-    return _marg_likelihood_helper(d, data, Vinv, v_scatter)
+def ln_Q(d, data, Vinvs, v_scatter=0.):
+    Vinvs = np.array(Vinvs)
+    if Vinvs.ndim < 3:
+        Vinvs = Vinvs[None]
+    return logsumexp([_marg_likelihood_helper(d, data, Vinv, v_scatter) for Vinv in Vinvs])
 
 def ln_H2_marg_v_likelihood(d1, d2, data1, data2, Vinv, v_scatter=0.):
     return (ln_Q(d1, data1, Vinv, v_scatter) + ln_Q(d2, data2, Vinv, v_scatter))

### 2016-10-05 mpi/hecate/submit.sh
new script

@@ -1,0 +1,29 @@
+#!/bin/bash
+#SBATCH -J likelihood-ratio           # job name
+#SBATCH -o h1h2.o%j             # output file name (%j expands to jobID)
+#SBATCH -e h1h2.e%j             # error file name (%j expands to jobID)
+#SBATCH -n 256                  # total number of mpi tasks requested
+#SBATCH -t 01:00:00             # run time (hh:mm:ss) - 1.5 hours
+#SBATCH --mail-user=adrn@princeton.edu
+#SBATCH --mail-type=begin       # email me when the job starts
+#SBATCH --mail-type=end         # email me when the job finishes
+
+cd /tigress/adrianp/projects/gaia-wide-binaries/scripts
+
+module load openmpi/gcc/1.10.2/64
+
+source activate gwb
+
+export SNR=8
+export NEIGH=64
+export DV=8
+
+python generate-pair-sample.py --tgas-file=../data/stacked_tgas.fits \
+--snr-cut=$SNR \
+--neighbors=$NEIGH \
+--deltav-cut=$DV
+
+srun python compute-likelihood-ratio.py --tgas-file=../data/stacked_tgas.fits -v --mpi \
+--snr-cut=$SNR \
+--pairs-file=../data/snr"$SNR"_n"$NEIGH"_dv"$DV".fits \
+--vscatter=0

### 2016-10-06 paper/ms.tex
author stuff
reworking some of the methods

@@
 \newcommand{\dr}{\acronym{DR1}}
+\newcommand{\tgas}{\acronym{TGAS}}
 
@@
 \newcommand{\kms}{\rm km~s^{-1}}
+\newcommand{\msun}{{\rm M}_\odot}
 \newcommand{\data}{\mathrm{data}}
@@
 \title{Co-moving stars in \textsl{Gaia DR1}}
-\author{Adrian M. Price-Whelan\altaffilmark{\pu,\adrn},
-        Semyeong Oh\altaffilmark{\pu},
+\author{Semyeong Oh\altaffilmark{\pu,\lead},
+        Adrian M. Price-Whelan\altaffilmark{\pu},
         David W. Hogg\altaffilmark{\ccpp,\mpia},
@@
 \newcommand{\pu}{1}
-\newcommand{\adrn}{2}
+\newcommand{\lead}{2}
 \newcommand{\ccpp}{3}
@@
                    Princeton University, Princeton, NJ 08544, USA}
-\altaffiltext{\adrn}{To whom correspondence should be addressed:
-                     \texttt{adrn@princeton.edu}}
+\altaffiltext{\lead}{To whom correspondence should be addressed:
+                     \texttt{semyeong@astro.princeton.edu}}
 \altaffiltext{\ccpp}{Center for Cosmology and Particle Physics,
@@
 The primary data set used in this \documentname\ is the Tycho-Gaia Astrometric
-Solution (TGAS), released as a part of Data Release 1 of the Gaia mission
-(\citealt{many}). TGAS contains astrometric measurements (sky position,
+Solution (\tgas), released as a part of Data Release 1 (\dr) of the Gaia mission
+(\citealt{many}). \tgas\ contains astrometric measurements (sky position,
 parallax, and proper motions) and associated covariance matrices for a large
 fraction of the \project{Tycho-2} catalog (\citealt{tycho2}) with median
-astrometric precision equivalent to the \project{Hipparcos} catalog ($\approx
-0.3~{\rm mas}$; \citealt{hipparcos}). In terms of parallax signal-to-noise
-($\snr = \varpi/\sigma_\varpi$), the TGAS catalog contains 1240114 stars with
-$\snr > 4$, 619618 stars with $\snr > 8$, and 197105 stars with $\snr > 16$. The
-increase in precision over Tycho-2 and increase in volume over Hipparcos makes
-the TGAS catalog a powerful data set to search for widely-separated, co-moving
-stars.
-
+astrometric precision comparable to that of the \project{Hipparcos} catalog
+($\approx 0.3~{\rm mas}$; \citealt{hipparcos}). In terms of parallax
+signal-to-noise ($\snr = \varpi/\sigma_\varpi$), the \tgas\ catalog contains
+42385 high-precision stars with $\snr > 32$: here we visualize quantities
+computed from this high-precsision sample to emphasize the quality of the data.
 In the sections that follow, we construct a statistical model that propagates
 the non-trivial uncertainties in the data to our beliefs about the likelihood
-that a given pair are co-moving. Here, we instead compute quantities directly
-from the astrometric measurements to visualize and emphasize the quality of the
-data. For each star with $\snr > 16$, we use the reported parallax and parallax
+that a given pair are co-moving.
+
+For each star with $\snr > 32$, we use the reported parallax and parallax
 error to compute a distance corrected for the Lutz-Kelker bias
@@
 \project{scikit-learn} (\citealt{scikit-learn}). We remove duplicates from
-this sample and are left with 138962 pairs of stars. For all pairs, we then
+this sample and are left with 29870 pairs of stars. For all pairs, we then
 compute the tangential velocity difference
@@
 \end{equation}
-\figname~\ref{fig:dv-sep} shows these quantities plotted for all pairs that pass
-the above signal-to-noise cut. Pairs with small physical separation also tend to
-have small velocity differences. Intriguingly, there appear to be an abundance
-of pairs with small velocity separation at separations $|\delta\bs{x}| \gtrsim
-1~{\rm pc}$.
+\figname~\ref{fig:dv-sep} shows the tangential velocity difference against the
+physical separation plotted for all pairs that pass the above signal-to-noise
+cut. Pairs with small physical separation also tend to have small velocity
+differences. Intriguingly, there appear to be an abundance of pairs with small
+velocity separation at separations $|\delta\bs{x}| \gtrsim 1~{\rm pc}$.
 
@@
 
-For a given pair of stars $(i,j)$, we compute the fully marginalized likelihood
-(FML) that the pair has the same 3-space velocity (hypothesis 1,
-$\mathcal{L}_1$) and the FML of the stars having different 3-space velocities
-(hypothesis 2, $\mathcal{L}_2$). We use the ratio of these FMLs as a scalar for
-selecting candidate widely-separated binary stars. \gaia\ \dr\ only contains
-parallax, $\varpi$, and two proper motion components, $\propm = (\mu_\alpha,
-\mu_\delta)$ for each star, so for the majority of sources we only have
-information about the tangential velocities of the stars. Where there is overlap
-with \rave\ or \apogee, some number of observed radial velocities, $v_r$, with
-observed variances $\sigma^2_{v_r}$ (independent of the \gaia\ data) may also be
-available. We assume that the uncertainties in these observed quantites are
-known with covariance matrix $\matrx{C}$. The likelihood functions $L_1, L_2$
-are marginalized over true 3-space velocity and distance for each star in the
-pair.
+The abundance of pairs of stars with small velocity difference in the
+high-precsision sample of \figname~\ref{fig:dv-sep} suggests that there are a
+significant number of co-moving stars in the \tgas\ data at separations as large
+as a few parsecs. Here we develop a method to select high-confidence co-moving
+stars that properly incorporates the complex uncertainties associated with the
+\gaia\ data. We make the following assumptions in order to construct a
+statistical model (a likelihood function with explicit priors on our
+parameters):
+\begin{itemize}
+  \item We assume that the uncertainties in the data---parallax, $\varpi$, and
+    two proper motion components, $\propm = (\begin{array}[t]{c c} \mu_\alpha &
+    \mu_\delta\end{array})^\mathsf{T}$---are Gaussian with known covariances.
+  \item We assume that the tangential velocities of a given pair of stars in the
+    \tgas\ sample (relative to the solar system barycenter) are either (1)
+    co-moving with velocities equal to the circular velocity of a $2~\msun$
+    binary star system with the given physical separation, (2) co-moving with
+    velocities drawn from a zero-mean Gaussian with velocity dispersion $\sigma
+    = 1~\kms$, or (3) drawn from a mixture of three isotropic, zero-mean
+    Gaussian distributions with velocity dispersions equal to $(\sigma_1,
+    \sigma_2, \sigma_3) = (8, 32, 128)~\kms$ meant to represent young disk
+    stars, old disk stars, and halo stars.
+\end{itemize}
+
+Under these assumptions, the likelihood function for the proper-motion data for
+a star given true distance, $d$, and true tangential velocity, $\bs{v}^\perp =
+(\begin{array}[t]{c c} v_\alpha & v_\delta\end{array})^\mathsf{T}$ is
+\begin{align}
+  L(\bs{v}^\perp, d) &=
+    \left[\det\left(\frac{\matrx{C}^{-1}}{2\pi}\right)\right]^{1/2} \,
+    \exp \left[ -\frac{1}{2} \transp{\left(\bs{\mu} - \bs{x}_\theta \right)} \,
+    \matrx{C}^{-1} \,
+    \left(\bs{\mu} - \bs{x}_\theta \right) \right] \\
+  \bs{x}_\theta &= d^{-1} \, \bs{v}^\perp
+\end{align}
+
+\todo{APW is working on this text...}
+We compute the fully
+marginalized likelihood (FML) that the pair has the same 3-space velocity
+(hypothesis 1, $\mathcal{L}_1$) and the FML of the stars having different
+3-space velocities (hypothesis 2, $\mathcal{L}_2$). We use the ratio of these
+FMLs as a scalar for selecting candidate co-moving stars, as described below in
+more detail. The likelihood functions $L_1, L_2$ are marginalized over true
+3-space velocity and distance for each star in the pair.
 \begin{align}
@@
 where
-\begin{align}
-  L(\bs{v}, d) &=
-    \left[\det\left(\frac{\matrx{C}^{-1}}{2\pi}\right)\right]^{1/2} \,
-    \exp \left[ -\frac{1}{2} \transp{\left(\bs{x} - \bs{x}_\theta \right)} \,
-    \matrx{C}^{-1} \,
-    \left(\bs{x} - \bs{x}_\theta \right) \right] \\
-  \bs{x} &= \transp{(\begin{array}{c c c} \mu_\alpha &
-    \mu_\delta & v_r \end{array})} \\
-  \bs{x}_\theta &= \transp{\left(\begin{array}{c c c} d^{-1}\,v_\alpha &
-    d^{-1}\,v_\delta & \tilde{v}_r \end{array}\right)}
-\end{align}
+
 and we compute the posterior pdfs for the true distances, $p(d\given\varpi)$,
@@
 signal-to-noise.
-
-\section{Results 1: A catalog of candidate wide binaries}
-
-\todo{Galactic context / orbits}
-
-\section{Results 2: Star clusters and cluster remnants}
-
-\section{Conclusions}
-
-\acknowledgements
-
-This research was partially supported by the \acronym{NSF} (grants
-  \acronym{IIS-1124794}, \acronym{AST-1312863}, \acronym{AST-1517237}),
-  \acronym{NASA} (grant \acronym{NNX12AI50G}),
-  and the Moore-Sloan Data Science Environment at \acronym{NYU}. The data
-analysis presented in this article was partially performed on computational
-resources supported by the Princeton Institute for Computational Science and
-Engineering (PICSciE) and the Office of Information Technology's High
-Performance Computing Center and Visualization Laboratory at Princeton
-University.
-
-\software{The code used in this project is available from
-\url{https://github.com/smoh/gaia-wide-binaries} under the MIT open-source
-software license. This version was generated at git commit
-\texttt{\githash\,(\gitdate)}.
-This research additionally utilized:
-    \texttt{Astropy} (\citealt{Astropy-Collaboration:2013}),
-    %\texttt{emcee} (\citealt{Foreman-Mackey:2013}),
-    \texttt{IPython} (\citealt{Perez:2007}),
-    \texttt{matplotlib} (\citealt{Hunter:2007}),
-    and \texttt{numpy} (\citealt{Van-der-Walt:2011}).}
-
-% \facility{\sdssiii, \apogee}
-
-\bibliographystyle{aasjournal}
-\bibliography{refs}
-
-\appendix
-
-\section{Some mathematics relevant to the above} \label{sec:appendix}
-
-\todo{clean this up}
-
-In what follows, all vectors are column vectors, unless we have transposed them.
-We indicate whether a symbol represents a scalar, vector, or tensor
-(matrix) by context, not typesetting.
-Properties of the squared exponential:
-\begin{eqnarray}
-  \ln\left[\int\exp(-\frac{1}{2}\,\transp{[x-\nu]}\,\inv{A}\,[x-\nu] - \Delta)\,\dd x\right]
-  &=& +\frac{1}{2}\ln ||2\pi\,A|| -\Delta
-  \quad ,
-\end{eqnarray}
-where $x$ and $\nu$ are $D$-dimensional vectors, $A$ is a positive definite
-matrix, $\Delta$ is a scalar, and the integral is implicitly over all
-of $D$-dimensional $x$-space.
-We will need to complete squares.
-If we equate
-\begin{eqnarray}
-  \frac{1}{2}\,\transp{[x-\nu]}\,\inv{A}\,[x-\nu] + \Delta
-  &=& \frac{1}{2}\,\transp{x}\,\inv{A}\,x + \transp{x}\,B\,b + C
-  \quad ,
-\end{eqnarray}
-where $B\,b$ is a $d$-vector, and $C$ is a scalar, then we find
-\begin{eqnarray}
-  \nu &=& -A\,B\,b
-  \\
-  \Delta & = & C - \frac{1}{2}\,\transp{\nu}\,\inv{A}\,\nu
-  \quad .
-\end{eqnarray}
 
@@
 
+\section{Results 1: A catalog of candidate wide binaries}
+
+\todo{Galactic context / orbits}
+
+\section{Results 2: Star clusters and cluster remnants}
+
+\section{Conclusions}
+
+\acknowledgements
+
+This research was partially supported by the \acronym{NSF} (grants
+  \acronym{IIS-1124794}, \acronym{AST-1312863}, \acronym{AST-1517237}),
+  \acronym{NASA} (grant \acronym{NNX12AI50G}),
+  and the Moore-Sloan Data Science Environment at \acronym{NYU}. The data
+analysis presented in this article was partially performed on computational
+resources supported by the Princeton Institute for Computational Science and
+Engineering (PICSciE) and the Office of Information Technology's High
+Performance Computing Center and Visualization Laboratory at Princeton
+University.
+
+\software{The code used in this project is available from
+\url{https://github.com/smoh/gaia-wide-binaries} under the MIT open-source
+software license. This version was generated at git commit
+\texttt{\githash\,(\gitdate)}.
+This research additionally utilized:
+    \texttt{Astropy} (\citealt{Astropy-Collaboration:2013}),
+    %\texttt{emcee} (\citealt{Foreman-Mackey:2013}),
+    \texttt{IPython} (\citealt{Perez:2007}),
+    \texttt{matplotlib} (\citealt{Hunter:2007}),
+    and \texttt{numpy} (\citealt{Van-der-Walt:2011}).}
+
+% \facility{\sdssiii, \apogee}
+
+\bibliographystyle{aasjournal}
+\bibliography{refs}
+
+\appendix
+
+\section{Some mathematics relevant to the above} \label{sec:appendix}
+
+\todo{clean this up}
+
+In what follows, all vectors are column vectors, unless we have transposed them.
+We indicate whether a symbol represents a scalar, vector, or tensor
+(matrix) by context, not typesetting.
+Properties of the squared exponential:
+\begin{eqnarray}
+  \ln\left[\int\exp(-\frac{1}{2}\,\transp{[x-\nu]}\,\inv{A}\,[x-\nu] - \Delta)\,\dd x\right]
+  &=& +\frac{1}{2}\ln ||2\pi\,A|| -\Delta
+  \quad ,
+\end{eqnarray}
+where $x$ and $\nu$ are $D$-dimensional vectors, $A$ is a positive definite
+matrix, $\Delta$ is a scalar, and the integral is implicitly over all
+of $D$-dimensional $x$-space.
+We will need to complete squares.
+If we equate
+\begin{eqnarray}
+  \frac{1}{2}\,\transp{[x-\nu]}\,\inv{A}\,[x-\nu] + \Delta
+  &=& \frac{1}{2}\,\transp{x}\,\inv{A}\,x + \transp{x}\,B\,b + C
+  \quad ,
+\end{eqnarray}
+where $B\,b$ is a $d$-vector, and $C$ is a scalar, then we find
+\begin{eqnarray}
+  \nu &=& -A\,B\,b
+  \\
+  \Delta & = & C - \frac{1}{2}\,\transp{\nu}\,\inv{A}\,\nu
+  \quad .
+\end{eqnarray}
+
 \end{document}

### 2016-10-06 gwb/likelihood.py
weights for priors

@@ -131,20 +131,26 @@ def _marg_likelihood_helper(ds, data, Vinv, v_scatter):
     assert sgn > 0
     return 0.5*log_detA - Delta
 
-def ln_H1_marg_v_likelihood(d1, d2, data1, data2, Vinvs, v_scatter=0.):
+def ln_H1_marg_v_likelihood(d1, d2, data1, data2, Vinvs, v_scatter=0., prior_weights=None):
     Vinvs = np.array(Vinvs)
     if Vinvs.ndim < 3:
         Vinvs = Vinvs[None]
+    if prior_weights is None:
+        prior_weights = np.ones(Vinv.shape[0])/Vinv.shape[0]
 
     ds = np.array([d1, d2])
     data = [data1, data2]
-    return logsumexp([_marg_likelihood_helper(ds, data, Vinv, v_scatter) for Vinv in Vinvs])
+    return logsumexp([_marg_likelihood_helper(ds, data, Vinv, v_scatter) for Vinv in Vinvs],
+                     b=prior_weights)
 
-def ln_Q(d, data, Vinvs, v_scatter=0.):
+def ln_Q(d, data, Vinvs, v_scatter=0., prior_weights=None):
     Vinvs = np.array(Vinvs)
     if Vinvs.ndim < 3:
         Vinvs = Vinvs[None]
-    return logsumexp([_marg_likelihood_helper(d, data, Vinv, v_scatter) for Vinv in Vinvs])
+    if prior_weights is None:
+        prior_weights = np.ones(Vinv.shape[0])/Vinv.shape[0]
+    return logsumexp([_marg_likelihood_helper(d, data, Vinv, v_scatter) for Vinv in Vinvs],
+                     b=prior_weights)
 
-def ln_H2_marg_v_likelihood(d1, d2, data1, data2, Vinv, v_scatter=0.):
-    return (ln_Q(d1, data1, Vinv, v_scatter) + ln_Q(d2, data2, Vinv, v_scatter))
+def ln_H2_marg_v_likelihood(d1, d2, data1, data2, Vinv, v_scatter=0., prior_weights=None):
+    return (ln_Q(d1, data1, Vinv, v_scatter, prior_weights) + ln_Q(d2, data2, Vinv, v_scatter, prior_weights))

### 2016-10-06 gwb/data.py
do not want this now

@@ -99,8 +99,6 @@ class TGASData(object):
 
         if lutz_kelker:
             snr = self._data['parallax'] / self._data['parallax_error']
-            if np.any(snr < 4):
-                raise ValueError("S/N is smaller than 4!")
             tmp = self._data['parallax'] * (0.5 + 0.5*np.sqrt(1 - 16/snr**2))
 
         else:

### 2016-10-06 gwb/likelihood.py
fix trivial mistake

@@ -136,7 +136,7 @@ def ln_H1_marg_v_likelihood(d1, d2, data1, data2, Vinvs, v_scatter=0., prior_wei
     if Vinvs.ndim < 3:
         Vinvs = Vinvs[None]
     if prior_weights is None:
-        prior_weights = np.ones(Vinv.shape[0])/Vinv.shape[0]
+        prior_weights = np.ones(Vinvs.shape[0])/Vinvs.shape[0]
 
     ds = np.array([d1, d2])
     data = [data1, data2]
@@ -148,7 +148,7 @@ def ln_Q(d, data, Vinvs, v_scatter=0., prior_weights=None):
     if Vinvs.ndim < 3:
         Vinvs = Vinvs[None]
     if prior_weights is None:
-        prior_weights = np.ones(Vinv.shape[0])/Vinv.shape[0]
+        prior_weights = np.ones(Vinvs.shape[0])/Vinvs.shape[0]
     return logsumexp([_marg_likelihood_helper(d, data, Vinv, v_scatter) for Vinv in Vinvs],
                      b=prior_weights)
 

### 2016-10-14 paper/ms.tex
methods section

@@
 % ----------
-
+% - check for TODO's and references to APW, HOGG, SMOH
 % style notes
@@
 \hypersetup{backref,breaklinks,colorlinks,urlcolor=cbblue,linkcolor=cbblue,citecolor=black}
+\usepackage{dsfont}
 
@@
 \newcommand{\snr}{[S/N]_\varpi}
+\newcommand{\eye}{\mathds{1}}
 
@@
     two proper motion components, $\propm = (\begin{array}[t]{c c} \mu_\alpha &
-    \mu_\delta\end{array})^\mathsf{T}$---are Gaussian with known covariances.
-  \item We assume that the tangential velocities of a given pair of stars in the
-    \tgas\ sample (relative to the solar system barycenter) are either (1)
-    co-moving with velocities equal to the circular velocity of a $2~\msun$
-    binary star system with the given physical separation, (2) co-moving with
-    velocities drawn from a zero-mean Gaussian with velocity dispersion $\sigma
-    = 1~\kms$, or (3) drawn from a mixture of three isotropic, zero-mean
-    Gaussian distributions with velocity dispersions equal to $(\sigma_1,
-    \sigma_2, \sigma_3) = (8, 32, 128)~\kms$ meant to represent young disk
-    stars, old disk stars, and halo stars.
+    \mu_\delta\end{array})^\mathsf{T}$---are Gaussian with known covariances
+    $\matrx{C}$.
+  \item We assume that the 3-space velocities of a given pair of stars
+    $(\bs{v}_i, \bs{v}_j)$ in the \tgas\ sample (relative to the solar system
+    barycenter) are either (1) co-moving with velocity drawn from a velocity
+    prior $p(\bs{v})$ and  velocity difference drawn from a zero-mean Gaussian
+    with velocity dispersion $s$, or (2) individually drawn from a velocity
+    prior $p(\bs{v})$.
 \end{itemize}
 
-Under these assumptions, the likelihood function for the proper-motion data for
-a star given true distance, $d$, and true tangential velocity, $\bs{v}^\perp =
+Under these assumptions, the likelihood of a proper motion measurement for a
+star given true distance, $d$, and true tangential velocity, $\bs{v}^\perp =
 (\begin{array}[t]{c c} v_\alpha & v_\delta\end{array})^\mathsf{T}$ is
 \begin{align}
-  L(\bs{v}^\perp, d) &=
-    \left[\det\left(\frac{\matrx{C}^{-1}}{2\pi}\right)\right]^{1/2} \,
+  L(\bs{v}, d, s^2) &=
+    \left[\det\left(\frac{\tilde{\matrx{C}}^{-1}}{2\pi}\right)\right]^{1/2} \,
     \exp \left[ -\frac{1}{2} \transp{\left(\bs{\mu} - \bs{x}_\theta \right)} \,
-    \matrx{C}^{-1} \,
+    \tilde{\matrx{C}}^{-1} \,
     \left(\bs{\mu} - \bs{x}_\theta \right) \right] \\
@@
 \end{align}
+where the tangential velocity $\bs{v}^\perp$ is related to the 3-space velocity
+$\bs{v}$ through projection onto a tangent plane at a given star's sky position
+$(\alpha, \delta)$
+\begin{align}
+  \bs{v}^\perp &= \matrx{M}\,\bs{v} \\
+  & = \left(
+      \begin{array}{c c c}
+        -\sin\alpha & \cos\alpha & 0 \\
+        -\sin\delta \, \cos\alpha & -\sin\delta \, \sin\alpha & \cos\delta
+      \end{array}
+    \right) \,
+    \left(\begin{array}{c} v_x \\ v_y \\ v_z \end{array}\right)
+\end{align}
+and the modified covariance matrix $\tilde{\matrx{C}}$ is
+\begin{equation}
+  \tilde{\matrx{C}} = \matrx{C} + \eye\,s^2
+\end{equation}
 
-\todo{APW is working on this text...}
-We compute the fully
-marginalized likelihood (FML) that the pair has the same 3-space velocity
-(hypothesis 1, $\mathcal{L}_1$) and the FML of the stars having different
-3-space velocities (hypothesis 2, $\mathcal{L}_2$). We use the ratio of these
-FMLs as a scalar for selecting candidate co-moving stars, as described below in
-more detail. The likelihood functions $L_1, L_2$ are marginalized over true
-3-space velocity and distance for each star in the pair.
+We compute the fully marginalized likelihood (FML) that a given pair of stars
+has the same 3-space velocity with a small velocity difference drawn from a
+zero-mean Gaussian with variance $s^2$ (hypothesis 1, $\mathcal{L}_1$), and
+the FML of the stars having different 3-space velocities (hypothesis 2,
+$\mathcal{L}_2$).
+We use the FML ratio $\mathcal{L}_1/\mathcal{L}_2$ as a scalar for selecting
+candidate co-moving stars, as described below in more detail.
+To compute these FMLs, the likelihood functions $L_i, L_j$ are marginalized
+over true 3-space velocity and distance for each star in the pair $(i,j)$.
 \begin{align}
@@
     \int \, \dd d_i \, \dd d_j \, \dd^3 \bs{v} \,
-    L_i(\bs{v}, d_i) \,
-    L_j(\bs{v}, d_j) \,
+    L_i(\bs{v}, d_i, s^2) \,
+    L_j(\bs{v}, d_j, s^2) \,
     p(\bs{v}) \, p(d_i \given \varpi_i) \, p(d_j \given \varpi_j) \\
@@
     \int \, \dd d_i \, \dd d_j \, \dd^3 \bs{v}_i \, \dd^3 \bs{v}_j \,
-    L_i(\bs{v}_i, d_i) \,
-    L_j(\bs{v}_j, d_j) \,
-    p(\bs{v}_i) \, p(\bs{v}_j) \, p(d_i \given \varpi_i) \, p(d_j \given \varpi_j) \label{eq:hyp2}
+    L_i(\bs{v}_i, d_i, 0) \,
+    L_j(\bs{v}_j, d_j, 0) \,
+    p(\bs{v}_i) \, p(\bs{v}_j) \, p(d_i \given \varpi_i) \, p(d_j \given \varpi_j). \label{eq:hyp2}
 \end{align}
-where
-
-and we compute the posterior pdfs for the true distances, $p(d\given\varpi)$,
-using a uniform space density prior \citep{Astraatmadja:2016} on heliocentric
-distance
+The marginalization integral for hypothesis 2 can be split into the product of
+two simpler integrals $\mathcal{L}_2 = Q_i \, Q_j$ where
 \begin{equation}
-  p(d) \propto d^2
+  Q = \int \, \dd d \, \dd^3 \bs{v} \, L(\bs{v}, d, 0) \, p(\bs{v}) \, p(d\given\varpi)
 \end{equation}
-We assume Gaussian priors on the true velocity components
+If the velocity prior $p(\bs{v})$ is also Gaussian, the integrals over velocity
+in both cases can be performed analytically:
+\todo{SMOH: is this what we do?}
+We use an equal-weight mixture of three isotropic, zero-mean Gaussian
+distributions
 \begin{equation}
-  p(\bs{v}) = \normal(0,25)~\kms
+  p(\bs{v}) = \frac{1}{3} \, \sum_{m=1}^3 \, \mathcal{N}(0, \sigma_m^2)
 \end{equation}
-
-For hypothesis 1, the integral over velocity is non-trivial because the same
-``true'' 3-space velocity vector projects into tangential velocity components
-differently depending on sky position ($v_\alpha = v_\alpha(\bs{v}, \alpha,
-\delta)$). The marginalization integral for hypothesis 2 is simpler because the
-integral in \eqname~\ref{eq:hyp2} can be split into the product of two simpler
-integrals $\mathcal{L}_2 = Q_i \, Q_j$ where
-\begin{equation}
-  Q = \int \, \dd d \, \dd^3 \bs{v} \, L(\bs{v}, d) \, p(\bs{v}) \, p(d\given\varpi)
-\end{equation}
-In either case, the marginalization over true 3-space velocity can be performed
-analytically; we derive the relevant expressions in
-\sectionname~\ref{sec:appendix}. After marginalizing over velocity, the
-likelihood integrands only depend on distance; we numerically compute the
-integrals over the true distances to each star in a pair, $d_1,d_2$, using Monte
-Carlo integration with $K$ samples from the distance posterior pdfs:
+with velocity dispersions $(\sigma_1, \sigma_2, \sigma_3) = (8, 32, 128)~\kms$
+meant to represent young disk stars, old disk stars, and halo stars.
+We derive the relevant expressions in Appendix~\ref{sec:appendix}.
+After marginalizing over velocity, the likelihood integrands only depend on
+distance; we numerically compute the integrals over the true distances to each
+star in a pair, $d_1,d_2$, using Monte Carlo integration with $K$ samples from
+the distance posterior pdfs:
 \begin{equation}
@@
 \end{equation}
+where $\tilde{L}(d)$ is the velocity-marginalized likelihood function.
 Through experimentation, we have found that $K=128$ samples are sufficient for
@@
 signal-to-noise.
+
+\section{Results 1: A catalog of candidate wide binaries}
+
+\todo{Galactic context / orbits}
+
+\section{Results 2: Star clusters and cluster remnants}
+
+\section{Conclusions}
+
+\acknowledgements
+
+This research was partially supported by the \acronym{NSF} (grants
+  \acronym{IIS-1124794}, \acronym{AST-1312863}, \acronym{AST-1517237}),
+  \acronym{NASA} (grant \acronym{NNX12AI50G}),
+  and the Moore-Sloan Data Science Environment at \acronym{NYU}. The data
+analysis presented in this article was partially performed on computational
+resources supported by the Princeton Institute for Computational Science and
+Engineering (PICSciE) and the Office of Information Technology's High
+Performance Computing Center and Visualization Laboratory at Princeton
+University.
+
+\software{The code used in this project is available from
+\url{https://github.com/smoh/gaia-wide-binaries} under the MIT open-source
+software license. This version was generated at git commit
+\texttt{\githash\,(\gitdate)}.
+This research additionally utilized:
+    \texttt{Astropy} (\citealt{Astropy-Collaboration:2013}),
+    %\texttt{emcee} (\citealt{Foreman-Mackey:2013}),
+    \texttt{IPython} (\citealt{Perez:2007}),
+    \texttt{matplotlib} (\citealt{Hunter:2007}),
+    and \texttt{numpy} (\citealt{Van-der-Walt:2011}).}
+
+% \facility{\sdssiii, \apogee}
+
+\bibliographystyle{aasjournal}
+\bibliography{refs}
+
+\appendix
+
+\section{Some mathematics relevant to the above} \label{sec:appendix}
+
+\todo{APW will clean this up}
+
+In what follows, all vectors are column vectors, unless we have transposed them.
+We indicate whether a symbol represents a scalar, vector, or tensor
+(matrix) by context, not typesetting.
+Properties of the squared exponential:
+\begin{eqnarray}
+  \ln\left[\int\exp(-\frac{1}{2}\,\transp{[x-\nu]}\,\inv{A}\,[x-\nu] - \Delta)\,\dd x\right]
+  &=& +\frac{1}{2}\ln ||2\pi\,A|| -\Delta
+  \quad ,
+\end{eqnarray}
+where $x$ and $\nu$ are $D$-dimensional vectors, $A$ is a positive definite
+matrix, $\Delta$ is a scalar, and the integral is implicitly over all
+of $D$-dimensional $x$-space.
+We will need to complete squares.
+If we equate
+\begin{eqnarray}
+  \frac{1}{2}\,\transp{[x-\nu]}\,\inv{A}\,[x-\nu] + \Delta
+  &=& \frac{1}{2}\,\transp{x}\,\inv{A}\,x + \transp{x}\,B\,b + C
+  \quad ,
+\end{eqnarray}
+where $B\,b$ is a $d$-vector, and $C$ is a scalar, then we find
+\begin{eqnarray}
+  \nu &=& -A\,B\,b
+  \\
+  \Delta & = & C - \frac{1}{2}\,\transp{\nu}\,\inv{A}\,\nu
+  \quad .
+\end{eqnarray}
 
@@
 
-\section{Results 1: A catalog of candidate wide binaries}
-
-\todo{Galactic context / orbits}
-
-\section{Results 2: Star clusters and cluster remnants}
-
-\section{Conclusions}
-
-\acknowledgements
-
-This research was partially supported by the \acronym{NSF} (grants
-  \acronym{IIS-1124794}, \acronym{AST-1312863}, \acronym{AST-1517237}),
-  \acronym{NASA} (grant \acronym{NNX12AI50G}),
-  and the Moore-Sloan Data Science Environment at \acronym{NYU}. The data
-analysis presented in this article was partially performed on computational
-resources supported by the Princeton Institute for Computational Science and
-Engineering (PICSciE) and the Office of Information Technology's High
-Performance Computing Center and Visualization Laboratory at Princeton
-University.
-
-\software{The code used in this project is available from
-\url{https://github.com/smoh/gaia-wide-binaries} under the MIT open-source
-software license. This version was generated at git commit
-\texttt{\githash\,(\gitdate)}.
-This research additionally utilized:
-    \texttt{Astropy} (\citealt{Astropy-Collaboration:2013}),
-    %\texttt{emcee} (\citealt{Foreman-Mackey:2013}),
-    \texttt{IPython} (\citealt{Perez:2007}),
-    \texttt{matplotlib} (\citealt{Hunter:2007}),
-    and \texttt{numpy} (\citealt{Van-der-Walt:2011}).}
-
-% \facility{\sdssiii, \apogee}
-
-\bibliographystyle{aasjournal}
-\bibliography{refs}
-
-\appendix
-
-\section{Some mathematics relevant to the above} \label{sec:appendix}
-
-\todo{clean this up}
-
-In what follows, all vectors are column vectors, unless we have transposed them.
-We indicate whether a symbol represents a scalar, vector, or tensor
-(matrix) by context, not typesetting.
-Properties of the squared exponential:
-\begin{eqnarray}
-  \ln\left[\int\exp(-\frac{1}{2}\,\transp{[x-\nu]}\,\inv{A}\,[x-\nu] - \Delta)\,\dd x\right]
-  &=& +\frac{1}{2}\ln ||2\pi\,A|| -\Delta
-  \quad ,
-\end{eqnarray}
-where $x$ and $\nu$ are $D$-dimensional vectors, $A$ is a positive definite
-matrix, $\Delta$ is a scalar, and the integral is implicitly over all
-of $D$-dimensional $x$-space.
-We will need to complete squares.
-If we equate
-\begin{eqnarray}
-  \frac{1}{2}\,\transp{[x-\nu]}\,\inv{A}\,[x-\nu] + \Delta
-  &=& \frac{1}{2}\,\transp{x}\,\inv{A}\,x + \transp{x}\,B\,b + C
-  \quad ,
-\end{eqnarray}
-where $B\,b$ is a $d$-vector, and $C$ is a scalar, then we find
-\begin{eqnarray}
-  \nu &=& -A\,B\,b
-  \\
-  \Delta & = & C - \frac{1}{2}\,\transp{\nu}\,\inv{A}\,\nu
-  \quad .
-\end{eqnarray}
-
 \end{document}

### 2016-10-14 paper/ms.tex
words about projection

@@
 differences. Intriguingly, there appear to be an abundance of pairs with small
-velocity separation at separations $|\delta\bs{x}| \gtrsim 1~{\rm pc}$.
+velocity separation at separations $|\delta\bs{x}| \gtrsim 1~{\rm pc}$. Of
+course, this figure really only highlights close pairs with small angular
+separations:
+At large angular separations, the tangential velocity differences will be large
+because of projection effects.
+To select a more representative population of co-moving stars, we must instead
+model the 3-space velocities of the stars given observations of just two
+components.
 
@@
 \end{figure*}
-
-\todo{More naive visualizations, or save for the full sample?}
 

### 2016-10-14 paper/ms.tex
todo about results
typo and consistency

@@
 % Results
-We find pairs of comoving stars out to very wide
+We find pairs of co-moving stars out to very wide
 separations, including separations of a few parsecs! There does not
@@
 The abundance of pairs of stars with small velocity difference in the
-high-precsision sample of \figname~\ref{fig:dv-sep} suggests that there are a
+high-precision sample of \figname~\ref{fig:dv-sep} suggests that there are a
 significant number of co-moving stars in the \tgas\ data at separations as large
@@
     $\matrx{C}$.
+    %SMOH: mention GAIA paper that says Gaussian is good approximation?
   \item We assume that the 3-space velocities of a given pair of stars
@@
 signal-to-noise.
+
+\todo{APW/SMOH: in results, specify what we set $s^2$ to:}
+To search for wide binaries, we set $s^2 = \frac{2 \, G \, \msun}{|\bs{x}_i-\bs{x}_j|}$.
+To search for dissolving open clusters, we set $s^2 = 1~(\kms)^2$.
 

### 2016-10-14 paper/ms.tex
still working on appendix

--- 2016-10-14 paper/ms.tex
@@
 \hypersetup{backref,breaklinks,colorlinks,urlcolor=cbblue,linkcolor=cbblue,citecolor=black}
-\usepackage{dsfont}
 
@@
 \newcommand{\propm}{\bs{\mu}}
-\newcommand{\matrx}[1]{\mathbf{#1}}
+\newcommand{\mat}[1]{\mathbf{#1}}
+\renewcommand{\vec}[1]{\bs{#1}}
 \newcommand{\kms}{\rm km~s^{-1}}
@@
 \newcommand{\snr}{[S/N]_\varpi}
-\newcommand{\eye}{\mathds{1}}
+\newcommand{\eye}{\mathbb{I}}
 
@@
 \begin{equation}
-  |\Delta \bs{v}_{\rm t}| = d_{\rm LK} \,
+  |\Delta \vec{v}_{\rm t}| = d_{\rm LK} \,
                             \sqrt{(\mu_{\alpha,1}-\mu_{\alpha,2})^2 +
@@
 differences. Intriguingly, there appear to be an abundance of pairs with small
-velocity separation at separations $|\delta\bs{x}| \gtrsim 1~{\rm pc}$. Of
+velocity separation at separations $|\delta\vec{x}| \gtrsim 1~{\rm pc}$. Of
 course, this figure really only highlights close pairs with small angular
@@
     \mu_\delta\end{array})^\mathsf{T}$---are Gaussian with known covariances
-    $\matrx{C}$.
+    $\mat{C}$.
   \item We assume that the 3-space velocities of a given pair of stars
-    $(\bs{v}_i, \bs{v}_j)$ in the \tgas\ sample (relative to the solar system
+    $(\vec{v}_i, \vec{v}_j)$ in the \tgas\ sample (relative to the solar system
     barycenter) are either (1) co-moving with velocity drawn from a velocity
-    prior $p(\bs{v})$ and  velocity difference drawn from a zero-mean Gaussian
+    prior $p(\vec{v})$ and  velocity difference drawn from a zero-mean Gaussian
     with velocity dispersion $s$, or (2) individually drawn from a velocity
-    prior $p(\bs{v})$.
+    prior $p(\vec{v})$.
 \end{itemize}
@@
 Under these assumptions, the likelihood of a proper motion measurement for a
-star given true distance, $d$, and true tangential velocity, $\bs{v}^\perp =
+star given true distance, $d$, and true tangential velocity, $\vec{v}^\perp =
 (\begin{array}[t]{c c} v_\alpha & v_\delta\end{array})^\mathsf{T}$ is
 \begin{align}
-  L(\bs{v}, d, s^2) &=
-    \left[\det\left(\frac{\tilde{\matrx{C}}^{-1}}{2\pi}\right)\right]^{1/2} \,
-    \exp \left[ -\frac{1}{2} \transp{\left(\bs{\mu} - \bs{x}_\theta \right)} \,
-    \tilde{\matrx{C}}^{-1} \,
-    \left(\bs{\mu} - \bs{x}_\theta \right) \right] \\
-  \bs{x}_\theta &= d^{-1} \, \bs{v}^\perp
+  L(\vec{v}, d, s^2) &=
+    \left[\det\left(\frac{\tilde{\mat{C}}^{-1}}{2\pi}\right)\right]^{1/2} \,
+    \exp \left[ -\frac{1}{2} \transp{\left(\vec{\mu} - \vec{x}_\theta \right)} \,
+    \tilde{\mat{C}}^{-1} \,
+    \left(\vec{\mu} - \vec{x}_\theta \right) \right] \label{eq:likefn} \\
+  \vec{x}_\theta &= d^{-1} \, \vec{v}^\perp
 \end{align}
-where the tangential velocity $\bs{v}^\perp$ is related to the 3-space velocity
-$\bs{v}$ through projection onto a tangent plane at a given star's sky position
+where the tangential velocity $\vec{v}^\perp$ is related to the 3-space velocity
+$\vec{v}$ through projection onto a tangent plane at a given star's sky position
 $(\alpha, \delta)$
 \begin{align}
-  \bs{v}^\perp &= \matrx{M}\,\bs{v} \\
+  \vec{v}^\perp &= \mat{M}\,\vec{v} \\
   & = \left(
@@
 \end{align}
-and the modified covariance matrix $\tilde{\matrx{C}}$ is
+and the modified covariance matrix $\tilde{\mat{C}}$ is
 \begin{equation}
-  \tilde{\matrx{C}} = \matrx{C} + \eye\,s^2
+  \tilde{\mat{C}} = \mat{C} + s^2 \, \eye
 \end{equation}
@@
   \mathcal{L}_1 &=
-    \int \, \dd d_i \, \dd d_j \, \dd^3 \bs{v} \,
-    L_i(\bs{v}, d_i, s^2) \,
-    L_j(\bs{v}, d_j, s^2) \,
-    p(\bs{v}) \, p(d_i \given \varpi_i) \, p(d_j \given \varpi_j) \\
+    \int \, \dd d_i \, \dd d_j \, \dd^3 \vec{v} \,
+    L_i(\vec{v}, d_i, s^2) \,
+    L_j(\vec{v}, d_j, s^2) \,
+    p(\vec{v}) \, p(d_i \given \varpi_i) \, p(d_j \given \varpi_j) \\
   \mathcal{L}_2 &=
-    \int \, \dd d_i \, \dd d_j \, \dd^3 \bs{v}_i \, \dd^3 \bs{v}_j \,
-    L_i(\bs{v}_i, d_i, 0) \,
-    L_j(\bs{v}_j, d_j, 0) \,
-    p(\bs{v}_i) \, p(\bs{v}_j) \, p(d_i \given \varpi_i) \, p(d_j \given \varpi_j). \label{eq:hyp2}
+    \int \, \dd d_i \, \dd d_j \, \dd^3 \vec{v}_i \, \dd^3 \vec{v}_j \,
+    L_i(\vec{v}_i, d_i, 0) \,
+    L_j(\vec{v}_j, d_j, 0) \,
+    p(\vec{v}_i) \, p(\vec{v}_j) \, p(d_i \given \varpi_i) \, p(d_j \given \varpi_j). \label{eq:hyp2}
 \end{align}
@@
 \begin{equation}
-  Q = \int \, \dd d \, \dd^3 \bs{v} \, L(\bs{v}, d, 0) \, p(\bs{v}) \, p(d\given\varpi)
+  Q = \int \, \dd d \, \dd^3 \vec{v} \, L(\vec{v}, d, 0) \, p(\vec{v}) \, p(d\given\varpi)
 \end{equation}
-If the velocity prior $p(\bs{v})$ is also Gaussian, the integrals over velocity
+If the velocity prior $p(\vec{v})$ is also Gaussian, the integrals over velocity
 in both cases can be performed analytically:
@@
 \begin{equation}
-  p(\bs{v}) = \frac{1}{3} \, \sum_{m=1}^3 \, \mathcal{N}(0, \sigma_m^2)
+  p(\vec{v}) = \frac{1}{3} \, \sum_{m=1}^3 \, \mathcal{N}(0, \sigma_{v,m}^2)
 \end{equation}
-with velocity dispersions $(\sigma_1, \sigma_2, \sigma_3) = (8, 32, 128)~\kms$
-meant to represent young disk stars, old disk stars, and halo stars.
+with velocity dispersions $(\sigma_{v,1}, \sigma_{v,2}, \sigma_{v,3}) = (8, 32,
+128)~\kms$ meant to represent young disk stars, old disk stars, and halo stars.
 We derive the relevant expressions in Appendix~\ref{sec:appendix}.
@@
 signal-to-noise.
+
+\todo{APW/SMOH: in results, specify what we set $s^2$ to:}
+To search for wide binaries, we set $s^2 = \frac{2 \, G \, \msun}{|\vec{x}_i-\vec{x}_j|}$.
+To search for dissolving open clusters, we set $s^2 = 1~(\kms)^2$.
 
@@
 
-\section{Some mathematics relevant to the above} \label{sec:appendix}
-
-\todo{APW will clean this up}
+\section{Relevant properties of Gaussian integrals}
+\label{sec:appendixA}
 
 In what follows, all vectors are column vectors, unless we have transposed them.
-We indicate whether a symbol represents a scalar, vector, or tensor
-(matrix) by context, not typesetting.
-Properties of the squared exponential:
+A relevant exponential integral solution is
 \begin{eqnarray}
-  \ln\left[\int\exp(-\frac{1}{2}\,\transp{[x-\nu]}\,\inv{A}\,[x-\nu] - \Delta)\,\dd x\right]
-  &=& +\frac{1}{2}\ln ||2\pi\,A|| -\Delta
+  \ln\left[\int\exp(-\frac{1}{2}\,
+    \transp{[\vec{x}-\vec{\nu}]} \,
+    \inv{\mat{A}} \,
+    [\vec{x}-\vec{\nu}] - \Delta) \, \dd \vec{x}\right]
+  &=& +\frac{1}{2}\ln ||2\pi\,\mat{A}|| -\Delta
+  \quad , \label{eq:gauss-int}
+\end{eqnarray}
+where $\vec{x}$ and $\vec{\nu}$ are $D$-dimensional vectors, $\mat{A}$ is a
+positive definite matrix, $\Delta$ is a scalar, and the integral is over all of
+$D$-dimensional $\vec{x}$-space.
+To cast our problem in this form, we will need to complete the square of the
+exponential argument.
+If we equate
+\begin{eqnarray}
+  \frac{1}{2}\,\transp{[\vec{x}-\vec{\nu}]}\,\inv{\mat{A}}\,[\vec{x}-\vec{\nu}] + \Delta
+  &=& \frac{1}{2}\,\transp{\vec{x}}\,\inv{\mat{A}}\,\vec{x} + \transp{\vec{x}}\,\mat{B}\,\vec{b} + C
   \quad ,
 \end{eqnarray}
-where $x$ and $\nu$ are $D$-dimensional vectors, $A$ is a positive definite
-matrix, $\Delta$ is a scalar, and the integral is implicitly over all
-of $D$-dimensional $x$-space.
-We will need to complete squares.
-If we equate
+where $\mat{B}\,\vec{b}$ is an $D$-vector, and $C$ is a scalar, then we find
 \begin{eqnarray}
-  \frac{1}{2}\,\transp{[x-\nu]}\,\inv{A}\,[x-\nu] + \Delta
-  &=& \frac{1}{2}\,\transp{x}\,\inv{A}\,x + \transp{x}\,B\,b + C
-  \quad ,
-\end{eqnarray}
-where $B\,b$ is a $d$-vector, and $C$ is a scalar, then we find
-\begin{eqnarray}
-  \nu &=& -A\,B\,b
+  \vec{\nu} &=& -\mat{A}\,\mat{B}\,\vec{b}
   \\
-  \Delta & = & C - \frac{1}{2}\,\transp{\nu}\,\inv{A}\,\nu
+  \Delta & = & C - \frac{1}{2}\,\transp{\vec{\nu}}\,\inv{\mat{A}}\,\vec{\nu}
   \quad .
 \end{eqnarray}
+We will identify terms in our likelihood functions with $\mat{A}$,
+$\mat{B}\,\vec{b}$, and $C$, convert to $\vec{\nu}$ and $\Delta$ and compute
+the marginalized likelihood using \eqname~\ref{eq:gauss-int}.
 
-Now in this case, we want to write down the probability density for the data
-given the distances and velocity (the likelihood), and the prior pdf for
-the velocity, and their product.
-The prior pdf for the velocity is Gaussian:
-\begin{eqnarray}
-  p(v) &=& \normal(v\given 0,V)
-  \\
-  \ln p(v) &=& -\frac{1}{2}\,\ln||2\pi\,V|| -\frac{1}{2}\,\transp{v}\,\inv{V}\,v
-  \\
-  V &=& \sigma_v^2\,I
-  \quad ,
-\end{eqnarray}
-where $v$ is the three-vector velocity, and $V$ is an isotropic
-variance tensor.
+\section{Expressions for the marginalized likelihoods}\label{sec:appendix}
+
+At given distance $d$, the velocity-marginalized likelihood can be computed
+analytically using the expressions in Appendix~\ref{sec:appendixA}.
+We will start by writing down expressions for the the likelihood multiplied by
+the prior pdf for the velocities.
+As described above, we use an isotropic, mixture-of-Gaussians prior on velocity
+where the velocity dispersion of a given mixture component is $\sigma_{v,m}$ and
+the 3-space variance tensor is $\mat{V}_m = \sigma_{v,m}^2 \, \eye$.
+The likelihood for the data is also a Gaussian, as shown in
+\eqname~\ref{eq:likefn}.
+
 Similarly, the likelihood for the data is also a Gaussian.

### 2016-10-14 paper/ms.tex

@@
     $\mat{C}$.
+    \todo{SMOH: mention GAIA paper that says Gaussian is good approximation?}
   \item We assume that the 3-space velocities of a given pair of stars

### 2016-10-15 paper/ms.tex
update appendix

@@
     \right) \,
-    \left(\begin{array}{c} v_x \\ v_y \\ v_z \end{array}\right)
+    \left(\begin{array}{c} v_x \\ v_y \\ v_z \end{array}\right) \label{eq:transformation}
 \end{align}
@@
 analytically using the expressions in Appendix~\ref{sec:appendixA}.
-We will start by writing down expressions for the the likelihood multiplied by
-the prior pdf for the velocities.
 As described above, we use an isotropic, mixture-of-Gaussians prior on velocity
@@
 \eqname~\ref{eq:likefn}.
+We will start by writing down expressions for the the likelihood multiplied by
+the prior pdf for the velocities.
+Here we slightly change the notation used in \sectionname~\ref{sec:methods} for
+simplicity.
 
-Similarly, the likelihood for the data is also a Gaussian.
-Before we write that down, we will make some definitions.
 We construct a velocity-space data vector $y$ as follows:
+\begin{equation}
+  \vec{y} =
+    \transp{\left(
+      \begin{array}{c c c c}
+        d_i\,\mu_{\alpha,i} &
+        d_i\,\mu_{\delta,i} &
+        d_j\,\mu_{\alpha,j} &
+        d_j\,\mu_{\delta,j}
+      \end{array}
+    \right)}
+\end{equation}
+where again the subscripts $i,j$ refer to the indices of each star in the pair
+and we have multiplied the observables (the proper motions) by the distances
+$d_i, d_j$, which is permitted because we are conditioning on the distances.
+Fundamentally, our hypothesis 1 model (the stars have the same velocity with a
+small difference) is
+\begin{equation}
+  \vec{y} = \mat{M} \, \vec{v} + \mathrm{noise}
+\end{equation}
+where now the $4 \times 3$ transformation matrix $\mat{M}$ is a stack of the
+transformation matrices for each star computed from the pair of sky positions
+and using \eqname~\ref{eq:transformation}.
+The noise (in the $y$ vector) is drawn from a $4 \times 4$ Gaussian with
+block-diagonal covariance matrix, $\mat{\Sigma}$, constructed from the
+covariance matrix of the observables, $\mat{C}_i, \mat{C}_j$, and the distances
+$d_i, d_j$:
+\begin{equation}
+  \mat{\Sigma} = \left(
+    \begin{array}{c c}
+      d_i^2 \, \mat{C}_i & 0 \\
+      0 & d_j^2 \, \mat{C}_j
+    \end{array}
+  \right)
+\end{equation}
+
+Given all these definitions, the likelihood function for hypothesis 1 is
 \begin{eqnarray}
-  \transp{y} &\equiv& [(d_1\,\mu_{\alpha1}), (d_1\,\mu_{\delta1}),
-                       (d_2\,\mu_{\alpha2}), (d_2\,\mu_{\delta2})] \quad \mbox{or}
-  \\
-  \transp{y} &\equiv& [(d_1\,\mu_{\alpha1}), (d_1\,\mu_{\delta1}), (v_{r1}),
-                       (d_2\,\mu_{\alpha2}), (d_2\,\mu_{\delta2}), (v_{r2})]
+  p(\data \given \vec{v}, d_i, d_j) &=& d_i^2\,d_j^2\,
+    \normal(\vec{y} \given \mat{M}\,\vec{v}, \mat{\Sigma}) \\
+  \ln p(\data \given \vec{v}, d_i, d_j) &=& 2\,\ln d_i + 2\,\ln d_j
+    -\frac{1}{2}\,\ln||2\pi\,\mat{\Sigma}|| \nonumber \\
+    && \quad -\frac{1}{2}\,\transp{[\vec{y}-\mat{M}\,\vec{v}]}\,
+      \inv{\mat{\Sigma}}\,
+      [\vec{y}-\mat{M}\,\vec{v}]
   \quad ,
 \end{eqnarray}
-where we have given two versions, a 4-component and an 6-component
-version (depending on whether there are measured radial velocities),
-the subscripts 1, 2 refer to the first and second stars (arbitrarily
-ordered) of the pair, we have multiplied the strict observables by
-distances $d_1, d_2$, which we are permitted to do (because we are
-\emph{conditioning} on $d_1, d_2$), and note the wacky minus-ones.
-The vector $y$ is a column vector that we have transposed to make it
-easy to write down.
-Fundamentally, our H1 model (same velocity) is
+where the factor of $d_i^2\,d_j^2$ converts to units of one over \gaia\ data
+from units of one over $y$, which includes the distances.
+
+Now we multiply this likelihood with the velocity prior (here shown only for one
+prior velocity mixture component) and complete the square.
+In the original parameterization, we find:
 \begin{eqnarray}
-  y &=& M\,v + \mathrm{noise}
+  \mat{A} &=& \inv{[\transp{\mat{M}}\,\inv{\mat{\Sigma}}\,\mat{M}+\inv{\mat{V}_m}]}
   \\
-  \transp{M} &=& [\hat{\alpha_1}, \hat{\delta_1},
-                  \hat{\alpha_2}, \hat{\delta_2}] \quad \mbox{or}
+  \vec{\nu} &=& -\mat{A}\,\transp{\mat{M}}\,\inv{\mat{\Sigma}}\,\vec{y}
   \\
-  \transp{M} &=& [\hat{\alpha}_1, \hat{\delta}_2, \hat{r}_1,
-                  \hat{\alpha}_2, \hat{\delta}_2, \hat{r}_2]
-  \quad,
-\end{eqnarray}
-where $M$ is a $6\times 3$ or $8\times 3$ design matrix, with zeros
-in two rows, and unit vectors along the other rows.
-The noise (in the $y$ vector) is drawn from a Gaussian with ($6\times 6$ or $8\times 8$)
-covariance matrix $\Sigma$, constructed from the distances $d_1, d_2$ and
-the \gaia-provided individual-object covariance matrices.
-Given all these definitions, the likelihood function is
-\begin{eqnarray}
-  p(\data\given v,d_1,d_2) &=& d_1^2\,d_2^2\,\normal(y\given M\,v, \Sigma)
-  \\
-  \ln p(\data\given v,d_1,d_2) &=& 3\,\ln d_1 + 3\,\ln d_2
-  -\frac{1}{2}\,\ln||2\pi\,\Sigma|| \nonumber \\ && \quad
-  -\frac{1}{2}\,\transp{[y-M\,v]}\,\inv{\Sigma}\,[y-M\,v]
+  \Delta &=& -2\,\ln d_i -2\,\ln d_j
+    +\frac{1}{2}\,\ln||2\pi\,\mat{\Sigma}|| +\frac{1}{2}\,\ln||2\pi\,\mat{V}_m|| \nonumber \\
+    && \quad +\frac{1}{2}\,\transp{\vec{y}}\,\inv{\mat{\Sigma}}\,\vec{y} -\frac{1}{2}\,\transp{\vec{\nu}}\,\inv{\mat{A}}\,\vec{\nu}
   \quad ,
 \end{eqnarray}
-where the factor of $d_1^2\,d_2^2$ converts to units of per \gaia\ data
-from units of per $y$.
+which we plug in to \eqname~\ref{eq:gauss-int} to get the marginalized
+likelihood conditioned on the two distances $d_i, d_j$.
 
-Now we multiply together (add in the log) and complete the square.
-We get, in the original parameterization:
+The marginalized likelihood for the hypothesis 2 model (the stars have
+independent velocities) is very similar.
+In this case, the marginalized likelihood is a product of two independent
+integrals $Q$, composed in the same way as the hypothesis 1 model but now for
+each star individually, where
 \begin{eqnarray}
-  \inv{A} &=& [\transp{M}\,\inv{\Sigma}\,M+\inv{V}]
+  \vec{y} &=&
+    \transp{\left(
+      \begin{array}{c c}
+        d\,\mu_{\alpha} &
+        d\,\mu_{\delta}
+      \end{array}
+    \right)}
   \\
-  \nu &=& -A\,\transp{M}\,\inv{\Sigma}\,y
+  \mat{\Sigma} &=& d^2 \, \mat{C}
   \\
-  \Delta &=& -2\,\ln d_1 -2\,\ln d_2
-  +\frac{1}{2}\,\ln||2\pi\,\Sigma|| +\frac{1}{2}\,\ln||2\pi\,V|| \nonumber \\ && \quad
-  +\frac{1}{2}\,\transp{y}\,\inv{\Sigma}\,y -\frac{1}{2}\,\transp{\nu}\,\inv{A}\,\nu
+\end{eqnarray}
+and $\mat{M}$ is now the transformation matrix for one star. Then, again,
+\begin{eqnarray}
+  \mat{A} &=& \inv{[\transp{\mat{M}}\,\inv{\mat{\Sigma}}\,\mat{M}+\inv{\mat{V}_m}]}
+  \\
+  \vec{\nu} &=& -\mat{A}\,\transp{\mat{M}}\,\inv{\mat{\Sigma}}\,\vec{y}
+  \\
+  \Delta &=& -2\,\ln d_i -2\,\ln d_j
+    +\frac{1}{2}\,\ln||2\pi\,\mat{\Sigma}|| +\frac{1}{2}\,\ln||2\pi\,\mat{V}_m|| \nonumber \\
+    && \quad +\frac{1}{2}\,\transp{\vec{y}}\,\inv{\mat{\Sigma}}\,\vec{y} -\frac{1}{2}\,\transp{\vec{\nu}}\,\inv{\mat{A}}\,\vec{\nu}
   \quad ,
 \end{eqnarray}
-where there are sign issues that Hogg doesn't like to think about.
-Finally, we marginalize out the velocity to get the marginalized
-likelihood conditioned on the two distances $d_1, d_2$:
-\begin{eqnarray}
-  \ln p(\data\given d_1,d_2)
-  &=& \frac{1}{2}\ln ||2\pi\,A|| -\Delta
-  \quad .
-\end{eqnarray}
-
-The story for the H2 model (independent velocities) is very
-similar. In this case, the marginalized likelihood is a product of
-two independent integrals $Q$ each constructed as follows:
-\begin{eqnarray}
-  \transp{y} &\equiv& [(d\,\mu_{\alpha}), (d\,\mu_{\delta})] \quad \mbox{or}
-  \\
-  \transp{y} &\equiv& [(d\,\mu_{\alpha}), (d\,\mu_{\delta}), (v_{r})]
-  \\
-  \transp{M} &=& [\hat{\alpha}, \hat{\delta}] \quad \mbox{or}
-  \\
-  \transp{M} &=& [\hat{\alpha}, \hat{\delta}_2, \hat{r}]
-  \\
-  \inv{A} &=& [\transp{M}\,\inv{\Sigma}\,M+\inv{V}]
-  \\
-  \nu &=& A\,\transp{M}\,\inv{\Sigma}\,y
-  \\
-  \Delta &=& -2\,\ln d
-  +\frac{1}{2}\,\ln||2\pi\,\Sigma|| +\frac{1}{2}\,\ln||2\pi\,V|| \nonumber \\ && \quad
-  +\frac{1}{2}\,\transp{y}\,\inv{\Sigma}\,y -\frac{1}{2}\,\transp{\nu}\,\inv{A}\,\nu
-  \\
-  Q &=& \frac{1}{2}\ln ||2\pi\,A|| -\Delta
-  \quad ,
-\end{eqnarray}
-where again there are two options for the dimensionality of $y$ and
-$M$, and once the vectors and $\Sigma$ are properly truncated to one
-star, everything else follows.
+and
+\begin{equation}
+  Q = \frac{1}{2}\ln ||2\pi\,\mat{A}|| -\Delta \quad .
+\end{equation}
 

### 2016-10-19 paper/ms.tex
update results and figures

@@
 Solution (\tgas), released as a part of Data Release 1 (\dr) of the Gaia mission
-(\citealt{many}). \tgas\ contains astrometric measurements (sky position,
+\citep{2016arXiv160904172G,2016arXiv160904303L}.
+\tgas\ contains astrometric measurements (sky position,
 parallax, and proper motions) and associated covariance matrices for a large
-fraction of the \project{Tycho-2} catalog (\citealt{tycho2}) with median
+fraction of the \project{Tycho-2} catalog \citep{2000A&A...355L..27H} with median
 astrometric precision comparable to that of the \project{Hipparcos} catalog
-($\approx 0.3~{\rm mas}$; \citealt{hipparcos}). In terms of parallax
+\citep[$\approx 0.3~{\rm mas}$;][]{2007ASSL..350.....V}. In terms of parallax
 signal-to-noise ($\snr = \varpi/\sigma_\varpi$), the \tgas\ catalog contains
 42385 high-precision stars with $\snr > 32$: here we visualize quantities
-computed from this high-precsision sample to emphasize the quality of the data.
+computed from this high-precision sample to emphasize the quality of the data.
 In the sections that follow, we construct a statistical model that propagates
 the non-trivial uncertainties in the data to our beliefs about the likelihood
-that a given pair are co-moving.
+that a given pair of stars is co-moving.
 
@@
 \section{Results 1: A catalog of candidate wide binaries}
+In this section, we examine the distribution of likelihood ratios
+$\ln \mathcal{L}_1 /\mathcal{L}_2$, and come up with a reasonable cut
+for this quantities to select co-moving pairs from candidate pairs list.
+We present the statistics of the resulting co-moving pairs.
+We also match the pairs with known open clusters in the Milky Way to
+provide information on cluster membership.
+
+Figure~\ref{fig:likelihoodratios} shows the likelihood ratios for all 2.6 million
+candidate pairs.
+Because we do not carry over the normalizations of all probabilities involved
+in $L_1$ and $L_2$, the absolute values of the likelihood ratio is not
+a meaningful quantity.
+Given our crude way of coming up with candidate pair list, we expect
+a huge number of pairs that are not genuinely co-moving, compared to
+the smaller number of pairs that are.
+In order to gauge an appropriate cut on the likelihood ratios,
+we compute the likelihood ratios for 200,000 random pairs of stars with the same
+signal-to-noise ratio cut as candidate pairs.
+Additionally, we select from candidate pair list, those with very small
+difference in tangential velocities ($\Delta v_t < 0.5$~km/s) in
+two minimum signal-to-noise ratio\footnote{lower of the two stars in a pair}
+bins (S/N$>32$ and S/N$<32$) that are highly probable
+to be genuine co-moving pairs.
+As expected, we see that the likelihood ratio $\ln \mathcal{L}_1 /\mathcal{L}_2$
+distribution peaks at higher values, favoring hypothesis 1, for highly probable
+co-moving pairs than the random pairs.
+We also see that the values are even higher for higher minimum signal-to-noise ratio pairs.
+
+Based on the comparison with random pairs, we can come up with a cut
+on the likelihood ratios that reduces the cumulative contamination due to random pairs
+to a given threshold.
+The value of the cut, the resulting number of pairs, and
+its fraction of the total number of candidate pairs
+depending on the contamination rate are listed in
+Table~\ref{tab:candidatecounts}.
+Because the likelihood ratio $\ln \mathcal{L}_1 /\mathcal{L}_2$ asymtotically
+is chi-square distribution, the power-law fall off makes the number of resulting
+pairs sensitive to the cut.
+
+Once we have decided the co-moving pairs from candidate pair list,
+we can count how many co-moving neighbors a star has.
+This is presented in Figure~\ref{fig:hist_Nneighbors}.
+It is clear from this figure that we have many aggregates of co-moving stars, most
+likely an open cluster or a moving group. We also see that there is still
+a large number of stars which only one or two co-moving neighbors.
+In order to select pairs that are not associated with any other pair,
+we separate a subset of co-moving pairs whose member is a mutually exclusive
+co-moving neighbor to the other member (``mutually exclusive pairs'').
+The number of mutually exclusive pairs is listed on the last column of
+Table~\ref{tab:candidatecounts}.
+
+\begin{table}[tbh]
+  \caption{Number of pairs}
+  \label{tab:candidatecounts}
+  \centering
+
+  \begin{tabular}{l|c|c|c|r}
+  \hline
+
+  \hline
+  \textbf{Contamination} & \textbf{$\ln \mathcal{L}_1 /\mathcal{L}_2$ cut} &
+    \textbf{Count} & \textbf{Fraction} & \textbf{$N_\textrm{me}$} \\
+  \hline
+    0.01    \% & 5.33 & 92458 & 0.462 \% & 5055 \\
+    0.001   \% & 6.53 & 13354 & 0.067 \% & 1549 \\
+    0.0001  \% & 7.29 &  2832 & 0.014 \% & 582  \\
+    0.00001 \% & 7.35 &  2495 & 0.012 \% & 532  \\
+  \hline
+  \end{tabular}
+\end{table}
+
+\begin{figure*}[p]
+  \begin{center}
+    \includegraphics[width=\textwidth]{figures/likelihoodratios.pdf}
+  \end{center}
+  \caption{%
+    Density histogram of likelihood ratios for candidate pairs and random pairs.
+    We also show the distribution for highly probable genuine co-moving pairs with
+    $\Delta v_t < 0.5$~km/s in high ($S/N>32$) and low ($S/N<32$) signal-to-noise ratio
+    bins. In both $S/N$ bins, the highly probable co-moving pairs are clearly
+    separated from random pairs while the distribution of $\ln \mathcal{L}_1 /\mathcal{L}_2$
+    for candidate pairs is expected to be a combination of
+    a huge number of ``random'' pairs that are not actually co-moving, and
+    a smaller number of genuine co-moving pairs with varying $S/N$.
+    \label{fig:likelihoodratios}}
+\end{figure*}
+
+We now examine the separation distribution of co-moving pairs in
+Figure~\ref{fig:hist_separation}.
+
+Finally, we show the distribution co-moving pairs in galactic longitude and
+distance in Figure~\ref{fig:glon_d_pairlines}.
+In the top panel, we show all co-moving pairs with contamination rate of $0.01~\%$
+(undersampled by a factor of 4 in order to unclutter the plot) as connected line segments.
+We find many aggregates of co-moving pairs, which coincides with the positions of
+known Milky Way star clusters \citep{Kharchenko:2016aa} indicated by red circles.
+Clumps of co-moving pairs at $l=300-360$~degrees and $d=100-200$~pc correspond
+to the location of OB associations
+Upper Scorpius, Upper Centaurus Lupus, and Lower Centaurus Crux
+\citep{de-Zeeuw:1999aa}.
+The strong correlation with known co-moving structures disappears when we only
+consider mutually exclusive pairs (bottom panel).
+
+\begin{figure*}[p]
+  \begin{center}
+    \includegraphics[width=\textwidth]{figures/hist_Nneighbors.pdf}
+  \end{center}
+  \caption{%
+    Histogram of the number of co-moving neighbors to a star for varying
+    degrees of contamination due to random pairs.
+    \todo{SMOH: What is the optimal bins for this histogram to separate
+    small number cases e.g., binary case (N=1) appropriately?}
+    \label{fig:hist_Nneighbors}}
+\end{figure*}
+
+\begin{figure*}[p]
+  \begin{center}
+    \includegraphics[width=\textwidth]{figures/hist_separation.pdf}
+    \includegraphics[width=\textwidth]{figures/hist_separation_mutexc.pdf}
+  \end{center}
+  \caption{%
+    Top: Separation distribution of co-moving pairs of stars for varying degrees
+    of contamination due to random pairs. The distribution peaks around 10~pc,
+    which is similar to the typical tidal radius of clusters in the Milky Way
+    \citep{Kharchenko:2013aa}.
+    Bottom: Same as the previous plot but only for mutually exclusive pairs.
+    \label{fig:hist_separation}}
+\end{figure*}
+
+\begin{figure*}[p]
+  \begin{center}
+    \includegraphics[width=\textwidth]{figures/glon_d_pairlines.pdf}
+    \includegraphics[width=\textwidth]{figures/glon_d_pairlines_mutexc.pdf}
+  \end{center}
+  \caption{%
+    Top: Distribution of co-moving pairs in galactic longitude and
+    distance. Each co-moving pair is connected by a line.
+    Red circles mark the locations of 135 known Milky Way open clusters within
+    $d<600$~pc from \citealt{Kharchenko:2016aa} while
+    stars in known OB associations \citep{de-Zeeuw:1999aa} are colored
+    by which association they belong.
+    We see that some of the aggregates of co-moving pairs are clearly
+    associated with known open clusters, and OB associations.
+    Bottom: Same as the previous plot but only for mutually exclusive pairs.
+    The mutually exclusive pairs no longer traces any known co-moving structures
+    strongly.
+    \label{fig:glon_d_pairlines}}
+\end{figure*}
 

### 2016-11-02 paper/ms.tex
rewrote data; added new figures

@@
 \renewcommand{\vec}[1]{\bs{#1}}
-\newcommand{\kms}{\rm km~s^{-1}}
+\newcommand{\kms}{\ensuremath{\rm km~s^{-1}}}
 \newcommand{\msun}{{\rm M}_\odot}
@@
 signal-to-noise ($\snr = \varpi/\sigma_\varpi$), the \tgas\ catalog contains
-42385 high-precision stars with $\snr > 32$: here we visualize quantities
-computed from this high-precision sample to emphasize the quality of the data.
-In the sections that follow, we construct a statistical model that propagates
+42385 high-precision stars with $\snr > 32$.
+
+We construct a sample of candidate pair list as follows.
+We first apply a global parallax S/N cut of 8 to \tgas, which leaves 619,618 stars.
+Then, for each star, we search a potential co-moving partner star with a
+difference in tangential velocity $\Delta v_t < 10$~\kms\ within 10 pc
+using RA, DEC, a point-estimate of distance, and the proper motions.
+Throughout this \documentname\, we use the following estimator for
+the point-estimate of distance applying a correction for the Lutz-Kelker bias
+(\citealt{lutzkelker}):
+\begin{equation}
+  d = 1000 \, \left[\frac{\varpi}{2} \,
+    \left(1 + \sqrt{1 - \frac{16}{\snr^2}} \right) \right]^{-1} \, {\rm pc}
+\end{equation}
+where $\varpi$ is the parallax in mas.
+The difference in tangential velocity between two stars is, then,
+\begin{equation}
+  |\Delta \vec{v}_{\rm t}| = |d_1 \vec\mu_1 - d_2 \vec\mu_2|
+\end{equation}
+where $\vec\mu = (\mu_\alpha, \mu_\delta)$.
+
+Figure~\ref{fig:dv-sep} shows $|\Delta v_t|$ against the physical separation
+for the resulting 271,232 unique candidate pairs.
+A few key observations can be made:
+
+\begin{itemize}
+  \item At small separation ($<1$~pc), there is a population of pairs with
+  very small tangential velocity difference ($<2$~km/s). Given that these pairs
+  are very close in both positions and proper motions,
+  it is highly probable that they are actually co-moving e.g., as wide binaries.
+
+  \item  However, it is not only (wide) binaries which are gravitationally
+  bound that are expected to appear as co-moving.
+  OB associations, moving groups, and open clusters are also of consideration.
+  Since the basic unit of our choice is a pair of stars, we can expect that these astrophysical objects may be detected as networks of pairs.
+  As the pair separation increases, the nature of co-moving pairs
+  will change from binaries to those related to these larger objects,
+  which will tend to subtend a larger angle in sky.
+  Under the assumption of having the same 3D velocity,
+  two stars of a co-moving pair can essentially be thought of as projections
+  of this velocity onto sphere at different viewing angles.
+  The larger the difference in viewing angles is, the larger the difference in tangential 
+  velocities will be.
+  Due to this projection effect, a population of genuine co-moving pairs
+  will extend to larger $\Delta v_t$ at larger separation.
+  This indeed can be seen in Figure~\ref{fig:dv-sep} as an over-density
+  in lower right corner that gets thinner as $\Delta v_t$ increases.
+
+  \item Finally, there is a population of “random” pairs of field stars
+  which are not actually co-moving, but still has $\Delta v_t < 10$~\kms\
+  by chance.
+  As $\Delta v_t$ increases, this population will dominate.
+  However, it is clear from this Figure that there is an overlap between
+  genuine co-moving pairs and “random” pairs.
+\end{itemize}
+
+In the following section, we construct a statistical model that propagates
 the non-trivial uncertainties in the data to our beliefs about the likelihood
@@
 
-For each star with $\snr > 32$, we use the reported parallax and parallax
-error to compute a distance corrected for the Lutz-Kelker bias
-(\citealt{lutzkelker}):
-\begin{equation}
-  d_{\rm LK} = 1000 \, \left[\frac{\varpi}{2} \,
-    \left(1 + \sqrt{1 - \frac{16}{\snr^2}} \right) \right]^{-1} \, {\rm pc}
-\end{equation}
-We use this distance and the sky positions of each source to compute 3D
-Cartesian positions, $\boldsymbol{x}$, for all stars and find the nearest
-neighbor to each source in 3-space using the KDTree implementation in
-\project{scikit-learn} (\citealt{scikit-learn}). We remove duplicates from
-this sample and are left with 29870 pairs of stars. For all pairs, we then
-compute the tangential velocity difference
-\begin{equation}
-  |\Delta \vec{v}_{\rm t}| = d_{\rm LK} \,
-                            \sqrt{(\mu_{\alpha,1}-\mu_{\alpha,2})^2 +
-                                  (\mu_{\delta,1}-\mu_{\delta,2})^2}
-\end{equation}
-\figname~\ref{fig:dv-sep} shows the tangential velocity difference against the
-physical separation plotted for all pairs that pass the above signal-to-noise
-cut. Pairs with small physical separation also tend to have small velocity
-differences. Intriguingly, there appear to be an abundance of pairs with small
-velocity separation at separations $|\delta\vec{x}| \gtrsim 1~{\rm pc}$. Of
-course, this figure really only highlights close pairs with small angular
-separations:
-At large angular separations, the tangential velocity differences will be large
-because of projection effects.
-To select a more representative population of co-moving stars, we must instead
-model the 3-space velocities of the stars given observations of just two
-components.
-
 \begin{figure*}[p]
   \begin{center}
-    \includegraphics[width=\textwidth]{figures/dv-sep.pdf}
+    \includegraphics[width=\textwidth]{figures/sep_dvtan.pdf}
   \end{center}
   \caption{%
-    Difference in tangential velocity vs. physical separation for pairs of stars
-    selected to be nearest neighbors in separation. Dashed line shows zero velocity
-    difference. Note the points with small velocity difference: these are likely
-    widely-separated binary stars.
+    Difference in tangential velocity vs. physical separation for candidate
+    pairs of stars with separation $< 10$~pc and $|\Delta v_t| < 10$~\kms.
+    % Note the points with small velocity difference: these are likely
+    % widely-separated binary stars.
     \label{fig:dv-sep}}
@@
 
-The abundance of pairs of stars with small velocity difference in the
-high-precision sample of \figname~\ref{fig:dv-sep} suggests that there are a
-significant number of co-moving stars in the \tgas\ data at separations as large
-as a few parsecs. Here we develop a method to select high-confidence co-moving
+The abundance of pairs of stars with small velocity difference in
+\figname~\ref{fig:dv-sep} suggests that there are a
+significant number of co-moving stars in the \tgas\ data at a range
+of separations.
+Here we develop a method to select high-confidence co-moving
 stars that properly incorporates the complex uncertainties associated with the
@@
 
-\section{Results 1: A catalog of candidate wide binaries}
+\section{Results}
+
+This section is divided into three parts. First, we discuss and justify a cut of
+the likelihood ratio to select candidate co-moving pairs.
+Second, we present the statistics and properties of co-moving pairs, and discuss its nature.
+Finally, we describe our catalog.
+
+\subsection{Selecting candidate co-moving pairs}
+
 In this section, we examine the distribution of likelihood ratios
@@
 for this quantities to select co-moving pairs from candidate pairs list.
-We present the statistics of the resulting co-moving pairs.
-We also match the pairs with known open clusters in the Milky Way to
-provide information on cluster membership.
+
+With these remarks in mind, we discuss the likelihood ratio distribution.
 
@@
 pairs sensitive to the cut.
+
+\subsection{Statistics, properties, and nature of co-moving pairs}
+
+We present the statistics of the resulting co-moving pairs.
+We also match the pairs with known open clusters in the Milky Way to
+provide information on cluster membership.
 
@@
   \begin{center}
-    \includegraphics[width=\textwidth]{figures/hist_Nneighbors.pdf}
+    \includegraphics[width=\textwidth]{figures/dist_networksize.pdf}
   \end{center}
   \caption{%
-    Histogram of the number of co-moving neighbors to a star for varying
-    degrees of contamination due to random pairs.
-    \todo{SMOH: What is the optimal bins for this histogram to separate
-    small number cases e.g., binary case (N=1) appropriately?}
+    Histogram of co-moving network sizes.
     \label{fig:hist_Nneighbors}}
@@
   \begin{center}
-    \includegraphics[width=\textwidth]{figures/glon_d_pairlines.pdf}
-    \includegraphics[width=\textwidth]{figures/glon_d_pairlines_mutexc.pdf}
+    \includegraphics[width=\textwidth]{figures/glon_d_pie.pdf}
   \end{center}
   \caption{%
-    Top: Distribution of co-moving pairs in galactic longitude and
-    distance. Each co-moving pair is connected by a line.
-    Red circles mark the locations of 135 known Milky Way open clusters within
-    $d<600$~pc from \citealt{Kharchenko:2016aa} while
-    stars in known OB associations \citep{de-Zeeuw:1999aa} are colored
-    by which association they belong.
-    We see that some of the aggregates of co-moving pairs are clearly
-    associated with known open clusters, and OB associations.
-    Bottom: Same as the previous plot but only for mutually exclusive pairs.
-    The mutually exclusive pairs no longer traces any known co-moving structures
-    strongly.
+    Panoramic view of co-moving pairs of stars. The angle is the Galactic
+    longitude, and the distance is in pc.
+    Pair networks of size 2 and 3 are connected by gray and blue lines.
+    Pair networks of size 4 and larger are plotted with a unique color.
+    Black circles mark the positions of known Milky Way star clusters
+    from \citet{Kharchenko:2016aa}.
     \label{fig:glon_d_pairlines}}
@@
 
+% \begin{figure*}[p]
+%   \begin{center}
+%     \includegraphics[width=\textwidth]{figures/glon_d_pairlines.pdf}
+%     \includegraphics[width=\textwidth]{figures/glon_d_pairlines_mutexc.pdf}
+%   \end{center}
+%   \caption{%
+%     Top: Distribution of co-moving pairs in galactic longitude and
+%     distance. Each co-moving pair is connected by a line.
+%     Red circles mark the locations of 135 known Milky Way open clusters within
+%     $d<600$~pc from \citealt{Kharchenko:2016aa} while
+%     stars in known OB associations \citep{de-Zeeuw:1999aa} are colored
+%     by which association they belong.
+%     We see that some of the aggregates of co-moving pairs are clearly
+%     associated with known open clusters, and OB associations.
+%     Bottom: Same as the previous plot but only for mutually exclusive pairs.
+%     The mutually exclusive pairs no longer traces any known co-moving structures
+%     strongly.
+%     \label{fig:glon_d_pie}}
+% \end{figure*}
+
 \todo{Galactic context / orbits}
 
-\section{Results 2: Star clusters and cluster remnants}
+\subsection{Catalog of candidate co-moving pairs}
 

### 2016-11-03 paper/ms.tex
worked on method and sec 4.1

@@
     pairs of stars with separation $< 10$~pc and $|\Delta v_t| < 10$~\kms.
-    % Note the points with small velocity difference: these are likely
-    % widely-separated binary stars.
     \label{fig:dv-sep}}
@@
     \todo{SMOH: mention GAIA paper that says Gaussian is good approximation?}
+    % http://adsabs.harvard.edu/abs/2012A&A...538A..78L
   \item We assume that the 3-space velocities of a given pair of stars
@@
 in both cases can be performed analytically:
-\todo{SMOH: is this what we do?}
-We use an equal-weight mixture of three isotropic, zero-mean Gaussian
-distributions
+We use a mixture of three isotropic, zero-mean Gaussian distributions
 \begin{equation}
-  p(\vec{v}) = \frac{1}{3} \, \sum_{m=1}^3 \, \mathcal{N}(0, \sigma_{v,m}^2)
+  p(\vec{v}) = \sum_{m=1}^3 \, w_m \mathcal{N}(0, \sigma_{v,m}^2)
 \end{equation}
-with velocity dispersions $(\sigma_{v,1}, \sigma_{v,2}, \sigma_{v,3}) = (8, 32,
-128)~\kms$ meant to represent young disk stars, old disk stars, and halo stars.
+with velocity dispersions $(\sigma_{v,1}, \sigma_{v,2}, \sigma_{v,3}) = (15, 30, 50)
+~\kms$ and weights $(w_1,w_2,w_3) = (0.3, 0.55, 0.15)$
+meant to encompass young thin disk stars to halo stars.
 We derive the relevant expressions in Appendix~\ref{sec:appendix}.
@@
 estimating the above integrals for stars with a wide range in parallax
-signal-to-noise.
+signal-to-noise. \todo{(SMOH: still unsure about this)}
+
+\todo{discuss vorbit}
 
@@
 $\ln \mathcal{L}_1 /\mathcal{L}_2$, and come up with a reasonable cut
-for this quantities to select co-moving pairs from candidate pairs list.
+for this quantity to select co-moving pairs from candidate pairs list.
 
-With these remarks in mind, we discuss the likelihood ratio distribution.
+\begin{figure*}[p]
+  \begin{center}
+    \includegraphics[width=\textwidth]{figures/likelihoodratios.pdf}
+  \end{center}
+  \caption{%
+    Left: Density histogram of likelihood ratios for candidate pairs and random pairs.
+    Along with the distribution for the entire candidate pairs (black thick line),
+    we show the distribution for subsets in slices of $\Delta v_t$.
+    Right: $\Delta v_t$ vs separation in slices of $\ln \mathcal{L}_1 /\mathcal{L}_2$.
+    \label{fig:likelihoodratios}}
+\end{figure*}
 
-Figure~\ref{fig:likelihoodratios} shows the likelihood ratios for all 2.6 million
-candidate pairs.
-Because we do not carry over the normalizations of all probabilities involved
-in $L_1$ and $L_2$, the absolute values of the likelihood ratio is not
-a meaningful quantity.
-Given our crude way of coming up with candidate pair list, we expect
-a huge number of pairs that are not genuinely co-moving, compared to
-the smaller number of pairs that are.
-In order to gauge an appropriate cut on the likelihood ratios,
+Figure~\ref{fig:likelihoodratios} shows the likelihood ratios for all
+271k candidate pairs examined.
+As discussed in \S~\ref{sec:data}, we expect a correlation between likelihood
+ratios of pairs, and their distribution in $\Delta v_t$ vs separation plane.
+Specifically, as we sweep through from small $\Delta v_t$ to large, we expect
+the population of pairs to change from genuinely co-moving to random.
+This becomes clear when we look at the distribution of likelihood ratios for
+slices of $\Delta v_t$ at 0.5, 5, and 9.5~\kms.
+Pairs with $|\Delta v_t-0.5|<0.5$ are most likely actual co-moving pairs, and
+their likelihood distribution is narrowly peaked above 6.
+The distribution peaks at lower values and gets broader as $\Delta v_t$ increases,
+and the number of random pairs increasingly dominate.
+On the right panel of Figure~\ref{fig:likelihoodratios}, we show
+how the distribution of pairs on $\Delta v_t$ vs separation plane changes
+with high to low $\ln \mathcal{L}_1 /\mathcal{L}_2$ ratios.
+This is in agreement with our discussion in \S~\ref{sec:data}.
+Finally, as a test,
 we compute the likelihood ratios for 200,000 random pairs of stars with the same
-signal-to-noise ratio cut as candidate pairs.
-Additionally, we select from candidate pair list, those with very small
-difference in tangential velocities ($\Delta v_t < 0.5$~km/s) in
-two minimum signal-to-noise ratio\footnote{lower of the two stars in a pair}
-bins (S/N$>32$ and S/N$<32$) that are highly probable
-to be genuine co-moving pairs.
-As expected, we see that the likelihood ratio $\ln \mathcal{L}_1 /\mathcal{L}_2$
-distribution peaks at higher values, favoring hypothesis 1, for highly probable
-co-moving pairs than the random pairs.
-We also see that the values are even higher for higher minimum signal-to-noise ratio pairs.
+parallax signal-to-noise ratio cut as candidate pairs ($S/N > 8$).
+Shown as gray shaded histogram, this distribution peaks at lower value ($\approx 2$),
+and is clearly separated from highly probable co-moving pairs.
 
-Based on the comparison with random pairs, we can come up with a cut
-on the likelihood ratios that reduces the cumulative contamination due to random pairs
-to a given threshold.
-The value of the cut, the resulting number of pairs, and
-its fraction of the total number of candidate pairs
-depending on the contamination rate are listed in
-Table~\ref{tab:candidatecounts}.
-Because the likelihood ratio $\ln \mathcal{L}_1 /\mathcal{L}_2$ asymtotically
-is chi-square distribution, the power-law fall off makes the number of resulting
-pairs sensitive to the cut.
+Based on these comparisons, we select pairs with
+$\ln \mathcal{L}_1 /\mathcal{L}_2 > 6$ as co-moving pair candidates.
+Out of 271,232 candidate pairs examined, 13,058 pairs (4.8\%)
+satisfy this condition.
 
@@
 Table~\ref{tab:candidatecounts}.
-
-\begin{table}[tbh]
-  \caption{Number of pairs}
-  \label{tab:candidatecounts}
-  \centering
-
-  \begin{tabular}{l|c|c|c|r}
-  \hline
-
-  \hline
-  \textbf{Contamination} & \textbf{$\ln \mathcal{L}_1 /\mathcal{L}_2$ cut} &
-    \textbf{Count} & \textbf{Fraction} & \textbf{$N_\textrm{me}$} \\
-  \hline
-    0.01    \% & 5.33 & 92458 & 0.462 \% & 5055 \\
-    0.001   \% & 6.53 & 13354 & 0.067 \% & 1549 \\
-    0.0001  \% & 7.29 &  2832 & 0.014 \% & 582  \\
-    0.00001 \% & 7.35 &  2495 & 0.012 \% & 532  \\
-  \hline
-  \end{tabular}
-\end{table}
-
-\begin{figure*}[p]
-  \begin{center}
-    \includegraphics[width=\textwidth]{figures/likelihoodratios.pdf}
-  \end{center}
-  \caption{%
-    Density histogram of likelihood ratios for candidate pairs and random pairs.
-    We also show the distribution for highly probable genuine co-moving pairs with
-    $\Delta v_t < 0.5$~km/s in high ($S/N>32$) and low ($S/N<32$) signal-to-noise ratio
-    bins. In both $S/N$ bins, the highly probable co-moving pairs are clearly
-    separated from random pairs while the distribution of $\ln \mathcal{L}_1 /\mathcal{L}_2$
-    for candidate pairs is expected to be a combination of
-    a huge number of ``random'' pairs that are not actually co-moving, and
-    a smaller number of genuine co-moving pairs with varying $S/N$.
-    \label{fig:likelihoodratios}}
-\end{figure*}
 

### 2016-11-09 paper/ms.tex
work on finishing results; filled in intro/conclusion

@@
 \newcommand{\apogee}{\project{\acronym{APOGEE}}}
+\newcommand{\tmass}{\project{\acronym{2MASS}}}
 \newcommand{\documentname}{\textsl{Article}}
@@
 We find pairs of co-moving stars out to very wide
-separations, including separations of a few parsecs! There does not
-seem to be any tidal-disruption feature in the separation distribution
-at sub-parsec scales. Pairs at separations beyond a
-parsec---whether they are bound in a multiple system or drifting
-apart---must be very short-lived. The photometric properties of the
-members of these pairs are consistent with youth. The prospects for
-testing stellar models and chemical abundance measurements are
-superficially discussed.
+separations up to 10~pc, the limit of our search.
+Some of these pairs form larger networks through mutual co-moving neighbors,
+many of which correspond to the known open clusters, and stars in the OB associations.
+However, there still is a huge number of very wide ($>3$~pc) separation pairs of
+stars that are co-moving.
+We discuss their nature, and the prospects for testing stellar models and chemical
+abundance measurements using our catalog.
+
+% HOGG:
+% There does not seem to be any tidal-disruption feature in the separation distribution
+% at sub-parsec scales. Pairs at separations beyond a
+% parsec---whether they are bound in a multiple system or drifting
+% apart---must be very short-lived. The photometric properties of the
+% members of these pairs are consistent with youth. The prospects for
+% testing stellar models and chemical abundance measurements are
+% superficially discussed.
 \end{abstract}
@@
 
-Two stars moving with similar full-space velocity vectors (``co-moving stars'')
-are either members of a widely-separated stellar multiplet, or remnants of a
-dispersing stellar system.
+Stars that are roughly co-located, and moving with similar space velocity (``co-moving stars'')
+are of special interest in many branches of astrophysics.
 
-Widely-separated binary stars are weakly bound binary star systems with
-semi-major axes larger than $a \gtrsim 10^{-3}~{\rm pc}$. It is generally
-assumed that the stars in a wide binary form from gas with the same chemical
-composition and are separated enough that there is no interaction or minimal
-cross-pollution as the more massive star evolves (\citealt{todo}). Because the
-binding energy of these systems is so low, if wide binary stars survive for many
-Gigayears their existence alone places limits on the tidal field of the Galaxy
-(\citealt{todo}) and the graininess of the gravitational field (\citealt{todo}).
-In this picture, a remaining puzzle is how these systems survive the dense
-stellar environments of the birth clusters in which they form. Wide binaries are
-therefore important tools for studying small-scale star formation and the
-large-scale Galactic mass distribution. These stars will appear to be co-moving
-with velocity differences comparable to the binary orbital velocity at the
-given separation.
+At small scales ($0.001-1$~pc), they correspond to wide binaries (and triples, or
+multiples) that are
+either weakly gravitationally bound or currently being dispersed.
+Because they have low binding energies,
+a sample of wide binaries is valuable for investigating
+the Galactic dynamical environment.
+These systems must have survived their dynamic birth environment which
+provides a window to the star formation process.
+They may also be disrupted by Galactic tides, or other massive objects
+such as other stars, molecular clouds, or MACHOs \citep{Yoo:2004aa,Allen:2014aa}.
+Wide binaries are also good testbeds for stellar evolution models.
+For example, main sequence-red giant wide binaries are especially valuable
+in testing consistency between stellar atmosphere models of these two types
+as the two stars were born at the same time with the same chemical compositions,
+but remained isolated since due to wide separation.
+Finally, calibration of stellar parameters of low-mass stars (e.g., M dwarfs),
+which dominate the stellar content of the Galaxy by number, can benefit from
+a larger sample of widely separated binaries containing a low mass star and a
+much brighter FGK star whose stellar parameters are easier to measure
+\citep[e.g.,][]{Rojas-Ayala:2012aa}.
 
-At the largest separations, $a \gtrsim 1~{\rm pc}$, co-moving stars are likely
-either remnants of disrupting stellar systems or young wide binary stars: at
-these extreme separations, the Galactic tidal field would disrupt binary stars
-within $t_{\rm disrupt} \lesssim XX~{\rm Gyr}$. Unbound open clusters with
-velocity dispersions $\sigma_v \lesssim 1~\kms$ in the Galactic disk disrupt
-quickly leaving small streams of stars with the same chemical abundances
-(\citealt{chemicaltaggingstuff}). Disrupting or disrupted globular clusters and
-dwarf galaxies also produce streams of stars with similar chemical properties
-but with larger velocity differences (\citealt{streampeeps}). Stars from such
-systems at similar orbital phases will appear to be co-moving with velocity
-differences comparable to the internal velocity dispersions, $\sigma_v \approx
-1$--$10~\kms$.
+At larger scales ($\gtrsim 1$~pc), they include moving groups, associations, and star clusters.
+The origin of moving groups is still under active debate: are they remnants of a coeval
+star formation event with similar chemical composition? Or are they formed by
+dynamical effects of non-axisymmetric features of the Galaxy such as spirals
+and bars? With the recent improvements in measuring chemical abundances
+using high and low resolution spectroscopy, we can now start to explore these questions
+in unprecedented width (statistics) and depth (detail). 
+The dynamics of cluster dissolution provides important clues to understanding
+the star formation history and the dynamical evolution of the Milky Way.
+In the halo, we have known the presence of disrupting globular clusters and dwarf galaxies
+\citep[e.g.,][]{Ibata:1994aa}.
+The tidal streams formed during this process are modeled to infer the
+parameters of the Galactic potential at large scales.
+Similar process is at work with star clusters in the disk. However, the dynamical
+time is much shorter, and the dynamics will be much more complex due to other
+perturbers in the disk.
 
 To date, thousands of candidate co-moving star pairs have been identified by
-searching for stars with common proper motions (\citealt{Luyten:1979,
-Poveda:1994, Allen:2000, Gould:2003, Chaname:2004, Lepine:2007,
-Alonso-Floriano:2015}).
-\todo{how is our search different?
-(1) we have distance, (2) bigger volume than Hipparcos, (3) we use the data properly, (4) we don't cut on distance or on magnitude of proper motion -- we cut on parallax S/N}
-% Cite:
-% Tokovinin & Lepine 2012
-% https://ui.adsabs.harvard.edu/#abs/2012ApJ...757..170A/abstract
-% https://ui.adsabs.harvard.edu/#abs/2012AJ....144..102T/abstract
-% https://ui.adsabs.harvard.edu/#abs/2014ApJ...790..158A/abstract
+searching for stars with common proper motions
+(\citealt{Poveda:1994aa,Allen:2000aa,Chaname:2004aa,Lepine:2007aa,Alonso-Floriano:2015aa}).
+Here, we use the recent first data release of \gaia\ which includes precise distances,
+enabling us to ask whether two stars share the same {\it physical} velocity rather than
+just the projections in the proper motion space.
+
+%TODO: highlight the merits of this search:
+% - we have distance,
+% - bigger volume than Hipparcos,
+% - we use the data properly, 
+% - we don't cut on distance or on magnitude of proper motion -- we cut on parallax S/N}
 
 % Many searches find things close in angular separation, but need pms / distance
-
-% Two stars moving with full-space velocity vector differences $\lesssim 1~\kms$
-% with orbital semi-major axes $a \gtrsim 10^-3~{\rm pc}$.
-
-Useful for:
-- star cluster disruption and dispersal
-- chemical abundance comparison (no interaction)
-- dynamics
-
-In either case, co-moving stars are of great interest
-for studying the destruction of star clusters and dwarf galaxies, the tidal
-field and small-scale properties of the mass distribution in the Milky Way, and
-chemical uniformity and star formation processes.
 
@@
 Solution (\tgas), released as a part of Data Release 1 (\dr) of the Gaia mission
-\citep{2016arXiv160904172G,2016arXiv160904303L}.
+\citep{Gaia-Collaboration:2016aa,Lindegren:2016aa}.
 \tgas\ contains astrometric measurements (sky position,
@@
 difference in tangential velocity $\Delta v_t < 10$~\kms\ within 10 pc
-using RA, DEC, a point-estimate of distance, and the proper motions.
+using RA, DEC, a point-estimate of the distance, and the proper motions.
 Throughout this \documentname\, we use the following estimator for
-the point-estimate of distance applying a correction for the Lutz-Kelker bias
-(\citealt{lutzkelker}):
+the point-estimate of the distance applying a correction for the Lutz-Kelker bias
+(\citealt{Lutz:1973aa}):
 \begin{equation}
@@
   will change from binaries to those related to these larger objects,
-  which will tend to subtend a larger angle in sky.
-  Under the assumption of having the same 3D velocity,
-  two stars of a co-moving pair can essentially be thought of as projections
-  of this velocity onto sphere at different viewing angles.
-  The larger the difference in viewing angles is, the larger the difference in tangential 
+  which generally subtend a larger angle in sky.
+  Since the proper motions of two stars with same 3D velocity
+  are essentially projections of this velocity onto sphere at
+  two different viewing angles,
+  the larger the difference in viewing angles is, the larger the difference in tangential 
   velocities will be.
@@
 
-\begin{figure*}[p]
+\begin{figure*}[htbp]
   \begin{center}
@@
     \mu_\delta\end{array})^\mathsf{T}$---are Gaussian with known covariances
-    $\mat{C}$.
-    \todo{SMOH: mention GAIA paper that says Gaussian is good approximation?}
-    % http://adsabs.harvard.edu/abs/2012A&A...538A..78L
+    $\mat{C}$. The values of covariances are provided by the \gaia\ data
+    \citep{Lindegren:2012aa,Lindegren:2016aa}.
   \item We assume that the 3-space velocities of a given pair of stars
@@
 zero-mean Gaussian with variance $s^2$ (hypothesis 1, $\mathcal{L}_1$), and
-the FML of the stars having different 3-space velocities (hypothesis 2,
+the FML of the stars having independent 3-space velocities (hypothesis 2,
 $\mathcal{L}_2$).
 We use the FML ratio $\mathcal{L}_1/\mathcal{L}_2$ as a scalar for selecting
-candidate co-moving stars, as described below in more detail.
+candidate co-moving stars, as described in $\S~\ref{sub:selection}$ in more detail.
 To compute these FMLs, the likelihood functions $L_i, L_j$ are marginalized
@@
 estimating the above integrals for stars with a wide range in parallax
-signal-to-noise. \todo{(SMOH: still unsure about this)}
+signal-to-noise.
+%TODO: SMOH: still unsure about this
 
-\todo{discuss vorbit}
+\begin{figure*}[htbp]
+  \begin{center}
+    \includegraphics[width=\textwidth]{figures/a_vorb.pdf}
+  \end{center}
+  \caption{%
+    Binary orbital velocity as a function of semi-major axis for $2~M_\odot$
+    binary.
+    \label{fig:vorbit}}
+\end{figure*}
 
-\todo{APW/SMOH: in results, specify what we set $s^2$ to:}
-To search for wide binaries, we set $s^2 = \frac{2 \, G \, \msun}{|\vec{x}_i-\vec{x}_j|}$.
-To search for dissolving open clusters, we set $s^2 = 1~(\kms)^2$.
+For wide binaries, the assumption of same 3D velocity can break down
+for high precision proper motion measurements because of the orbital
+velocity (Figure~\ref{fig:vorbit}).
+To account for this, we set $s^2 = \frac{2 \, G \, \msun}{|\vec{x}_i-\vec{x}_j|}$.
 
@@
 Second, we present the statistics and properties of co-moving pairs, and discuss its nature.
-Finally, we describe our catalog.
+Finally, we describe our catalog of candidate co-moving pairs, the main product
+of this study.
 
 \subsection{Selecting candidate co-moving pairs}
+\label{sub:selection}
 
@@
 This becomes clear when we look at the distribution of likelihood ratios for
-slices of $\Delta v_t$ at 0.5, 5, and 9.5~\kms.
+slices of $\Delta v_t$ centered at 0.5, 5, and 9.5~\kms.
 Pairs with $|\Delta v_t-0.5|<0.5$ are most likely actual co-moving pairs, and
@@
 how the distribution of pairs on $\Delta v_t$ vs separation plane changes
-with high to low $\ln \mathcal{L}_1 /\mathcal{L}_2$ ratios.
+with decreasing $\ln \mathcal{L}_1 /\mathcal{L}_2$ ratios.
 This is in agreement with our discussion in \S~\ref{sec:data}.
@@
 parallax signal-to-noise ratio cut as candidate pairs ($S/N > 8$).
-Shown as gray shaded histogram, this distribution peaks at lower value ($\approx 2$),
+Shown as gray shaded histogram in Figure~\ref{fig:likelihoodratios},
+this distribution peaks at a much lower value ($\approx 2$),
 and is clearly separated from highly probable co-moving pairs.
@@
 
-We present the statistics of the resulting co-moving pairs.
-We also match the pairs with known open clusters in the Milky Way to
-provide information on cluster membership.
-
-Once we have decided the co-moving pairs from candidate pair list,
-we can count how many co-moving neighbors a star has.
-This is presented in Figure~\ref{fig:hist_Nneighbors}.
-It is clear from this figure that we have many aggregates of co-moving stars, most
-likely an open cluster or a moving group. We also see that there is still
-a large number of stars which only one or two co-moving neighbors.
-In order to select pairs that are not associated with any other pair,
-we separate a subset of co-moving pairs whose member is a mutually exclusive
-co-moving neighbor to the other member (``mutually exclusive pairs'').
-The number of mutually exclusive pairs is listed on the last column of
-Table~\ref{tab:candidatecounts}.
-
-We now examine the separation distribution of co-moving pairs in
-Figure~\ref{fig:hist_separation}.
-
-Finally, we show the distribution co-moving pairs in galactic longitude and
-distance in Figure~\ref{fig:glon_d_pairlines}.
-In the top panel, we show all co-moving pairs with contamination rate of $0.01~\%$
-(undersampled by a factor of 4 in order to unclutter the plot) as connected line segments.
-We find many aggregates of co-moving pairs, which coincides with the positions of
-known Milky Way star clusters \citep{Kharchenko:2016aa} indicated by red circles.
-Clumps of co-moving pairs at $l=300-360$~degrees and $d=100-200$~pc correspond
-to the location of OB associations
-Upper Scorpius, Upper Centaurus Lupus, and Lower Centaurus Crux
-\citep{de-Zeeuw:1999aa}.
-The strong correlation with known co-moving structures disappears when we only
-consider mutually exclusive pairs (bottom panel).
-
-\begin{figure*}[p]
+\begin{figure*}[htbp]
   \begin{center}
@@
 
-\begin{figure*}[p]
+\begin{figure*}[htbp]
   \begin{center}
-    \includegraphics[width=\textwidth]{figures/hist_separation.pdf}
-    \includegraphics[width=\textwidth]{figures/hist_separation_mutexc.pdf}
+    \includegraphics[width=\textwidth]{figures/graphviz_examples.pdf}
   \end{center}
   \caption{%
-    Top: Separation distribution of co-moving pairs of stars for varying degrees
-    of contamination due to random pairs. The distribution peaks around 10~pc,
-    which is similar to the typical tidal radius of clusters in the Milky Way
-    \citep{Kharchenko:2013aa}.
-    Bottom: Same as the previous plot but only for mutually exclusive pairs.
-    \label{fig:hist_separation}}
+    Visualizations of a few example networks of co-moving pairs of stars.
+    Each star (node) is marked as red circles, and a line (edge) is drawn
+    between two stars if they are co-moving by our selection criteria
+    (see \S~\ref{sub:selection}). On left, we show the largest network
+    found in this study corresponding to the Pleiades star cluster.
+    \label{fig:graphviz_examples}}
 \end{figure*}
 
-\begin{figure*}[p]
+\begin{figure*}[htbp]
   \begin{center}
@@
     longitude, and the distance is in pc.
-    Pair networks of size 2 and 3 are connected by gray and blue lines.
-    Pair networks of size 4 and larger are plotted with a unique color.
-    Black circles mark the positions of known Milky Way star clusters
-    from \citet{Kharchenko:2016aa}.
+    Pair networks of sizes less 6 are connected by gray lines.
+    Pair networks of size larger than 6 are plotted with a unique color per network.
+    We also show the positions of known Milky Way star clusters
+    from \citet{Kharchenko:2016aa} as empty purple circles,
+    and stars in OB associations from \citet{de-Zeeuw:1999aa} as
+    black circles.
     \label{fig:glon_d_pairlines}}
@@
 
-% \begin{figure*}[p]
-%   \begin{center}
-%     \includegraphics[width=\textwidth]{figures/glon_d_pairlines.pdf}
-%     \includegraphics[width=\textwidth]{figures/glon_d_pairlines_mutexc.pdf}
-%   \end{center}
-%   \caption{%
-%     Top: Distribution of co-moving pairs in galactic longitude and
-%     distance. Each co-moving pair is connected by a line.
-%     Red circles mark the locations of 135 known Milky Way open clusters within
-%     $d<600$~pc from \citealt{Kharchenko:2016aa} while
-%     stars in known OB associations \citep{de-Zeeuw:1999aa} are colored
-%     by which association they belong.
-%     We see that some of the aggregates of co-moving pairs are clearly
-%     associated with known open clusters, and OB associations.
-%     Bottom: Same as the previous plot but only for mutually exclusive pairs.
-%     The mutually exclusive pairs no longer traces any known co-moving structures
-%     strongly.
-%     \label{fig:glon_d_pie}}
-% \end{figure*}
+Once we have decided the co-moving pairs from the candidate pair list,
+these pairs form an undirected graph where stars are nodes, and
+edges between the nodes exist for co-moving pairs of stars.
+A star may have multiple co-moving neighbors, and two stars may either
+be directly or indirectly connected by a set of edges.
+We divide the network into connected components\footnote{A connected component of an undirected graph $G$
+is a subgraph of $G$ in which any two nodes are connected to each other by a path.
+We use the term ``network'' loosely to refer to a graph or its connected components.},
+and show the distribution of their sizes in Figure~\ref{fig:hist_Nneighbors}.
+The most common are size 2 networks, which mean mutually exclusive co-moving
+pairs. However, it is clear that there are many aggregates of co-moving stars
+discovered by looking for co-moving pairs.
+These aggregates are likely moving groups, OB associations or star clusters.
+There are 4555 connected components among 10,606 unique stars.
+The maximum size of connected components is 151.
+We show this largest network along with four randomly selected
+examples of networks with sizes greater than 4 in Figure~\ref{fig:graphviz_examples}.
+The largest network corresponds to the Pleiades open cluster.
+One particular star in this network, TYC 1799-272-1, has 94 directly connected
+co-moving neighbors, and is located at the center of the network.
 
-\todo{Galactic context / orbits}
+We show the distribution co-moving pairs in galactic longitude and
+distance in Figure~\ref{fig:glon_d_pairlines}.
+The connection between known co-moving structures and the networks of co-moving
+pairs found in this work becomes immediately clear when
+we overplot the positions of known Milky Way star clusters \citep{Kharchenko:2016aa},
+and stars in OB associations \citep{de-Zeeuw:1999aa}.
+Many of the larger networks are clearly associated with known star clusters.
+The most prominent is Melotte 111 at
+$(l,d)\approx (220\arcdeg, 87~\textrm{pc})$ otherwise known as the Coma star cluster.
+Clumps of co-moving pairs at $(l,d)\approx (300-360\arcdeg, 100-200~\textrm{pc})$
+seem to strongly correlate with the locations of OB associations
+Upper Scorpius, Upper Centaurus Lupus, and Lower Centaurus Crux
+\citep{de-Zeeuw:1999aa}.
+
+Ultimately, any candidate co-moving pair found in this work
+needs to be verified using radial velocities.
+Here, we use 210,368 cross-matches of \tgas\ with
+the Radial Velocity Experiment (\rave) to
+assess the false positive rates of our selection.
+We have 283 pairs with both stars matched with \rave.
+Figure~\ref{fig:raverv} shows the difference in radial velocities
+between the two stars in a pair, $\Delta v_r$, as a function of their physical separation.
+We show $\Delta v_r$ in units of $\sigma_{\Delta v_r}$ which we estimate
+as the quadrature sum of $\sigma _{v_r}$ for each star.
+The fraction of pairs with good agreement in radial velocity decreases with
+increasing separation.
+This, after all, is not surprising because we are only using
+2D velocity information (proper motions) with errors.
+However, the contamination becomes significant only at $>1$~pc
+(which depends on the local stellar number density and velocity dispersion).
+Given the excellent correspondence between co-moving pair networks of larger sizes
+and known genuine co-moving structures (Figure~\ref{fig:glon_d_pairlines}),
+we may expect that pairs in these larger networks which will often have
+separations $>1$~pc to have less contamination.
+We divide pairs into those mutually exclusively connected (i.e., in a network of
+size 2), and those in a larger network.
+We indeed find that many pairs in larger networks are at $>1$~pc, yet
+the fraction of pairs that have identical radial velocities within $3\sigma$ is
+higher than mutually exclusive pairs, and remains high ($>80\%$) to $\sim 10$~pc.
+
+\begin{figure*}[htbp]
+  \begin{center}
+    \includegraphics[width=\textwidth]{figures/raverv.pdf}
+  \end{center}
+  \caption{%
+    Radial velocity differences of 283 pairs with \rave\ measurements.
+    Left: $\Delta v_r$ in units of $\sigma_{\Delta v_r} = \sqrt{\sigma_{v_r,1}^2 + \sigma_{v_r,2}^2}$
+    as a function of physical separation.
+    We highlight $3\sigma$ limit in gray shades.
+    Right: Fraction of pairs with $\Delta v_r/\sigma_{\Delta v_r} <3$
+    as a function of physical separation. We calculate the fraction with a running
+    bin containing 31 data points at a time.
+    The median, and the minimum and maximum separation of pairs in each bin are
+    indicated with a marker, and its errorbars.
+    We separate pairs with network size of 2 (i.e., mutually exclusively connected)
+    from those in larger networks.
+    \label{fig:raverv}}
+\end{figure*}
+
+We now examine the separation distribution of co-moving pairs in
+Figure~\ref{fig:hist_separation}.
+As expected, pairs in larger networks are mostly found with separations
+larger than 1~pc.
+Surprisingly, however, we also find a huge number of mutually exclusive pairs
+at $>1$~pc as well. Even if we consider the increasing false positive rate at large
+separations, the distribution is not significantly changed as the number of pairs
+at $>1$~pc is in fact increasing much faster (as a power-law)
+than the decrease due to the false positives (right panel of Figure~\ref{fig:raverv}).
+The nature of these very large separation, mutually exclusive pairs,
+which cannot be gravitationally bound to each other, needs further investigation.
+Can they be remnants of escaped binaries that are drifting apart?
+In a study of the evolution of wide binaries including the Galactic tidal field
+as well as passing field stars, \citet{Jiang:2010aa} found that we expect 
+to find a peak at $\sim 100-300$~pc in the projected separation due to
+stars that were once in a wide binary system, but are drifting apart with small
+relative velocities ($\sim 0.1$~\kms).
+Our search which stops on the maximum radius of 10~pc is obviously insufficient
+to clearly detect this peak, but it shows such hint, and demonstrates
+that such large scale phase-space correlations may now be looked for and studied within the \gaia\ data.
+
+\begin{figure*}[htbp]
+  \begin{center}
+    \includegraphics[width=\textwidth]{figures/hist_sep.pdf}
+  \end{center}
+  \caption{%
+    Separation distribution of co-moving pairs of stars. 
+    We divide the pairs into those mutually exclusively connected
+    (network size $=2$), and those connected to larger networks (size $>2$).
+    \label{fig:hist_separation}
+    }
+\end{figure*}
+
+Finally, we present the color-magnitude diagrams of co-moving pairs
+using the cross-matches with \tmass.
+A more detailed study of stellar parameters using photometry from various sources
+will follow. Figure~\ref{fig:cmd_large} and \ref{fig:cmd_me} shows $G-J$ vs $G$
+color-magnitude diagrams for stars in larger networks (size$>2$) and in mutually
+exclusive pairs (networks of size 2), respectively.
+The networks shown in Figure~\ref{fig:cmd_large} correspond to those visualized
+in Figure~\ref{fig:graphviz_examples}.
+For stars in larger networks, there is a noticeable lack of evolved, off-main sequence
+stars, in agreement with these kinematic structures being young.
+For mutually exclusive pairs, we divide the pairs by separation at 1~pc
+above which the false positive rate due to random pairs starts to increase.
+While many pairs are located along the main sequence, we also find
+quite a few of main sequence-red giant pairs,
+which will be valuable to anchoring stellar atmospheric models together.
+%TODO: SMOH: what else can i discuss?
+
+\begin{figure*}[htbp]
+  \begin{center}
+    \includegraphics[width=\textwidth]{figures/gjg_graphviz_examples.png}
+  \end{center}
+  \caption{
+    Color-magnitude diagrams of stars in larger networks presented in
+    Figure~\ref{fig:graphviz_examples}. Each panel corresponds to the panel
+    at the same position in Figure~\ref{fig:graphviz_examples}.
+    We show the distribution of all stars common in \tgas\ and \tmass\
+    with parallax $S/N>16$ in blue for reference.
+    \label{fig:cmd_large}}
+\end{figure*}
+
+\begin{figure*}[htbp]
+  \begin{center}
+    \includegraphics[width=\textwidth]{figures/gjg_mepairs.png}
+  \end{center}
+  \caption{
+    Color-magnitude diagrams of stars in networks of size 2 (i.e., mutually exclusive)
+    in two separation bins.
+    We connect each pair by a line on the left for pairs with separations smaller than
+    1~pc.
+    We show the distribution of all stars common in \tgas\ and \tmass\
+    with parallax $S/N>16$ in blue for reference same as in Figure~\ref{fig:cmd_large}.
+    \label{fig:cmd_me}}
+\end{figure*}
 
@@
 
+In this section, we describe our catalog of candidate co-moving pairs of stars.
+For 13,058 pairs (10,606 unique stars) found in this work,
+we provide the following information:
+%
+\begin{itemize}
+  \item \tgas\ source id for the two stars in a pair; these may be used to easily
+  retrieve cross matches of \gaia\ with other surveys such as WISE, \tmass, and \rave
+  using the \gaia\ data archive.
+  \item physical separation between the two stars in pc
+  \item likelihood ratio, $\ln \mathcal{L}_1 /\mathcal{L}_2$
+  \item size of the network that the pair belongs to
+  \item unique ID of the network that the pair belongs to
+\end{itemize}
+%
+Additionally, given the strong correlation between the larger networks of pairs
+and the known star clusters and OB associations, we provide for each star
+%
+\begin{itemize}
+  \item index of the closest cluster in the Milky Way Star Cluster catalog of
+  \citet{Kharchenko:2016aa} using 3D separation
+  \item distance to the closest cluster in pc
+  \item the ID of the OB association (A-L) if the star is in the catalog of
+  stars in the known OB associations by \citet{de-Zeeuw:1999aa}
+\end{itemize}
+%
+A part of the entire table is presented in Table~\ref{tab:catalog}.
+
+\input{stub}
+
+We note a caveat on the completeness of the networks found in our catalog.
+Because we applied a simple cut in the likelihood ratio ($\ln \mathcal{L}_1 /\mathcal{L}_2>6$),
+there is a possibility that, for example, a star in a mutually exclusive pair in our catalog may
+still have another possibly co-moving companion which has been dropped because the likelihood ratio
+is slightly less than 6.
+
 \section{Conclusions}
 
+In this \documentname, we looked for co-moving pairs of stars in the
+\tgas\ data released as part of the \gaia\ DR1.
+Our method is to compare
+the fully marginalized likelihoods between the two hypotheses, that a pair of
+stars shares the same 3D velocity, and that the two stars have two independent
+3D velocities, incorporating the errors in parallax and proper motions.
+We argued for a reasonable cut of the likelihood ratio, and found
+13,058 candidate pairs of co-moving stars among 10,606 unique stars
+with separation ranging from 0.005~pc to 10~pc, the limit of our search.
+
+We examined the network of co-moving stars, and found that many of the larger
+networks of co-moving stars naturally corresponds to some of the known co-moving structures
+such as open clusters and stellar associations.
+We have also found a surprisingly large number of very wide separation ($>3$~pc)
+co-moving pairs which are not clearly associated to any known existing clusters.
+While their nature is yet unclear, they may be remnants of dissolving wide binaries and
+open clusters. If so, this population should still be relatively young compared
+to the general disk field population. Modeling the color-magnitude diagram
+distribution of these stars can shed some light on this issue.
+If they are truly dissolving stars that were born coeval,
+the sample of very wide separation co-moving pairs can potentially be used
+to measure the recent ($\lesssim 1$~Gyr)
+star formation history in the Solar neighborhood.
+Co-moving stars with separation less than 1~pc are very promising candidates
+for wide binaries. They are found to be pairs of stars of varying stellar types.
+Some of these pairs, such as main sequence-red giant or FGK-M dwarfs,
+will be particularly valuable for testing theoretical
+stellar models and calibrating observational measurements of low mass stars.
+
 \acknowledgements
+
+This work has made use of data from the European Space Agency (ESA)
+mission {\it Gaia} (\url{http://www.cosmos.esa.int/gaia}), processed by
+the {\it Gaia} Data Processing and Analysis Consortium (DPAC,
+\url{http://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding
+for the DPAC has been provided by national institutions, in particular
+the institutions participating in the {\it Gaia} Multilateral Agreement.
+This project was developed in part at the 2016 NYC Gaia Sprint,
+hosted by the Center for Computational Astrophysics at the Simons Foundation in New York City.
 

### 2016-11-09 paper/stub.tex
work on finishing results; filled in intro/conclusion

\begin{deluxetable*}{cccccc}
\tablecaption{Catalog of candidate co-moving pairs \label{tab:catalog}}
\tablehead{\colhead{star1 source id} & \colhead{star2 source id} & \colhead{separation} & \colhead{$\ln \mathcal{L}_1 /\mathcal{L}_2$} & \colhead{ID$_\mathrm{network}$} & \colhead{$N_\mathrm{network}$}\\ \colhead{ } & \colhead{ } & \colhead{$\mathrm{pc}$} & \colhead{ } & \colhead{ } & \colhead{ }}
\startdata
249282341403694592 & 441371523899475840 & 4.9 & 7.0 & 1 & 125 \\
249087281167662464 & 441901694662492928 & 8.3 & 7.1 & 1 & 125 \\
64933755122821120 & 66786500935624320 & 7.8 & 7.5 & 0 & 151 \\
436536249718223744 & 441356921010671232 & 6.5 & 6.5 & 1 & 125 \\
3953625835302703488 & 4008706729289355520 & 6.1 & 6.3 & 8 & 47 \\
63730305286697600 & 65188085906203520 & 7.0 & 6.6 & 0 & 151 \\
1873311936758998016 & 1873312074197947392 & 7.5 & 7.2 & 4013 & 2 \\
5814505765886192512 & 5814997419380722944 & 6.1 & 6.2 & 14 & 20 \\
350903157411208832 & 352510643410053632 & 9.9 & 6.8 & 565 & 2 \\
64114241002810496 & 67618281484716544 & 7.2 & 6.1 & 0 & 151
\enddata
\tablecomments{Table 1 is published in its entirety in the machine-readable format. A portion is shown here for guidance regarding its form and content.}
\end{deluxetable*}

### 2016-11-09 paper/ms.tex
tweak in figure captions

@@
     Visualizations of a few example networks of co-moving pairs of stars.
-    Each star (node) is marked as red circles, and a line (edge) is drawn
+    Each star (node) is marked as a red circle, and a line (edge) is drawn
     between two stars if they are co-moving by our selection criteria
@@
     longitude, and the distance is in pc.
-    Pair networks of sizes less 6 are connected by gray lines.
-    Pair networks of size larger than 6 are plotted with a unique color per network.
+    Pair networks of sizes less than 5 are connected by gray lines.
+    Pair networks of size $\geq 5$ are plotted with a unique color per network.
     We also show the positions of known Milky Way star clusters
@@
     Color-magnitude diagrams of stars in larger networks presented in
-    Figure~\ref{fig:graphviz_examples}. Each panel corresponds to the panel
-    at the same position in Figure~\ref{fig:graphviz_examples}.
+    Figure~\ref{fig:graphviz_examples}. Each panel shows the same network
+    visualized in the panel at the same position
+    in Figure~\ref{fig:graphviz_examples}.
     We show the distribution of all stars common in \tgas\ and \tmass\

### 2016-11-10 paper/ms.tex
Modify figure 1 and 3
- new macro for consistant dvtan symbol
- combine figure 1 and 2 and revmoe 2
- change "candidate pair list" to "initial sample of pairs"
- change "network" to "connected component"

@@
 \newcommand{\eye}{\mathbb{I}}
+\newcommand{\absdvtan}{\ensuremath{|\Delta\vec v_\mathrm{t}|}}
 
@@
 Some of these pairs form larger networks through mutual co-moving neighbors,
-many of which correspond to the known open clusters, and stars in the OB associations.
+many of which correspond to the known open clusters, and stars in OB associations.
 However, there still is a huge number of very wide ($>3$~pc) separation pairs of
@@
 abundance measurements using our catalog.
+
+% change to probabilistic statement "co-moving"
+% we present a table of co-moving pairs
+% add dynamical inference 
+% the last line = the most important thing
 
@@
 
-We construct a sample of candidate pair list as follows.
+We construct the initial sample of pairs as follows.
 We first apply a global parallax S/N cut of 8 to \tgas, which leaves 619,618 stars.
 Then, for each star, we search a potential co-moving partner star with a
-difference in tangential velocity $\Delta v_t < 10$~\kms\ within 10 pc
+difference in tangential velocity $\absdvtan < 10$~\kms\ within 10 pc
 using RA, DEC, a point-estimate of the distance, and the proper motions.
@@
 \begin{equation}
-  |\Delta \vec{v}_{\rm t}| = |d_1 \vec\mu_1 - d_2 \vec\mu_2|
+  \absdvtan = |d_1 \vec\mu_1 - d_2 \vec\mu_2|
 \end{equation}
@@
 
-Figure~\ref{fig:dv-sep} shows $|\Delta v_t|$ against the physical separation
-for the resulting 271,232 unique candidate pairs.
+Figure~\ref{fig:dv-sep} shows \absdvtan\ against the physical separation
+for the resulting 271,232 unique pairs in the initial sample.
 A few key observations can be made:
@@
   Due to this projection effect, a population of genuine co-moving pairs
-  will extend to larger $\Delta v_t$ at larger separation.
+  will extend to larger \absdvtan\ at larger separation.
   This indeed can be seen in Figure~\ref{fig:dv-sep} as an over-density
-  in lower right corner that gets thinner as $\Delta v_t$ increases.
+  in lower right corner that gets thinner as \absdvtan\ increases.
 
   \item Finally, there is a population of “random” pairs of field stars
-  which are not actually co-moving, but still has $\Delta v_t < 10$~\kms\
+  which are not actually co-moving, but still has $\absdvtan 10$~\kms\
   by chance.
-  As $\Delta v_t$ increases, this population will dominate.
+  As \absdvtan\ increases, this population will dominate.
   However, it is clear from this Figure that there is an overlap between
@@
   \caption{%
-    Difference in tangential velocity vs. physical separation for candidate
-    pairs of stars with separation $< 10$~pc and $|\Delta v_t| < 10$~\kms.
+    Initial sample of pairs on the
+    difference in tangential velocity vs. physical separation plane.
+    For the initial sample, we look for stars
+    with separation $< 10$~pc and $\absdvtan < 10$~\kms\ for
+    every star in \tgas.
+    The blue solid line shows the magnitude of the 3D
+    orbital velocity as a function of semi-major axis for a $2~M_\odot$ binary system.
+    Note the stream of points that starts at small separation ($\lesssim 0.01$~pc),
+    small $\absdvtan$ ($\lesssim 2$~\kms),
+    but climbs to larger $\absdvtan$ at $|\Delta \vec{x}|\gtrsim 1$~pc, which
+    eventually merges with random pairs of field stars dominating in the upper right corner
+    (see \S~\ref{sec:data} for details).
     \label{fig:dv-sep}}
@@
 
-\begin{figure*}[htbp]
-  \begin{center}
-    \includegraphics[width=\textwidth]{figures/a_vorb.pdf}
-  \end{center}
-  \caption{%
-    Binary orbital velocity as a function of semi-major axis for $2~M_\odot$
-    binary.
-    \label{fig:vorbit}}
-\end{figure*}
-
 For wide binaries, the assumption of same 3D velocity can break down
 for high precision proper motion measurements because of the orbital
-velocity (Figure~\ref{fig:vorbit}).
+velocity (blue solid line in Figure~\ref{fig:dv-sep}).
 To account for this, we set $s^2 = \frac{2 \, G \, \msun}{|\vec{x}_i-\vec{x}_j|}$.

### 2016-11-10 paper/ms.tex
Modify figure 4 graph visualization and caption

@@
   \caption{%
-    Histogram of co-moving network sizes.
-    \label{fig:hist_Nneighbors}}
+    Histogram of the sizes of connected components.
+    \label{fig:hist_ccsize}}
 \end{figure*}
@@
   \caption{%
-    Visualizations of a few example networks of co-moving pairs of stars.
+    Visualizations of a few example connected components of co-moving pairs of stars.
     Each star (node) is marked as a red circle, and a line (edge) is drawn
@@
     found in this study corresponding to the Pleiades star cluster.
+    On right, we show four examples of connected components with varying sizes.
+    The connected component on the upper left panel with a size of 58 corresponds
+    to NGC~2632, also known as the Beehive cluster.
     \label{fig:graphviz_examples}}
@@
 We use the term ``network'' loosely to refer to a graph or its connected components.},
-and show the distribution of their sizes in Figure~\ref{fig:hist_Nneighbors}.
+and show the distribution of their sizes in Figure~\ref{fig:hist_ccsize}.
 The most common are size 2 networks, which mean mutually exclusive co-moving

### 2016-11-10 paper/ms.tex
Modify raverv and hist_sep figures and captions

@@
     \left(1 + \sqrt{1 - \frac{16}{\snr^2}} \right) \right]^{-1} \, {\rm pc}
+    \label{eq:dist}
 \end{equation}
@@
   \caption{%
-    Panoramic view of co-moving pairs of stars. The angle is the Galactic
-    longitude, and the distance is in pc.
-    Pair networks of sizes less than 5 are connected by gray lines.
-    Pair networks of size $\geq 5$ are plotted with a unique color per network.
+    Panoramic view of co-moving pairs of stars. This is a cylindrical
+    projection of onto the Galactic plane. The angle is the Galactic
+    longitude, and the radius is $d\cos b$ in pc where $d$ is the point-estimate
+    of distance using Equation~\ref{eq:dist}.
+    Pairs in connected components of sizes less than 5 are connected by gray lines.
+    Pair in connected components of size $\geq 5$ are plotted with a unique color per network,
+    and their nodes (stars) are highlighted with small black circles.
     We also show the positions of known Milky Way star clusters
@@
     and stars in OB associations from \citet{de-Zeeuw:1999aa} as
-    black circles.
+    gray circles.
     \label{fig:glon_d_pairlines}}
@@
 
-\begin{figure*}[htbp]
+\begin{figure}[htbp]
   \begin{center}
-    \includegraphics[width=\textwidth]{figures/raverv.pdf}
+    % \includegraphics[width=\linewidth]{figures/raverv.pdf}
+    \includegraphics[width=.5\linewidth]{figures/raverv.pdf}
   \end{center}
   \caption{%
-    Radial velocity differences of 283 pairs with \rave\ measurements.
-    Left: $\Delta v_r$ in units of $\sigma_{\Delta v_r} = \sqrt{\sigma_{v_r,1}^2 + \sigma_{v_r,2}^2}$
+    Validation of candidate co-moving pairs using radial velocities from \rave.
+    Top: Radial velocity differences, $\Delta v_r$, of 283 pairs with \rave\ measurements.
+    $\Delta v_r$ is plotted in units of $\sigma_{\Delta v_r} = \sqrt{\sigma_{v_r,1}^2 + \sigma_{v_r,2}^2}$
     as a function of physical separation.
-    We highlight $3\sigma$ limit in gray shades.
-    Right: Fraction of pairs with $\Delta v_r/\sigma_{\Delta v_r} <3$
+    We highlight $3\sigma$ limit with two horizontal lines.
+    Bottom: Fraction of pairs with $\Delta v_r/\sigma_{\Delta v_r} <3$
     as a function of physical separation. We calculate the fraction with a running
@@
     indicated with a marker, and its errorbars.
-    We separate pairs with network size of 2 (i.e., mutually exclusively connected)
-    from those in larger networks.
+    We separate pairs in connected components of size 2 (i.e., mutually exclusively connected)
+    from those in larger connected components.
     \label{fig:raverv}}
-\end{figure*}
+\end{figure}
 
@@
     Separation distribution of co-moving pairs of stars.
-    We divide the pairs into those mutually exclusively connected
-    (network size $=2$), and those connected to larger networks (size $>2$).
+    As with Figure~\ref{fig:raverv},
+    we divide the pairs into those mutually exclusively connected
+    (connected component size $=2$), and those connected to larger connected components (size $>2$).
     \label{fig:hist_separation}

### 2016-11-16 paper/ms.tex
fixed CCA affiliation, added to Hogg. fixed #21
Modify color magnitude diagrams
- Fix y-axis label
- Output to rasterized pdf

@@
         Adrian M. Price-Whelan\altaffilmark{\pu},
-        David W. Hogg\altaffilmark{\ccpp,\mpia},
+        David W. Hogg\altaffilmark{\ccpp,\mpia,\cca},
         Timothy D. Morton\altaffilmark{\pu},
@@
                      K\"onigstuhl 17, D-69117 Heidelberg, Germany}
-\altaffiltext{\cca}{Simons Center for Computational Astrophysics, ...,
-                    New York, NY XXXXX, USA}
+\altaffiltext{\cca}{Center for Computational Astrophysics, 162 5th Ave, New York, NY 10003, USA}
 
@@
   \begin{center}
-    \includegraphics[width=\textwidth]{figures/gjg_graphviz_examples.png}
+    \includegraphics[width=\textwidth]{figures/gjg_graphviz_examples.pdf}
   \end{center}

### 2016-11-16 paper/ms.tex
some modifications

@@
 
-At small scales ($0.001-1$~pc), they correspond to wide binaries (and triples, or
-multiples) that are
-either weakly gravitationally bound or currently being dispersed.
-Because they have low binding energies,
-a sample of wide binaries is valuable for investigating
-the Galactic dynamical environment.
+At small separations ($0.001-1$~pc), they correspond to wide binaries (and
+multiples) that are either weakly gravitationally bound or currently being
+dispersed.
+Because they have low binding energies, a sample of wide binaries is valuable
+for investigating the Galactic dynamical environment.
 These systems must have survived their dynamic birth environment which
@@
 Wide binaries are also good testbeds for stellar evolution models.
-For example, main sequence-red giant wide binaries are especially valuable
-in testing consistency between stellar atmosphere models of these two types
-as the two stars were born at the same time with the same chemical compositions,
-but remained isolated since due to wide separation.
+For example, main sequence-red giant wide binaries are especially valuable in
+testing consistency between stellar atmosphere models of these two types as the
+two stars were likely born at the same time with the same chemical compositions,
+but evolve indepdently becayse if their wide separation.
 Finally, calibration of stellar parameters of low-mass stars (e.g., M dwarfs),
@@
 
-At larger scales ($\gtrsim 1$~pc), they include moving groups, associations, and star clusters.
+At larger separation ($\gtrsim 1$~pc), co-moving stars are likely members of
+moving groups, associations, and star clusters.
 The origin of moving groups is still under active debate: are they remnants of a coeval
@@
 using high and low resolution spectroscopy, we can now start to explore these questions
-in unprecedented width (statistics) and depth (detail).
+in unprecedented width (statistics) and depth (detail) \todo{What do you mean by this?}.
 The dynamics of cluster dissolution provides important clues to understanding
 the star formation history and the dynamical evolution of the Milky Way.
-In the halo, we have known the presence of disrupting globular clusters and dwarf galaxies
-\citep[e.g.,][]{Ibata:1994aa}.
-The tidal streams formed during this process are modeled to infer the
-parameters of the Galactic potential at large scales.
-Similar process is at work with star clusters in the disk. However, the dynamical
-time is much shorter, and the dynamics will be much more complex due to other
-perturbers in the disk.
+In the halo, we know of more than 20 disrupting globular clusters and dwarf galaxies
+(see, e.g., \citealt{Grillmair:2016} for a summary of known streams).
+These tidal streams are modeled to infer the parameters of the Galactic
+potential (e.g., \citealt{Kupper:2015}).
+Similar processes are at work with star clusters in the disk.
+However, the dynamical time is much shorter, and the dynamics will be much more
+complex because of the existence of other perturbers in the disk.
 
@@
 just the projections in the proper motion space.
+
+This paper proceeds as follows:
+In \sectionname~\ref{sec:data}, we briefly describe the data set (TGAS) used in
+this work.
+In \sectionname~\ref{sec:methods}, we develop a statistical method to identify
+high-confidence co-moving stars in this catalog.
+In \sectionname~\ref{sec:results}, we present and discuss our resulting catalog
+of co-moving stars.
+We conclude in \sectionname~\ref{sec:conclusions}.
 
@@
 
-We construct the initial sample of pairs as follows.
-We first apply a global parallax S/N cut of 8 to \tgas, which leaves 619,618 stars.
-Then, for each star, we search a potential co-moving partner star with a
-difference in tangential velocity $\absdvtan < 10$~\kms\ within 10 pc
-using RA, DEC, a point-estimate of the distance, and the proper motions.
-Throughout this \documentname\, we use the following estimator for
-the point-estimate of the distance applying a correction for the Lutz-Kelker bias
+We construct an initial sample of star pairs to search for co-moving pairs as
+follows.
+We first apply a global parallax signal-to-noise cut, $\snr < 8$,  to \tgas,
+which leaves 619,618 stars.
+For each surviving star, we then establish an initial sample of possible
+co-moving partners by selecting all other stars within 10 pc with a difference
+in tangential velocity below a threshold, $\absdvtan < 10$~\kms.
+We ultimately build a statistical model that incorporates the covariances of the
+data, but for this position cut and for visualizations we use a point-estimate
+of the distance by applying a correction for the Lutz-Kelker bias
 (\citealt{Lutz:1973aa}):
 \begin{equation}
-  d = 1000 \, \left[\frac{\varpi}{2} \,
+  \hat{d} = 1000 \, \left[\frac{\varpi}{2} \,
     \left(1 + \sqrt{1 - \frac{16}{\snr^2}} \right) \right]^{-1} \, {\rm pc}
@@
 where $\varpi$ is the parallax in mas.
-The difference in tangential velocity between two stars is, then,
+An estimate for the difference in tangential velocity between two stars is,
+then,
 \begin{equation}
-  \absdvtan = |d_1 \vec\mu_1 - d_2 \vec\mu_2|
+  \absdvtan = |\hat{d}_1 \vec\mu_1 - \hat{d}_2 \vec\mu_2|
 \end{equation}
@@
 \begin{itemize}
-  \item At small separation ($<1$~pc), there is a population of pairs with
+  \item At small separations ($<1$~pc), there is a population of pairs with
   very small tangential velocity difference ($<2$~km/s). Given that these pairs
-  are very close in both positions and proper motions,
-  it is highly probable that they are actually co-moving e.g., as wide binaries.
+  are very close in both position and proper motion,
+  it is highly probable that they are actually co-moving wide binaries.
 
-  \item  However, it is not only (wide) binaries which are gravitationally
-  bound that are expected to appear as co-moving.
+  \item  However, it is not only (wide) binaries, which are gravitationally
+  bound, that are expected to appear as co-moving.
   OB associations, moving groups, and open clusters are also of consideration.
@@
   which generally subtend a larger angle in sky.
-  Since the proper motions of two stars with same 3D velocity
+  Since the proper motions of two stars with the same 3D velocity
   are essentially projections of this velocity onto sphere at
@@
   This indeed can be seen in Figure~\ref{fig:dv-sep} as an over-density
-  in lower right corner that gets thinner as \absdvtan\ increases.
+  in the lower right corner that gets thinner as \absdvtan\ increases.
 
   \item Finally, there is a population of “random” pairs of field stars
-  which are not actually co-moving, but still has $\absdvtan 10$~\kms\
+  which are not actually co-moving, but still have $\absdvtan < 10$~\kms\
   by chance.
@@
   \caption{%
-    Initial sample of pairs on the
-    difference in tangential velocity vs. physical separation plane.
-    For the initial sample, we look for stars
-    with separation $< 10$~pc and $\absdvtan < 10$~\kms\ for
-    every star in \tgas.
+    Point estimates of tangential velocity and physical separation computed for
+    all 271,232 unique pairs in the initial sample of pairs (black points).
+    For this sample, we consider stars
+    with separation $< 10$~pc and $\absdvtan < 10$~\kms\
+    computed relative to every other star in \tgas.
     The blue solid line shows the magnitude of the 3D
@@
     eventually merges with random pairs of field stars dominating in the upper right corner
-    (see \S~\ref{sec:data} for details).
+    (see \sectionname~\ref{sec:data} for details).
     \label{fig:dv-sep}}
@@
     prior $p(\vec{v})$ and  velocity difference drawn from a zero-mean Gaussian
-    with velocity dispersion $s$, or (2) individually drawn from a velocity
-    prior $p(\vec{v})$.
+    with velocity dispersion $s$, or (2) individually drawn from the same
+    velocity prior $p(\vec{v})$.
 \end{itemize}
 
-Under these assumptions, the likelihood of a proper motion measurement for a
-star given true distance, $d$, and true tangential velocity, $\vec{v}^\perp =
+Under these assumptions, the likelihood of a proper motion measurement,
+$\bs{\mu}$, for a star given true distance, $d$, and true tangential velocity,
+$\vec{v}^\perp =
 (\begin{array}[t]{c c} v_\alpha & v_\delta\end{array})^\mathsf{T}$ is
@@
 We use the FML ratio $\mathcal{L}_1/\mathcal{L}_2$ as a scalar for selecting
-candidate co-moving stars, as described in $\S~\ref{sub:selection}$ in more detail.
-To compute these FMLs, the likelihood functions $L_i, L_j$ are marginalized
-over true 3-space velocity and distance for each star in the pair $(i,j)$.
+candidate co-moving stars, as described in $\sectionname~\ref{sub:selection}$ in
+more detail.
+To compute these FMLs, the likelihood functions for each star in a pair, $L_i,
+L_j$, are marginalized over true 3-space velocity and distance for each star in
+the pair $(i,j)$.
 \begin{align}
@@
 star in a pair, $d_1,d_2$, using Monte Carlo integration with $K$ samples from
-the distance posterior pdfs:
+the distance posterior pdfs \todo{Comment on the choice of distance prior}:
 \begin{equation}
@@
 
-For wide binaries, the assumption of same 3D velocity can break down
-for high precision proper motion measurements because of the orbital
+For wide binaries, the assumption of the stars having the same 3D velocity can
+break down for high precision proper motion measurements because of the orbital
 velocity (blue solid line in Figure~\ref{fig:dv-sep}).
@@
 
-\section{Results}
+\section{Results} \label{sec:results}
 
@@
 the likelihood ratio to select candidate co-moving pairs.
-Second, we present the statistics and properties of co-moving pairs, and discuss its nature.
+Second, we present the statistics and properties of our candidate co-moving
+pairs.
 Finally, we describe our catalog of candidate co-moving pairs, the main product
@@
 
-In this section, we examine the distribution of likelihood ratios
+In this section, we examine the distribution of (log-)likelihood ratios
 $\ln \mathcal{L}_1 /\mathcal{L}_2$, and come up with a reasonable cut
-for this quantity to select co-moving pairs from candidate pairs list.
+for this quantity to select co-moving pairs from the initial pairs sample.
 
@@
     \label{fig:likelihoodratios}}
+    \vspace{1em}
 \end{figure*}
@@
 Figure~\ref{fig:likelihoodratios} shows the likelihood ratios for all
-271k candidate pairs examined.
-As discussed in \S~\ref{sec:data}, we expect a correlation between likelihood
+$\approx$271k initial pairs examined.
+As discussed in \sectionname~\ref{sec:data}, we expect a correlation between likelihood
 ratios of pairs, and their distribution in $\Delta v_t$ vs separation plane.
@@
 slices of $\Delta v_t$ centered at 0.5, 5, and 9.5~\kms.
+\todo{Fix in below and in the figure, $\Delta v_t$ should be $|\Delta v_t|$}
 Pairs with $|\Delta v_t-0.5|<0.5$ are most likely actual co-moving pairs, and
@@
 with decreasing $\ln \mathcal{L}_1 /\mathcal{L}_2$ ratios.
-This is in agreement with our discussion in \S~\ref{sec:data}.
+This is in agreement with our discussion in \sectionname~\ref{sec:data}.
 Finally, as a test,
@@
     between two stars if they are co-moving by our selection criteria
-    (see \S~\ref{sub:selection}). On left, we show the largest network
+    (see \sectionname~\ref{sub:selection}). On left, we show the largest network
     found in this study corresponding to the Pleiades star cluster.
@@
 
-Once we have decided the co-moving pairs from the candidate pair list,
-these pairs form an undirected graph where stars are nodes, and
-edges between the nodes exist for co-moving pairs of stars.
+Once we have identified candidate co-moving star pairs from the initial pair
+sample, these pairs form an undirected graph where stars are nodes, and edges
+between the nodes exist for co-moving pairs of stars.
 A star may have multiple co-moving neighbors, and two stars may either
@@
 discovered by looking for co-moving pairs.
-These aggregates are likely moving groups, OB associations or star clusters.
+These aggregates are likely moving groups, OB associations, or star clusters.
 There are 4555 connected components among 10,606 unique stars.
@@
   \begin{center}
-    \includegraphics[width=\textwidth]{figures/gjg_graphviz_examples.pdf}
+    \includegraphics[width=\textwidth]{figures/gjg_graphviz_examples.png}
   \end{center}
@@
 
-\section{Conclusions}
+\section{Conclusions}\label{sec:conclusions}
 

### 2016-11-17 paper/ms.tex
Minor typo fix and edits

@@
 two stars were likely born at the same time with the same chemical compositions,
-but evolve indepdently becayse if their wide separation.
+but evolved indepdently because of their wide separation.
 Finally, calibration of stellar parameters of low-mass stars (e.g., M dwarfs),
@@
 follows.
-We first apply a global parallax signal-to-noise cut, $\snr < 8$,  to \tgas,
+We first apply a global parallax signal-to-noise cut, $\snr > 8$,  to \tgas,
 which leaves 619,618 stars.
@@
 Figure~\ref{fig:likelihoodratios} shows the likelihood ratios for all
-$\approx$271k initial pairs examined.
+$\sim 271$k initial pairs examined.
 As discussed in \sectionname~\ref{sec:data}, we expect a correlation between likelihood

### 2016-11-17 paper/ms.tex
Use macro for consistent symbol of dvtan

@@
   \caption{%
+    \todo{
     Left: Density histogram of likelihood ratios for candidate pairs and random pairs.
@@
     we show the distribution for subsets in slices of $\Delta v_t$.
-    Right: $\Delta v_t$ vs separation in slices of $\ln \mathcal{L}_1 /\mathcal{L}_2$.
+    Right: $\Delta v_t$ vs separation in slices of $\ln \mathcal{L}_1 /\mathcal{L}_2$.}
     \label{fig:likelihoodratios}}
@@
 As discussed in \sectionname~\ref{sec:data}, we expect a correlation between likelihood
-ratios of pairs, and their distribution in $\Delta v_t$ vs separation plane.
-Specifically, as we sweep through from small $\Delta v_t$ to large, we expect
+ratios of pairs, and their distribution in \absdvtan\ vs separation plane.
+Specifically, as we sweep through from small \absdvtan\ to large, we expect
 the population of pairs to change from genuinely co-moving to random.
 This becomes clear when we look at the distribution of likelihood ratios for
-slices of $\Delta v_t$ centered at 0.5, 5, and 9.5~\kms.
-\todo{Fix in below and in the figure, $\Delta v_t$ should be $|\Delta v_t|$}
-Pairs with $|\Delta v_t-0.5|<0.5$ are most likely actual co-moving pairs, and
+slices of \absdvtan\ centered at 0.5, 5, and 9.5~\kms.
+Pairs with $|\absdvtan -0.5|<0.5$ are most likely actual co-moving pairs, and
 their likelihood distribution is narrowly peaked above 6.
-The distribution peaks at lower values and gets broader as $\Delta v_t$ increases,
+The distribution peaks at lower values and gets broader as \absdvtan\ increases,
 and the number of random pairs increasingly dominate.
 On the right panel of Figure~\ref{fig:likelihoodratios}, we show
-how the distribution of pairs on $\Delta v_t$ vs separation plane changes
+how the distribution of pairs on \absdvtan\ vs separation plane changes
 with decreasing $\ln \mathcal{L}_1 /\mathcal{L}_2$ ratios.

### 2016-11-17 paper/ms.tex
Edit abstract and text (from DNS)

@@
 The primary sample of the \gaia\ \textsl{First Data Release} is the brand-new
-\textsl{Tycho-Gaia Astrometric Solution}. The precision and size of this sample
-creates an opportunity to find new binary stars and moving groups, especially
+\textsl{Tycho-Gaia Astrometric Solution} (\tgas). The precision and size of this sample
+presents an opportunity to find new binary stars and moving groups, especially
 rare binaries or sparsely populated groups.
 % Aims
-Here we seize this opportunity.
+Here, we look for high-confidence co-moving pairs of stars using this data.
 % Methods
 We use a justified marginalized likelihood ratio test to separate
-pairs of stars with surprisingly similar three-space velocities from
+pairs of stars with similar three-space velocities from
 those consistent with being drawn independently from the field
@@
 many of which correspond to the known open clusters, and stars in OB associations.
-However, there still is a huge number of very wide ($>3$~pc) separation pairs of
-stars that are co-moving.
-We discuss their nature, and the prospects for testing stellar models and chemical
-abundance measurements using our catalog.
-
-% change to probabilistic statement "co-moving"
-% we present a table of co-moving pairs
-% add dynamical inference
-% the last line = the most important thing
+However, there still is a surprisingly large number (\todo{SMOH: give number})
+of very wide ($>3$~pc) separation pairs of stars that are co-moving.
+We present a catalog of high-confidence co-moving pairs of stars in \tgas\
+with auxiliary information,
+and discuss its potential usage
+for making dynamical inferences, and testing stellar models and chemical
+abundance measurements.
 
@@
 
-At small separations ($0.001-1$~pc), they correspond to wide binaries (and
-multiples) that are either weakly gravitationally bound or currently being
-dispersed.
+At small separations ($0.001-1$~pc), they are wide binaries (and
+multiples) that are either weakly gravitationally bound or
+slowly separating.
 Because they have low binding energies, a sample of wide binaries is valuable
 for investigating the Galactic dynamical environment.
-These systems must have survived their dynamic birth environment which
-provides a window to the star formation process.
-They may also be disrupted by Galactic tides, or other massive objects
-such as other stars, molecular clouds, or MACHOs \citep{Yoo:2004aa,Allen:2014aa}.
+These systems must have survived both their dynamic birth environment and 
+have avoided tidal destruction along their orbit.
+Thus, the statistical properties of wide binaries provide a window into both the star formation
+process and the Galactic dynamics
+including Galactic tides, and other massive perturbers such as molecular clouds
+and MACHOs \citep{Yoo:2004aa,Allen:2014aa}.
 Wide binaries are also good testbeds for stellar evolution models.
 For example, main sequence-red giant wide binaries are especially valuable in
-testing consistency between stellar atmosphere models of these two types as the
+testing consistency between stellar atmosphere models as the
 two stars were likely born at the same time with the same chemical compositions,
@@
   OB associations, moving groups, and open clusters are also of consideration.
-  Since the basic unit of our choice is a pair of stars, we can expect that these astrophysical objects may be detected as networks of pairs.
+  Since the basic unit of our choice is a pair of stars, we can expect that
+  these astrophysical objects may be detected as networks of pairs.
   As the pair separation increases, the nature of co-moving pairs
@@
     \mu_\delta\end{array})^\mathsf{T}$---are Gaussian with known covariances
-    $\mat{C}$. The values of covariances are provided by the \gaia\ data
+    $\mat{C}$. The values of covariances are provided as part of the \gaia\ data
     \citep{Lindegren:2012aa,Lindegren:2016aa}.
@@
 
-We show the distribution co-moving pairs in galactic longitude and
+We show the distribution of co-moving pairs in galactic longitude and
 distance in Figure~\ref{fig:glon_d_pairlines}.

### 2016-11-17 paper/ms.tex
Fix typo

@@
 two stars were likely born at the same time with the same chemical compositions,
-but evolved indepdently because of their wide separation.
+but evolved independently because of their wide separation.
 Finally, calibration of stellar parameters of low-mass stars (e.g., M dwarfs),

### 2016-11-20 paper/ms.tex
Improve likelihood ratio figure and apply style
- New version of likelihood ratio figure
- apply consistent style and labels to all figures

@@
 many of which correspond to the known open clusters, and stars in OB associations.
-However, there still is a surprisingly large number (\todo{SMOH: give number})
+However, there still is a surprisingly large number
 of very wide ($>3$~pc) separation pairs of stars that are co-moving.
@@
   \item At small separations ($<1$~pc), there is a population of pairs with
-  very small tangential velocity difference ($<2$~km/s). Given that these pairs
-  are very close in both position and proper motion,
-  it is highly probable that they are actually co-moving wide binaries.
+    very small tangential velocity difference ($<2$~km/s). Given that these
+    pairs are very close in both position and proper motion, it is highly
+    probable that they are actually co-moving wide binaries.
 
-  \item  However, it is not only (wide) binaries, which are gravitationally
-  bound, that are expected to appear as co-moving.
-  OB associations, moving groups, and open clusters are also of consideration.
-  Since the basic unit of our choice is a pair of stars, we can expect that
-  these astrophysical objects may be detected as networks of pairs.
+  \item  A sample of co-moving stars also include stars that are part of OB associations,
+    moving groups, and open clusters.
+    These astrophysical objects may be detected as a network of co-moving pairs.
     As the pair separation increases, the nature of co-moving pairs
@@
   \item Finally, there is a population of “random” pairs of field stars
-  which are not actually co-moving, but still have $\absdvtan < 10$~\kms\
+    that are not co-moving, but still have $\absdvtan < 10$~\kms\
     by chance.
     As \absdvtan\ increases, this population will dominate.
-  However, it is clear from this Figure that there is an overlap between
+    Figure~\ref{fig:dv-sep} shows that there is an overlap between
     genuine co-moving pairs and “random” pairs.
@@
 star in a pair, $d_1,d_2$, using Monte Carlo integration with $K$ samples from
-the distance posterior pdfs \todo{Comment on the choice of distance prior}:
+the distance posterior pdfs assuming a uniform density prior
+\citep{Bailer-Jones:2015aa}
 \begin{equation}
@@
   \caption{%
-    \todo{
-    Left: Density histogram of likelihood ratios for candidate pairs and random pairs.
-    Along with the distribution for the entire candidate pairs (black thick line),
-    we show the distribution for subsets in slices of $\Delta v_t$.
-    Right: $\Delta v_t$ vs separation in slices of $\ln \mathcal{L}_1 /\mathcal{L}_2$.}
+    Justification of the likelihood ratio cut.
+    On the top row, we show how the distribution of pairs on \absdvtan\ vs. separation changes
+    in slices of $\ln \mathcal{L}_1 /\mathcal{L}_2$ (sequentially color-coded in blue).
+    On the bottom row, we show how the distribution of $\ln \mathcal{L}_1 /\mathcal{L}_2$
+    changes in slices of \absdvtan\ (sequentially color-coded in red).
     \label{fig:likelihoodratios}}
@@
 We divide the network into connected components\footnote{A connected component of an undirected graph $G$
-is a subgraph of $G$ in which any two nodes are connected to each other by a path.
-We use the term ``network'' loosely to refer to a graph or its connected components.},
+is a subgraph of $G$ in which any two nodes are connected to each other by a path.},
 and show the distribution of their sizes in Figure~\ref{fig:hist_ccsize}.
@@
   \begin{center}
-    \includegraphics[width=\textwidth]{figures/gjg_mepairs.png}
+    \includegraphics[width=\textwidth]{figures/gjg_mepairs.pdf}
   \end{center}

### 2016-11-21 paper/ms.tex
More edits to text and figures
- change row order of the likelihood ratio figure
- change angle origin and orientation of pie plot
- consistent usage of "connected component"
- consistent use of symbols
- more emphasis on the novel results

@@
 
-\title{Co-moving stars in \textsl{Gaia DR1}}
+\title{Co-moving stars in \textsl{Gaia DR1}:\\
+  Abundance of very wide separation co-moving pairs}
 \author{Semyeong Oh\altaffilmark{\pu,\lead},
@@
 We use a justified marginalized likelihood ratio test to separate
-pairs of stars with similar three-space velocities from
+pairs of stars with similar 3D velocities from
 those consistent with being drawn independently from the field
@@
 % Results
-We find pairs of co-moving stars out to very wide
-separations up to 10~pc, the limit of our search.
-Some of these pairs form larger networks through mutual co-moving neighbors,
+We find 13,085 pairs of co-moving stars among 10,606 unique stars
+out to very wide separations up to 10~pc, the limit of our search.
+Some of these pairs form a larger group through mutual co-moving neighbors,
 many of which correspond to the known open clusters, and stars in OB associations.
-However, there still is a surprisingly large number
-of very wide ($>3$~pc) separation pairs of stars that are co-moving.
-We present a catalog of high-confidence co-moving pairs of stars in \tgas\
-with auxiliary information,
-and discuss its potential usage
+However, we also discover many co-moving groups of pairs for the first time.
+More surprisingly, we find a large number
+of very wide ($>3$~pc) separation co-moving pairs of stars.
+The number of these very wide separation pairs only increase with increasing
+separation, and cannot be merely due to false positives.
+We present a catalog of high-confidence co-moving pairs of stars in the \tgas\
+with auxiliary information, and discuss its potential usage
 for making dynamical inferences, and testing stellar models and chemical
@@
 a larger sample of widely separated binaries containing a low mass star and a
-much brighter FGK star whose stellar parameters are easier to measure
+much brighter F/G/K star whose stellar parameters are easier to measure
 \citep[e.g.,][]{Rojas-Ayala:2012aa}.
@@
 using high and low resolution spectroscopy, we can now start to explore these questions
-in unprecedented width (statistics) and depth (detail) \todo{What do you mean by this?}.
+with unprecedented statistics and in unexplored detail.
 The dynamics of cluster dissolution provides important clues to understanding
@@
 In the halo, we know of more than 20 disrupting globular clusters and dwarf galaxies
-(see, e.g., \citealt{Grillmair:2016} for a summary of known streams).
+(``streams''; see, e.g., \citealt{Grillmair:2016} for a summary of known streams).
 These tidal streams are modeled to infer the parameters of the Galactic
@@
 \citep{Gaia-Collaboration:2016aa,Lindegren:2016aa}.
-\tgas\ contains astrometric measurements (sky position,
+The \tgas\ contains astrometric measurements (sky position,
 parallax, and proper motions) and associated covariance matrices for a large
@@
 follows.
-We first apply a global parallax signal-to-noise cut, $\snr > 8$,  to \tgas,
+We first apply a global parallax signal-to-noise cut, $\snr > 8$,  to the \tgas,
 which leaves 619,618 stars.
@@
 We ultimately build a statistical model that incorporates the covariances of the
-data, but for this position cut and for visualizations we use a point-estimate
+data, but for these initial cuts and for visualizations we use a point-estimate
 of the distance by applying a correction for the Lutz-Kelker bias
@@
     very small tangential velocity difference ($<2$~km/s). Given that these
-    pairs are very close in both position and proper motion, it is highly
+    pairs are very close in both 3D position and tangential velocities, it is highly
     probable that they are actually co-moving wide binaries.
@@
     moving groups, and open clusters.
-    These astrophysical objects may be detected as a network of co-moving pairs.
+    These astrophysical objects may be detected as a network of co-moving pairs,
+    sharing some mutual co-moving neighbors.
     As the pair separation increases, the nature of co-moving pairs
@@
     Since the proper motions of two stars with the same 3D velocity
-    are essentially projections of this velocity onto sphere at
+    are essentially projections of this velocity onto the celestial sphere at
     two different viewing angles,
@@
 \figname~\ref{fig:dv-sep} suggests that there are a
-significant number of co-moving stars in the \tgas\ data at a range
+significant number of co-moving pairs in the \tgas\ data at a range
 of separations.
@@
     Justification of the likelihood ratio cut.
-    On the top row, we show how the distribution of pairs on \absdvtan\ vs. separation changes
-    in slices of $\ln \mathcal{L}_1 /\mathcal{L}_2$ (sequentially color-coded in blue).
-    On the bottom row, we show how the distribution of $\ln \mathcal{L}_1 /\mathcal{L}_2$
-    changes in slices of \absdvtan\ (sequentially color-coded in red).
+    On the top row, we show how the distribution of $\ln \mathcal{L}_1 /\mathcal{L}_2$
+    changes (left) in slices of \absdvtan\ (sequentially color-coded in pink on the right panel).
+    On the bottom row, we show how the distribution of pairs on \absdvtan\ vs. separation changes (right)
+    in slices of $\ln \mathcal{L}_1 /\mathcal{L}_2$ (sequentially color-coded in blue on the left).
+    The black dots (upper right) and lines (lower left) correspond to the entire initial sample of pairs.
+    For comparision, we also present the likelihood ratio distribution of random pairs
+    of stars in the \tgas\ with parallax $S/N>8$ in gray filled histogram.
+    Based on this, we choose $\ln \mathcal{L}_1 /\mathcal{L}_2 > 6$ as our high-confidence
+    candidate co-moving pairs.
     \label{fig:likelihoodratios}}
@@
 As discussed in \sectionname~\ref{sec:data}, we expect a correlation between likelihood
-ratios of pairs, and their distribution in \absdvtan\ vs separation plane.
+ratios of pairs, and their distribution on \absdvtan\ vs separation plane.
 Specifically, as we sweep through from small \absdvtan\ to large, we expect
 the population of pairs to change from genuinely co-moving to random.
-This becomes clear when we look at the distribution of likelihood ratios for
-slices of \absdvtan\ centered at 0.5, 5, and 9.5~\kms.
-Pairs with $|\absdvtan -0.5|<0.5$ are most likely actual co-moving pairs, and
-their likelihood distribution is narrowly peaked above 6.
+This becomes clear when we look at the distribution of the likelihood ratios of pairs
+in slices of \absdvtan\ (top row of Figure~\ref{fig:likelihoodratios}).
+Pairs with $\absdvtan \in (0,2.5)~\kms$ are most likely actual co-moving pairs, and
+their likelihood ratio distribution is narrowly peaked at $\gtrsim 5$
+(darkest pink histogram in the upper left panel of Figure~\ref{fig:likelihoodratios}).
 The distribution peaks at lower values and gets broader as \absdvtan\ increases,
 and the number of random pairs increasingly dominate.
-On the right panel of Figure~\ref{fig:likelihoodratios}, we show
+On the bottom row of Figure~\ref{fig:likelihoodratios}, we show
 how the distribution of pairs on \absdvtan\ vs separation plane changes
@@
 we compute the likelihood ratios for 200,000 random pairs of stars with the same
-parallax signal-to-noise ratio cut as candidate pairs ($S/N > 8$).
-Shown as gray shaded histogram in Figure~\ref{fig:likelihoodratios},
+parallax signal-to-noise ratio cut as the initial sample ($S/N > 8$).
+Shown as the gray filled histogram in Figure~\ref{fig:likelihoodratios},
 this distribution peaks at a much lower value ($\approx 2$),
@@
 Based on these comparisons, we select pairs with
-$\ln \mathcal{L}_1 /\mathcal{L}_2 > 6$ as co-moving pair candidates.
-Out of 271,232 candidate pairs examined, 13,058 pairs (4.8\%)
+$\ln \mathcal{L}_1 /\mathcal{L}_2 > 6$ as candidate co-moving pair.
+Out of 271,232 pairs in the initial sample, 13,058 pairs (4.8\%)
 satisfy this condition.
@@
     projection of onto the Galactic plane. The angle is the Galactic
-    longitude, and the radius is $d\cos b$ in pc where $d$ is the point-estimate
+    longitude, and the radius is $\hat d\cos b$ in pc where $\hat d$ is the point-estimate
     of distance using Equation~\ref{eq:dist}.
     Pairs in connected components of sizes less than 5 are connected by gray lines.
-    Pair in connected components of size $\geq 5$ are plotted with a unique color per network,
+    Pair in connected components of size $\geq 5$ are plotted with a unique random color,
     and their nodes (stars) are highlighted with small black circles.
     We also show the positions of known Milky Way star clusters
-    from \citet{Kharchenko:2016aa} as empty purple circles,
+    from \citet{Kharchenko:2016aa} as purple circles,
     and stars in OB associations from \citet{de-Zeeuw:1999aa} as
     gray circles.
+    Some of the larger connected components are clearly associated with known
+    clusters, but we also discover quite a few for the first time.
     \label{fig:glon_d_pairlines}}
@@
 
-Once we have identified candidate co-moving star pairs from the initial pair
+Once we have identified candidate co-moving pairs from the initial
 sample, these pairs form an undirected graph where stars are nodes, and edges
@@
 A star may have multiple co-moving neighbors, and two stars may either
-be directly or indirectly connected by a set of edges.
-We divide the network into connected components\footnote{A connected component of an undirected graph $G$
+be directly or indirectly connected by a sequence of edges (``path'').
+We divide the graph into connected components\footnote{A connected component of an undirected graph $G$
 is a subgraph of $G$ in which any two nodes are connected to each other by a path.},
 and show the distribution of their sizes in Figure~\ref{fig:hist_ccsize}.
-The most common are size 2 networks, which mean mutually exclusive co-moving
+The most common are connected components of size 2, which mean mutually exclusive co-moving
 pairs. However, it is clear that there are many aggregates of co-moving stars
@@
 These aggregates are likely moving groups, OB associations, or star clusters.
-There are 4555 connected components among 10,606 unique stars.
-The maximum size of connected components is 151.
-We show this largest network along with four randomly selected
-examples of networks with sizes greater than 4 in Figure~\ref{fig:graphviz_examples}.
-The largest network corresponds to the Pleiades open cluster.
-One particular star in this network, TYC 1799-272-1, has 94 directly connected
-co-moving neighbors, and is located at the center of the network.
+In 13,058 co-moving pairs that we identified,
+there are 4555 connected components among 10,606 unique stars.
+The maximum size of the connected components is 151.
+We show this largest connected component along with four other
+examples of varying sizes in Figure~\ref{fig:graphviz_examples}.
+The largest connected component corresponds to the Pleiades open cluster (left panel of Figure~\ref{fig:graphviz_examples})
+while the one on the upper left panel of the right column of Figure~\ref{fig:graphviz_examples}
+is NGC~2632, another known open cluster of the Milky Way \citep{Kharchenko:2016aa}.
+%One particular star in this network, TYC 1799-272-1, has 94 directly connected
+%co-moving neighbors, and is located at the center of the network.
 
@@
 distance in Figure~\ref{fig:glon_d_pairlines}.
-The connection between known co-moving structures and the networks of co-moving
-pairs found in this work becomes immediately clear when
+The connection between known co-moving structures and
+the larger connected components found in this work becomes immediately clear when
 we overplot the positions of known Milky Way star clusters \citep{Kharchenko:2016aa},
 and stars in OB associations \citep{de-Zeeuw:1999aa}.
-Many of the larger networks are clearly associated with known star clusters.
-The most prominent is Melotte 111 at
-$(l,d)\approx (220\arcdeg, 87~\textrm{pc})$ otherwise known as the Coma star cluster.
+Many of the larger connected components are
+clearly associated with known star clusters:
+Melotte 22 (Pleiades) at $(l,d)\approx (167\arcdeg, 130~\textrm{pc})$,
+Melotte 20 at $(l,d)\approx (147\arcdeg, 175~\textrm{pc})$,
+Melotte 25 (Hyades) at $(l,d)\approx (180\arcdeg, 50~\textrm{pc})$, and
+NGC 2632 (Beehive) at $(l,d)\approx (206\arcdeg, 187~\textrm{pc})$
+to name a few.
 Clumps of co-moving pairs at $(l,d)\approx (300-360\arcdeg, 100-200~\textrm{pc})$
@@
 
+However, there are still many larger connected components that we
+discover for the first time.
+If we define a condition to associate a connected component to a known cluster
+as having more than 3 members within 10~pc of the nominal position of
+the cluster, for the 61 connected components with size larger than 5,
+we find that only 10 are associated with a cluster cataloged in
+\citet{Kharchenko:2016aa}.
+It is also worth noting that the positions of some of the known clusters
+are offset from those of the connected components associated with them,
+indicating that the \tgas\ data improves the distance estimates of these clusters.
+Finally, not all known star clusters are recovered in our search.
+This is primarily because of the non-uniform coverage of \tgas\ data.
+
 Ultimately, any candidate co-moving pair found in this work
 needs to be verified using radial velocities.
-Here, we use 210,368 cross-matches of \tgas\ with
+Here, we use 210,368 cross-matches of the \tgas\ with
 the Radial Velocity Experiment (\rave) to
@@
 (which depends on the local stellar number density and velocity dispersion).
-Given the excellent correspondence between co-moving pair networks of larger sizes
+Given the excellent correspondence between
+aggregates of co-moving pairs (connected components)
 and known genuine co-moving structures (Figure~\ref{fig:glon_d_pairlines}),
-we may expect that pairs in these larger networks which will often have
-separations $>1$~pc to have less contamination.
-We divide pairs into those mutually exclusively connected (i.e., in a network of
-size 2), and those in a larger network.
-We indeed find that many pairs in larger networks are at $>1$~pc, yet
+we may expect that pairs in these larger connected components,
+which will often have separations $>1$~pc,
+to have less contamination from false positives.
+We divide the co-moving pairs into those
+mutually exclusively connected (i.e., in a connected component of
+size 2), and those in a larger group.
+We indeed find that many pairs in larger groups are at $>1$~pc, yet
 the fraction of pairs that have identical radial velocities within $3\sigma$ is
-higher than mutually exclusive pairs, and remains high ($>80\%$) to $\sim 10$~pc.
+higher than the mutually exclusive pairs, and remains high ($>80\%$) to $\sim 10$~pc.
 
@@
     We highlight $3\sigma$ limit with two horizontal lines.
-    Bottom: Fraction of pairs with $\Delta v_r/\sigma_{\Delta v_r} <3$
+    Bottom: Fraction of pairs with $|\Delta v_r/\sigma_{\Delta v_r}| <3$
     as a function of physical separation. We calculate the fraction with a running
@@
 Figure~\ref{fig:hist_separation}.
-As expected, pairs in larger networks are mostly found with separations
-larger than 1~pc.
-Surprisingly, however, we also find a huge number of mutually exclusive pairs
+As expected, pairs in larger connected components are mostly
+found with separations larger than 1~pc.
+Surprisingly, however, we also find a large number of mutually exclusive co-moving
+pairs
 at $>1$~pc as well. Even if we consider the increasing false positive rate at large
@@
 at $>1$~pc is in fact increasing much faster (as a power-law)
-than the decrease due to the false positives (right panel of Figure~\ref{fig:raverv}).
+than the decrease due to the false positives
+(bottom panel of Figure~\ref{fig:raverv}).
 The nature of these very large separation, mutually exclusive pairs,
@@
 will follow. Figure~\ref{fig:cmd_large} and \ref{fig:cmd_me} shows $G-J$ vs $G$
-color-magnitude diagrams for stars in larger networks (size$>2$) and in mutually
-exclusive pairs (networks of size 2), respectively.
-The networks shown in Figure~\ref{fig:cmd_large} correspond to those visualized
+color-magnitude diagrams for stars in larger connected components (size$>2$)
+and in mutually exclusive pairs, respectively.
+The connected components shown in Figure~\ref{fig:cmd_large}
+correspond to those visualized
 in Figure~\ref{fig:graphviz_examples}.
-For stars in larger networks, there is a noticeable lack of evolved, off-main sequence
+For stars in larger co-moving groups, there is a noticeable
+lack of evolved, off-main sequence
 stars, in agreement with these kinematic structures being young.
@@
   \caption{
-    Color-magnitude diagrams of stars in larger networks presented in
-    Figure~\ref{fig:graphviz_examples}. Each panel shows the same network
-    visualized in the panel at the same position
+    Color-magnitude diagrams of stars in larger connected components presented in
+    Figure~\ref{fig:graphviz_examples}. Each panel shows the same group
+    of stars visualized in the panel at the same position
     in Figure~\ref{fig:graphviz_examples}.
-    We show the distribution of all stars common in \tgas\ and \tmass\
-    with parallax $S/N>16$ in blue for reference.
+    We show a reference distribution of stars randomly drawn to have
+    a matching distribution in distance as the stars in our co-moving pairs
+    in blue.
     \label{fig:cmd_large}}
@@
   \caption{
-    Color-magnitude diagrams of stars in networks of size 2 (i.e., mutually exclusive)
+    Color-magnitude diagrams of stars in mutually exclusive co-moving pairs
     in two separation bins.
-    We connect each pair by a line on the left for pairs with separations smaller than
-    1~pc.
-    We show the distribution of all stars common in \tgas\ and \tmass\
-    with parallax $S/N>16$ in blue for reference same as in Figure~\ref{fig:cmd_large}.
+    We connect each pair by a line on the left for pairs
+    with separations smaller than 1~pc.
+    We show a reference distribution of stars randomly drawn to have
+    a matching distribution in distance as the stars in our co-moving pairs
+    in blue, same as Figure~\ref{fig:cmd_large}.
     \label{fig:cmd_me}}
@@
   \item \tgas\ source id for the two stars in a pair; these may be used to easily
-  retrieve cross matches of \gaia\ with other surveys such as WISE, \tmass, and \rave
-  using the \gaia\ data archive.
+    retrieve cross matches of \gaia\ with other surveys such as WISE, \tmass, and
+    \rave\ using the \gaia\ data archive.
   \item physical separation between the two stars in pc
   \item likelihood ratio, $\ln \mathcal{L}_1 /\mathcal{L}_2$
-  \item size of the network that the pair belongs to
-  \item unique ID of the network that the pair belongs to
+  \item size of the connected component that the pair belongs to
+  \item unique ID of the connected component that the pair belongs to; we assigned
+    a unique ID in descending order of size, thus ID~0 is the largest connected component with 151 stars that corresponds to the Pleiades.
 \end{itemize}
 %
-Additionally, given the strong correlation between the larger networks of pairs
+Additionally, given the strong correlation between the larger connected components
+of co-moving pairs
 and the known star clusters and OB associations, we provide for each star
@@
 
-We note a caveat on the completeness of the networks found in our catalog.
-Because we applied a simple cut in the likelihood ratio ($\ln \mathcal{L}_1 /\mathcal{L}_2>6$),
-there is a possibility that, for example, a star in a mutually exclusive pair in our catalog may
-still have another possibly co-moving companion which has been dropped because the likelihood ratio
-is slightly less than 6.
+We note a caveat on the completeness of a connected component of co-moving
+pairs found in our catalog. Because we applied a simple cut in the likelihood
+ratio ($\ln \mathcal{L}_1 /\mathcal{L}_2>6$), there is a possibility that, for
+example, a star in a mutually exclusive pair in our catalog may still have
+another possibly co-moving companion which has been dropped because the
+likelihood ratio is slightly less than 6.
 
@@
 
-We examined the network of co-moving stars, and found that many of the larger
-networks of co-moving stars naturally corresponds to some of the known co-moving structures
+We examined the network of co-moving stars which forms an undirected graph,
+and found that many of the larger groups of 
+co-moving stars sharing one or more co-moving neighbor (connected components)
+naturally correspond to some of the known co-moving structures
 such as open clusters and stellar associations.
+Yet, still larger number of co-moving groups are discovered for the first time.
+
 We have also found a surprisingly large number of very wide separation ($>3$~pc)
 co-moving pairs which are not clearly associated to any known existing clusters.
-While their nature is yet unclear, they may be remnants of dissolving wide binaries and
+These are most likely remnants of dissolving wide binaries and
 open clusters. If so, this population should still be relatively young compared

### 2016-11-21 paper/stub.tex
More edits to text and figures
- change row order of the likelihood ratio figure
- change angle origin and orientation of pie plot
- consistent usage of "connected component"
- consistent use of symbols
- more emphasis on the novel results

@@
 \tablecaption{Catalog of candidate co-moving pairs \label{tab:catalog}}
-\tablehead{\colhead{star1 source id} & \colhead{star2 source id} & \colhead{separation} & \colhead{$\ln \mathcal{L}_1 /\mathcal{L}_2$} & \colhead{ID$_\mathrm{network}$} & \colhead{$N_\mathrm{network}$}\\ \colhead{ } & \colhead{ } & \colhead{$\mathrm{pc}$} & \colhead{ } & \colhead{ } & \colhead{ }}
+\tablehead{\colhead{star1 source id} & \colhead{star2 source id} & \colhead{separation} & \colhead{$\ln \mathcal{L}_1 /\mathcal{L}_2$} & \colhead{ID$_\mathrm{CC}$} & \colhead{$N_\mathrm{CC}$}\\ \colhead{ } & \colhead{ } & \colhead{$\mathrm{pc}$} & \colhead{ } & \colhead{ } & \colhead{ }}
 \startdata

### 2016-11-21 paper/ms.tex

@@
 for investigating the Galactic dynamical environment.
-These systems must have survived both their dynamic birth environment and 
-have avoided tidal destruction along their orbit.
-Thus, the statistical properties of wide binaries provide a window into both the star formation
-process and the Galactic dynamics
-including Galactic tides, and other massive perturbers such as molecular clouds
+These systems must have both survived their dynamic birth environment and 
+avoided tidal destruction along their orbit.
+Thus, the statistical properties of wide binaries provide a window into both star formation
+processes and Galactic dynamics,
+including Galactic tides and other massive perturbers such as molecular clouds
 and MACHOs \citep{Yoo:2004aa,Allen:2014aa}.
+
 Wide binaries are also good testbeds for stellar evolution models.
 For example, main sequence-red giant wide binaries are especially valuable in
-testing consistency between stellar atmosphere models as the
+testing consistency between stellar atmosphere models, as the
 two stars were likely born at the same time with the same chemical compositions,

### 2016-11-21 paper/ms.tex
Hogg comments on title and abstract

@@
 \title{Co-moving stars in \textsl{Gaia DR1}:\\
-  Abundance of very wide separation co-moving pairs}
+  An abundance of very wide separation co-moving pairs}
 \author{Semyeong Oh\altaffilmark{\pu,\lead},
@@
 out to very wide separations up to 10~pc, the limit of our search.
-Some of these pairs form a larger group through mutual co-moving neighbors,
+Some of these pairs form larger groups through mutual co-moving neighbors,
 many of which correspond to the known open clusters, and stars in OB associations.
-However, we also discover many co-moving groups of pairs for the first time.
+However, we also discover many new co-moving groups of pairs.
 More surprisingly, we find a large number
 of very wide ($>3$~pc) separation co-moving pairs of stars.
-The number of these very wide separation pairs only increase with increasing
-separation, and cannot be merely due to false positives.
+The number of these very wide separation pairs increase with increasing
+separation; and cannot be explained purely by false-positives contamination.
 We present a catalog of high-confidence co-moving pairs of stars in the \tgas\
-with auxiliary information, and discuss its potential usage
+with auxiliary information, and discuss its potential use
 for making dynamical inferences, and testing stellar models and chemical
@@
     Some of the larger connected components are clearly associated with known
-    clusters, but we also discover quite a few for the first time.
+    clusters, but we also discover quite a few new co-moving groups.
     \label{fig:glon_d_pairlines}}
@@
 
-However, there are still many larger connected components that we
-discover for the first time.
+However, there are still many new larger connected components that we
+discover.
 If we define a condition to associate a connected component to a known cluster
@@
 the Radial Velocity Experiment (\rave) to
-assess the false positive rates of our selection.
+assess the false-positive rates of our selection.
 We have 283 pairs with both stars matched with \rave.
@@
 which will often have separations $>1$~pc,
-to have less contamination from false positives.
+to have less contamination from false-positives.
 We divide the co-moving pairs into those
@@
 pairs
-at $>1$~pc as well. Even if we consider the increasing false positive rate at large
+at $>1$~pc as well. Even if we consider the increasing false-positive rate at large
 separations, the distribution is not significantly changed as the number of pairs
 at $>1$~pc is in fact increasing much faster (as a power-law)
-than the decrease due to the false positives
+than the decrease due to the false-positives
 (bottom panel of Figure~\ref{fig:raverv}).
@@
 For mutually exclusive pairs, we divide the pairs by separation at 1~pc
-above which the false positive rate due to random pairs starts to increase.
+above which the false-positive rate due to random pairs starts to increase.
 While many pairs are located along the main sequence, we also find

### 2016-12-01 paper/ms.tex
Edits on text and figures
- update 2mass cross match as in Gaia data archive
- better colors in glon_d_pie
- text edit on summary

@@
 \begin{abstract}
-% Context
 The primary sample of the \gaia\ \textsl{First Data Release} is the brand-new
@@
 rare binaries or sparsely populated groups.
-% Aims
 Here, we look for high-confidence co-moving pairs of stars using this data.
-% Methods
 We use a justified marginalized likelihood ratio test to separate
@@
 responsibly.
-% Results
 We find 13,085 pairs of co-moving stars among 10,606 unique stars
@@
 many of which correspond to the known open clusters, and stars in OB associations.
-However, we also discover many new co-moving groups of pairs.
+We also report many new co-moving groups of pairs.
 More surprisingly, we find a large number
@@
 abundance measurements.
-
-% HOGG:
-% There does not seem to be any tidal-disruption feature in the separation distribution
-% at sub-parsec scales. Pairs at separations beyond a
-% parsec---whether they are bound in a multiple system or drifting
-% apart---must be very short-lived. The photometric properties of the
-% members of these pairs are consistent with youth. The prospects for
-% testing stellar models and chemical abundance measurements are
-% superficially discussed.
 \end{abstract}
@@
 of co-moving stars.
-We conclude in \sectionname~\ref{sec:conclusions}.
+We conclude in \sectionname~\ref{sec:summary}.
 
@@
 For each surviving star, we then establish an initial sample of possible
-co-moving partners by selecting all other stars within 10 pc with a difference
-in tangential velocity below a threshold, $\absdvtan < 10$~\kms.
+co-moving partners by selecting all other stars that satisfy two criteria:
+separation less than 10~pc and difference
+in tangential velocity less than $\absdvtan < 10$~\kms.
 We ultimately build a statistical model that incorporates the covariances of the
@@
 
-\begin{figure*}[htbp]
+\begin{figure}[htbp]
   \begin{center}
-    \includegraphics[width=\textwidth]{figures/sep_dvtan.pdf}
+    \includegraphics[width=\linewidth]{figures/sep_dvtan.pdf}
   \end{center}
@@
     \label{fig:dv-sep}}
-\end{figure*}
+\end{figure}
 
@@
 
-\begin{figure*}[htbp]
+\begin{figure}[htbp]
   \begin{center}
-    \includegraphics[width=\textwidth]{figures/dist_networksize.pdf}
+    \includegraphics[width=\linewidth]{figures/dist_networksize.pdf}
   \end{center}
@@
     \label{fig:hist_ccsize}}
-\end{figure*}
+\end{figure}
 
@@
     We also show the positions of known Milky Way star clusters
-    from \citet{Kharchenko:2016aa} as purple circles,
+    from \citet{Kharchenko:2016aa} as light blue circles,
     and stars in OB associations from \citet{de-Zeeuw:1999aa} as
@@
 
-\section{Conclusions}\label{sec:conclusions}
+\section{Summary}\label{sec:summary}
 
@@
 stars shares the same 3D velocity, and that the two stars have two independent
-3D velocities, incorporating the errors in parallax and proper motions.
+3D velocities, incorporating the covariances of parallax and proper motions.
 We argued for a reasonable cut of the likelihood ratio, and found
@@
 
-We examined the network of co-moving stars which forms an undirected graph,
-and found that many of the larger groups of 
-co-moving stars sharing one or more co-moving neighbor (connected components)
+We found that some co-moving pairs that we have identified are connected
+by sharing a common co-moving neighbor.
+This network of co-moving pairs, which forms an undirected graph,
+can be decomposed into connected components in which any two stars are connected
+by a path.
+The entire 13,058 candidate co-moving pairs are grouped into 4555 connected
+components. The most common is a size-2 connected component i.e., the two
+stars in these pairs are mutually exclusively linked.
+Many of the larger connected components 
 naturally correspond to some of the known co-moving structures
 such as open clusters and stellar associations.
-Yet, still larger number of co-moving groups are discovered for the first time.
+Some of these co-moving groups of stars are newly discovered.
 
 We have also found a surprisingly large number of very wide separation ($>3$~pc)
-co-moving pairs which are not clearly associated to any known existing clusters.
+mutually exclusive co-moving pairs, in which the stars are the only co-moving neighbor
+of each other, and not part of large connected components.
 These are most likely remnants of dissolving wide binaries and
@@
 for wide binaries. They are found to be pairs of stars of varying stellar types.
-Some of these pairs, such as main sequence-red giant or FGK-M dwarfs,
+Some of these pairs, such as main sequence-red giant or F/G/K-M dwarfs,
 will be particularly valuable for testing theoretical
@@
 
+We note that a similar search for wide binaries using the \tgas\ data is performed
+in a recent work by \citet{2016arXiv161107883O}. They find $\approx 1900$ wide binaries
+with separation typically less than 1.5~pc, and 256 pairs with separation larger than
+$\sim 1$~pc.
+We emphasize that our method is based on a probabilistic model for the assumptions
+on the \emph{3D velocities} of the two stars in a pair, and that we
+marginalize over the (unknown) three components of the velocities 
+in contrast to just working in the proper motion space.
+
+Finally, we make our catalog of 13,058 candidate co-moving pair available to the
+community.
+What we find using the \tgas\ is only a taste of what we will discover with
+the future releases of \gaia.
+
 \acknowledgements
 
-This work has made use of data from the European Space Agency (ESA)
-mission {\it Gaia} (\url{http://www.cosmos.esa.int/gaia}), processed by
-the {\it Gaia} Data Processing and Analysis Consortium (DPAC,
-\url{http://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding
-for the DPAC has been provided by national institutions, in particular
-the institutions participating in the {\it Gaia} Multilateral Agreement.
-This project was developed in part at the 2016 NYC Gaia Sprint,
-hosted by the Center for Computational Astrophysics at the Simons Foundation in New York City.
+This work has made use of data from the European Space Agency (ESA) mission
+{\it Gaia} (\url{http://www.cosmos.esa.int/gaia}), processed by the {\it Gaia}
+Data Processing and Analysis Consortium (DPAC,
+\url{http://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding for the DPAC
+has been provided by national institutions, in particular the institutions
+participating in the {\it Gaia} Multilateral Agreement. This project was
+developed in part at the 2016 NYC Gaia Sprint, hosted by the Center for
+Computational Astrophysics at the Flatiron Institute in New York City. The
+Flatiron Institute is supported by the Simons Foundation.
 

### 2016-12-02 paper/ms.tex
Refs to intro and minor edits; method incomplete

@@
 
-Stars that are roughly co-located, and moving with similar space velocity (``co-moving stars'')
-are of special interest in many branches of astrophysics.
+Stars that are roughly co-located, and moving with similar space velocities
+(``co-moving stars'') are of special interest in many branches of astrophysics.
 
@@
 Thus, the statistical properties of wide binaries provide a window into both star formation
-processes and Galactic dynamics,
+processes (e.g., \citealt{Parker:2009aa}) and Galactic dynamics (\citealt{Heggie:1975aa}),
 including Galactic tides and other massive perturbers such as molecular clouds
-and MACHOs \citep{Yoo:2004aa,Allen:2014aa}.
+and MACHOs \citep{Weinberg:1987aa,Jiang:2010aa,Yoo:2004aa,Allen:2014aa}.
 
@@
 moving groups, associations, and star clusters.
-The origin of moving groups is still under active debate: are they remnants of a coeval
+The origin of moving groups is still under active debate (e.g., \citealt{Bovy:2010aa}:
+are they remnants of a coeval
 star formation event with similar chemical composition? Or are they formed by
 dynamical effects of non-axisymmetric features of the Galaxy such as spirals
-and bars? With the recent improvements in measuring chemical abundances
-using high and low resolution spectroscopy, we can now start to explore these questions
+and bars? With the recent advances in measuring chemical abundances of
+a large volume of stars using high and low resolution spectroscopy
+(e.g., \citealt{Steinmetz:2006aa,Majewski:2015aa,Gilmore:2012aa} to name a few),
+we can now start to explore these questions
 with unprecedented statistics and in unexplored detail.
@@
 Here, we use the recent first data release of \gaia\ which includes precise distances,
-enabling us to ask whether two stars share the same {\it physical} velocity rather than
+enabling us to ask whether two stars share the same \emph{physical (3D) velocity} rather than
 just the projections in the proper motion space.
@@
 This paper proceeds as follows:
-In \sectionname~\ref{sec:data}, we briefly describe the data set (TGAS) used in
+In \sectionname~\ref{sec:data}, we briefly describe the data set used in
 this work.
 In \sectionname~\ref{sec:methods}, we develop a statistical method to identify
-high-confidence co-moving stars in this catalog.
+high-confidence co-moving pairs in this catalog.
 In \sectionname~\ref{sec:results}, we present and discuss our resulting catalog
-of co-moving stars.
-We conclude in \sectionname~\ref{sec:summary}.
-
-%TODO: highlight the merits of this search:
-% - we have distance,
-% - bigger volume than Hipparcos,
-% - we use the data properly,
-% - we don't cut on distance or on magnitude of proper motion -- we cut on parallax S/N}
+of co-moving pairs.
+We summarize in \sectionname~\ref{sec:summary}.
 
@@
 which leaves 619,618 stars.
-For each surviving star, we then establish an initial sample of possible
+Then, for each star we establish an initial sample of possible
 co-moving partners by selecting all other stars that satisfy two criteria:
@@
 
-  \item  A sample of co-moving stars also include stars that are part of OB associations,
-    moving groups, and open clusters.
+  \item  A sample of co-moving stars also include stars that are part of e.g.,
+    OB associations, moving groups, and open clusters.
     These astrophysical objects may be detected as a network of co-moving pairs,
@@
     Since the proper motions of two stars with the same 3D velocity
-    are essentially projections of this velocity onto the celestial sphere at
+    are projections of this velocity onto the celestial sphere at
     two different viewing angles,
@@
 of separations.
-Here we develop a method to select high-confidence co-moving
-stars that properly incorporates the complex uncertainties associated with the
+Here, we develop a method to select high-confidence co-moving
+pairs that properly incorporates the uncertainties associated with the
 \gaia\ data. We make the following assumptions in order to construct a
@@
     \citep{Lindegren:2012aa,Lindegren:2016aa}.
-  \item We assume that the 3-space velocities of a given pair of stars
-    $(\vec{v}_i, \vec{v}_j)$ in the \tgas\ sample (relative to the solar system
-    barycenter) are either (1) co-moving with velocity drawn from a velocity
-    prior $p(\vec{v})$ and  velocity difference drawn from a zero-mean Gaussian
-    with velocity dispersion $s$, or (2) individually drawn from the same
-    velocity prior $p(\vec{v})$.
+  \item We assume that the 3D velocities of stars in a given pair
+    $(\vec{v}_i, \vec{v}_j)$ (relative to the solar system
+    barycenter) are either (1) the same with a small (Gaussian) dispersion $s$
+    or (2) independent.
+    In both cases, velocity is drawn from
+    the velocity prior $p(\vec{v})$ that we describe later.
 \end{itemize}
@@
 Under these assumptions, the likelihood of a proper motion measurement,
-$\bs{\mu}$, for a star given true distance, $d$, and true tangential velocity,
-$\vec{v}^\perp =
+$\bs{\mu}$, for a star with true distance, $d$, and true tangential velocity,
+$\vec{v}_t =
 (\begin{array}[t]{c c} v_\alpha & v_\delta\end{array})^\mathsf{T}$ is
 \begin{align}
-  L(\vec{v}, d, s^2) &=
-    \left[\det\left(\frac{\tilde{\mat{C}}^{-1}}{2\pi}\right)\right]^{1/2} \,
+  L(\vec{\mu} \given \vec{v}, d) &=
+    \left[\det\left(\frac{{\mat{C}}^{-1}}{2\pi}\right)\right]^{1/2} \,
     \exp \left[ -\frac{1}{2} \transp{\left(\vec{\mu} - \vec{x}_\theta \right)} \,
-    \tilde{\mat{C}}^{-1} \,
+    {\mat{C}}^{-1} \,
     \left(\vec{\mu} - \vec{x}_\theta \right) \right] \label{eq:likefn} \\
-  \vec{x}_\theta &= d^{-1} \, \vec{v}^\perp
+  \vec{x}_\theta &= d^{-1} \, \vec{v}_t
 \end{align}
-where the tangential velocity $\vec{v}^\perp$ is related to the 3-space velocity
-$\vec{v}$ through projection onto a tangent plane at a given star's sky position
+where the tangential velocity $\vec{v}_t$ is related to the 3D velocity
+$\vec{v}$ by projection matrix $\mat{M}$ at the star's sky position
 $(\alpha, \delta)$
 \begin{align}
-  \vec{v}^\perp &= \mat{M}\,\vec{v} \\
+  \vec{v}_t &= \mat{M}\,\vec{v} \\
   & = \left(
@@
 \end{align}
-and the modified covariance matrix $\tilde{\mat{C}}$ is
-\begin{equation}
-  \tilde{\mat{C}} = \mat{C} + s^2 \, \eye
-\end{equation}
+%and the modified covariance matrix $\tilde{\mat{C}}$ is
+%\begin{equation}
+%  \tilde{\mat{C}} = \mat{C} + s^2 \, \eye
+%\end{equation}
 
-We compute the fully marginalized likelihood (FML) that a given pair of stars
-has the same 3-space velocity with a small velocity difference drawn from a
-zero-mean Gaussian with variance $s^2$ (hypothesis 1, $\mathcal{L}_1$), and
-the FML of the stars having independent 3-space velocities (hypothesis 2,
-$\mathcal{L}_2$).
-We use the FML ratio $\mathcal{L}_1/\mathcal{L}_2$ as a scalar for selecting
-candidate co-moving stars, as described in $\sectionname~\ref{sub:selection}$ in
+For a given pair, we compute the fully marginalized likelihood (FML) for the hypotheses (1) and (2),
+$\mathcal{L}_1$ and $\mathcal{L}_2$.
+We use the FML ratio $\mathcal{L}_1/\mathcal{L}_2$ as the scalar quantity to select
+candidate co-moving pairs, as described in $\sectionname~\ref{sub:selection}$ in
 more detail.
 To compute these FMLs, the likelihood functions for each star in a pair, $L_i,
-L_j$, are marginalized over true 3-space velocity and distance for each star in
+L_j$, are marginalized over the (unknown) true 3D velocity and distance for each star in
 the pair $(i,j)$.
@@
 \end{align}
-The marginalization integral for hypothesis 2 can be split into the product of
+Here, $p(d_i \given \varpi_i)$ is the posterior distribution of distance given parallax measurement
+$\varpi_i$ and its Gaussian error $\sigma_{\varpi,i}$.
+Note that the FML for the hypothesis (1) involves integration over one velocity $\vec{v}$ that
+generates the likelihoods for both stars, $L_i$ and $L_j$.
+The marginalization integral for hypothesis (2) can be split into the product of
 two simpler integrals $\mathcal{L}_2 = Q_i \, Q_j$ where
 \begin{equation}
-  Q = \int \, \dd d \, \dd^3 \vec{v} \, L(\vec{v}, d, 0) \, p(\vec{v}) \, p(d\given\varpi)
+  Q_i = \int \, \dd d \, \dd^3 \vec{v} \, L(\vec{v}, d, 0) \, p(\vec{v}) \, p(d\given\varpi_i)
 \end{equation}
+
 If the velocity prior $p(\vec{v})$ is also Gaussian, the integrals over velocity
@@
 After marginalizing over velocity, the likelihood integrands only depend on
-distance; we numerically compute the integrals over the true distances to each
-star in a pair, $d_1,d_2$, using Monte Carlo integration with $K$ samples from
-the distance posterior pdfs assuming a uniform density prior
-\citep{Bailer-Jones:2015aa}
+distance; we numerically compute the integrals over the true distances of each
+star in a pair using Monte Carlo integration with $K$ samples from
+the distance posterior.
 \begin{equation}
@@
 where $\tilde{L}(d)$ is the velocity-marginalized likelihood function.
+In order to generate a sample of distances from the distance posterior $p(d_i \given \varpi_i)$,
+we need to assume a distance prior. We adopt the uniform density prior
+\citep{Bailer-Jones:2015aa} with a maximum distance of 1~kpc.
 Through experimentation, we have found that $K=128$ samples are sufficient for
@@
 signal-to-noise.
-%TODO: SMOH: still unsure about this
 
-For wide binaries, the assumption of the stars having the same 3D velocity can
+For wide binaries, the assumption of the stars having the same 3D velocity for 
+hypothesis (1) can
 break down for high precision proper motion measurements because of the orbital
 velocity (blue solid line in Figure~\ref{fig:dv-sep}).
-To account for this, we set $s^2 = \frac{2 \, G \, \msun}{|\vec{x}_i-\vec{x}_j|}$.
+To account for this, we modify set $s^2 = \frac{2 \, G \, \msun}{|\vec{x}_i-\vec{x}_j|}$.
 
@@
 is NGC~2632, another known open cluster of the Milky Way \citep{Kharchenko:2016aa}.
-%One particular star in this network, TYC 1799-272-1, has 94 directly connected
-%co-moving neighbors, and is located at the center of the network.
 
@@
   \begin{center}
-    % \includegraphics[width=\linewidth]{figures/raverv.pdf}
-    \includegraphics[width=.5\linewidth]{figures/raverv.pdf}
+    \includegraphics[width=0.6\linewidth]{figures/raverv.pdf}
   \end{center}

### 2016-12-02 paper/ms.tex
more text edits

@@
 those consistent with being drawn independently from the field
-population.  Although we perform some visualizations using a
-(bias-corrected) inverse parallax as a distance, the likelihood test
-works in the observable space and uses the \gaia\ noise model
-responsibly.
+population.  Although we perform some visualizations using
+(bias-corrected) inverse parallax as point-estimate of distance,
+the likelihood ratio test
+is based on a probabilistic model that includes the covariances
+of parallax and proper motions, and marginalizes the unknown
+distances and 3D velocities of stars.
 We find 13,085 pairs of co-moving stars among 10,606 unique stars
@@
 Some of these pairs form larger groups through mutual co-moving neighbors,
-many of which correspond to the known open clusters, and stars in OB associations.
+many of which correspond to the known open clusters, and OB associations.
 We also report many new co-moving groups of pairs.
@@
 The number of these very wide separation pairs increase with increasing
-separation; and cannot be explained purely by false-positives contamination.
+separation; and cannot be explained purely by false-positive contamination.
 We present a catalog of high-confidence co-moving pairs of stars in the \tgas\
@@
     In both cases, velocity is drawn from
-    the velocity prior $p(\vec{v})$ that we describe later.
+    the velocity prior $p(\vec{v})$.
 \end{itemize}
@@
 Under these assumptions, the likelihood of a proper motion measurement,
-$\bs{\mu}$, for a star with true distance, $d$, and true tangential velocity,
-$\vec{v}_t =
-(\begin{array}[t]{c c} v_\alpha & v_\delta\end{array})^\mathsf{T}$ is
+$\bs{\mu}$, for a star with true distance, $d$, and true 3D velocity $\vec{v}$ is
 \begin{align}
-  L(\vec{\mu} \given \vec{v}, d) &=
-    \left[\det\left(\frac{{\mat{C}}^{-1}}{2\pi}\right)\right]^{1/2} \,
+  L(\vec{\mu} \given \vec{v}, d, s^2) &=
+    \left[\det\left(\frac{\tilde{\mat{C}}^{-1}}{2\pi}\right)\right]^{1/2} \,
     \exp \left[ -\frac{1}{2} \transp{\left(\vec{\mu} - \vec{x}_\theta \right)} \,
-    {\mat{C}}^{-1} \,
+    \tilde{\mat{C}}^{-1} \,
     \left(\vec{\mu} - \vec{x}_\theta \right) \right] \label{eq:likefn} \\
@@
 \end{align}
-where the tangential velocity $\vec{v}_t$ is related to the 3D velocity
+where the tangential velocity $\vec{v}_t = (\begin{array}[t]{c c} v_\alpha & v_\delta\end{array})^\mathsf{T}$
+is related to the 3D velocity
 $\vec{v}$ by projection matrix $\mat{M}$ at the star's sky position
@@
 \end{align}
-%and the modified covariance matrix $\tilde{\mat{C}}$ is
-%\begin{equation}
-%  \tilde{\mat{C}} = \mat{C} + s^2 \, \eye
-%\end{equation}
+and the modified covariance matrix $\tilde{\mat{C}}$ is
+\begin{equation}
+  \tilde{\mat{C}} = \mat{C} + (s/d)^2 \, \eye
+\end{equation}
+The parameter $s$ is added to allow for small tolerance in velocities which we discuss below.
 
@@
 To compute these FMLs, the likelihood functions for each star in a pair, $L_i,
-L_j$, are marginalized over the (unknown) true 3D velocity and distance for each star in
+L_j$, are marginalized over the (unknown) true distance and 3D velocity for each star in
 the pair $(i,j)$.
@@
     \int \, \dd d_i \, \dd d_j \, \dd^3 \vec{v} \,
-    L_i(\vec{v}, d_i, s^2) \,
-    L_j(\vec{v}, d_j, s^2) \,
+    L_i(\vec{\mu_i} \given \vec{v}, d_i, s^2) \,
+    L_j(\vec{\mu_j} \given \vec{v}, d_j, s^2) \,
     p(\vec{v}) \, p(d_i \given \varpi_i) \, p(d_j \given \varpi_j) \\
@@
     \int \, \dd d_i \, \dd d_j \, \dd^3 \vec{v}_i \, \dd^3 \vec{v}_j \,
-    L_i(\vec{v}_i, d_i, 0) \,
-    L_j(\vec{v}_j, d_j, 0) \,
+    L_i(\vec{\mu_i} \given \vec{v}_i, d_i, s^2) \,
+    L_j(\vec{\mu_j} \given \vec{v}_j, d_j, s^2) \,
     p(\vec{v}_i) \, p(\vec{v}_j) \, p(d_i \given \varpi_i) \, p(d_j \given \varpi_j). \label{eq:hyp2}
@@
 The marginalization integral for hypothesis (2) can be split into the product of
-two simpler integrals $\mathcal{L}_2 = Q_i \, Q_j$ where
+two simpler integrals $\mathcal{L}_2 = Q(\vec{\mu_i}, \varpi_i) \, Q(\vec{\mu_j}, \varpi_j)$ where
 \begin{equation}
-  Q_i = \int \, \dd d \, \dd^3 \vec{v} \, L(\vec{v}, d, 0) \, p(\vec{v}) \, p(d\given\varpi_i)
+  Q(\vec{\mu}, \varpi) = \int \, \dd d \, \dd^3 \vec{v} \, L(\vec{\mu} \given \vec{v}, d, s^2) \,
+  p(\vec{v}) \, p(d \given \varpi)
 \end{equation}
@@
 velocity (blue solid line in Figure~\ref{fig:dv-sep}).
-To account for this, we modify set $s^2 = \frac{2 \, G \, \msun}{|\vec{x}_i-\vec{x}_j|}$.
+To account for this, we set $s^2 = \frac{2 \, G \, \msun}{|\vec{x}_i-\vec{x}_j|}$.
+Because $s^2$ is much smaller than the velocity dispersions of the velocity prior
+$p(\vec{v})$, it has minimal effect on the hypothesis (2).
 
@@
 $\ln \mathcal{L}_1 /\mathcal{L}_2$, and come up with a reasonable cut
-for this quantity to select co-moving pairs from the initial pairs sample.
+for this quantity to select co-moving pairs from the initial sample.
 
@@
     in slices of $\ln \mathcal{L}_1 /\mathcal{L}_2$ (sequentially color-coded in blue on the left).
-    The black dots (upper right) and lines (lower left) correspond to the entire initial sample of pairs.
+    The black dots in the upper right panel and
+    the black line in the lower lef panel correspond to the entire initial sample of pairs.
     For comparision, we also present the likelihood ratio distribution of random pairs
@@
 Figure~\ref{fig:likelihoodratios} shows the likelihood ratios for all
-$\sim 271$k initial pairs examined.
-As discussed in \sectionname~\ref{sec:data}, we expect a correlation between likelihood
+$\sim 271$k pairs in the initial sample.
+As discussed in \sectionname~\ref{sec:data}, we expect a correlation between the likelihood
 ratios of pairs, and their distribution on \absdvtan\ vs separation plane.
@@
     longitude, and the radius is $\hat d\cos b$ in pc where $\hat d$ is the point-estimate
-    of distance using Equation~\ref{eq:dist}.
+    of distance using \eqname~\ref{eq:dist}.
     Pairs in connected components of sizes less than 5 are connected by gray lines.
@@
 If we define a condition to associate a connected component to a known cluster
-as having more than 3 members within 10~pc of the nominal position of
-the cluster, for the 61 connected components with size larger than 5,
+as having more than 3 members within 10~pc from the nominal position of
+the cluster, for the 61 connected components with sizes larger than 5,
 we find that only 10 are associated with a cluster cataloged in
@@
 Finally, not all known star clusters are recovered in our search.
-This is primarily because of the non-uniform coverage of \tgas\ data.
+This is primarily because of the non-uniform coverage of the \tgas\ data.
 
@@
 Here, we use 210,368 cross-matches of the \tgas\ with
-the Radial Velocity Experiment (\rave) to
+Radial Velocity Experiment (\rave) to
 assess the false-positive rates of our selection.
@@
 (bottom panel of Figure~\ref{fig:raverv}).
-The nature of these very large separation, mutually exclusive pairs,
+The nature of these very wide separation, mutually exclusive pairs,
 which cannot be gravitationally bound to each other, needs further investigation.
@@
 to clearly detect this peak, but it shows such hint, and demonstrates
-that such large scale phase-space correlations may now be looked for and studied within the \gaia\ data.
+that such large scale phase-space correlations may now be looked for and studied using the \gaia\ data.
 
@@
 which will be valuable to anchoring stellar atmospheric models together.
-%TODO: SMOH: what else can i discuss?
 
@@
     \rave\ using the \gaia\ data archive.
-  \item physical separation between the two stars in pc
+  \item physical separation between the two stars in pc. Note that this is calculated
+    using the point-estimates of distances (\eqname~\ref{eq:dist}) of the two stars.
   \item likelihood ratio, $\ln \mathcal{L}_1 /\mathcal{L}_2$
@@
   \item distance to the closest cluster in pc
-  \item the ID of the OB association (A-L) if the star is in the catalog of
+  \item ID of the OB association (A-L) if the star is in the catalog of
   stars in the known OB associations by \citet{de-Zeeuw:1999aa}
@@
 We argued for a reasonable cut of the likelihood ratio, and found
-13,058 candidate pairs of co-moving stars among 10,606 unique stars
+13,058 candidate co-moving pairs among 10,606 stars
 with separation ranging from 0.005~pc to 10~pc, the limit of our search.
@@
 of each other, and not part of large connected components.
-These are most likely remnants of dissolving wide binaries and
-open clusters. If so, this population should still be relatively young compared
+These are most likely remnants of dissolving wide binaries.
+If so, this population should still be relatively young compared
 to the general disk field population. Modeling the color-magnitude diagram
 distribution of these stars can shed some light on this issue.
-If they are truly dissolving stars that were born coeval,
+If they are remnants of dissolving systems that were born coeval,
 the sample of very wide separation co-moving pairs can potentially be used
@@
 on the \emph{3D velocities} of the two stars in a pair, and that we
-marginalize over the (unknown) three components of the velocities 
-in contrast to just working in the proper motion space.
+marginalize over the (unknown) distances and velocities of the stars
+in contrast to just applying a cut in the proper motion space.
 

### 2016-12-04 paper/ms.tex
text edits

@@
 \begin{abstract}
-The primary sample of the \gaia\ \textsl{First Data Release} is the brand-new
-\textsl{Tycho-Gaia Astrometric Solution} (\tgas). The precision and size of this sample
-presents an opportunity to find new binary stars and moving groups, especially
-rare binaries or sparsely populated groups.
-Here, we look for high-confidence co-moving pairs of stars using this data.
-We use a justified marginalized likelihood ratio test to separate
-pairs of stars with similar 3D velocities from
-those consistent with being drawn independently from the field
-population.  Although we perform some visualizations using
-(bias-corrected) inverse parallax as point-estimate of distance,
-the likelihood ratio test
-is based on a probabilistic model that includes the covariances
-of parallax and proper motions, and marginalizes the unknown
-distances and 3D velocities of stars.
-We find 13,085 pairs of co-moving stars among 10,606 unique stars
-out to very wide separations up to 10~pc, the limit of our search.
-Some of these pairs form larger groups through mutual co-moving neighbors,
-many of which correspond to the known open clusters, and OB associations.
-We also report many new co-moving groups of pairs.
-More surprisingly, we find a large number
-of very wide ($>3$~pc) separation co-moving pairs of stars.
-The number of these very wide separation pairs increase with increasing
-separation; and cannot be explained purely by false-positive contamination.
-We present a catalog of high-confidence co-moving pairs of stars in the \tgas\
-with auxiliary information, and discuss its potential use
-for making dynamical inferences, and testing stellar models and chemical
-abundance measurements.
+The primary sample of the \gaia\ Data Release 1 is the
+\textsl{Tycho-Gaia Astrometric Solution} (\tgas): $\approx$ 2 million Tycho-2
+sources with improved parallaxes and proper motions relative to the initial
+catalog.
+This increased astrometric precision presents an opportunity to find new binary
+stars and moving groups.
+We search for high-confidence co-moving pairs of stars in \tgas\ by identifying
+pairs of stars consistent with having the same 3D velocity using a marginalized
+likelihood ratio test to discriminate candidate co-moving pairs from the field
+population.
+Although we perform some visualizations using (bias-corrected) inverse-parallax
+as a point-estimate of distance, the likelihood ratio is computed with a
+probabilistic model that includes the covariances of parallax and proper motions
+and marginalizes the (unknown) true distances and 3D velocities of the stars.
+We find 13,085 co-moving star pairs among 10,606 unique stars with separations
+as large as 10 pc (our search limit).
+Some of these pairs form larger groups through mutual co-moving neighbors:
+many of these pair networks correspond to known open clusters and OB
+associations, but we also report the discovery of several new co-moving groups.
+Most surprisingly, we find a large number of very wide ($>1$ pc) separation
+co-moving star pairs, the number of which increases with increasing separation
+and cannot be explained purely by false-positive contamination.
+Our key result is a catalog of high-confidence co-moving pairs of stars in
+\tgas.
+We discuss the utility of this catalog for making dynamical inferences about the
+Galaxy, testing stellar atmosphere models, and validating chemical abundance
+measurements.
 \end{abstract}
@@
 
-Stars that are roughly co-located, and moving with similar space velocities
+Stars that are roughly co-located and moving with similar space velocities
 (``co-moving stars'') are of special interest in many branches of astrophysics.
 
-At small separations ($0.001-1$~pc), they are wide binaries (and
+At small separations ($0.001$--$1$~pc), they are wide binaries (and
 multiples) that are either weakly gravitationally bound or
@@
 
-Wide binaries are also good testbeds for stellar evolution models.
-For example, main sequence-red giant wide binaries are especially valuable in
-testing consistency between stellar atmosphere models, as the
-two stars were likely born at the same time with the same chemical compositions,
-but evolved independently because of their wide separation.
+Wide binaries are also good testbeds for stellar models and age indicators:
+the constituent stars were likely born at the same time with the same chemical
+compositions, but evolved independently because of their wide separation.
+These pairs are therefore useful for validating gyrochronology relations
+(e.g., \citealt{Chaname:2012}) and may be valuable for testing consistency
+between stellar atmosphere models.
 Finally, calibration of stellar parameters of low-mass stars (e.g., M dwarfs),
@@
 
-At larger separation ($\gtrsim 1$~pc), co-moving stars are likely members of
-moving groups, associations, and star clusters.
+At larger separations ($\gtrsim 1$~pc), co-moving stars are likely members of
+(potentially dissolving) moving groups, associations, and star clusters or
+disrupted wide binaries.
 The origin of moving groups is still under active debate (e.g., \citealt{Bovy:2010aa}:
@@
 In the halo, we know of more than 20 disrupting globular clusters and dwarf galaxies
-(``streams''; see, e.g., \citealt{Grillmair:2016} for a summary of known streams).
+(``stellar streams''; see, e.g., \citealt{Grillmair:2016} for a summary of known streams).
 These tidal streams are modeled to infer the parameters of the Galactic
@@
 separation less than 10~pc and difference
-in tangential velocity less than $\absdvtan < 10$~\kms.
+in (point-estimate) tangential velocity less than $\absdvtan < 10$~\kms.
 We ultimately build a statistical model that incorporates the covariances of the
@@
 
-  \item  A sample of co-moving stars also include stars that are part of e.g.,
+  \item  A sample of co-moving stars also include stars that are part of, e.g.,
     OB associations, moving groups, and open clusters.
@@
 \begin{equation}
-  p(\vec{v}) = \sum_{m=1}^3 \, w_m \mathcal{N}(0, \sigma_{v,m}^2)
+  p(\vec{v}) = \sum_{m=1}^3 \, w_m \, \mathcal{N}(0, \sigma_{v,m}^2)
 \end{equation}
@@
 
-For wide binaries, the assumption of the stars having the same 3D velocity for 
+For small-separation binaries, the assumption that the stars having the same 3D velocity for
 hypothesis (1) can
-break down for high precision proper motion measurements because of the orbital
+break down for high-precision proper motion measurements because of the orbital
 velocity (blue solid line in Figure~\ref{fig:dv-sep}).
@@
 Because $s^2$ is much smaller than the velocity dispersions of the velocity prior
-$p(\vec{v})$, it has minimal effect on the hypothesis (2).
+$p(\vec{v})$, it has minimal effect on the hypothesis (2) FML.
 
@@
 
-Based on these comparisons, we select pairs with
-$\ln \mathcal{L}_1 /\mathcal{L}_2 > 6$ as candidate co-moving pair.
+Based on these comparisons, we select candidate co-moving pairs with
+$\ln \mathcal{L}_1 /\mathcal{L}_2 > 6$.
 Out of 271,232 pairs in the initial sample, 13,058 pairs (4.8\%)
@@
 
-\subsection{Statistics, properties, and nature of co-moving pairs}
+\subsection{Statistics and properties of the identified co-moving pairs}
 
@@
 In 13,058 co-moving pairs that we identified,
-there are 4555 connected components among 10,606 unique stars.
+there are 4,555 connected components among 10,606 unique stars.
 The maximum size of the connected components is 151.
@@
 The largest connected component corresponds to the Pleiades open cluster (left panel of Figure~\ref{fig:graphviz_examples})
-while the one on the upper left panel of the right column of Figure~\ref{fig:graphviz_examples}
-is NGC~2632, another known open cluster of the Milky Way \citep{Kharchenko:2016aa}.
+while the upper left panel of the right column of Figure~\ref{fig:graphviz_examples}
+is NGC~2632, another known Milky Way open cluster \citep{Kharchenko:2016aa}.
 
@@
 Finally, not all known star clusters are recovered in our search.
-This is primarily because of the non-uniform coverage of the \tgas\ data.
+This is primarily because of the non-uniform coverage and magnitude limit of the
+\tgas\ data.
 
@@
 relative velocities ($\sim 0.1$~\kms).
-Our search which stops on the maximum radius of 10~pc is obviously insufficient
-to clearly detect this peak, but it shows such hint, and demonstrates
-that such large scale phase-space correlations may now be looked for and studied using the \gaia\ data.
+In future work, we will increase the maximum search limit (in this work, 10 pc)
+to identify and study these large scale phase-space correlations.
 
@@
 
-In this \documentname, we looked for co-moving pairs of stars in the
-\tgas\ data released as part of the \gaia\ DR1.
-Our method is to compare
-the fully marginalized likelihoods between the two hypotheses, that a pair of
-stars shares the same 3D velocity, and that the two stars have two independent
-3D velocities, incorporating the covariances of parallax and proper motions.
+In this \documentname, we searched for co-moving pairs of stars in the
+\tgas\ catalog released as part of the \gaia\ DR1.
+Our method is to compare the fully marginalized likelihoods between two
+hypotheses: (1) that a pair of
+stars shares the same 3D velocity, and (2) that the two stars have independent
+3D velocities, in both cases incorporating the covariances of parallax and
+proper motions.
 We argued for a reasonable cut of the likelihood ratio, and found
 13,058 candidate co-moving pairs among 10,606 stars
-with separation ranging from 0.005~pc to 10~pc, the limit of our search.
+with separations ranging from 0.005~pc to 10~pc, the limit of our search.
 
@@
 by a path.
-The entire 13,058 candidate co-moving pairs are grouped into 4555 connected
-components. The most common is a size-2 connected component i.e., the two
+The entire 13,058 candidate co-moving pairs are grouped into 4,555 connected
+components. The most common is a size-2 connected component, i.e., the two
 stars in these pairs are mutually exclusively linked.
@@
 
-We have also found a surprisingly large number of very wide separation ($>3$~pc)
+We have also found a surprisingly large number of very wide separation ($>1$~pc)
 mutually exclusive co-moving pairs, in which the stars are the only co-moving neighbor
-of each other, and not part of large connected components.
+of each other and not part of large connected components.
 These are most likely remnants of dissolving wide binaries.
-If so, this population should still be relatively young compared
-to the general disk field population. Modeling the color-magnitude diagram
-distribution of these stars can shed some light on this issue.
+If confirmed with radial velocity measurements, this population should still be
+relatively young compared to the general disk field population.
+Modeling the color-magnitude diagram distribution of these stars can shed some
+light on this issue.
 If they are remnants of dissolving systems that were born coeval,
@@
 We note that a similar search for wide binaries using the \tgas\ data is performed
-in a recent work by \citet{2016arXiv161107883O}. They find $\approx 1900$ wide binaries
+in a recent work by \citet{2016arXiv161107883O}. They find $\approx$1,900 wide binaries
 with separation typically less than 1.5~pc, and 256 pairs with separation larger than
@@
 on the \emph{3D velocities} of the two stars in a pair, and that we
-marginalize over the (unknown) distances and velocities of the stars
+marginalize over the (unknown) true distances and velocities of the stars
 in contrast to just applying a cut in the proper motion space.
@@
 What we find using the \tgas\ is only a taste of what we will discover with
-the future releases of \gaia.
+the future releases of the \gaia\ mission.
 

### 2016-12-04 scripts/make-catalog.py
add machine readable table

@@ -1,0 +1,143 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+
+"""
+File: make-catalog.py
+Author: Semyeong Oh
+Github: https://github.com/smoh
+Description: Make the final catalog of co-moving pairs
+"""
+
+import numpy as np
+import networkx as nx
+import h5py
+from astropy.io import fits
+from astropy import table
+from astropy import coordinates as coords
+import astropy.units as u
+
+import gwb
+
+tgas = gwb.TGASData('data/stacked_tgas.fits')
+pairidx = fits.getdata('output/23560/snr8_r10_dv10.fits')
+with h5py.File("output/23560/snr8_r10_dv10_vscatter0-lratio.h5") as f:
+    lnH1 = f['lnH1'].value
+    lnH2 = f['lnH2'].value
+    llr = lnH1 - lnH2
+print('number of pairs', pairidx.size)
+
+parallax_snr = tgas.parallax_snr
+vtan = tgas.get_vtan().value
+c = tgas.get_coord()
+d = tgas.get_distance().value
+
+star1, star2 = pairidx['star1'], pairidx['star2']
+min_snr = np.min(np.vstack((parallax_snr[star1], parallax_snr[star2])), axis=0)
+dvtan = np.linalg.norm(vtan[star1]-vtan[star2], axis=1)
+vtanmean = (vtan[star1] + vtan[star2])*0.5
+sep = c[star1].separation_3d(c[star2]).value
+sep_sky = c[star1].separation(c[star2])
+
+c1 = c[star1]
+c2 = c[star2]
+ra1, dec1 = c1.ra.value, c1.dec.value
+ra2, dec2 = c2.ra.value, c2.dec.value
+l1, b1 = c1.transform_to(coords.Galactic).l.value, c1.transform_to(coords.Galactic).b.value
+l2, b2 = c2.transform_to(coords.Galactic).l.value, c2.transform_to(coords.Galactic).b.value
+d1 = d[star1]
+d2 = d[star2]
+dmean = (d1+d2)*0.5
+
+cond_lr_cut = llr>6
+print('number of pairs ln(L1/L2)>6 = %i' % (cond_lr_cut.sum()))
+print('number of pairs ln(L1/L2)>6 + sep<1pc = %i' % (sum((sep<1)&cond_lr_cut)))
+cmpairs = pairidx[cond_lr_cut]
+
+# sorted list of subgraphs from largest to smallest
[…]

### 2016-12-05 paper/ms.tex
final edits
- thickened lines in histogram figures
- lowererd alpha for the initial sample plot

@@
 as a point-estimate of distance, the likelihood ratio is computed with a
-probabilistic model that includes the covariances of parallax and proper motions
+probabilistic model that includes the covariances of parallax and proper motions,
 and marginalizes the (unknown) true distances and 3D velocities of the stars.
@@
 disrupted wide binaries.
-The origin of moving groups is still under active debate (e.g., \citealt{Bovy:2010aa}:
+The origin of moving groups is still under active debate (e.g., \citealt{Bovy:2010aa}):
 are they remnants of a coeval

### 2016-12-05 setup.py
repo name change

@@ -29,7 +29,7 @@ setup(
     author="Semyeong Oh, Adrian Price-Whelan, Timothy D. Morton, David W. Hogg",
     author_email="semyeong@astro.princeton.edu, adrn@princeton.edu",
     packages=["gwb", "gwb.tests"],
-    url="https://github.com/smoh/gaia-wide-binaries",
+    url="https://github.com/smoh/gaia-comoving-stars",
     license="MIT",
     description="Gaia wide binaries.",
     long_description=rd("README.md"),

### 2017-01-03 paper/ms.tex
change repo name
Hack to have subdir in repo but not for submission
deal with some of the comments

@@
 \hypersetup{backref,breaklinks,colorlinks,urlcolor=cbblue,linkcolor=cbblue,citecolor=black}
+\graphicspath{{figures/}}
 
@@
 \newcommand{\tmass}{\project{\acronym{2MASS}}}
-\newcommand{\documentname}{\textsl{Article}}
+\newcommand{\documentname}{Article}
 \newcommand{\sectionname}{Section}
@@
   \begin{center}
-    \includegraphics[width=\linewidth]{figures/sep_dvtan.pdf}
+    \includegraphics[width=\linewidth]{sep_dvtan.pdf}
   \end{center}
@@
 \end{equation}
+where $\eye$ is the identity matrix.
 The parameter $s$ is added to allow for small tolerance in velocities which we discuss below.
@@
   \begin{center}
-    \includegraphics[width=\textwidth]{figures/likelihoodratios.pdf}
+    \includegraphics[width=\textwidth]{likelihoodratios.pdf}
   \end{center}
@@
   \begin{center}
-    \includegraphics[width=\linewidth]{figures/dist_networksize.pdf}
+    \includegraphics[width=\linewidth]{dist_networksize.pdf}
   \end{center}
@@
   \begin{center}
-    \includegraphics[width=\textwidth]{figures/graphviz_examples.pdf}
+    \includegraphics[width=\textwidth]{graphviz_examples.pdf}
   \end{center}
@@
   \begin{center}
-    \includegraphics[width=\textwidth]{figures/glon_d_pie.pdf}
+    \includegraphics[width=\textwidth]{glon_d_pie.pdf}
   \end{center}
@@
 Here, we use 210,368 cross-matches of the \tgas\ with
-Radial Velocity Experiment (\rave) to
+Radial Velocity Experiment fifth data release (\rave DR~5; \citealt{2016arXiv160903210K}) to
 assess the false-positive rates of our selection.
@@
   \begin{center}
-    \includegraphics[width=0.6\linewidth]{figures/raverv.pdf}
+    \includegraphics[width=0.6\linewidth]{raverv.pdf}
   \end{center}
@@
   \begin{center}
-    \includegraphics[width=\textwidth]{figures/hist_sep.pdf}
+    \includegraphics[width=\textwidth]{hist_sep.pdf}
   \end{center}
@@
   \begin{center}
-    \includegraphics[width=\textwidth]{figures/gjg_graphviz_examples.pdf}
+    \includegraphics[width=\textwidth]{gjg_graphviz_examples.pdf}
   \end{center}
@@
   \begin{center}
-    \includegraphics[width=\textwidth]{figures/gjg_mepairs.pdf}
+    \includegraphics[width=\textwidth]{gjg_mepairs.pdf}
   \end{center}
@@
 \software{The code used in this project is available from
-\url{https://github.com/smoh/gaia-wide-binaries} under the MIT open-source
+\url{https://github.com/smoh/gaia-comoving-stars} under the MIT open-source
 software license. This version was generated at git commit

### 2017-01-17 paper/ms.tex
change distance symbol from d to r

@@
 \begin{equation}
-  \hat{d} = 1000 \, \left[\frac{\varpi}{2} \,
+  \hat{r} = 1000 \, \left[\frac{\varpi}{2} \,
     \left(1 + \sqrt{1 - \frac{16}{\snr^2}} \right) \right]^{-1} \, {\rm pc}
@@
 \begin{equation}
-  \absdvtan = |\hat{d}_1 \vec\mu_1 - \hat{d}_2 \vec\mu_2|
+  \absdvtan = |\hat{r}_1 \vec\mu_1 - \hat{r}_2 \vec\mu_2|
 \end{equation}
@@
 Under these assumptions, the likelihood of a proper motion measurement,
-$\bs{\mu}$, for a star with true distance, $d$, and true 3D velocity $\vec{v}$ is
+$\bs{\mu}$, for a star with true distance, $r$, and true 3D velocity $\vec{v}$ is
 \begin{align}
-  L(\vec{\mu} \given \vec{v}, d, s^2) &=
+  L(\vec{\mu} \given \vec{v}, r, s^2) &=
     \left[\det\left(\frac{\tilde{\mat{C}}^{-1}}{2\pi}\right)\right]^{1/2} \,
@@
     \left(\vec{\mu} - \vec{x}_\theta \right) \right] \label{eq:likefn} \\
-  \vec{x}_\theta &= d^{-1} \, \vec{v}_t
+  \vec{x}_\theta &= r^{-1} \, \vec{v}_t
 \end{align}
@@
 \begin{equation}
-  \tilde{\mat{C}} = \mat{C} + (s/d)^2 \, \eye
+  \tilde{\mat{C}} = \mat{C} + (s/r)^2 \, \eye
 \end{equation}
@@
   \mathcal{L}_1 &=
-    \int \, \dd d_i \, \dd d_j \, \dd^3 \vec{v} \,
-    L_i(\vec{\mu_i} \given \vec{v}, d_i, s^2) \,
-    L_j(\vec{\mu_j} \given \vec{v}, d_j, s^2) \,
-    p(\vec{v}) \, p(d_i \given \varpi_i) \, p(d_j \given \varpi_j) \\
+    \int \, \dd r_i \, \dd r_j \, \dd^3 \vec{v} \,
+    L_i(\vec{\mu_i} \given \vec{v}, r_i, s^2) \,
+    L_j(\vec{\mu_j} \given \vec{v}, r_j, s^2) \,
+    p(\vec{v}) \, p(r_i \given \varpi_i) \, p(r_j \given \varpi_j) \\
   \mathcal{L}_2 &=
-    \int \, \dd d_i \, \dd d_j \, \dd^3 \vec{v}_i \, \dd^3 \vec{v}_j \,
-    L_i(\vec{\mu_i} \given \vec{v}_i, d_i, s^2) \,
-    L_j(\vec{\mu_j} \given \vec{v}_j, d_j, s^2) \,
-    p(\vec{v}_i) \, p(\vec{v}_j) \, p(d_i \given \varpi_i) \, p(d_j \given \varpi_j). \label{eq:hyp2}
+    \int \, \dd r_i \, \dd r_j \, \dd^3 \vec{v}_i \, \dd^3 \vec{v}_j \,
+    L_i(\vec{\mu_i} \given \vec{v}_i, r_i, s^2) \,
+    L_j(\vec{\mu_j} \given \vec{v}_j, r_j, s^2) \,
+    p(\vec{v}_i) \, p(\vec{v}_j) \, p(r_i \given \varpi_i) \, p(r_j \given \varpi_j). \label{eq:hyp2}
 \end{align}
-Here, $p(d_i \given \varpi_i)$ is the posterior distribution of distance given parallax measurement
+Here, $p(r_i \given \varpi_i)$ is the posterior distribution of distance given parallax measurement
 $\varpi_i$ and its Gaussian error $\sigma_{\varpi,i}$.
@@
 \begin{equation}
-  Q(\vec{\mu}, \varpi) = \int \, \dd d \, \dd^3 \vec{v} \, L(\vec{\mu} \given \vec{v}, d, s^2) \,
-  p(\vec{v}) \, p(d \given \varpi)
+  Q(\vec{\mu}, \varpi) = \int \, \dd r \, \dd^3 \vec{v} \, L(\vec{\mu} \given \vec{v}, r, s^2) \,
+  p(\vec{v}) \, p(r \given \varpi)
 \end{equation}
@@
 \begin{equation}
-  \int \, \dd d \, \tilde{L}(d) \, p(d\given\varpi) \approx
-    \frac{1}{K} \, \sum_k^K \, \tilde{L}(d_k)
+  \int \, \dd r \, \tilde{L}(r) \, p(r \given \varpi) \approx
+    \frac{1}{K} \, \sum_k^K \, \tilde{L}(r_k)
 \end{equation}
-where $\tilde{L}(d)$ is the velocity-marginalized likelihood function.
-In order to generate a sample of distances from the distance posterior $p(d_i \given \varpi_i)$,
+where $\tilde{L}(r)$ is the velocity-marginalized likelihood function.
+In order to generate a sample of distances from the distance posterior $p(r_i \given \varpi_i)$,
 we need to assume a distance prior. We adopt the uniform density prior
@@
 
-At given distance $d$, the velocity-marginalized likelihood can be computed
+At given distance $r$, the velocity-marginalized likelihood can be computed
 analytically using the expressions in Appendix~\ref{sec:appendixA}.
@@
       \begin{array}{c c c c}
-        d_i\,\mu_{\alpha,i} &
-        d_i\,\mu_{\delta,i} &
-        d_j\,\mu_{\alpha,j} &
-        d_j\,\mu_{\delta,j}
+        r_i\,\mu_{\alpha,i} &
+        r_i\,\mu_{\delta,i} &
+        r_j\,\mu_{\alpha,j} &
+        r_j\,\mu_{\delta,j}
       \end{array}
@@
 and we have multiplied the observables (the proper motions) by the distances
-$d_i, d_j$, which is permitted because we are conditioning on the distances.
+$r_i, r_j$, which is permitted because we are conditioning on the distances.
 Fundamentally, our hypothesis 1 model (the stars have the same velocity with a
@@
 covariance matrix of the observables, $\mat{C}_i, \mat{C}_j$, and the distances
-$d_i, d_j$:
+$r_i, r_j$:
 \begin{equation}
@@
     \begin{array}{c c}
-      d_i^2 \, \mat{C}_i & 0 \\
-      0 & d_j^2 \, \mat{C}_j
+      r_i^2 \, \mat{C}_i & 0 \\
+      0 & r_j^2 \, \mat{C}_j
     \end{array}
@@
 \begin{eqnarray}
-  p(\data \given \vec{v}, d_i, d_j) &=& d_i^2\,d_j^2\,
+  p(\data \given \vec{v}, r_i, r_j) &=& r_i^2\,r_j^2\,
     \normal(\vec{y} \given \mat{M}\,\vec{v}, \mat{\Sigma}) \\
-  \ln p(\data \given \vec{v}, d_i, d_j) &=& 2\,\ln d_i + 2\,\ln d_j
+  \ln p(\data \given \vec{v}, r_i, r_j) &=& 2\,\ln r_i + 2\,\ln r_j
     -\frac{1}{2}\,\ln||2\pi\,\mat{\Sigma}|| \nonumber \\
@@
 \end{eqnarray}
-where the factor of $d_i^2\,d_j^2$ converts to units of one over \gaia\ data
+where the factor of $r_i^2\,r_j^2$ converts to units of one over \gaia\ data
 from units of one over $y$, which includes the distances.
@@
   \\
-  \Delta &=& -2\,\ln d_i -2\,\ln d_j
+  \Delta &=& -2\,\ln r_i -2\,\ln r_j
     +\frac{1}{2}\,\ln||2\pi\,\mat{\Sigma}|| +\frac{1}{2}\,\ln||2\pi\,\mat{V}_m|| \nonumber \\
@@
 which we plug in to \eqname~\ref{eq:gauss-int} to get the marginalized
-likelihood conditioned on the two distances $d_i, d_j$.
+likelihood conditioned on the two distances $r_i, r_j$.
 
@@
       \begin{array}{c c}
-        d\,\mu_{\alpha} &
-        d\,\mu_{\delta}
+        r\,\mu_{\alpha} &
+        r\,\mu_{\delta}
       \end{array}
@@
   \\
-  \mat{\Sigma} &=& d^2 \, \mat{C}
+  \mat{\Sigma} &=& r^2 \, \mat{C}
   \\

### 2017-02-01 paper/ms.tex
Work on catalog improvment

@@
 \usepackage{amsmath}
+\usepackage{tikz}
 \hypersetup{backref,breaklinks,colorlinks,urlcolor=cbblue,linkcolor=cbblue,citecolor=black}
@@
     L_j(\vec{\mu_j} \given \vec{v}, r_j, s^2) \,
-    p(\vec{v}) \, p(r_i \given \varpi_i) \, p(r_j \given \varpi_j) \\
+    p(\vec{v}) \, p(r_i \given \varpi_i) \, p(r_j \given \varpi_j) \label{eq:hyp1}\\
   \mathcal{L}_2 &=
@@
 meant to encompass young thin disk stars to halo stars.
+These numbers are empirically chosen to account for the distribution of one
+component velocities of the \tgas\ stars.
 We derive the relevant expressions in Appendix~\ref{sec:appendix}.
@@
 Here, we use 210,368 cross-matches of the \tgas\ with
-Radial Velocity Experiment fifth data release (\rave DR~5; \citealt{2016arXiv160903210K}) to
+Radial Velocity Experiment fifth data release (\rave\ DR~5; \citealt{2016arXiv160903210K}) to
 assess the false-positive rates of our selection.
@@
 In this section, we describe our catalog of candidate co-moving pairs of stars.
-For 13,058 pairs (10,606 unique stars) found in this work,
-we provide the following information:
+For 13,058 pairs (10,606 unique stars) found in this work.
+The catalog is composed of three tables of stars, pairs, and groups.
+We summarize the content of each table in Table~\ref{tab:table_meta}, and
+the relationships between the tables in Figure~\ref{fig:schema}.
+The star table contains all 10,606 stars that has at least one co-moving neighbor
+by our selection. We provide the TGAS source id, which may be used to easily
+retrieve cross matches of \gaia\ with other surveys
+using the \gaia\ data archive (\todo{cite}).
+For each star, we also include the positional measurements from \tgas,
+\gaia\ $G$-band and \tmass\ $J$, $H$-band magnitudes as well as
+\rave\ radial velocities.
+The co-moving relationship between the stars is described in the pair table.
+We also list the angular and physical separation of pairs, and the likelihood ratio,
+$\ln \mathcal{L}_1 /\mathcal{L}_2>6$ (see \eqname~\ref{eq:hyp1} and \ref{eq:hyp2})
+computed in this work.
+
+Finally, the information about the connected components found in these co-moving
+star graph is in the group table. We assign a unique index to each group
+in descending order of its size. Thus, group 0 is the largest group
+that contains 151 stars.
+Each star or pair is associated with a group that it is a member of, listed
+in the group id column of the star and pair table.
+Additionally, given the strong correlation between the larger connected 
+components of co-moving pairs and the known star clusters,
+we provided
 %
 \begin{itemize}
-  \item \tgas\ source id for the two stars in a pair; these may be used to easily
-    retrieve cross matches of \gaia\ with other surveys such as WISE, \tmass, and
-    \rave\ using the \gaia\ data archive.
-  \item physical separation between the two stars in pc. Note that this is calculated
-    using the point-estimates of distances (\eqname~\ref{eq:dist}) of the two stars.
-  \item likelihood ratio, $\ln \mathcal{L}_1 /\mathcal{L}_2$
-  \item size of the connected component that the pair belongs to
-  \item unique ID of the connected component that the pair belongs to; we assigned
-    a unique ID in descending order of size, thus ID~0 is the largest connected component with 151 stars that corresponds to the Pleiades.
-\end{itemize}
-%
-Additionally, given the strong correlation between the larger connected components
-of co-moving pairs
-and the known star clusters and OB associations, we provide for each star
-%
-\begin{itemize}
-  \item index of the closest cluster in the Milky Way Star Cluster catalog of
+  \item index and name of the closest cluster in the Milky Way Star Cluster catalog of
   \citet{Kharchenko:2016aa} using 3D separation
   \item distance to the closest cluster in pc
-  \item ID of the OB association (A-L) if the star is in the catalog of
-  stars in the known OB associations by \citet{de-Zeeuw:1999aa}
+  % \item ID of the OB association (A-L) if the star is in the catalog of
+  % stars in the known OB associations by \citet{de-Zeeuw:1999aa}
 \end{itemize}
-%
-A part of the entire table is presented in Table~\ref{tab:catalog}.
-
-\input{stub}
+
+% The distances and physical separations are calculated using the point-estimate
+% of distances (\eqname~\ref{eq:dist}).
+
+% \input{stub}
+\input{table_meta}
 
@@
 the future releases of the \gaia\ mission.
+
+\begin{figure}[htb]
+  \centering
+  \input{schema.tex}
+  \caption{Schematic diagram of the relationship between the tables.}
+  \label{fig:schema}
+\end{figure}
 

### 2017-02-01 paper/table_meta.tex
Work on catalog improvment

\begin{table}[htb]
\centering
\caption{Candidate co-moving pairs catalog description} \label{tab:table_meta}
\begin{tabular}{l|l|l}
\hline\hline
Column                                & Unit   & Description                         \\
\hline
\multicolumn{3}{c}{Table: Star (10,606 rows)}                                         \\
\hline
row id                                &        & zero-based row index                \\
TGAS source id                        &        & unique source id from \tgas          \\
RA                         & deg       & right ascension from \tgas          \\
DEC                        & deg       & declination from \tgas          \\
parallax                       & mas     & unique source id from \tgas          \\
distance                       & pc     & unique source id from \tgas          \\
$G$                               & mag    & \gaia\ $G$-band magnitudes            \\
$J$                               & mag    & \tmass\ $J$-band magnitudes            \\
$H$                               & mag    & \tmass\ $H$-band magnitudes            \\
RAVE OBS ID                           &        & unique id of the \rave\ match            \\
RV                                    & \kms   & radial velocity from \rave\           \\
eRV                                   & \kms   & uncertainty of radial velocity from \rave            \\
group id & & Id of the group this star belongs to\\
\hline
\multicolumn{3}{c}{Table: Pair (13,058 rows)}                                            \\
\hline
angsep                                & arcmin & angular separation                  \\
separation                            & pc     & physical separation                 \\
$\ln\mathcal{L}_1/\mathcal{L}_2$ &        & likelihood ratio                    \\
group id                              &        & id of the group the pair belongs to \\
\hline
\multicolumn{3}{c}{Table: Group (4,555 rows)}                                         \\
\hline
id                                    &        & unique group id                     \\
size                                  &        & number of stars in a group          \\
                                      &        &                                     \\
\hline\hline
\end{tabular}
\end{table}

### 2017-02-03 paper/ms.tex
more work on comments

@@
 \end{equation}
-where $\vec\mu = (\mu_\alpha, \mu_\delta)$.
+where $\vec\mu = (\mu_{\alpha^*}, \mu_\delta)$.
+\footnote{$\mu_{\alpha^*}$ is proper motion
+in right ascension direction, $\mu_{\alpha^*} = \mu_\alpha \cos\delta$}
 
@@
   \item We assume that the uncertainties in the data---parallax, $\varpi$, and
-    two proper motion components, $\propm = (\begin{array}[t]{c c} \mu_\alpha &
+    two proper motion components, $\propm = (\begin{array}[t]{c c} \mu_{\alpha^*} &
     \mu_\delta\end{array})^\mathsf{T}$---are Gaussian with known covariances

### 2017-02-03 paper/ms.tex
figure 5 updated

@@
 searching for stars with common proper motions
-(\citealt{Poveda:1994aa,Allen:2000aa,Chaname:2004aa,Lepine:2007aa,Alonso-Floriano:2015aa}).
+(\citealt{Poveda:1994aa,Allen:2000aa,Chaname:2004aa,
+Lepine:2007aa,Shaya:2011aa,Alonso-Floriano:2015aa}).
 Here, we use the recent first data release of \gaia\ which includes precise distances,
@@
   \caption{%
-    Panoramic view of co-moving pairs of stars. This is a cylindrical
-    projection of onto the Galactic plane. The angle is the Galactic
-    longitude, and the radius is $\hat d\cos b$ in pc where $\hat d$ is the point-estimate
-    of distance using \eqname~\ref{eq:dist}.
+    Panoramic view of co-moving pairs of stars around the Sun. This is a cylindrical
+    projection of onto the Galactic plane with the Sun at the origin.
     Pairs in connected components of sizes less than 5 are connected by gray lines.
@@
     We also show the positions of known Milky Way star clusters
-    from \citet{Kharchenko:2016aa} as light blue circles,
+    from \citet{Kharchenko:2016aa} as light blue circles of $10$~pc radius,
     and stars in OB associations from \citet{de-Zeeuw:1999aa} as
-    gray circles.
+    blue crosses.
     Some of the larger connected components are clearly associated with known
@@
 retrieve cross matches of \gaia\ with other surveys
-using the \gaia\ data archive (\todo{cite}).
+using the \gaia\ data archive.
 For each star, we also include the positional measurements from \tgas,

### 2017-02-14 paper/table_meta.tex
new catalog

@@
 \hline\hline
-Column                                & Unit   & Description                         \\
+Column Name    & Unit & Description                         \\
 \hline
@@
 eRV            & \kms & uncertainty of radial velocity from \rave \\
-group id & & Id of the group this star belongs to\\
+group id       &      & Id of the group this star belongs to      \\
 \hline

### 2017-02-14 gwb/mle.py

@@ -1,4 +1,4 @@
-"""codes related to maximum likelihood esimation"""
+"""codes related to maximum likelihood estimation"""
 import numpy as np
 from numpy import pi, log, deg2rad, rad2deg
 from astropy import units as u
@@ -156,13 +156,15 @@ class FitMCMC_samev(object):
         self.samples_d1 = np.vstack([
             1./d1,
             vra1/d1/4.74,
-            vdec1/d1/4.74]).T
+            vdec1/d1/4.74,
+            vr1]).T
         A2 = get_A(star2)
         vra2, vdec2, vr2 = A2.dot(samples[:,2:].T)
         self.samples_d2 = np.vstack([
             1./d2,
             vra2/d2/4.74,
-            vdec2/d2/4.74]).T
+            vdec2/d2/4.74,
+            vr2]).T
 
 class FitMCMCn_samev(object):
     ndim = None

### 2017-02-14 scripts/make-catalog.py

@@ -4,7 +4,6 @@
 """
 File: make-catalog.py
 Author: Semyeong Oh
-Github: https://github.com/smoh
 Description: Make the final catalog of co-moving pairs
 """
 
@@ -18,126 +17,96 @@ import astropy.units as u
 
 import gwb
 
-tgas = gwb.TGASData('data/stacked_tgas.fits')
-pairidx = fits.getdata('output/23560/snr8_r10_dv10.fits')
-with h5py.File("output/23560/snr8_r10_dv10_vscatter0-lratio.h5") as f:
-    lnH1 = f['lnH1'].value
-    lnH2 = f['lnH2'].value
-    llr = lnH1 - lnH2
-print('number of pairs', pairidx.size)
+if 'data_loaded' not in dir():
+    data_loaded = True
+    tgas = gwb.TGASData('data/stacked_tgas.fits')
+    pairidx = fits.getdata('output/23560/snr8_r10_dv10.fits')
+    with h5py.File("output/23560/snr8_r10_dv10_vscatter0-lratio.h5") as f:
+        lnH1 = f['lnH1'].value
+        lnH2 = f['lnH2'].value
+        llr = lnH1 - lnH2
+    print('number of pairs', pairidx.size)
 
-parallax_snr = tgas.parallax_snr
-vtan = tgas.get_vtan().value
-c = tgas.get_coord()
-d = tgas.get_distance().value
+    parallax_snr = tgas.parallax_snr
+    vtan = tgas.get_vtan().value
+    tgas_c = tgas.get_coord()
+    tgas_d = tgas.get_distance().to(u.pc).value
 
-star1, star2 = pairidx['star1'], pairidx['star2']
-min_snr = np.min(np.vstack((parallax_snr[star1], parallax_snr[star2])), axis=0)
-dvtan = np.linalg.norm(vtan[star1]-vtan[star2], axis=1)
-vtanmean = (vtan[star1] + vtan[star2])*0.5
-sep = c[star1].separation_3d(c[star2]).value
-sep_sky = c[star1].separation(c[star2])
+    cond_lr_cut = llr>6
+    print('number of pairs ln(L1/L2)>6 = %i' % (cond_lr_cut.sum()))
+    cmpairs = pairidx[cond_lr_cut]
 
-c1 = c[star1]
-c2 = c[star2]
-ra1, dec1 = c1.ra.value, c1.dec.value
-ra2, dec2 = c2.ra.value, c2.dec.value
[…]

### 2017-02-14 scripts/test_vshuffle.py

@@ -60,6 +60,22 @@ def prepare(tgasfile, outname, snr_cut, radius_cut, radius_cut_vshuf, dvtan_cut)
     vtan = tgas.get_vtan().value
 
     tree = KDTree(xyz)
+
+    Ntarget = 100000
+    batchsize = 100000
+    
+    pairi = np.random.randint(high=len(tgas), size=Ntarget)
+    pairjk = []
+    for i  in pairi:
+        treei = tree.query_radius([xyz[i]], 100)[0]
+        j = np.random.choice(treei)
+        k = np.random.choice(treei)
+        pairjk.append((j,k))
+
+        
+        
+
+
     treei = tree.query_radius(xyz, radius_cut)
     sizes = np.array([x.size for x in treei])
     logger.info("total number of pairs for r<%.1f = %i" % (radius_cut, sizes.sum()))

### 2017-02-24 paper/ms.tex
update catalog; other minor tweaks

@@
     The black dots in the upper right panel and
-    the black line in the lower lef panel correspond to the entire initial sample of pairs.
-    For comparision, we also present the likelihood ratio distribution of random pairs
+    the black line in the lower left panel correspond to the entire initial sample of pairs.
+    For comparison, we also present the likelihood ratio distribution of random pairs
     of stars in the \tgas\ with parallax $S/N>8$ in gray filled histogram.
@@
 \subsection{Catalog of candidate co-moving pairs}
+\label{sub:catalog}
+
+\input{table_meta}
+
+\begin{figure}[htbp]
+  \centering
+  \input{schema.tex}
+  \caption{Schematic diagram of the relationship between the tables.
+    See \sectionname~\ref{sub:catalog} and Table~\ref{tab:table_meta} for details.
+    \label{fig:schema}}
+\end{figure}
 
@@
 For each star, we also include the positional measurements from \tgas,
-\gaia\ $G$-band and \tmass\ $J$, $H$-band magnitudes as well as
+\gaia\ $G$-band and \tmass\ $J$-band magnitudes as well as
 \rave\ radial velocities.
@@
 Finally, the information about the connected components found in these co-moving
-star graph is in the group table. We assign a unique index to each group
+star pairs is in the group table. We assign a unique index to each group
 in descending order of its size. Thus, group 0 is the largest group
@@
 in the group id column of the star and pair table.
-Additionally, given the strong correlation between the larger connected 
-components of co-moving pairs and the known star clusters,
-we provided
-%
-\begin{itemize}
-  \item index and name of the closest cluster in the Milky Way Star Cluster catalog of
-  \citet{Kharchenko:2016aa} using 3D separation
-  \item distance to the closest cluster in pc
-  % \item ID of the OB association (A-L) if the star is in the catalog of
-  % stars in the known OB associations by \citet{de-Zeeuw:1999aa}
-\end{itemize}
-
-% The distances and physical separations are calculated using the point-estimate
-% of distances (\eqname~\ref{eq:dist}).
-
-% \input{stub}
-\input{table_meta}
 
@@
 the future releases of the \gaia\ mission.
-
-\begin{figure}[htb]
-  \centering
-  \input{schema.tex}
-  \caption{Schematic diagram of the relationship between the tables.}
-  \label{fig:schema}
-\end{figure}
 

### 2017-02-24 paper/table_meta.tex
update catalog; other minor tweaks

@@
 \hline
-row id         &      & zero-based row index                      \\
-TGAS source id &      & unique source id from \tgas               \\
-RA             & deg  & right ascension from \tgas                \\
-DEC            & deg  & declination from \tgas                    \\
-parallax       & mas  & unique source id from \tgas               \\
-distance       & pc   & unique source id from \tgas               \\
+row id         &      & Zero-based row index                      \\
+TGAS source id &      & Unique source id from \tgas               \\
+RA             & deg  & Right ascension from \tgas                \\
+DEC            & deg  & Declination from \tgas                    \\
+parallax       & mas  & Unique source id from \tgas               \\
+distance       & pc   & Unique source id from \tgas               \\
 $G$            & mag  & \gaia\ $G$-band magnitudes                \\
 $J$            & mag  & \tmass\ $J$-band magnitudes               \\
-$H$            & mag  & \tmass\ $H$-band magnitudes               \\
-RAVE OBS ID    &      & unique id of the \rave\ match             \\
-RV             & \kms & radial velocity from \rave\               \\
-eRV            & \kms & uncertainty of radial velocity from \rave \\
+RAVE OBS ID    &      & Unique id of the \rave\ match             \\
+RV             & \kms & Radial velocity from \rave\               \\
+eRV            & \kms & Uncertainty of radial velocity from \rave \\
 group id       &      & Id of the group this star belongs to      \\
@@
 \hline
-angsep                           & arcmin & angular separation                  \\
-separation                       & pc     & physical separation                 \\
-$\ln\mathcal{L}_1/\mathcal{L}_2$ &        & likelihood ratio                    \\
-group id                         &        & id of the group the pair belongs to \\
+star 1                           &        & Index of star 1 in the star table   \\
+star 2                           &        & Index of star 2 in the star table   \\
+angsep                           & arcmin & Angular separation                  \\
+separation                       & pc     & Physical separation                 \\
+$\ln\mathcal{L}_1/\mathcal{L}_2$ &        & Likelihood ratio                    \\
+group id                         &        & Id of the group the pair belongs to \\
 \hline
@@
 \hline
-id     &        & unique group id                     \\
-size   &        & number of stars in a group          \\
-       &        &                                     \\
+id            &        & Unique group id                     \\
+size          &        & Number of stars in a group          \\
+mean RA       &  deg   & Mean right ascension of members     \\
+mean DEC      &  deg   & Mean declination of members     \\
+mean distance &  pc    & Mean distance of members     \\
 \hline\hline

### 2017-02-24 scripts/make-catalog.py

@@ -59,8 +59,9 @@ if 'data_loaded' not in dir():
     tgas_rave_hrv = rave['HRV']
     tgas_rave_ehrv = rave['eHRV']
 
-tStar = table.Table()
+tStar = table.Table(masked=True)
 tStar['tgas_row'] = graph.nodes()
+tStar['row_id'] = np.arange(len(tStar))
 tStar['tgas_source_id'] = tgas.source_id[tStar['tgas_row']]
 tStar['tgas_ra'] = tgas.ra.value[tStar['tgas_row']]
 tStar['tgas_dec'] = tgas.dec.value[tStar['tgas_row']]
@@ -70,43 +71,73 @@ tStar['tgas_gmag'] = tgas._data['phot_g_mean_mag'][tStar['tgas_row']]
 tStar['tmass_jmag'] = tmass['j_m'][tStar['tgas_row']]
 tStar['rave_obs_id'] = tgas_rave_obs_id[tStar['tgas_row']]
 tStar['rave_hrv'] = tgas_rave_hrv[tStar['tgas_row']]
+tStar['rave_hrv'].mask = np.isnan(tStar['rave_hrv'])
 tStar['rave_ehrv'] = tgas_rave_ehrv[tStar['tgas_row']]
-tStar['group_id'] =[graph.node.get(n)['group_id'] for n in tStar['tgas_row']]
+tStar['rave_ehrv'].mask = np.isnan(tStar['rave_ehrv'])
+tStar['group_id'] = [graph.node.get(n)['group_id'] for n in tStar['tgas_row']]
 tStar['group_size'] = [len(nx.node_connected_component(graph, node)) for node in tStar['tgas_row']]
 
 star_c = tgas_c[tStar['tgas_row']]
 
 tPair = table.Table()
 edge_tgas_row = np.array(graph.edges())
-edge_star_row = array([list(map(lambda node: where(tStar['tgas_row'] == node)[0][0], [star1, star2])) for star1, star2 in edge_tgas_row])
+edge_star_row = np.array([list(map(lambda node: np.where(tStar['tgas_row'] == node)[0][0], [star1, star2])) for star1, star2 in edge_tgas_row])
 tPair['tgas_row1'] = edge_tgas_row[:,0]
 tPair['tgas_row2'] = edge_tgas_row[:,1]
-tPair['star_row1'] = edge_star_row[:,0]
-tPair['star_row2'] = edge_star_row[:,1]
[…]

### 2017-04-04 paper/ms.tex
Finish up the report
update bib and other tweaks

@@
 
-\include{gitstuff}
 \include{aastexmods}
@@
 meant to encompass young thin disk stars to halo stars.
-These numbers are empirically chosen to account for the distribution of one
-component velocities of the \tgas\ stars.
+These numbers are empirically chosen to account for the distribution of
+velocities of the \tgas\ stars.
 We derive the relevant expressions in Appendix~\ref{sec:appendix}.
@@
 Here, we use 210,368 cross-matches of the \tgas\ with
-Radial Velocity Experiment fifth data release (\rave\ DR~5; \citealt{2016arXiv160903210K}) to
+Radial Velocity Experiment fifth data release (\rave\ DR~5; \citealt{2017AJ....153...75K}) to
 assess the false-positive rates of our selection.
@@
 higher than the mutually exclusive pairs, and remains high ($>80\%$) to $\sim 10$~pc.
+Finally, we note that the false-positive rate for mutually exclusive pairs
+with large angular separation may have been over-estimated due to projection.
 
@@
 
-We have also found a surprisingly large number of very wide separation ($>1$~pc)
+We have also found a large number of very wide separation ($>1$~pc)
 mutually exclusive co-moving pairs, in which the stars are the only co-moving neighbor
 of each other and not part of large connected components.
-These are most likely remnants of dissolving wide binaries.
+These are most likely remnants of dissolving wide binaries (\citealt{Jiang:2010aa}).
+The abundance of highly probable wide separation co-moving pairs
+conclusively shows that there is no strict cut-off semi-major axis for wide binary systems
+(\citealt{Wasserman:1987aa}).
+The presence of these pairs and similar separation distribution have already
+been noticed by \citealt{Shaya:2011aa} using the \project{Hipparcos} data.
 If confirmed with radial velocity measurements, this population should still be
@@
 
-Finally, we make our catalog of 13,058 candidate co-moving pair available to the
+Finally, we make our catalog of 13,058 candidate co-moving pairs available to the
 community.
@@
 \acknowledgements
+
+We thank Scott Tremaine for useful discussions.
+This project was developed in part at the 2016 NYC Gaia Sprint, hosted by the Center for
+Computational Astrophysics at the Flatiron Institute in New York City. The
+Flatiron Institute is supported by the Simons Foundation.
 
@@
 has been provided by national institutions, in particular the institutions
-participating in the {\it Gaia} Multilateral Agreement. This project was
-developed in part at the 2016 NYC Gaia Sprint, hosted by the Center for
-Computational Astrophysics at the Flatiron Institute in New York City. The
-Flatiron Institute is supported by the Simons Foundation.
+participating in the {\it Gaia} Multilateral Agreement.
 
@@
 \url{https://github.com/smoh/gaia-comoving-stars} under the MIT open-source
-software license. This version was generated at git commit
-\texttt{\githash\,(\gitdate)}.
+software license.
 This research additionally utilized:
     \texttt{Astropy} (\citealt{Astropy-Collaboration:2013}),
-    %\texttt{emcee} (\citealt{Foreman-Mackey:2013}),
     \texttt{IPython} (\citealt{Perez:2007}),

### 2017-04-04 paper/ms.tex
clean up appendix

@@
   p(\vec{v}) = \sum_{m=1}^3 \, w_m \, \mathcal{N}(0, \sigma_{v,m}^2)
+  \label{eq:vprior}
 \end{equation}
@@
 conclusively shows that there is no strict cut-off semi-major axis for wide binary systems
-(\citealt{Wasserman:1987aa}).
+(e.g., \citealt{Wasserman:1987aa}).
 The presence of these pairs and similar separation distribution have already
@@
 \end{eqnarray}
-where $\mat{B}\,\vec{b}$ is an $D$-vector, and $C$ is a scalar, then we find
+where $\mat{B}\,\vec{b}$ is a $D$-vector, and $C$ is a scalar, then we find
 \begin{eqnarray}
-  \vec{\nu} &=& -\mat{A}\,\mat{B}\,\vec{b}
+  \vec{\nu} &=& -\mat{A}\,\mat{B}\,\vec{b} \label{eq:convnu}
   \\
-  \Delta & = & C - \frac{1}{2}\,\transp{\vec{\nu}}\,\inv{\mat{A}}\,\vec{\nu}
+  \Delta & = & C - \frac{1}{2}\,\transp{\vec{\nu}}\,\inv{\mat{A}}\,\vec{\nu} \label{eq:convDelta}
   \quad .
@@
 analytically using the expressions in Appendix~\ref{sec:appendixA}.
-As described above, we use an isotropic, mixture-of-Gaussians prior on velocity
-where the velocity dispersion of a given mixture component is $\sigma_{v,m}$ and
-the 3-space variance tensor is $\mat{V}_m = \sigma_{v,m}^2 \, \eye$.
-The likelihood for the data is also a Gaussian, as shown in
-\eqname~\ref{eq:likefn}.
 We will start by writing down expressions for the the likelihood multiplied by
 the prior pdf for the velocities.
-Here we slightly change the notation used in \sectionname~\ref{sec:methods} for
-simplicity.
-
-We construct a velocity-space data vector $y$ as follows:
+The likelihood for the data (proper motions of the two stars in a pair)
+is a Gaussian (\eqname~\ref{eq:likefn}).
+In order to simplify our notation, 
+we construct a velocity-space data vector $\vec{y}$ as follows:
 \begin{equation}
@@
 \end{equation}
-where again the subscripts $i,j$ refer to the indices of each star in the pair
+where the subscripts $i,j$ refer to the indices of each star in the pair
 and we have multiplied the observables (the proper motions) by the distances
@@
 and using \eqname~\ref{eq:transformation}.
-The noise (in the $y$ vector) is drawn from a $4 \times 4$ Gaussian with
-block-diagonal covariance matrix, $\mat{\Sigma}$, constructed from the
-covariance matrix of the observables, $\mat{C}_i, \mat{C}_j$, and the distances
+The noise (in $\vec{y}$) is drawn from a $4 \times 4$ Gaussian with
+block-diagonal covariance matrix, $\mat{\Sigma}$, constructed from
+the proper motion covariance matrix of each star,
+$\mat{C}_i, \mat{C}_j$, and the distances
 $r_i, r_j$:
@@
 
-Given all these definitions, the likelihood function for hypothesis 1 is
+Given these definitions, the likelihood function for hypothesis 1 is
 \begin{eqnarray}
@@
 \end{eqnarray}
-where the factor of $r_i^2\,r_j^2$ converts to units of one over \gaia\ data
-from units of one over $y$, which includes the distances.
-
-Now we multiply this likelihood with the velocity prior (here shown only for one
-prior velocity mixture component) and complete the square.
-In the original parameterization, we find:
+where the factor of $r_i^2\,r_j^2$ is the Jacobian from $\vec{y}$
+to data ($\transp{(\mu_{\alpha,i}\ \,\mu_{\delta,i}\ \,\mu_{\alpha,j}\ \,\mu_{\delta,j})}$) space.
+
+Now we multiply this likelihood with the velocity prior.
+As described in \sectionname~\ref{sec:methods}
+we use an isotropic, mixture-of-Gaussians prior on velocity (\eqname~\ref{eq:vprior}).
+For simplicity here let us work out the marginalization for one component of
+the mixture so that $\vec{v} \sim \normal(\vec{0}, \mat{V}_m)$.
+Then,
+
+\begin{eqnarray}
+  \ln p(\vec{v}) = -\frac{1}{2}\,\ln||2\pi\,\mat{V}_m||
+  -\frac{1}{2} \transp{\vec{v}}\,\inv{\mat{V}_m}\,\vec{v}
+\end{eqnarray}
+We can identify $\mat{A}$, $\mat{B}\,\vec{b}$, and $C$ in
+$\ln p(\vec{v}) + \ln p(\data \given \vec{v}, r_i, r_j)$, and write down
+$\vec{\nu}$ and $\Delta$ using \eqname~\ref{eq:convnu} and \ref{eq:convDelta} as
+
 \begin{eqnarray}
@@
 \end{eqnarray}
-and $\mat{M}$ is now the transformation matrix for one star. Then, again,
+and $\mat{M}$ is now the transformation matrix for one star. Then,
 \begin{eqnarray}
@@
   \\
-  \Delta &=& -2\,\ln r_i -2\,\ln r_j
+  \Delta &=& -2\,\ln r
     +\frac{1}{2}\,\ln||2\pi\,\mat{\Sigma}|| +\frac{1}{2}\,\ln||2\pi\,\mat{V}_m|| \nonumber \\

### 2017-04-04 paper/ms.tex
simplify statement

@@
 \end{eqnarray}
-We can identify $\mat{A}$, $\mat{B}\,\vec{b}$, and $C$ in
-$\ln p(\vec{v}) + \ln p(\data \given \vec{v}, r_i, r_j)$, and write down
-$\vec{\nu}$ and $\Delta$ using \eqname~\ref{eq:convnu} and \ref{eq:convDelta} as
+We can identify $\vec{\nu}$ and $\Delta$ in
+$\ln p(\vec{v}) + \ln p(\data \given \vec{v}, r_i, r_j)$
+using \eqname~\ref{eq:convnu} and \ref{eq:convDelta} as
 

### 2017-04-04 paper/ms.tex
minor tweaks and rule for diff pdf

@@
 
-  \item Finally, there is a population of “random” pairs of field stars
+  \item Finally, there is a population of ``random'' pairs of field stars
     that are not co-moving, but still have $\absdvtan < 10$~\kms\
@@
     Figure~\ref{fig:dv-sep} shows that there is an overlap between
-    genuine co-moving pairs and “random” pairs.
+    genuine co-moving pairs and ``random'' pairs.
 \end{itemize}
@@
   \item We assume that the uncertainties in the data---parallax, $\varpi$, and
-    two proper motion components, $\propm = (\begin{array}[t]{c c} \mu_{\alpha^*} &
-    \mu_\delta\end{array})^\mathsf{T}$---are Gaussian with known covariances
+    two proper motion components, $\propm = \transp{(\mu_{\alpha^*}\, \mu_\delta)}$
+    ---are Gaussian with known covariances
     $\mat{C}$. The values of covariances are provided as part of the \gaia\ data
@@
 We also list the angular and physical separation of pairs, and the likelihood ratio,
-$\ln \mathcal{L}_1 /\mathcal{L}_2>6$ (see \eqname~\ref{eq:hyp1} and \ref{eq:hyp2})
+$\ln \mathcal{L}_1 /\mathcal{L}_2(>6)$ (see \eqname~\ref{eq:hyp1} and \ref{eq:hyp2})
 computed in this work.

### 2017-04-07 paper/ms.tex
some minor revisions to the text

@@
 where $\vec\mu = (\mu_{\alpha^*}, \mu_\delta)$.
-\footnote{$\mu_{\alpha^*}$ is proper motion
-in right ascension direction, $\mu_{\alpha^*} = \mu_\alpha \cos\delta$}
+\footnote{$\mu_{\alpha^*}$ is the proper motion component
+in the right ascension direction, $\mu_{\alpha^*} = \mu_\alpha \cos\delta$}
 
@@
 the relationships between the tables in Figure~\ref{fig:schema}.
-The star table contains all 10,606 stars that has at least one co-moving neighbor
+The star table contains all 10,606 stars that have at least one co-moving neighbor
 by our selection. We provide the TGAS source id, which may be used to easily
-retrieve cross matches of \gaia\ with other surveys
+retrieve cross matches between \gaia\ and other surveys
 using the \gaia\ data archive.
 For each star, we also include the positional measurements from \tgas,
-\gaia\ $G$-band and \tmass\ $J$-band magnitudes as well as
-\rave\ radial velocities.
+\gaia\ $G$-band, \tmass\ $J$-band magnitudes, and
+\rave\ radial velocities where they exist.
 The co-moving relationship between the stars is described in the pair table.
-We also list the angular and physical separation of pairs, and the likelihood ratio,
+We also list the angular and physical separation of each pair, and the likelihood ratio,
 $\ln \mathcal{L}_1 /\mathcal{L}_2(>6)$ (see \eqname~\ref{eq:hyp1} and \ref{eq:hyp2})
@@
 \end{eqnarray}
-where the factor of $r_i^2\,r_j^2$ is the Jacobian from $\vec{y}$
-to data ($\transp{(\mu_{\alpha,i}\ \,\mu_{\delta,i}\ \,\mu_{\alpha,j}\ \,\mu_{\delta,j})}$) space.
+where the factor of $r_i^2\,r_j^2$ is the Jacobian of the transformation from $\vec{y}$
+to the data space ($\transp{(\mu_{\alpha,i}\ \,\mu_{\delta,i}\ \,\mu_{\alpha,j}\ \,\mu_{\delta,j})}$).
 
@@
       \begin{array}{c c}
-        r\,\mu_{\alpha} &
+        r\,\mu_{\alpha} & \quad
         r\,\mu_{\delta}

### 2017-04-18 paper/ms.tex
consistent notation for pmra

@@
       \begin{array}{c c c c}
-        r_i\,\mu_{\alpha,i} &
+        r_i\,\mu_{\alpha,i}^* &
         r_i\,\mu_{\delta,i} &
-        r_j\,\mu_{\alpha,j} &
+        r_j\,\mu_{\alpha,j}^* &
         r_j\,\mu_{\delta,j}
@@
 where the factor of $r_i^2\,r_j^2$ is the Jacobian of the transformation from $\vec{y}$
-to the data space ($\transp{(\mu_{\alpha,i}\ \,\mu_{\delta,i}\ \,\mu_{\alpha,j}\ \,\mu_{\delta,j})}$).
+to the data space ($\transp{(\mu_{\alpha,i}^*\ \,\mu_{\delta,i}\ \,\mu_{\alpha,j}^*\ \,\mu_{\delta,j})}$).
 
@@
       \begin{array}{c c}
-        r\,\mu_{\alpha} & \quad
+        r\,\mu_{\alpha}^* & \quad
         r\,\mu_{\delta}

### 2017-04-18 scripts/make-catalog.py
set reasonable digits on rave rvs

@@ -63,17 +63,17 @@ tStar = table.Table(masked=True)
 tStar['tgas_row'] = graph.nodes()
 tStar['row_id'] = np.arange(len(tStar))
 tStar['tgas_source_id'] = tgas.source_id[tStar['tgas_row']]
-tStar['tgas_ra'] = tgas.ra.value[tStar['tgas_row']]
-tStar['tgas_dec'] = tgas.dec.value[tStar['tgas_row']]
-tStar['tgas_parallax'] = tgas.parallax.value[tStar['tgas_row']]
-tStar['tgas_distance'] = tgas_d[tStar['tgas_row']]
-tStar['tgas_gmag'] = tgas._data['phot_g_mean_mag'][tStar['tgas_row']]
-tStar['tmass_jmag'] = tmass['j_m'][tStar['tgas_row']]
+tStar['ra'] = tgas.ra.value[tStar['tgas_row']]
+tStar['dec'] = tgas.dec.value[tStar['tgas_row']]
+tStar['parallax'] = tgas.parallax.value[tStar['tgas_row']]
+tStar['distance'] = tgas_d[tStar['tgas_row']]
+tStar['G'] = tgas._data['phot_g_mean_mag'][tStar['tgas_row']]
+tStar['J'] = tmass['j_m'][tStar['tgas_row']]
 tStar['rave_obs_id'] = tgas_rave_obs_id[tStar['tgas_row']]
-tStar['rave_hrv'] = tgas_rave_hrv[tStar['tgas_row']]
-tStar['rave_hrv'].mask = np.isnan(tStar['rave_hrv'])
-tStar['rave_ehrv'] = tgas_rave_ehrv[tStar['tgas_row']]
-tStar['rave_ehrv'].mask = np.isnan(tStar['rave_ehrv'])
+tStar['rv'] = tgas_rave_hrv[tStar['tgas_row']]
+tStar['rv'].mask = np.isnan(tStar['rv'])
+tStar['erv'] = tgas_rave_ehrv[tStar['tgas_row']]
+tStar['erv'].mask = np.isnan(tStar['erv'])
 tStar['group_id'] = [graph.node.get(n)['group_id'] for n in tStar['tgas_row']]
 tStar['group_size'] = [len(nx.node_connected_component(graph, node)) for node in tStar['tgas_row']]
 
@@ -96,9 +96,9 @@ tPair['group_size'] = tStar['group_size'][tPair['star1']]
 tGroup = table.Table()
[…]

### 2017-04-18 scripts/make-catalog.py
add name column

@@ -63,6 +63,8 @@ tStar = table.Table(masked=True)
 tStar['tgas_row'] = graph.nodes()
 tStar['row_id'] = np.arange(len(tStar))
 tStar['tgas_source_id'] = tgas.source_id[tStar['tgas_row']]
+tStar['name'] = ['HIP {:d}'.format(hip) if hip != 999999 else 'TYC {:s}'.format(tyc)
+        for hip, tyc in zip(tgas.hip[tStar['tgas_row']], tgas.tycho2_id[tStar['tgas_row']])]
 tStar['ra'] = tgas.ra.value[tStar['tgas_row']]
 tStar['dec'] = tgas.dec.value[tStar['tgas_row']]
 tStar['parallax'] = tgas.parallax.value[tStar['tgas_row']]

### 2017-04-18 paper/table_meta.tex
put in new columns and make schema standalone

@@
 TGAS source id &      & Unique source id from \tgas               \\
+Name           &      & \project{Hipparcos} or \project{Tycho-2} identifier \\
 RA             & deg  & Right ascension from \tgas                \\
@@
 group id       &      & Id of the group this star belongs to      \\
+group size     &      & Size of the group this star belongs to    \\
 \hline
@@
 group id                         &        & Id of the group the pair belongs to \\
+group size                       &        & Size of the group the pair belongs to \\
 \hline

### 2017-06-01 paper/ms.tex
put in new columns and make schema standalone
add flatiron
remove sentence
minor punctuations edit from the proof

@@
 
-\title{Co-moving stars in \textsl{Gaia DR1}:\\
-  An abundance of very wide separation co-moving pairs}
+\title{Comoving stars in \textsl{Gaia DR1}:\\
+  An abundance of very wide separation comoving pairs}
 \author{Semyeong Oh\altaffilmark{\pu,\lead},
@@
                      K\"onigstuhl 17, D-69117 Heidelberg, Germany}
-\altaffiltext{\cca}{Center for Computational Astrophysics, 162 5th Ave, New York, NY 10003, USA}
+\altaffiltext{\cca}{Center for Computational Astrophysics, Flatiron Institute,
+                    162 5th Ave, New York, NY 10003, USA}
 
@@
 stars and moving groups.
-We search for high-confidence co-moving pairs of stars in \tgas\ by identifying
+We search for high-confidence comoving pairs of stars in \tgas\ by identifying
 pairs of stars consistent with having the same 3D velocity using a marginalized
-likelihood ratio test to discriminate candidate co-moving pairs from the field
+likelihood ratio test to discriminate candidate comoving pairs from the field
 population.
-Although we perform some visualizations using (bias-corrected) inverse-parallax
-as a point-estimate of distance, the likelihood ratio is computed with a
-probabilistic model that includes the covariances of parallax and proper motions,
+Although we perform some visualizations using (bias-corrected) inverse parallax
+as a point estimate of distance, the likelihood ratio is computed with a
+probabilistic model that includes the covariances of parallax and proper motions
 and marginalizes the (unknown) true distances and 3D velocities of the stars.
-We find 13,085 co-moving star pairs among 10,606 unique stars with separations
+We find 13,085 comoving star pairs among 10,606 unique stars with separations
 as large as 10 pc (our search limit).
-Some of these pairs form larger groups through mutual co-moving neighbors:
+Some of these pairs form larger groups through mutual comoving neighbors:
 many of these pair networks correspond to known open clusters and OB
-associations, but we also report the discovery of several new co-moving groups.
+associations, but we also report the discovery of several new comoving groups.
 Most surprisingly, we find a large number of very wide ($>1$ pc) separation
-co-moving star pairs, the number of which increases with increasing separation
+comoving star pairs, the number of which increases with increasing separation
 and cannot be explained purely by false-positive contamination.
-Our key result is a catalog of high-confidence co-moving pairs of stars in
+Our key result is a catalog of high-confidence comoving pairs of stars in
 \tgas.
@@
 Stars that are roughly co-located and moving with similar space velocities
-(``co-moving stars'') are of special interest in many branches of astrophysics.
+(``comoving stars'') are of special interest in many branches of astrophysics.
 
@@
 
-Wide binaries are also good testbeds for stellar models and age indicators:
+Wide binaries are also good test beds for stellar models and age indicators:
 the constituent stars were likely born at the same time with the same chemical
@@
 which dominate the stellar content of the Galaxy by number, can benefit from
-a larger sample of widely separated binaries containing a low mass star and a
+a larger sample of widely separated binaries containing a low-mass star and a
 much brighter F/G/K star whose stellar parameters are easier to measure
@@
 
-At larger separations ($\gtrsim 1$~pc), co-moving stars are likely members of
+At larger separations ($\gtrsim 1$~pc), comoving stars are likely members of
 (potentially dissolving) moving groups, associations, and star clusters or
@@
 star formation event with similar chemical composition? Or are they formed by
-dynamical effects of non-axisymmetric features of the Galaxy such as spirals
+dynamical effects of nonaxisymmetric features of the Galaxy such as spirals
 and bars? With the recent advances in measuring chemical abundances of
-a large volume of stars using high and low resolution spectroscopy
+a large volume of stars using high- and low- resolution spectroscopy
 (e.g., \citealt{Steinmetz:2006aa,Majewski:2015aa,Gilmore:2012aa} to name a few),
@@
 
-To date, thousands of candidate co-moving star pairs have been identified by
+To date, thousands of candidate comoving star pairs have been identified by
 searching for stars with common proper motions
@@
 In \sectionname~\ref{sec:methods}, we develop a statistical method to identify
-high-confidence co-moving pairs in this catalog.
+high-confidence comoving pairs in this catalog.
 In \sectionname~\ref{sec:results}, we present and discuss our resulting catalog
-of co-moving pairs.
+of comoving pairs.
 We summarize in \sectionname~\ref{sec:summary}.
@@
 
-We construct an initial sample of star pairs to search for co-moving pairs as
+We construct an initial sample of star pairs to search for comoving pairs as
 follows.
@@
 Then, for each star we establish an initial sample of possible
-co-moving partners by selecting all other stars that satisfy two criteria:
+comoving partners by selecting all other stars that satisfy two criteria:
 separation less than 10~pc and difference
@@
     pairs are very close in both 3D position and tangential velocities, it is highly
-    probable that they are actually co-moving wide binaries.
-
-  \item  A sample of co-moving stars also include stars that are part of, e.g.,
+    probable that they are actually comoving wide binaries.
+
+  \item  A sample of comoving stars also include stars that are part of, e.g.,
     OB associations, moving groups, and open clusters.
-    These astrophysical objects may be detected as a network of co-moving pairs,
-    sharing some mutual co-moving neighbors.
-    As the pair separation increases, the nature of co-moving pairs
+    These astrophysical objects may be detected as a network of comoving pairs,
+    sharing some mutual comoving neighbors.
+    As the pair separation increases, the nature of comoving pairs
     will change from binaries to those related to these larger objects,
@@
     velocities will be.
-    Due to this projection effect, a population of genuine co-moving pairs
+    Due to this projection effect, a population of genuine comoving pairs
     will extend to larger \absdvtan\ at larger separation.
@@
   \item Finally, there is a population of ``random'' pairs of field stars
-    that are not co-moving, but still have $\absdvtan < 10$~\kms\
+    that are not comoving, but still have $\absdvtan < 10$~\kms\
     by chance.
@@
     Figure~\ref{fig:dv-sep} shows that there is an overlap between
-    genuine co-moving pairs and ``random'' pairs.
+    genuine comoving pairs and ``random'' pairs.
 \end{itemize}
@@
 the non-trivial uncertainties in the data to our beliefs about the likelihood
-that a given pair of stars is co-moving.
+that a given pair of stars is comoving.
 
@@
 \figname~\ref{fig:dv-sep} suggests that there are a
-significant number of co-moving pairs in the \tgas\ data at a range
+significant number of comoving pairs in the \tgas\ data at a range
 of separations.
-Here, we develop a method to select high-confidence co-moving
+Here, we develop a method to select high-confidence comoving
 pairs that properly incorporates the uncertainties associated with the
@@
 We use the FML ratio $\mathcal{L}_1/\mathcal{L}_2$ as the scalar quantity to select
-candidate co-moving pairs, as described in $\sectionname~\ref{sub:selection}$ in
+candidate comoving pairs, as described in $\sectionname~\ref{sub:selection}$ in
 more detail.
@@
 This section is divided into three parts. First, we discuss and justify a cut of
-the likelihood ratio to select candidate co-moving pairs.
-Second, we present the statistics and properties of our candidate co-moving
+the likelihood ratio to select candidate comoving pairs.
+Second, we present the statistics and properties of our candidate comoving
 pairs.
-Finally, we describe our catalog of candidate co-moving pairs, the main product
+Finally, we describe our catalog of candidate comoving pairs, the main product
 of this study.
 
-\subsection{Selecting candidate co-moving pairs}
+\subsection{Selecting candidate comoving pairs}
 \label{sub:selection}
@@
 $\ln \mathcal{L}_1 /\mathcal{L}_2$, and come up with a reasonable cut
-for this quantity to select co-moving pairs from the initial sample.
+for this quantity to select comoving pairs from the initial sample.
 
@@
     Based on this, we choose $\ln \mathcal{L}_1 /\mathcal{L}_2 > 6$ as our high-confidence
-    candidate co-moving pairs.
+    candidate comoving pairs.
     \label{fig:likelihoodratios}}
@@
 Specifically, as we sweep through from small \absdvtan\ to large, we expect
-the population of pairs to change from genuinely co-moving to random.
+the population of pairs to change from genuinely comoving to random.
 This becomes clear when we look at the distribution of the likelihood ratios of pairs
 in slices of \absdvtan\ (top row of Figure~\ref{fig:likelihoodratios}).
-Pairs with $\absdvtan \in (0,2.5)~\kms$ are most likely actual co-moving pairs, and
+Pairs with $\absdvtan \in (0,2.5)~\kms$ are most likely actual comoving pairs, and
 their likelihood ratio distribution is narrowly peaked at $\gtrsim 5$
@@
 this distribution peaks at a much lower value ($\approx 2$),
-and is clearly separated from highly probable co-moving pairs.
-
-Based on these comparisons, we select candidate co-moving pairs with
+and is clearly separated from highly probable comoving pairs.
+
+Based on these comparisons, we select candidate comoving pairs with
 $\ln \mathcal{L}_1 /\mathcal{L}_2 > 6$.
@@
 
-\subsection{Statistics and properties of the identified co-moving pairs}
+\subsection{Statistics and properties of the identified comoving pairs}
 
@@
   \caption{%
-    Visualizations of a few example connected components of co-moving pairs of stars.
+    Visualizations of a few example connected components of comoving pairs of stars.
     Each star (node) is marked as a red circle, and a line (edge) is drawn
-    between two stars if they are co-moving by our selection criteria
+    between two stars if they are comoving by our selection criteria
     (see \sectionname~\ref{sub:selection}). On left, we show the largest network
@@
   \caption{%
-    Panoramic view of co-moving pairs of stars around the Sun. This is a cylindrical
+    Panoramic view of comoving pairs of stars around the Sun. This is a cylindrical
     projection of onto the Galactic plane with the Sun at the origin.
@@
     Some of the larger connected components are clearly associated with known
-    clusters, but we also discover quite a few new co-moving groups.
+    clusters, but we also discover quite a few new comoving groups.
     \label{fig:glon_d_pairlines}}
@@
 
-Once we have identified candidate co-moving pairs from the initial
+Once we have identified candidate comoving pairs from the initial
 sample, these pairs form an undirected graph where stars are nodes, and edges
-between the nodes exist for co-moving pairs of stars.
-A star may have multiple co-moving neighbors, and two stars may either
+between the nodes exist for comoving pairs of stars.
+A star may have multiple comoving neighbors, and two stars may either
 be directly or indirectly connected by a sequence of edges (``path'').
@@
 and show the distribution of their sizes in Figure~\ref{fig:hist_ccsize}.
-The most common are connected components of size 2, which mean mutually exclusive co-moving
-pairs. However, it is clear that there are many aggregates of co-moving stars
-discovered by looking for co-moving pairs.
+The most common are connected components of size 2, which mean mutually exclusive comoving
+pairs. However, it is clear that there are many aggregates of comoving stars
+discovered by looking for comoving pairs.
 These aggregates are likely moving groups, OB associations, or star clusters.
-In 13,058 co-moving pairs that we identified,
+In 13,058 comoving pairs that we identified,
 there are 4,555 connected components among 10,606 unique stars.
@@
 
-We show the distribution of co-moving pairs in galactic longitude and
+We show the distribution of comoving pairs in galactic longitude and
 distance in Figure~\ref{fig:glon_d_pairlines}.
-The connection between known co-moving structures and
+The connection between known comoving structures and
 the larger connected components found in this work becomes immediately clear when
@@
 to name a few.
-Clumps of co-moving pairs at $(l,d)\approx (300-360\arcdeg, 100-200~\textrm{pc})$
+Clumps of comoving pairs at $(l,d)\approx (300-360\arcdeg, 100-200~\textrm{pc})$
 seem to strongly correlate with the locations of OB associations
@@
 
-Ultimately, any candidate co-moving pair found in this work
+Ultimately, any candidate comoving pair found in this work
 needs to be verified using radial velocities.
@@
 Given the excellent correspondence between
-aggregates of co-moving pairs (connected components)
-and known genuine co-moving structures (Figure~\ref{fig:glon_d_pairlines}),
+aggregates of comoving pairs (connected components)
+and known genuine comoving structures (Figure~\ref{fig:glon_d_pairlines}),
 we may expect that pairs in these larger connected components,
@@
 to have less contamination from false-positives.
-We divide the co-moving pairs into those
+We divide the comoving pairs into those
 mutually exclusively connected (i.e., in a connected component of
@@
   \caption{%
-    Validation of candidate co-moving pairs using radial velocities from \rave.
+    Validation of candidate comoving pairs using radial velocities from \rave.
     Top: Radial velocity differences, $\Delta v_r$, of 283 pairs with \rave\ measurements.
@@
 
-We now examine the separation distribution of co-moving pairs in
+We now examine the separation distribution of comoving pairs in
 Figure~\ref{fig:hist_separation}.
@@
 found with separations larger than 1~pc.
-Surprisingly, however, we also find a large number of mutually exclusive co-moving
+Surprisingly, however, we also find a large number of mutually exclusive comoving
 pairs
@@
   \caption{%
-    Separation distribution of co-moving pairs of stars.
+    Separation distribution of comoving pairs of stars.
     As with Figure~\ref{fig:raverv},
@@
 
-Finally, we present the color-magnitude diagrams of co-moving pairs
+Finally, we present the color-magnitude diagrams of comoving pairs
 using the cross-matches with \tmass.
@@
 in Figure~\ref{fig:graphviz_examples}.
-For stars in larger co-moving groups, there is a noticeable
+For stars in larger comoving groups, there is a noticeable
 lack of evolved, off-main sequence
@@
     We show a reference distribution of stars randomly drawn to have
-    a matching distribution in distance as the stars in our co-moving pairs
+    a matching distribution in distance as the stars in our comoving pairs
     in blue.
@@
   \caption{
-    Color-magnitude diagrams of stars in mutually exclusive co-moving pairs
+    Color-magnitude diagrams of stars in mutually exclusive comoving pairs
     in two separation bins.
@@
     We show a reference distribution of stars randomly drawn to have
-    a matching distribution in distance as the stars in our co-moving pairs
+    a matching distribution in distance as the stars in our comoving pairs
     in blue, same as Figure~\ref{fig:cmd_large}.
@@
 
-\subsection{Catalog of candidate co-moving pairs}
+\subsection{Catalog of candidate comoving pairs}
 \label{sub:catalog}
-
-\input{table_meta}
 
@@
   \centering
-  \input{schema.tex}
+  \includegraphics{schema.pdf}
   \caption{Schematic diagram of the relationship between the tables.
@@
 
-In this section, we describe our catalog of candidate co-moving pairs of stars.
-For 13,058 pairs (10,606 unique stars) found in this work.
+\input{table_meta}
+
+In this section, we describe our catalog of candidate comoving pairs of stars.
 The catalog is composed of three tables of stars, pairs, and groups.
@@
 the relationships between the tables in Figure~\ref{fig:schema}.
-The star table contains all 10,606 stars that have at least one co-moving neighbor
+The star table contains all 10,606 stars that have at least one comoving neighbor
 by our selection. We provide the TGAS source id, which may be used to easily
@@
 \rave\ radial velocities where they exist.
-The co-moving relationship between the stars is described in the pair table.
+The comoving relationship between the stars is described in the pair table.
 We also list the angular and physical separation of each pair, and the likelihood ratio,
@@
 
-Finally, the information about the connected components found in these co-moving
+Finally, the information about the connected components found in these comoving
 star pairs is in the group table. We assign a unique index to each group
@@
 
-We note a caveat on the completeness of a connected component of co-moving
+We note a caveat on the completeness of a connected component of comoving
 pairs found in our catalog. Because we applied a simple cut in the likelihood
@@
 example, a star in a mutually exclusive pair in our catalog may still have
-another possibly co-moving companion which has been dropped because the
+another possibly comoving companion which has been dropped because the
 likelihood ratio is slightly less than 6.
@@
 
-In this \documentname, we searched for co-moving pairs of stars in the
+In this \documentname, we searched for comoving pairs of stars in the
 \tgas\ catalog released as part of the \gaia\ DR1.
@@
 We argued for a reasonable cut of the likelihood ratio, and found
-13,058 candidate co-moving pairs among 10,606 stars
+13,058 candidate comoving pairs among 10,606 stars
 with separations ranging from 0.005~pc to 10~pc, the limit of our search.
 
-We found that some co-moving pairs that we have identified are connected
-by sharing a common co-moving neighbor.
-This network of co-moving pairs, which forms an undirected graph,
+We found that some comoving pairs that we have identified are connected
+by sharing a common comoving neighbor.
+This network of comoving pairs, which forms an undirected graph,
 can be decomposed into connected components in which any two stars are connected
 by a path.
-The entire 13,058 candidate co-moving pairs are grouped into 4,555 connected
+The entire 13,058 candidate comoving pairs are grouped into 4,555 connected
 components. The most common is a size-2 connected component, i.e., the two
@@
 Many of the larger connected components
-naturally correspond to some of the known co-moving structures
+naturally correspond to some of the known comoving structures
 such as open clusters and stellar associations.
-Some of these co-moving groups of stars are newly discovered.
+Some of these comoving groups of stars are newly discovered.
 
 We have also found a large number of very wide separation ($>1$~pc)
-mutually exclusive co-moving pairs, in which the stars are the only co-moving neighbor
+mutually exclusive comoving pairs, in which the stars are the only comoving neighbor
 of each other and not part of large connected components.
 These are most likely remnants of dissolving wide binaries (\citealt{Jiang:2010aa}).
-The abundance of highly probable wide separation co-moving pairs
+The abundance of highly probable wide separation comoving pairs
 conclusively shows that there is no strict cut-off semi-major axis for wide binary systems
@@
 If they are remnants of dissolving systems that were born coeval,
-the sample of very wide separation co-moving pairs can potentially be used
+the sample of very wide separation comoving pairs can potentially be used
 to measure the recent ($\lesssim 1$~Gyr)
 star formation history in the Solar neighborhood.
-Co-moving stars with separation less than 1~pc are very promising candidates
+Comoving stars with separation less than 1~pc are very promising candidates
 for wide binaries. They are found to be pairs of stars of varying stellar types.
@@
 
-Finally, we make our catalog of 13,058 candidate co-moving pairs available to the
+Finally, we make our catalog of 13,058 candidate comoving pairs available to the
 community.
